### 2025-04-03 MAIN.tex
[MGa] basic structure

% interactnlmsample.tex
% v1.05 - August 2017

\documentclass[]{interact}
\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
\usepackage[caption=false]{subfig}% Support for small, `sub' figures and tables
\usepackage[overload]{empheq}
%\usepackage[nolists,tablesfirst]{endfloat}% To `separate' figures and tables from text if required
%\usepackage[doublespacing]{setspace}% To produce a `double spaced' document if required
%\setlength\parindent{24pt}% To increase paragraph indentation when line spacing is doubled

\usepackage[numbers,sort&compress]{natbib}% Citation support using natbib.sty
\bibpunct[, ]{[}{]}{,}{n}{,}{,}% Citation support using natbib.sty
\renewcommand\bibfont{\fontsize{10}{12}\selectfont}% Bibliography support using natbib.sty
\makeatletter% @ becomes a letter
\def\NAT@def@citea{\def\@citea{\NAT@separator}}% Suppress spaces between citations using natbib.sty
\makeatother% @ becomes a symbol again

\theoremstyle{plain}% Theorem-like structures provided by amsthm.sty
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newtheorem{notation}{Notation}

\usepackage{setspace}
%\usepackage[latin1]{inputenc}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{float}
\usepackage{siunitx}
%\usepackage{units}
\usepackage{listings}
\usepackage{pdfpages}
\usepackage{keyval}
\usepackage{eso-pic}
\usepackage{atbegshi}
\usepackage{pdflscape}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{url}
\usepackage{array}
\usepackage{fancyhdr}

\usepackage[np]{numprint}
\npdecimalsign{\ensuremath{.}}

\usepackage{caption}

\captionsetup[figure]{labelfont={bf},name={Fig.},labelsep=space}
\captionsetup[table]{labelfont={bf},name={Table},labelsep=space}

\newcommand {\fncyblank }{\fancyhf{}}

\newcommand{\cedpage}{\newpage{\thispagestyle{empty}\cleardoublepage}}
\usepackage{algorithm,algpseudocode}
\usepackage{tikz,tikz-3dplot}\usetikzlibrary{arrows.meta,angles}

\DeclareMathOperator{\mf}{F_{\text{MF}}}

%\usepackage{hyperref}

%%% Local definitions %%%%%%
\input{mymacros}
%%% End Local definitions %%

\begin{document}

%\articletype{ARTICLE TEMPLATE}% Specify the article type or omit as appropriate

\title{Disturbance-aware minimum-time planning strategies for motorsport vehicles with probabilistic safety certificates}

\author{Martino Gulisano\textsuperscript{a}, Matteo Masoni\textsuperscript{a}, and \name{Marco Gabiccini,\textsuperscript{a,*}
\thanks{* Contact: Marco Gabiccini. Email: marco.gabiccini@unipi.it}
}\affil{
\textsuperscript{a} Dipartimento di Ingegneria Civile e Industriale,	Universit\`{a} di Pisa, Pisa, Italy.}
}

\maketitle

\begin{abstract}
This paper analyzes ...
\end{abstract}

\begin{keywords}
Minimum lap-time trajectory planning; stochastic vehicle dynamics; probabilistic safety certificates.
\end{keywords}

%\tableofcontents
\input{intro}
\input{stochastic_vehicle_dynamics}
\input{open_loop_planning}
%for back-off calculations
\input{closed_loop_planning}
\input{performance_comparison}
\input{conclusions}

\section*{Acknowledgement(s)}

The authors would like to thank Dr. Lorenzo Bartali for his significant contributions to the initial version of the optimization framework developed and utilized in this paper.

\section*{Disclosure statement}

No potential conflict of interest was reported by the author(s).

\section*{Funding}

This work is supported by project PRIN 2022 PNRR ``Global Stability of road vehicle motion - STAVE'' CUP I53D23005670001.

%\section*{Notes on contributor(s)}
%
%An unnumbered section, e.g.\ \verb"\section*{Notes on contributors}", may be included \emph{in the non-anonymous version} if required. A photograph may be added if requested.
%
%
%\section*{Nomenclature/Notation}
%
%An unnumbered section, e.g.\ \verb"\section*{Nomenclature}" (or \verb"\section*{Notation}"), may be included if required, before any Notes or References.
%
%
%\section*{Notes}
%
%An unnumbered `Notes' section may be included before the References (if using the \verb"endnotes" package, use the command \verb"\theendnotes" where the notes are to appear, instead of creating a \verb"\section*").

\bibliographystyle{tfnlm}
\bibliography{biblio}

%\bigskip
%
%
%\section{Appendices}
%
%\input{Appendices}

%\section{Troubleshooting}

%\section{Obtaining the template and class file}
%
%\subsection{Via the Taylor \& Francis website}
%
%...
%
%
%\subsection{Via e-mail}

\end{document}

### 2025-04-03 closed_loop_planning.tex
[MGa] basic structure

\section{Closed-loop uncertainty-aware planning over the full horizon}
\label{sec:closed_loop_planning}

Building upon our previous paper that presented a 3D multibody model suitable for embedding within an MLTP problem~\cite{Domenighini:Designs:2023} - treated here as the \emph{baseline} model - two models with increasing complexity are presented. The first represents a vehicle capable of independently steering all four wheels (AS) and applying four independent torques to each wheel (AIWD, aka, torque vectoring), whereas the second model additionally incorporates an active camber variation system for each wheel (DCC).
%The kinematic tree of the two models is slightly different, but the procedure used to build the dynamic model is similar apart from a non-trivial aspect that will be discussed later on. In the remainder of this study, unless otherwise specified, the descriptions provided apply to both models. The figures related to Sections~\ref{sec:suspanal}-\ref{sec:tyremodel} are courtesy of the authors.

%% CHANGE THESE SECTIONS TO MERGE BOTH MODELS DESCRIPTIONS
\subsection{Kinematic Tree Structure of the Vehicle}
\begin{figure}[b]\centering
	\begin{tikzpicture}[scale=1.2]\small
		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw,gray](g)at(-0.1,-0.1){};
		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](b)at(0,2){};
		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](h11)at(-0.9,2.9){};
		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](h12)at(0.9,2.95){};
		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](h21)at(-1,1){};
		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](h22)at(0.95,1){};
		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](k11)at(-2.0,3){};
		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](k12)at(2.0,3){};
		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](k21)at(-2.1,1.1){};
		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](k22)at(2.0,1.05){};
		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](w11)at(-3.0,2.9){};
		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](w12)at(3.0,2.9){};
		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](w21)at(-3.1,1.){};
		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](w22)at(3.0,.95){};
		\draw[gray](g)to node(2){}(b);
		\draw(b)to node(4){}(h11);
		\draw(b)to(h12);
		\draw(b)to(h21);
		\draw(b)to(h22);
		\draw(h11)to node(8){}(k11);
		\draw(h12)to(k12);
		\draw(h21)to(k21);
		\draw(h22)to(k22);
		\draw(k11)to node(6){}(w11);
		\draw(k12)to(w12);
		\draw(k21)to(w21);
		\draw(k22)to(w22);
		\draw(g)node[below,gray](1){$\{G\}$}(b)+(0,0.1)node[above](3){$\{B\}$}(h11)node[above](5){$\{H_1\}$}(h12)node[above]{$\{H_2\}$}(h21)node[below]{$\{H_3\}$}(h22)node[below]{$\{H_4\}$}
		(k11)node[above](9){$\{K_1\}$}(k12)node[above]{$\{K_2\}$}(k21)node[below]{$\{K_3\}$}(k22)node[below]{$\{K_4\}$}
		(w11)node[above](7){$\{W_1\}$}(w12)node[above]{$\{W_2\}$}(w21)node[below]{$\{W_3\}$}(w22)node[below]{$\{W_4\}$};
		\node[gray,inner sep=1pt](world)at(1,-0.3){world};
		\node[gray,inner sep=0pt,align=center](6DoF)at(-1.25,0.15){floating base\\(virtual 6-DoF)};
		\node[inner sep=1pt](chassis)at(0.05,3){chassis};
		\node[inner sep=0pt,align=center](susp)at(-2.2,1.6){suspension linkages\\(generalized 1-DoF)};
		\node[inner sep=1pt](knuckle1)at(-.4,3.7){knuckles 1};
		\node[inner sep=1pt](knuckle2)at(-2.05,3.8){knuckles 2};
		\node[inner sep=0pt,align=center](hub)at(-4.5,2.25){hub bearings\\(revolute 1-DoF)};
		\node[inner sep=0pt,align=center](camber)at(-2.,2.4){camber joint\\(revolute 1-DoF)};
		\node[inner sep=1pt](rim)at(-3.4,3.65){rims};
		\draw[gray](world.west)to(1.east);
		\draw[gray](6DoF.north east)to(2);
		\draw(chassis.south)to(3);
		\draw(susp.north east)to(4);
		\draw(knuckle1.south)to(5);
		\draw(knuckle2.south)to(9);
		\draw(hub.east)to(6);
		\draw(camber.north)to(8);
		\draw(rim.south east)to(7);
	\end{tikzpicture}
	\caption{Kinematic tree representing the interconnections between bodies. Bodies are referred to by their attached frame.}
	\label{fig:tree}
\end{figure}
The vehicle is modeled as an articulated rigid-body system with chassis, knuckles, and wheels as rigid bodies. The chassis represents all sprung masses rigidly connected, while the knuckles (zero-mass bodies) decouple steering and vertical suspension motion from the wheel spindle. In the second model, knuckles are split into two zero-mass bodies, introducing an additional degree of freedom per wheel to simulate dynamic camber control (DCC). Wheels include all rotating unsprung masses, with non-rotating parts (e.g., brake calipers) incorporated into the wheel-tire mass.

Each body has a right-handed barycentric frame: $\sref{B}$ (chassis), $\sref{H}$ and $\sref{K}$ (knuckles), and $\sref{W}$ (wheels). Additional frames $\sref{G}$ (ground) and $\sref{S}$ (centerline) are defined, with \(x\)-axis of $\sref{S}$ tangent to the centerline, pointing forward. Frame relationships are described by homogeneous transformation matrices $ g(d, \vPhi) \in SE(3) $, with $ d $ as position and \(\vPhi\) (Euler angles) parameterizing the rotation.

As shown in Fig.~\ref{fig:tree}, bodies are connected by joints in a kinematic tree. The second model generalizes the first, where $\sref{H}$ and $\sref{K}$ coincide, so only the second model's kinematic tree is shown for brevity.
%% START ORIGINAL
%The vehicle is modeled as an articulated rigid-body system, with chassis, knuckles, and wheels treated as rigid bodies. The chassis includes all sprung masses, rigidly connected, while the knuckles (modeled as zero-mass bodies) decouple steering and vertical suspension motion from the wheel spindle. In the second model, the knuckles are split into two zero-mass bodies, adding one degree of freedom per wheel to simulate dynamic camber control (DCC). The wheels encompass all rotating unsprung masses, with non-rotating parts (e.g., brake calipers) included in the wheel-tire mass.
%
%Each body has a right-handed barycentric frame: $\sref{B}$ for the chassis, $\sref{H}$ (and $\sref{K}$ in the second model) for the knuckles, and $\sref{W}$ for the wheels. Two additional frames, $\sref{G}$ (ground) and $\sref{S}$ (attached to the vehicle's centerline), are also defined. The frame $\sref{S}$ lies in the tangent plane, with its $x$-axis tangent to the centerline, pointing forward.
%
%Relative positions and orientations between frames are described by homogeneous transformation matrices \( g(d, \vPhi) \in SE(3) \), where \( d \in \mathbb{R}^3 \) is the Cartesian position and \( \vPhi \in \mathbb{R}^3 \) are the Euler angles parameterizing the rotation matrix \( R \in SO(3) \).
%
%The bodies are connected by joints, forming the kinematic tree shown in Fig.~\ref{fig:tree}. The first model is a special case of the second, where $\sref{H}$ and $\sref{K}$ are always coincident, so only the kinematic tree for the second model is presented to avoid redundancy.

%% END ORIGINAL

%The vehicle is modeled as an articulated rigid-body system. In particular, chassis, knuckles and wheels are considered as rigid bodies. The chassis comprises all the sprung masses, including the driver, which are rigidly attached to one another. The knuckles are assumed here as zero-mass bodies, and their function is to decouple the motion of the suspension due to steering and vertical travel to the wheel spindle. In the second model, the knuckles are split into two zero-mass bodies, introducing one additional degree of freedom (DoF) for each wheel that allows to simulate the active camber variation system, here named dynamic camber control (DCC). The wheels include all the rotating unsprung masses, while all non-rotating parts of the unsprung masses, such as the brake calipers are included in the wheel-tire mass.
%
%Each body has a right-handed barycentral reference frame attached. The frames are denoted as $\sref{B}$, $\sref{H}$ (and $\sref{K}$) and $\sref{W}$ for the chassis, knuckles and wheels, respectively, where $\sref{K}$ is present only in the second model. Two additional frames are introduced: $\sref{G}$, fixed to the ground, and $\sref{S}$, that follows the vehicle on the ground. More precisely, $\sref{S}$ is attached to the centerline curve with its $x$ and $y$ axes lying in the tangent plane, and the $x$ axis also tangent to the centerline curve pointing forward.
%
%The relative positions and orientations between reference frames are described by homogeneous transformation matrices $g(d,\vPhi)\in SE(3)$, where $d\in \bbR^3$ is the relative Cartesian coordinates vector and $\vPhi \in \bbR^3$ are the three Euler angles that parameterize the rotation matrix $R\in SO(3)$.
%
%The bodies are connected by joints and the kinematic tree of the second model is shown in Figure \ref{fig:tree}. It is worth noting that the first model can be seen as a particular case of the second, in which the reference frames $\sref{H}$ and $\sref{K}$ are constrained to be always overlapping. For this reason, and to avoid cluttering the document, has been chosen to show the kinematic tree only for the second model.

\begin{figure}[t]
	\centering
	\includegraphics[scale = 0.8]{Images/vehicle_frames}
	\caption{Reference frames: $\sref{G}$ is a fixed inertial reference frame with the $z$-axis antiparallel to the gravity vector; $\sref{S}$ is a reference frame following the vehicle on track; $\sref{B}$ and $\sref{K}$ are FLU (Forward Left Up) reference frames aligned with the principal axes of inertia of the chassis and wheel, respectively; $\sref{H}$ is an auxiliary reference frame necessary to add the camber DoF; and $\sref{W}$ rotates about the $y$-axis with respect to $\sref{K}$ by the wheel rotation angle.}
	\label{fig:frames}
\end{figure}

The joints modeled are: hub bearings, connecting the rims $\sref{W}$ to the knuckles $\sref{H}$ (or $\sref{K}$ for the second model), and the suspension linkages, connecting the knuckles $\sref{H}$ to the chassis $\sref{B}$. For the second model a 1 DoF revolute joint connecting the two parts of the knuckles is also present. The chassis is a floating-base and its position and orientation with respect to the ground are unconstrained. This situations is conveniently encoded by a virtual six-DoF joint connecting $\{B\}$ to $\{G\}$. The resulting systems have 4+4+6=14 and 4+4+4+6=18 degrees of freedom, respectively.
%After defining the system's topology, a set of suitable state variables needs to be defined to describe the configuration and motion of each body.

%To fully describe the position and orientation of the frame $\sref{B}$ w.r.t. $\sref{G}$ we define vector $q_{gb}=(d_{gb};\varPhi_{gb})$.
%Its first-order kinematics is encoded in twist $V_{gb}^b = (v_{gb};\omega_{gb})$. In general, a twist $V_{ij}^k$ represents the motion of the frame $\sref{J}$ w.r.t. the frame $\sref{I}$ in $\sref{K}$ components. Unless otherwise specified, we employ \emph{body} coordinates, therefore $k=j$; this choice aligns to ABA convention~\cite{Featherstone:book:2008}. For each of the 4 DoF's associated to the suspension motion, a variable $z$ (and its derivative $\dot{z}$) is defined representing spring's length variation with respect to the nominal value. With these definitions, along with variables $\delta$ and its derivatives, defined in Sec.\ref{sec:suspanal}, one can find the six-component vector $q_{bh}$ representing the pose of $\sref{H}$ w.r.t. $\sref{B}$ and the twist $V_{bh}$. Assuming the wheel is an axis-symmetrical rigid body, and given that the relative orientation of the frame $\sref{W}$ w.r.t. $\sref{H}$ (or $\sref{K}$) is not relevant for the scope, only the angular velocity $\omega$ is introduced. For the second model, it is also necessary to introduce the variable $q_{hk}$ to describe the relative position of the frame $\sref{K}$ w.r.t. $\sref{H}$.
\subsection{Suspension Analysis}
\label{sec:suspanal}
The suspensions connect the chassis to the unsprung masses. This connection cannot be described by an elementary joint, such as prismatic or revolute. Therefore, a set of six variables $q_{bh}$, that parameterize a homogeneous transformation matrix, is necessary. The transformation can be realized by composing three translations and three rotations (the ZXY Euler sequence was selected) using the global Product of Exponentials (PoE) formula~\cite{Murray:book:1994}, which reads: \begin{equation}\label{eqn:PoE}
	g_{bh}=e^{\hat{Y}_1q_{bh,1}}\cdots e^{\hat{Y}_6q_{bh,6}} g_{bh}(0),
\end{equation}
where the offset $g_{bh(0)}=I$ and $Y_i=e_i$, with $e_i\in\bbR^6$ are the six normalized screw vectors (canonical basis elements in $\bbR^6$) associated with the Cartesian coordinates and the Euler ZXY angles in $q_{bh}$.
Clearly, the six components of $q_{bh}$ are not independent; the modeled suspension is, in fact, a double wishbone that has: 2 DoFs for wheels that can steer, and 1 DoF otherwise.
In this case, both axles have wheels that can steer; hence, a pair of independent variables ($z;\delta $) is associated with each wheel.
A detailed analysis is omitted here for the sake of brevity. The interested reader is referred to~\cite{Domenighini:Designs:2023} for a similar treatment using the PoE formalism.

%As mentioned in the previous sections, $z$ represents the variation of the distance between the two attachment points of the spring-damper system, positive when the distance increases and negative when it decreases, while $\delta$ represents the motion of the steering rod attachment point to the chassis along its local $y$ axis (frame $\sref{B}$). Considering that the type of the suspension is known, six scalar equations of constraints can be expressed implicitly as $F(q_{bh};z,\delta)=0\in \bbR^6$.
%
%To find an explicit relation between inputs $(z,\delta)$ and outputs $q_{bh}$, the constraint equations are solved on a grid of pairs $(z^{(p)},\delta^{(l)})$, where $p = 1,\dots,N$ and $l = 1,\dots,M$. Then, the data are fitted (in the least-squares sense) with a 2D regression polynomial $q_{bh}(z,\delta)$ as follows:
%\begin{equation}\label{eqn:poly}
%	q_{bh,i}(z,\delta)=\sum_{r=0}^{3} \sum_{s=0}^{3-r} a_{{rs},i} z^r \delta^s.
%\end{equation}
%For the purpose a polynomial of degree three is sufficiently accurate. Now is possible to express the homogeneous transformation $g_{bh}\bigl(q_{bh}(z,\delta)\bigr)$ and to compute the twist $V_{bh}$ and $\dot{V}_{bh}$ as functions of $(z,\delta)$ and their derivatives. We start from:
%\begin{equation}\label{eqn:Jbh}
%	V_{bh}=\sum_{i=1}^6J_{bh,i}\dot{q}_{bh,i},
%\end{equation}
%where $J_{bh}$ is the \emph{geometric Jacobian} of the suspension joint relative to $q_{bh}$. Then, following the procedure thoroughly described in~\cite[Sec. 2.2]{Domenighini:Designs:2023}, it is possible to obtain %
%assumed with independent components. Each column of $J_{bh}$ can be computed based on the PoE formula \eqref{eqn:PoE}. For the $i$-th column, we have $J_{bh,i}=A_i^{-1}Y_i$, with $A_6 = I_{6\times 6}$ and $A_i=\Ad_{e^{\hat{Y}_{i}q_{bh,i}}\cdots e^{\hat{Y}_6q_{bh,6}}}$, $i = 1, \dots, 5$. The \emph{Adjoint operator} $\Ad_g$ has been defined as the $6\times6$ matrix satisfying $Y=\Ad_gX$ whenever $\hat{Y}=g\hat{X}g^{-1}$, so that, if $g=\left[\begin{smallmatrix}R&d\\0^T&1\end{smallmatrix}\right]$, then $\Ad_g=\left[\begin{smallmatrix}R&\hat{d}R\\0&R\end{smallmatrix}\right]$.\\
%As a function of $(z,\delta)$, $\dot{q}_{bh,i}$ can be expressed as follows:
%\begin{equation}\label{eqn:dt_qbh}
%	\dot{q}_{bh,i}=\frac{\partial}{\partial z}q_{bh,i}\dot{z}+\frac{\partial}{\partial\delta}q_{bh,i}\dot{\delta},
%\end{equation}
%where the (geometric) partial derivatives $\frac{\partial}{\partial z}q_{bh,i}$ and $\frac{\partial}{\partial\delta}q_{bh,i}$ are obtained by differentiating Eq.~\eqref{eqn:poly}.
%Then, by defining the Jacobians with respect to the true independent variables $z$ and $\delta$ as follows
%\begin{align}\label{eqn:Jbh,z}
%	J_{bh,z}=\sum_{i=1}^6J_{bh,i}\frac{\partial}{\partial z}q_{bh,i}\quad \textrm{and}\quad J_{bh,\delta}=\sum_{i=1}^6J_{bh,i}\frac{\partial}{\partial\delta}q_{bh,i},
%\end{align}
%it possible to obtain the rigid-body velocity $V_{bh}$ of the constrained motion from Eq.~\eqref{eqn:Jbh}, in terms of $\dot{z}$ and $\dot{\delta}$, as follows
%\begin{equation}\label{eqn:Vbh}
%	V_{bh}=J_{bh,z}\dot{z}+J_{bh,\delta}\dot{\delta}.
%\end{equation}
%
%To write the dynamic equations the time derivative $\dot{V}_{bh}$ of the twist $V_{bh}$ is necessary. To this sake, it is sufficient to differentiate Eqs.~\eqref{eqn:Jbh}--\eqref{eqn:Vbh}.
%\begin{equation}\label{eqn:dt_Jbh}
%	\dot{V}_{bh}=\sum_{i=1}^6\dot{J}_{bh,i}\dot{q}_{bh,i}+\sum_{i=1}^6J_{bh,i}\ddot{q}_{bh,i}.
%\end{equation}
%
%Here, following~\cite{Gabiccini:ICRA:2012}, the derivatives $\dot{J}_{bh,i}$ can computed using the formula
%\begin{align}\label{eqn:dt_Jbhi}
%	\dot{J}_{bh,i}=-\sum\limits_{i < j \leq 6}\ad_{J_{bh,j}\dot{q}_{bh,j}} J_{bh,i}
%\end{align}
%using Lie derivatives. In the matrix case, the Lie derivative $\hat{Z}$ of the vector field $\hat{X}$ with respect to vector field $\hat{Y}$ is calculated by as a commutator operation $\hat{Z}=\hat{Y}\hat{X}-\hat{X}\hat{Y}$; in the vector case, the $6\times6$ matrix $\ad_Y$ is used, defined such that $Z=\ad_Y X$. If $v$ and $\omega$ are the translational and rotational component of $V$, then $\ad_V=\left[\begin{smallmatrix}\hat{\omega}&\hat{v}\\0&\hat{\omega}\end{smallmatrix}\right]$.
%
%In Eq.~\eqref{eqn:dt_Jbh}, the accelerations $\ddot{q}_{bh,i}$ are obtained by differentiating the velocities in Eq.~\eqref{eqn:dt_qbh}:
%\begin{equation}\label{eqn:dt_dt_qbh}
%	\ddot{q}_{bh,i}=\frac{\partial^2}{\partial z^2}q_{bh,i}\dot{z}^2+2\frac{\partial^2}{\partial z\partial\delta}q_{bh,i}\dot{z}\dot{\delta}+\frac{\partial^2}{\partial\delta^2}q_{bh,i}\dot{\delta}^2+\frac{\partial}{\partial z}q_{bh,i}\ddot{z}+\frac{\partial}{\partial\delta}q_{bh,i}\ddot{\delta},
%\end{equation}
%where the partial derivatives are computed by differentiating the regression polynomials~in~\eqref{eqn:poly}.
%After simple but tedious calculations, the explicit expression for the derivatives of the Jacobians in \eqref{eqn:Jbh,z} appear as follows:
%\begin{align}
%	\dot{J}_{bh,z}&=\sum_{i=1}^6\dot{J}_{bh,i}\frac{\partial}{\partial z}q_{bh,i}+\sum_{i=1}^6J_{bh,i}\Bigl(\frac{\partial^2}{\partial z^2}q_{bh,i}\dot{z}+\frac{\partial^2}{\partial z\partial\delta}q_{bh,i}\dot{\delta}\Bigr)\label{eqn:dt_Jbh,z}\quad\textrm{and}\\\dot{J}_{bh,\delta}&=\sum_{i=1}^6\dot{J}_{bh,i}\frac{\partial}{\partial\delta}q_{bh,i}+\sum_{i=1}^6J_{bh,i}\Bigl(\frac{\partial^2}{\partial z\partial\delta}q_{bh,i}\dot{\delta}+\frac{\partial^2}{\partial\delta^2}q_{bh,i}\dot{\delta}\Bigr)\label{eqn:dt_Jbh,d}.
%\end{align}
%Substituting Eqs.~\eqref{eqn:dt_Jbhi} and~\eqref{eqn:dt_dt_qbh} in Eqs.~\eqref{eqn:dt_Jbh}, and casting some intermediate results as~\eqref{eqn:dt_Jbh,z} and~\eqref{eqn:dt_Jbh,d},
%the constrained rigid-body acceleration $\dot{V}_{bh}$ in the following form:
%\begin{equation}\label{eqn:dt_Vbh}
%	\dot{V}_{bh}=\dot{J}_{bh,z}\dot{z}+\dot{J}_{bh,\delta}\dot{\delta}+J_{bh,z}\ddot{z}+J_{bh,\delta}\ddot{\delta}.
%\end{equation}
%The explicit expressions for auxiliary Jacobians $J_{bh,z}$, $J_{bh,\delta}$, $\dot{J}_{bh,z}$ and $\dot{J}_{bh,\delta}$ are omitted here for brevity. The interested reader can find the detailed analytical derivation in~\cite[Eqs. (5)--(8)]{Domenighini:Designs:2023}.
%
%A generalized force $\tau\in\bbR$ is introduced that depends only on the independent variable $z\in\bbR$. $\tau$ is assumed to be generated only by the spring and the damper, considered aligned along the same axis. It is worth noting that, using the defined $z$, we are allowed to compute the force of the suspension without calculate the associated $q_{bh}$ nor using the Principle of Virtual Works. In this expression, also the contribution of the anti-roll bar is considered, which depends on the difference between the left and right trim, and the contribution of the bump-stops, which activate when the trim reaches its lower or upper limits. The intensity of $\tau$, for each wheel, is expressed as follow:
%\begin{equation}
%	\tau_{ij}=k_{s,i}(l_{c,i}-l_{0,i}+z_{ij})+c_{d,i}\dot{z}_{ij}+(-1)^j k_{a,i}(z_{i,1}-z_{i,2}) + \left(\frac{2z-z_{\text{max}}-z_{\text{min}}}{|z_{\text{max}}-z_{\text{min}}|}\right)^{b},
%\end{equation}
%where $l_{c,i}$ is the spring's length in the reference position, $l_{0,i}$ is the spring's length at rest, $k_{s,i}$ and $c_{d,i}$ are the stiffness and the damping coefficients, $k_{a,i}$ is the anti-roll bar stiffness, $b$ influences the stiffness of the bump-stops, and $z_{\text{min}}$ and $z_{\text{max}}$ are the knees of the bump-stops' force characteristics.
\section{Tire Model}
\label{sec:tyremodel}
\begin{figure}[t]\centering
	\tdplotsetmaincoords{0}{0}
	\begin{tikzpicture}[tdplot_main_coords,scale=3.5]\small
		\tdplotsetrotatedcoords{-166.5322}{-65.2292}{-166.5322}\draw[tdplot_rotated_coords,dashed,gray!75,fill=gray!10](-0.6,0,0)--(-0.6,1.1,0)--(0.6,1.1,0)--(0.6,0,0);\draw[tdplot_rotated_coords,very thick,fill=white](0,0)coordinate(C)..controls(0.25,0)and(0.5,0)..(0.5,0.45)arc(0:180:0.5)..controls(-0.5,0)and(-0.25,0)..(0,0);\draw[tdplot_rotated_coords,very thick,fill=gray!25](0,0.45)coordinate(O)circle(0.25);\draw[tdplot_rotated_coords,very thick,dashed](0.5,0.45)arc(0:-180:0.5);\draw[tdplot_rotated_coords,very thin,dashed](0,0,0)--(0,0.2,0)(0,0.7,0)--(0,0.95,0)(-0.5,0.45,0)--(-0.25,0.45,0)(0.25,0.45,0)--(0.5,0.45,0);\draw[tdplot_rotated_coords,very thin](0,0.2,0)--(0,0.7,0)(0,0.95,0)--(0,1.25,0)(-1,0.45,0)--(-0.5,0.45,0)(-0.25,0.45,0)--(0.25,0.45,0)(0.5,0.45,0)--(1,0.45,0);\draw[tdplot_rotated_coords,{Stealth[inset=0pt]}-{Stealth[inset=0pt]},very thin](0.75,0,0)--node[right]{$r$}(0.75,0.45,0);\tdplotsetrotatedcoordsorigin{(O)}\tdplotsetrotatedcoords{-166.5322}{-65.2292}{-146.5322}\draw[tdplot_rotated_coords,dashed](-0.25,0,0)--(-0.5,0,0);\draw[tdplot_rotated_coords,dashed](0,0.25,0)--(0,0.5,0);\draw[tdplot_rotated_coords,->](0,0,0)--(-0.25,0,0)(-0.5,0,0)--(-1,0,0)node[below left]{$x_h$};\draw[tdplot_rotated_coords,->](0,0,0)--(0,0.25,0)(0,0.5,0)--(0,0.8,0)node[above left]{$z_h$};\draw[tdplot_rotated_coords,->](0,0,0)--(0,0,0.8)node[right]{$y_h$};\draw[tdplot_rotated_coords,{Stealth[inset=0pt]}-{Stealth[inset=0pt]},very thin](-0.8,0)arc(180:160:0.8)node[midway,left]{$\sigma$};\tdplotsetrotatedcoordsorigin{(C)}\tdplotsetrotatedcoords{43.2192}{-27.9909}{-46.7808}\draw[tdplot_rotated_coords,dashed,gray!75,fill=gray!25](0,0,-0.6)--(0.6,0,-0.6)--(0.6,0,0.6)--(0,0,0.6);\draw[tdplot_rotated_coords,->](0,0,0)--(0.8,0,0)node[right]{$y_n$};\draw[tdplot_rotated_coords,->](0,0,0)node[circle,fill=black,inner sep=0pt,minimum size=3pt,draw]{}--(0,1.25,0)node[above]{$z_n$};\draw[tdplot_rotated_coords,->](0,0,-1)--(0,0,1)node[below left]{$x_n$};\draw[tdplot_rotated_coords,{Stealth[inset=0pt]}-{Stealth[inset=0pt]},very thin](0,1.1)arc(90:110:1.1)node[midway,above]{$\gamma$};\draw(0,-0.02)..controls(0,-0.3)and(0.3,-0.1)..(0.3,-0.3)node[below]{contact point};\draw(-0.5,0.7)..controls(-0.35,0.55)and(-0.6,0.7)..(-0.55,0.5)node[below,align=center]{wheel\\plane};\draw(0.4,0.15)..controls(0.3,0.25)and(0.55,0.2)..(0.5,0.3)node[above,align=center]{road\\plane};\draw(O)node[below right]{$O_h$}(C)node[below right]{$O_n$};\draw[tdplot_rotated_coords,thick,-{Stealth[inset=0pt]}](C)--+(0,0,0.8)node[below right]{$F_x$};\draw[tdplot_rotated_coords,thick,-{Stealth[inset=0pt]}](C)--+(0.5,0,0)node[above]{$F_y$};\draw[tdplot_rotated_coords,thick,-{Stealth[inset=0pt]}](C)--+(0,1,0)node[right]{$F_z$};
	\end{tikzpicture}
	\hfill
	\begin{tikzpicture}[scale=4.25]\small
		\coordinate(C)at(0,0);\coordinate(-)at(-0.375,0);\coordinate(+)at(0.325,0);\draw[->](-0.7,0)--(0.6,0)node[right]{$y_n$};\draw[->](0,0)node[circle,fill=black,inner sep=0pt,minimum size=3pt,draw]{}--(0,1.25)node[above]{$z_n$};\tdplotsetrotatedcoords{10}{0}{0}\draw[tdplot_rotated_coords,dash dot](0,-0.15)--(0,1.25)(-0.35,0.45)--(0.6,0.45);\draw[tdplot_rotated_coords,very thick](-0.25,0.2)--(-0.25,0.7)..controls(-0.25,0.95)and(-0.25,0.95)..(0,0.95)..controls(0.25,0.95)and(0.25,0.95)..(0.25,0.7)--(0.25,0.2)(-0.25,0.2)..controls(-0.25,0.025)and(-)..(C)..controls(+)and(0.25,-0.025)..(0.25,0.2);\draw[tdplot_rotated_coords,very thick,dashed](-0.25,0.2)..controls(-0.25,-0.05)and(-0.25,-0.05)..(0,-0.05)node(B){}..controls(0.25,-0.05)and(0.25,-0.05)..(0.25,0.2);\node[tdplot_rotated_coords](O)at(0,0.45){};\draw[very thin](O)--({(-0.55,0)}|-O);\draw[very thin,{Stealth[inset=0pt]}-{Stealth[inset=0pt]}](-0.475,0)--node[left]{$h$}({(-0.475,0)}|-O);\draw[very thin](B)--({(-0.7,0)}|-B);\draw[very thin,-{Stealth[inset=0pt]}](-0.625,0.3)--node[left]{$d$}(-0.625,0);\draw[very thin](-0.625,0)--({(-0.625,0)}|-B);\draw[very thin,-{Stealth[inset=0pt]}](-0.625,-0.15)--({(-0.625,0)}|-B);\draw[tdplot_rotated_coords,very thin](0.05,0)--(0.45,0);\draw[tdplot_rotated_coords,very thin](0.05,-0.05)--(0.6,-0.05);\draw[tdplot_rotated_coords,very thin,{Stealth[inset=0pt]}-{Stealth[inset=0pt]}](0.375,0)--node[right]{$r$}(0.375,0.45);\draw[tdplot_rotated_coords,very thin,{Stealth[inset=0pt]}-{Stealth[inset=0pt]}](0.525,-0.05)--node[right]{$r_0$}(0.525,0.45);\draw[very thin,{Stealth[inset=0pt]}-{Stealth[inset=0pt]}](0,1.125)arc(90:100:1.125)node[midway,above]{$\gamma$};\draw(0,-0.4);
	\end{tikzpicture}
	\caption{Decomposition of the tyre forces along the axes of the auxiliary frame $\{N\}$. The frame $\{N\}$ is such that: $x_n$ lies along the intersection between the wheel plane ($x_hz_h$-plane) and the road plane ($x_s y_s$-plane), pointing forward; $y_n$ points left (w.r.t. the driver) along the intersection line between the road plane ($x_s y_s$-plane) and the transverse plane ($y_h z_s$-plane) through the wheel center $O_h$; $z_n$ is normal to the road surface. The contact point is estimated in correspondence to the origin $O_n$. On the right, we report a detail of the tyre vertical deformation.}
	\label{fig:tyre_wheel}
\end{figure}

The tangential forces exchanged between the tire and road are computed using \mbox{Pacejka's} \emph{Magic Formula}:
\begin{equation}\label{eqn_mf}
	[F_x,F_y]=\mf(\kappa,\alpha,F_z,\gamma),
\end{equation}
where $ F_x, F_y $ are the longitudinal and lateral forces, dependent on tire slips $( \kappa, \alpha )$, vertical load $( F_z )$, and camber angle $( \gamma )$.
As shown in Fig.~\ref{fig:tyre_wheel}, an auxiliary frame $\sref{N}$ is introduced to calculate $ F_z $. This depends upon the interpenetration $ d $ of the tire into the ground, following a penalty-based compliant tire model. This model treats the wheel as a radial spring-damper system with constant stiffness $ k_t $ and damping $ c_t $. The vertical force is expressed as:
\begin{equation}\label{eqn_Fz}
	F_z= F_0\log_2\left(1 + 2^{\frac{k_t d + c_t \dot{d}}{F_0}}\right)
\end{equation}
which is differentiable for $ d = 0 $ and $\dot{d} = 0$, improving numerical stability. The model ensures $ F_z $ approaches zero when the tire loses contact with the ground, while maintaining a non-zero gradient to guide optimization algorithms in scenarios where the tire detaches from the ground or encounters sudden changes.

%To compute the tangential components of the force exchanged by tyre and road, a formulation of the Pacejka's \emph{Magic Formula} ~\cite[Section~4.3.2]{Pacejka:book:2012} is used, which reads:
%\begin{equation}\label{eqn_mf}
%	[F_x,F_y]=\mf(\kappa,\alpha,F_z,\gamma).
%\end{equation}
%The Magic Formula $\mf$ computes the longitudinal and lateral forces $F_x,F_y$ as a function of the \emph{tyre slips} $\kappa,\alpha$~\cite[Section~1.2.1]{Pacejka:book:2012}. The formula is also sensitive to variations of vertical load $F_z$ and camber angle $\gamma$.
%
%The auxiliary frame $\sref{N}$ is introduced to compute quantities used in Eq.~\eqref{eqn_mf} and the (attempted) interpenetration $d$ (see Figure.~\ref{fig:tyre_wheel}) of the tyre in the ground, necessary to compute the dynamic vertical load $F_z$. The construction of the frame $\sref{N}$ is reported in Figure~\ref{fig:tyre_wheel}. It is worth noting that the frame $\sref{N}$ is fully described if $\sref{H}$ and $\sref{S}$ are known.
%The model used for the contact is a unilateral, penalty-based compliant tyre model, that considers the wheel as a radial spring with constant stiffness $k_t$ and constant damping coefficient $c_t$\footnote{The introduction of a more refined model with nonlinear spring and damping characteristics would not pose particular difficulties.}. The traditional role of compression in a spring is conducted by $d$, which is computed considering the road and the lowermost point of the non-deformed tyre. The resultant function that relate $d$, $\dot{d}$ and $F_z$ is:
%\begin{equation}\label{eqn_Fz}
%	F_z= F_0\log_2\left(1 + 2^{\frac{k_t d + c_t \dot{d}}{F_0}}\right)
%\end{equation}
%This function, with $F_0$ tuned to be negligible, is differentiable for $d=0$ and $\dot{d}=0$, enhancing numerical performance, and behaves as a spring-damper system, as requested, when the tyre is in contact with the ground. When the tyre detaches completely from the ground, the resultant vertical forces tends to zero. However, since its gradient is not zero, this formulation helps the optimization algorithm to figure out proper search directions. This allows to deal with jumps and other situations in which the tyre may lose contact with the ground.

\subsection{Camber System Modeling}
\label{sec:cambersystemmodeling}
The dynamic camber control (DCC) system is modeled as a simplified actuated revolute joint connecting the two virtual halves of the knuckle, though a real implementation might require a more complex 1-DoF joint. While this simplification may influence joint values from the optimizer, it results in nearly the same optimal wheel posture. The choice was driven by the limited literature on such systems for FSAE vehicles, which is the vehicle type employed in our analysis. A realistic model, if available, could be characterized using the same process applied to the suspension. During the design phase of the actual joint, some parameters could be treated as time-invariant optimization variables, constrained within dimensional bounds, to achieve the best performance configuration.
%The dynamic camber control (DCC) system is modeled here as a simple actuated revolute joint connecting the two virtual halves of the knuckle. A real implementation could require a more complex generalized 1-DoF joint. Changing the joint characteristics might affect the joint values obtained from the optimizer but would result in approximately the same optimal wheel posture. This simplification was necessary due to the limited open literature on such systems for FSAE vehicles. If a realistic model of the system were available, the process used for the suspension could easily be repeated to characterize this joint. It is worth observing that, in the design phase of the actual controlled joint, some model parameters could be treated as time-invariant optimization variables, appropriately constrained to meet dimensional bounds, and optimized to find the best performance configuration.

In Table~\ref{tab:camber_prop}, we report the range of variation of function $q_{hk}(t)$ and its first and second derivatives. Considering the scarcity of open literature available, reasonable values have been selected. In Figure~\ref{fig:camber_showing} the wheel is shown in a general posture with zero camber (shaded black) and with maximum positive camber (solid black). Due to the convention adopted for rotations, this posture corresponds to the minimum value of $q_{hk}$.
\begin{table}[h]
	\centering
	\begin{tabular}{|c|c|c|}
		\hline
		$q_{hk}$ (rad) & $\dot{q}_{hk}$ (rad/s)& $\ddot{q}_{hk}$ (rad/s$^2$)\\\hline
		$\pm \pi/20$ & $\pm \pi/20$ & $\pm \pi/10$ \\\hline
	\end{tabular}
	\caption{Upper and lower bounds for camber angle value $q_{hk}(t)$ and its derivatives.}
	\label{tab:camber_prop}
\end{table}

\begin{figure}[h]
	\centering
	\includegraphics[scale = .6]{Images/camber_showing}
	\caption{Real scale representation of the wheel in two configurations: $q_{hk}=0$ (shaded surface) and $q_{hk}=-\pi/20$ (solid surface).}
	\label{fig:camber_showing}
\end{figure}
\subsection{External Wrenches}
\label{sec:extwren}
The external wrenches acting on each body, from wheels to chassis, are expressed in body-fixed components. Only gravity, aerodynamic forces, and road-tyre interactions are considered. Following~\cite{Guiggiani:book:2023}, the aerodynamic wrench in frame $\{B\}$ is $W_a=-\frac{1}{2}\rho S v_{gb,1}^2 [C_x, 0, C_z, 0, -h_0 C_x - a_f C_{zf} + a_r C_{zr}, 0]^T$, where $\rho$ is air density, $S$ is the frontal area, $v_{gb,1}$ the forward velocity (component of \emph{distal rigid-body twist} $V_{gb}$, see~\cite[p.~4]{Domenighini:Designs:2023}), and $C_x, C_z$ are drag/lift coefficients. $C_z$ is split into $C_{zf}=k_a C_z$ and $C_{zr}=(1-k_a)C_z$ by the aerodynamic balance coefficient $k_a$.

The total external wrench on a wheel is $W_{we}=\Ad_{g_{hn}}^*W_t+W_{w_h}$,
%\begin{equation}\label{eqn:Wwe}
%W_{we}=\Ad_{g_{hn}}^*W_t+W_{w_h},
%\end{equation}
where $W_t=[F_x, F_y, F_z, 0, 0, 0]^T$ is the ground wrench transferred from $\sref{N}$ to $\sref{H}$ by the \emph{starred} Adjoint operator, and $W_{w_h}=[F_{w_h}^T R_{hg}^T, 0, 0, 0]^T$ is the weight contribution with $F_{w_h}=[0, 0, -m_h g]^T$ in frame $\{G\}$, where $m_h$ is the wheel mass.

Knuckles experience no external forces, therefore:
\begin{equation}\label{eqn:Whe}
	W_{he}=[0,0,0,0,0,0]^T\quad\textrm{and}\quad W_{ke}=[0,0,0,0,0,0]^T.
\end{equation}
The chassis's external wrench is $W_{be}=W_a+W_{w_b}$, %\begin{equation}\label{eqn:Wbe}
%	W_{be}=W_a+W_{w_b},
%\end{equation}
where $W_a$ and $W_{w_b}$ are aerodynamic and weight contributions, respectively. For $W_{w_b}=[F_{w_b}^T R_{bg}^T, 0, 0, 0]^T$, the weight force is $F_{w_b}=[0, 0, -m_b g]^T$ in frame $\{G\}$, with $m_b$ as the chassis mass.
%The purpose of this section is to describe the external wrenches acting on each body, from the wheels to the chassis, each expressed in body-fixed components.
%Besides gravity, aerodynamic forces and the interactions between road and tyres are the only external forces considered.
%Adopting the notation in~\cite{Guiggiani:book:2023}, the aerodynamic contribution to the wrenches in frame $\{B\}$ is $W_a=-\frac{1}{2}\rho Sv_{gb,1}^2[C_x,0,C_z,0,-h_0C_x-a_fC_{zf}+a_rC_{zr},0]^T$, where $\rho$ is the air density, $S$ the frontal area of the vehicle, $v_{gb,1}$ the forward velocity, $h_0$ the nominal height of the CoM from ground and $a_f,a_r$ the distances of the CoM from the front and rear axle, respectively. The drag and lift coefficient $C_x$ and $C_z$ are dimensionless parameters that depend on the shape of the vehicle's body. $C_z$ can be further partitioned into $C_{zf}=k_a C_z$ and $C_{zr}=(1-k_a) C_z$ according to the \emph{aerodynamic balance coefficient} $k_a$, which is a parameter of the vehicle set-up.
%
%The total external wrench acting on a (general) wheel is:
%\begin{equation}\label{eqn:Wwe}
%	W_{we}=\Ad_{g_{hn}}^*W_t+W_{w_h},
%\end{equation}
%where $W_t=[F_x,F_y,F_z,0,0,0]^T$ are the ground forces computed in $\sref{N}$ and properly transferred to $\sref{H}$ components by the \emph{starred} Adjoint operator, while $W_{w_h} = [ F_{w_h}^T R^T_{hg}, 0,0,0 ]^T$, with $F_{w_h}=[0,0,-m_h g]^T$ in frame $\{G\}$, represents the weight contribution and $m_{h}$ is the wheel mass.
%
%Under our hypotheses, the knuckles are not subjected to external forces, thus \begin{equation}\label{eqn:Whe}
%	W_{he}=[0,0,0,0,0,0]^T\quad\textrm{and}\quad W_{ke}=[0,0,0,0,0,0]^T.
%\end{equation}
%
%The resultant external wrench acting on the chassis is therefore:
%\begin{equation}\label{eqn:Wbe}
%	W_{be}=W_a+W_{w_b},
%\end{equation}
%where $W_a$ and $W_{w_b}$ are, respectively, the contributions of aerodynamic and chassis weight forces, whose general expression is of the form $W_{w_b} = [ F_{w_b}^T R^T_{bg}, 0,0,0 ]^T$. Explicitly we have $F_{w_b}=[0,0,-m_b g]^T$ in frame $\{G\}$, where $m_b$ is the mass of the chassis.

\subsection{Recursive computation of the dynamics equations with ABA}
\label{sec:aba}
Thanks to the ABA algorithm, assuming known joint positions and velocities and joint forces and torques, it is possible to compute recursively and efficiently the accelerations of the joint variables. Following~\cite{Domenighini:Designs:2023}, twists are streamlined as follows: $V_{gb}^b = V_{gb} = V_b$, $V_{bh}^h=V_{bh}$, $V_{hw}^h=V_{hw}$. For the second model (DCC) the last expression is replaced with $V_{kw}^k=V_{kw}$ and $V_{hk}^k=V_{hk}$ is added. Given that the first model (without camber control) is a particular case of the DCC one, to avoid duplications only the algorithm for DCC is reported here.

The traditional ABA requires knowledge of the active forces and torques, denoted as $\tau_{ij}$, exerted by the parent body ($i$) on the child body ($j$). Its purpose is to solve the \emph{Forward Dynamics} problem of an articulated systems of rigid bodies. In our floating-base system, given the twist $V_b$ of the parent body $\{B\}$, the values and time derivatives of all joint variables ($q$ and $\dot{q}$), and the torque vector ($\tau$), the objective is to compute the joint accelerations ($\dot{V}_{b}$ and $\ddot{q}$).
%The fundamental concept is that of the inertia of an articulated body. This is represented by that of the root rigid body (original parent) plus the portion of the sub-trees, composed by the children bodies, whose inertia that can be structurally transferred backwards through the joints.
%This depends on the architecture, the pose and the joints' types.
The overall computation is performed in three sweeps defined recursively. One of the greatest computational benefits when it is employed to assemble symbolic dynamic equations, as in our case, is that \emph{no matrix inversion is required}, thus contributing to streamline the algebraic expressions.

The first step of the algorithm, as described in \textbf{Step 1}, performs the forward propagation of rigid-body velocities: from the root to the leaves of the tree.
%Twists are calculated as the sum of parent twist and the relative twist which is expressed consistently with the joint type. The interesting Jacobians are listed here: $J_{hk}=[0,0,0,1,0,0]^T$ and $J_{kw}=[0,0,0,0,1,0]^T$, and represent two revolute joints with axes $x$ and $y$ respectively.
In Algorithm~\ref{alg:step1} the terms ``Knuckle 1'' and ``Knuckle 2'' are used to identify the two halves of the knuckle, where, using kinematic trees terminology, the first is the parent of the second.
The Adjoint operator is defined as follows:
\begin{equation}\label{eq:Adjoint} \Ad_{g}=\left[\begin{array}{cc}R&\hat{d}R\\0&R\end{array}\right].
\end{equation}
In the case $g_{bh}\in SE(3)$ is considered, $R_{bh}\in SO(3)$ and $d_{bh}\in \bbR^3$ are functions of $(z,\delta)$. In the case of $g_{hk}$, the expressions are simply $d_{hk}=[0,0,0]^T$ and $R = \rotX(q_{hk})$, the latter being an elementary rotation about the common $x$-axis of the angle $q_{hk}$.

\begin{algorithm}[h]
	\floatname{algorithm}{Step}
	\caption{Forward Propagation of Velocity}\label{alg:step1}
	\begin{algorithmic}[1]
		\vspace{1mm}
		\For{$h=h_1,h_2,h_3,h_4$}
		\vspace{1mm}
		\State{$V_{bh}=J_{bh,z}\dot{z}+J_{bh,\delta}\dot{\delta}$}
		\vspace{1mm}
		\State{$V_h=\Ad_{g_{bh}}^{-1}V_b+V_{bh}$}\Comment{Knuckle 1 Rigid-Body Velocity}
		\vspace{1mm}
		\State{$V_{hk} = J_{hk}\dot{q}_{hk}$}
		\vspace{1mm}
		\State{$V_k=\Ad_{g_{hk}}^{-1}V_h+V_{hk}$}\Comment{Knuckle 2 Rigid-Body Velocity}
		\vspace{1mm}
		\State{$V_{kw}=J_{kw}\omega$}
		\vspace{1mm}
		\State{$V_w=V_k+V_{kw}$}\Comment{Rim Rigid-Body Velocity}
		\vspace{1mm}
		\EndFor
		\vspace{1mm}
	\end{algorithmic}
\end{algorithm}

The second step of the algorithm, as detailed in \textbf{Step 2}, performs the backward propagation of inertia and bias terms.
\begin{algorithm}[h]
	\floatname{algorithm}{Step}
	\caption{Backward Propagation of Articulated Inertia and Bias}\label{alg:step2}
	\begin{algorithmic}[1]
		\vspace{1mm}
		\For{$h=h_1,h_2,h_3,h_4$}
		\vspace{1mm}
		\State{$\hat{M}_w=M_w$}\Comment{Rim Articulated Inertia}
		\vspace{1mm}
		\State{$\hat{B}_w=-W_{we}+\ad_{V_w}^*M_wV_w$}\Comment{Rim Articulated Bias}
		\vspace{1mm}
		\State{$\bar{M}_w=\hat{M}_w-\dfrac{\hat{M}_wJ_{hw}J_{hw}^T\hat{M}_w}{J_{hw}^T\hat{M}_wJ_{hw}}$}
		\vspace{1mm}
		\State{$\bar{B}_w=\hat{B}_w-\dfrac{\hat{M}_wJ_{hw}J_{hw}^T\hat{B}_w}{J_{hw}^T\hat{M}_wJ_{hw}}-\bar{M}_w\ad_{V_{hw}}V_w$}
		\vspace{1mm}
		\State{$\hat{M}_k= \bar{M}_w$
		}\Comment{Knuckle 2 Articulated Inertia}
		\vspace{1mm}
		\State{$\hat{B}_k= \bar{B}_w$
		}\Comment{Knuckle 2 Articulated Bias}
		\vspace{1mm}
		\State{$\hat{M}_h= \Ad_{g_{hk}}^*\hat{M}_k\Ad_{g_{hk}}^{-1}$
		}\Comment{Knuckle 1 Articulated Inertia}
		\vspace{1mm}
		\State{$\hat{B}_h= \Ad_{g_{hk}}^*\bigl(\hat{M}_k(J_{hk}\ddot{q}_{hk}-\ad_{V_{hk}}\Ad_{g_{kh}}V_k)+\hat{B}_k\bigr)$
		}\Comment{Knuckle 1 Articulated Bias}
		\vspace{1mm}
		\State{$\bar{M}_h=\hat{M}_h-\dfrac{\hat{M}_hJ_{bh,z}J_{bh,z}^T\hat{M}_h}{J_{bh,z}^T\hat{M}_hJ_{bh,z}}$}
		\vspace{1mm}
		\State{$\bar{B}_h=\hat{B}_h-\dfrac{\hat{M}_hJ_{bh,z}(J_{bh,z}^T\hat{B}_h-\tau)}{J_{bh,z}^T\hat{M}_hJ_{bh,z}}-\bar{M}_h(\ad_{V_{bh}}V_h-\dot{J}_{bh,z}\dot{z}-\dot{J}_{bh,\delta}\dot{\delta}-J_{bh,\delta}\ddot{\delta})$}
		\vspace{1mm}
		\EndFor
		\vspace{1mm}
		\State{$\hat{M}_b=M_b+\displaystyle\sum_h\Ad_{g_{bh}}^*\bar{M}_h\Ad_{g_{bh}}^{-1}$}\Comment{Chassis Articulated Inertia}
		\vspace{1mm}
		\State{$\hat{B}_b=-W_{be}+\ad_{V_b}^*M_bV_b+\displaystyle\sum_h\Ad_{g_{bh}}^*\bar{B}_h$}\Comment{Chassis Articulated Bias}
		\vspace{1mm}
	\end{algorithmic}
\end{algorithm}

%Both inertia and bias are expressed in the body-fixed components.
In the case the DCC architecture is considered, the camber angle acceleration $\ddot{q}_{hk}$ is known, since $q_{hk(t)}$ is the assumed input, while the torque $\tau_{hk}$ has to be computed together with the rest of unknown accelerations, which are the outcome of the third sweep. This leads to a \emph{Hybrid Dynamics} problem that requires to carefully revisit the ABA's second and third steps.
The approach used is inspired by~\cite[Section~2.3]{Kim:notes:2012} where the the notation and conventions here employed were adapted according to~\cite{Domenighini:Designs:2023}.
%Specifically, apart from the labels assigned to the quantities, the primary difference between the two methods is the ordering of the twist vector. In this work, the twist vector is defined as $V = [v,\omega]^T$, whereas in the other method, it is defined as $V=[\omega,v]^T$. This difference might seem minor but leads to significant changes in the definitions of Adjoint operators, inertia matrices, bias and wrench vectors.

%For a generic articulated body the Newton-Euler equations can be expressed in the form:
%\begin{equation}\label{eqn:NE}
%	W=\hat{M}\dot{V}+\hat{B}.
%\end{equation}
%For a leaf node of the kinematic tree, having no children, the articulated inertia $\hat{M}$ is simply the generalized inertia $M$ of the body itself. Consistently, the bias $\hat{B}=B$, where:
%\begin{equation}
%	B=-W_e+\ad_V^*MV,
%\end{equation}
%and where $W_e$ is the resultant external wrench applied to the body and $\ad_V^*MV$ accounts for generalized gyroscopic and centrifugal forces, which are bi-linear in $V$. According to~\cite[Section~4.3.3]{Murray:book:1994} $\ad^*=-\ad^T$ has been defined. (If $v$ and $\omega$ are the translational and rotational component of $V$, then $\ad^*_V=\left[\begin{smallmatrix}\hat{\omega}&0\\\hat{v}&\hat{\omega}\end{smallmatrix}\right]$).

Starting from the leaf nodes, which in our case are represented by the wheels ($\sref{W}$), after providing information such as external wrenches, twists, joint geometry, body masses, joint forces / torques for traditional connections, and time evolution of variables defining kinematically imposed joints, the backward propagation of inertia and bias can be performed.

Finally the last step, denoted as \textbf{Step 3}, starting from the root node, computes joint variable accelerations and consequently body twists derivatives. For the kinematically controlled camber joints, the torques $\tau_{hk}$ needed to maintain the desired state of motion (DCC architecture) are computed as a by-product of the local inverse dynamics. This information is crucial to prevent actuators from  saturating or operating outside their bandwidth in the planned trajectories.

The root twist acceleration is determined by solving equation $W=\hat{M}\dot{V}+\hat{B}$, considering that the chassis is free-floating. The other unknown quantities, such as variable accelerations and torques, are computed projecting the Newton-Euler equations of each body along the respective Jacobian and then solving for the variable. See also~\cite{Domenighini:Designs:2023} for more details in the derivation of the equations.

\begin{algorithm}[h]
	\floatname{algorithm}{Step}
	\caption{Forward Propagation of Acceleration}\label{alg:step3}
	\begin{algorithmic}[1]
		\State{$\dot{V}_b=-\hat{M}_b^{-1}\hat{B}_b$}\Comment{Chassis Rigid-Body Acceleration}
		\vspace{1mm}
		\For{$h=h_1,h_2,h_3,h_4$}
		\vspace{1mm}
		\State{$\ddot{z}=-\dfrac{J_{bh,z}^T\bigl(\hat{B}_h+\hat{M}_h(\Ad_{g_{bh}}^{-1}\dot{V}_b-\ad_{V_{bh}}V_h+\dot{J}_{bh,z}\dot{z}+\dot{J}_{bh,\delta}\dot{\delta}+J_{bh,\delta}\ddot{\delta})\bigr)-\tau}{J_{bh,z}^T\hat{M}_hJ_{bh,z}}$}
		\vspace{1mm}
		\State{$\dot{V}_{bh}=\dot{J}_{bh,z}\dot{z}+\dot{J}_{bh,\delta}\dot{\delta}+J_{bh,z}\ddot{z}+J_{bh,\delta}\ddot{\delta}$}
		\vspace{1mm}
		\State{$\dot{V}_h=\Ad_{g_{bh}}^{-1}\dot{V}_b-\ad_{V_{bh}}V_h+\dot{V}_{bh}$}\Comment{Knuckle 1 Rigid-Body Acceleration}
		\vspace{1mm}
		\State{$\tau_{hk} = J_{hk}^T\bigl(\hat{B}_k+\hat{M}_k(\Ad_{g_{kh}}\dot{V}_h-\ad_{V_{hk}}V_k + J_{hk}\ddot{q}_{hk})\bigr)$}
		\vspace{1mm}
		\State{$\dot{V}_{bh} = J_{hk}\ddot{q}_{hk}$}
		\vspace{1mm}
		\State{$\dot{V}_k=\Ad_{g_{hk}}^{-1}\dot{V}_h-\ad_{V_{hk}}V_k+\dot{V}_{bh}$}\Comment{Knuckle 2 Rigid-Body Acceleration}
		\State{$\dot{\omega}=-\dfrac{J_{kw}^T\bigl(\hat{B}_w+\hat{M}_w(\dot{V}_k-\ad_{V_{kw}}V_w)\bigr)}{J_{kw}^T\hat{M}_wJ_{kw}}$}
		\vspace{1mm}
		\EndFor
		\vspace{1mm}
	\end{algorithmic}
\end{algorithm}

\subsection{Dynamic Model}
The dynamics equations, assembled via the procedure described  in Sec.~\ref{sec:aba}, are implemented in the software code by using a state-space representation $\dot{x}=f(x,u)$.
%\begin{equation}\label{eqn:f}
%	\dot{x}=f(x,u).
%\end{equation}
The state vector is $x = (q_{gb};V_{gb};z;\dot{z};\omega;q_{hk};\dot{q}_{hk})\in\bbR^{32}$, with components defined in previous sections. As already mentioned, in the model the camber joint is kinematically controlled as per the DCC architecture. Hence, the associated control vector is $u=(T_a;T_b;\delta;\ddot{q}_{hk})\in\bbR^{16}$, where each element is in $\bbR^{4}$ and collects the controls of the branch to each wheel. The torque is split into acceleration and braking components, allowing for minor modifications to model on-board motors coupled with traditional brakes and/or in-wheel motors as well. These components have their reaction torques in two different bodies in case of on-board motors and traditional brakes: the chassis and the knuckles, respectively. The $\delta$ vector collects the four translations of the steering rod attachment points, as described in Section~\ref{sec:suspanal}, while for the vector $\ddot{q}_{hk}$ no further explanation should be necessary. Clearly, this representation is valid for the model with active camber. The simpler model (no DCC) shares with the former the first 24 components of the state vector and the first 12 of the control vector.

The time derivative of the state vector has also been expressed previously except for the first element, $q_{gb}$. Its derivative is a function of the state and can be recovered using a standard procedure~\cite{Domenighini:Designs:2023} as follows:
\begin{equation}\label{eqn:dotqgb}
	\dot{q}_{gb}=J_{gb}^{-1}(q_{gb})V_{gb},
\end{equation}
where $J_{gb}(q_{gb})\in\bbR^{6\times6}$ is the Jacobian of the virtual joint connecting the ground to the chassis, and it is computed via the PoE parameterization of the transformation $g_{gb}$.
%, similarly to $J_{bh}$ in Eq.~\eqref{eqn:Jbh}.

### 2025-04-03 conclusions.tex
[MGa] basic structure

\section{Conclusions}
\label{sec:conclusions}

In this work we investigated the potential of some new technologies in empowering next generation autonomous race vehicles. In particular, the upgrades considered here were: four independent steering wheels (All Steering - AS), four independent wheel torques (All Independent Wheel Drive - AIWD), and independent control of the four camber angles (Dynamic Camber Control - DCC).

Firstly, we discussed the impact of enabling these control inputs on the vehicle performance, in relation to the mass increase associated to the new components needed for the systems.
The results confirmed that the vehicle endowed with all the features (All Independent Wheel Drive, All Steering, Dynamic Camber Control - AIWD AS DCC) is the most performant and forgiving to added mass. Compared to a traditional vehicle (Rear Wheel Drive, Forward Steering - RWD FS), its advantage in lap time persists until 123.5\,kg of added mass.

Then, lap times are presented for the baseline (Rear Wheel Drive, Forward Steering - RWD FS), and for vehicles with single upgrades added, keeping constant the total mass.
In particular, the configuration analyzed are: Rear Wheel Drive, All Steering - RWD AS, All Independent Wheel Drive, Forward Steering - AIWD FS, and Rear Wheel Drive, Forward Steering, Dynamic Camber Control - RWD FS DCC.
From these results it is evident that the single feature that improves the performance the most is the AIWD FS in Siena circuit, and the RWD FS DCC in N\"{u}rburgring circuit.

Focusing on trajectories, we can highlight the following features. The All Independent Wheel Drive, Forward Steering (AIWD FS) vehicle, being able to perform torque vectoring, travels turns closer to the inner boundary of the track, compared to a traditional vehicle, and exits with an higher forward speed.

A Rear Wheel Drive, All Steering (RWD AS) vehicle experiences in-phase rear steering for high speed turns, and counter-phase rear steering for low speed turns, hence improving cornering in each condition.

Being able to control the camber dynamically, a Rear Wheel Drive Vehicle, Forward Steering, Dynamic Camber Control (RWD FS DCC) vehicle, positions each wheel w.r.t. the road to raise the force limits, thus maximizing the ground forces.

To highlight the impact of the torque vectoring and camber control, a vehicle with these characteristics but \emph{without steering abilities} (All Independent Wheel Drive, Zero Steering, Dynamic Camber Control - AIWD 0S DCC) has been analyzed.
This vehicle completes a lap in Siena 3.66 seconds faster than a traditional vehicle (RWD FS), and corners effectively by applying higher torques on outer wheels than the inner wheels.

Introducing a temperature-dependent tyre model, we observed that the least affected vehicle is the traditional RWD FS, which stresses the wheels the least and thus loses only 11 ms due to the drop in grip factor. In contrast, All Independent Wheel Drive (AIWD) vehicles tend to stress all four wheels much more. As a result, when the tire thermal model in the MLTP takes into account the reduction in grip with increasing temperature, AIWD vehicles are penalized more than RWD FS vehicles, losing approximately 100 ms each in the considered sector.

\section{Future Work}

While this study provides a broad overview of the possibilities and implications of introducing certain features to a new generation of autonomous vehicles, some aspects require further investigation.

A more complete tire model, which could account for tread wear and the consequent inevitable performance drop due to tire usage, should be implemented. To achieve this, either a testing campaign on the actual modeled tires to fine-tune the parameters of the wear-dependent phenomena, or a new dataset of fully characterized tires, is essential.

The introduction of realistic aerodynamic maps could indeed lead to a refinement of the model, enhancing the accuracy of the vehicle dynamics representation.

Modeling a non-developable 3D track that incorporates bumps and holes would represent a significant advancement in fully understanding the potential and limitations of these features. This approach would enhance the physical realism of the road-vehicle interaction, providing deeper insights into its dynamics.

Additionally, future developments could include an evaluation of design solutions for the various subsystems that control the additional degrees of freedom. These solutions must, however, take into account their practical feasibility for real-world systems, considering current technological limitations in sensors and actuators, as well as the challenges posed by their integration.

### 2025-04-03 intro.tex
[MGa] basic structure

\section{Introduction}
\label{sec:intro}

As the automotive industry advances towards higher levels of automation, ...

\subsection{Related work}
In recent years, extensive research has been conducted on ... 
\subsection{Paper's contributions}
In this study, we investigate ...

### 2025-04-03 open_loop_planning.tex
[MGa] basic structure

\section{Open-loop planning via n-steps ahead predictions}
\label{sec:open_loop_planning}

The model described in Sec.~\ref{chap:vehicle_model} is embedded into a minimum lap time problem which is briefly outlined below.

\subsection{Direct Collocation}
The objective is to recast the MLTP into a Non Linear Program (NLP) using a direct orthogonal collocation approach, specifically based on Gauss-Legendre collocation points. The problem is formulated within the CasADi framework~\cite{Andersson:MPC:2019} and solved using the interior point method implemented in IPOPT~\cite{Wachter:MP:2006}. The resulting NLP can be expressed in the form:
\begin{align}
	\underset{x,v,u,z}{\textrm{minimize}}&\quad\sum_{k=1}^NL_k(x_k,v_k,u_k,z_k)\label{eqn:functional}\\
	\textrm{subject to}&\quad g_k(x_{k-1},x_k,v_k,u_k,z_k)=0,\ k=1,\ldots,N,\label{eqn:equality}\\
	&\quad h_k(x_k,u_k,z_k)\le0,\ k=1,\ldots,N.\label{eqn:inequality}
\end{align}
In Eqs.~\eqref{eqn:functional}, \eqref{eqn:equality} and \eqref{eqn:inequality}, in addition to the state and control vectors, $v_k$ and $z_k$ are included, representing the state vector at the collocation points and the algebraic quantities, respectively. The controls and the algebraic variables are approximated as piecewise constant, while the states are approximated as quadratic polynomials within each finite element. Specifically, the algebraic variables in the problem are the time steps $\nu_k$ and the lateral and vertical displacements of the Center of Mass relative to the origin of the $k$-th frame $\sref{S}$, attached to the track's centerline ($e_y$ and $e_z$). Their function is to serve as handle variables in the optimization process to promote problem sparsity.

The track centerline is represented by a curvilinear abscissa $\alpha\in[0,1]$, sampled at $N+1$ points $\alpha_0,\ldots,\alpha_N$. Each grid point is associated with a reference frame $\sref{S_k}$, whose origin lies at the corresponding point on the track centerline, as illustrated in Figure~\ref{fig:track_frames}. A more detailed description of the 3D track modeling can be found in~\cite{Bartali:Meccanica:2023}. Specifically, this representation allows us to capture track slope and banking, but it does not support non-developable track surfaces.

\begin{figure}[h]\centering
	\includegraphics[width=0.8\textwidth]{Images/track_frames.pdf}
	\caption{Track frames corresponding to adjacent nodes of the grid over $\alpha$. Note that the grid points need not be equispaced. If desired, they can be chosen irregularly, so as to capture more features of the track where it is more critical (e.g. in proximity to corners) and save space where it is more regular (e.g. on the straights).}
	\label{fig:track_frames}
\end{figure}

%According the direct collocation method, the state evolution on the $k$-th element of the grid is approximated with a polynomial $p_k$. The number of collocation points depends on the degree $d$ of the polynomial chosen and their position inside the interval can be selected based on different criteria; here the Gauss-Legendre collocation points are used. The polynomial is built on the unit interval and then scaled to match the length $\nu_k$ of the relative time step.
%
%In Eq.~\eqref{eqn:equality}, besides specific constraints inserted to bias the solution towards desired behaviours, two equations are paramount. The first  one is:
%\begin{equation}\label{eqn:collocation}
%	\frac{1}{\nu_k}\frac{\partial}{\partial\tau}p_k(\tau_i)=f(v_{k,i},u_k),
%\end{equation}
%the collocation equation, that constrains the state at each collocation point to comply with the system dynamics. The second one, which is $p(1)=x_{k+1}$, ensures continuity across time intervals.

In addition to the classical collocation and continuity constraints~\cite{Domenighini:Designs:2023}, the inequalities in Eq.~\eqref{eqn:inequality} for our MLTP ensure the vehicle remains within track bounds. These include the requirement for the CoM to lie on the $yz$ plane of the frame $\sref{S_k}$, as well as power and adherence limits.

In our optimizations, the cost function at each step is expressed as follows:
\begin{equation}\label{eqn:cost}
	L_k=\nu_k^2+k_\delta(\delta_k-\delta_{k-1})^2+k_T(T_k-T_{k-1})^2+k_v x_{k,2}^2 + l_k(\ddot{q}_{hk,k},\delta_k,x_k),
\end{equation}
where the first term penalizes travel time, the second and third prevent abrupt variations of the inputs and avoid singular arcs. The quadratic penalization ensures a well-defined dependence of the Hamiltonian on the control variables, preventing indeterminate solutions. The last two serve to shape the driving behaviors according to the designer preferences.

Their primary goal is to improve the optimization landscape, ensuring that the solver is less likely trapped by undesirable local minima. As an example, the fourth term, which includes the lateral velocity $x_{k,2}$, has been included to prevent undesirable behaviors which could emerge as a result of the increased system's DoFs in combination with the simplified aerodynamic modeling adopted in this work. Some local minima could automatically be avoided by incorporating more realistic aerodynamic maps and tyre wear models. These enhancements will be explored in future works. In this context, an interesting direction is presented in~\cite{Masouleh:IEEE:2016}, where the authors introduce an aerodynamic model that explicitly accounts for the influence of suspension and vehicle configuration. This approach offers a more comprehensive representation of the aero forces and could be effectively used to refine the current aerodynamic map. By incorporating such a model, we expect not only a more physically consistent representation of the aerodynamics but also a smoother and more meaningful optimization landscape, reducing the occurrence of non-physical local minima and improving overall solution robustness.%More in details, the all steering vehicle, having the capability to steer independently with all four wheels, and considering that the aerodynamic drag only depends on the forward velocity $x_1$, could run with excessively large $\beta$ angles, even on straights paths. Even if this behavior could be appropriate in certain track sectors, is undesirable from an engineering perspective.
%In some cases, it could also be the outcome of the solver being stuck in local minima: in these events the above term helps discourage such behaviours.

In the current context, we imposed coefficient $k_v$ to be a function of the (absolute value of) track curvature $c_k$ according to Eq.~\eqref{eqn:lateral_cost}. Here, the two constant $\bar{k}_v$ and $k_c$ are selected heuristically for each track; the first increases/decreases the weight of the cost term, while the second varies the hardness of the intervention in the range $[0,1]$.
\begin{equation}\label{eqn:lateral_cost}
	k_v = \bar{k}_v(1-\tanh(k_c c_k))
\end{equation}

To ensure reproducibility, we report the values of the main penalty weights for the cost function terms used in the Siena track simulations: $k_{\delta} = 0.05$, $k_T = 0.005$, $\bar{k}_v = 0.005$, and $k_c = 100$.

### 2025-04-03 performance_comparison.tex
[MGa] basic structure

\section{Performance comparison}
\label{sec:performance_comparison}

In this section, some of the new vehicle architectures are compared to the traditional vehicle to enlighten the different driving styles that lead to a faster or a slower lap. The track selected for the comparisons is Siena, described previously and shown in Figure~\ref{fig:siena_track}. Siena track presents low-speed curves in which dynamic camber control (DCC) architecture or four independent steering wheels (AS) impact substantially. The track is run clockwise from the start point, represented by the bold red line in Figure~\ref{fig:siena_track}. The thin red lines are spaced by a $\De\al=0.1$, and the black lines, also spaced by $\De\al=0.1$, are shifted from the red ones by $\al_s=0.05$. Additionally, the plot shows the enumeration of the turns, from 1 to 7, which will be useful for further observations.

First of all, a traditional vehicle (RWD FS) is compared with an AIWD FS vehicle, capable of applying four independent torques to the wheels. As already mentioned, this feature provides the highest improvement in performance. In the analyzed case, the vehicle is able to apply not only four driving torques, each one completely independent of the others, but also four independent braking torques. This is possible if we assume the vehicle has four separate oil lines or a different braking system. For this type of vehicle, usually driven by a human, this design is never applied due to the driver's inability to manage four control inputs, such as pedal forces. The decision to maintain the four independent braking torques is made here considering that the vehicles under study are not (yet) existing models and can be thought of as new architectures envisioned for future autonomous vehicle competitions. In this case, having more inputs would not be a problem unless some limits in the real-time computing power of the ECU are encountered --- a situation not discussed here. In this work, in fact, we focused on identifying and analyzing these solutions, which serve as a benchmark for maximizing vehicle performance, while intentionally postponing their implementation in a real vehicle to future designs.
%(Questa parte sotto e' poco chiara. Quello che facciamo noi fornisce un riferimento su come si dovrebbe guidare. E' chiaro che poi ti serve un controllore in feedback per fare tracking oppure per fare una pianificazione dinamica con re-planning tipo MPC che gestisce le discrepanze. Tuttavia questi sono aspetti che nella analisi corrente non sono considerati perche' andrebbero valutati in seconda battuta. Pero' non terrei il tono un po' negativo che traspare da questa parte di testo. Direi semplicemente che questi aspetti sarebbero da analizzare in 'future work' e per 'mettere a terra' una soluzione che noi analizziamo a livello funzionale e di potenzialita'.)
%In general, in worth bearing in mind that these solutions cannot be implemented ``as is'' for feed-forward control for multiple reasons. The most important is that the errors arising from the modeling process and discretization lead to a solution that is almost correct, and thus applying the resulting inputs would generate a different evolution of the system, resulting in a significant final cumulative error. Therefore, having a controller would be mandatory, and the high number of control channels would call for appropriate analyses. %The results of this case will justify the decision to separately compare RWD and AWD vehicles.

\begin{figure}[]
	\centering
	\includegraphics[scale = .85]{Images/DrivingStyles/siena_track.pdf}
	\caption{Siena Circuit. The bold red line indicates the starting point, while the other red lines are equally spaced by a $\De\al=0.1$. The black lines are also spaced like the red ones but are shifted by $\al_s=0.05$. The track is run in a clockwise direction. The curves, consisting of 2 to the left and 5 to the right, are sequentially numbered based on the order of encounter.}
	\label{fig:siena_track}
\end{figure}

\subsection{All Independent Wheel Drive, Forward Steering Vehicle - AIWD FS}
\label{sec:awdfs}
To perform a valid comparison between vehicles with different powertrain, they are assumed to have the same maximum (and maximum negative, i.e. braking) power. While the RWD vehicle, modeled with an open differential, will have identical torques on the driving wheels to generate power, the AIWD vehicle will  optimally distribute the power among four wheels, maintaining the same total value of the RWD vehicle.

For most turns, the two vehicles follow different trajectories, with the AIWD staying significantly closer to the inside boundary of the track during turns, as it can be observed in Figure~\ref{fig:rwdfs_awdfs_uayey}, third panel. See also the accompanying video~\cite{video:AWDFS:2024}.
The third plot, in fact, depicts the evolution of the variable $e_y (m)$, which represents the lateral distance, described in $\sref{S}$ components, of the CoM with respect to the center-line, at the corresponding $\al$. In a left turn, positive values of $e_y$ correspond to points that are closer to the inner boundary. Vice versa, in a right turn, negative values correspond to points that are closer to the inner boundary. Each of the three plots shown in Figure~\ref{fig:rwdfs_awdfs_uayey} includes green dash-dot lines marking the turns, while the forward speed plot (first panel) additionally displays the turn numbers, with labels placed over double-ended arrows. The enumeration follows the one described in Figure~\ref{fig:siena_track}, and each label indicates a letter that identifies the direction of the curve (L for left turns, R for right turns). Again from Figure~\ref{fig:rwdfs_awdfs_uayey}, third plot, we can observe that the AIWD vehicle begins to move toward the inner boundary of the track before the RWD vehicle. This behavior is evident in turn 2, which starts at $\al=0.2$, where the AIWD vehicle (blue line) reaches the inner boundary before the RWD (red line), with a difference of $\de\al=0.01$. On this track, this difference in curvilinear abscissa corresponds to 12.9\,m. Consequently, as the vehicles approach turn 2, the AIWD vehicle reaches the inner boundary approximately 13\,m ahead of the RWD vehicle, indicating a significantly different driving style.
%Di cosa stai parlando? dove devo guardare? a quale $\al$? non si segue...
%(quale e' la turn 2? cosa deve guardare il lettore? cosa sono le 1R, 2R, 3L sopra nel primo grafico? non sono mai state introdotte o spiegate)
\begin{figure}[h]
	\centering
	\includegraphics[scale = .8]{Images/DrivingStyles/rwdfsaiwdfs_p.pdf}
	\caption{
		Forward speed $u$, lateral acceleration in body components $a_y$ and $e_y$ evolution for RWD FS and AIWD FS vehicle, in red and blue respectively, over curvilinear abscissa $\al$. The quantity $e_y$ represents the lateral distance, in $\sref{S}$ reference, of the CoM with respect to the center-line at the corresponding $\al$.
		%Positive values in a left turn correspond to points that are closer to the inner limit, and vice versa in a right turn, positive values correspond to points that are closer to the outer limit.
		The turns are indicated in the forward speed plot (first panel) with the corresponding number (see Figure~\ref{fig:siena_track}) and a letter that identifies the direction (L for left turns and R for right turns). The more positive the values, the closer the trajectory to the left edge of the track, while the more negative the values, the closer the trajectory to the right edge.}
	\label{fig:rwdfs_awdfs_uayey}
\end{figure}

Still referring to Figure~\ref{fig:rwdfs_awdfs_uayey}, in particular to the first plot, which shows the forward speed $u$ over the curvilinear abscissa $\al$, we notice that, during cornering, the AIWD vehicle (blue line) approaches the point of minimum speed earlier than the RWD vehicle (red line).
The combination of this trend, and the previously analyzed tendency to remain closer to the inner boundary while cornering, allows the AIWD vehicle to exit with higher forward speed than the RWD vehicle. This affects positively the lap time.

In the second plot of Figure~\ref{fig:rwdfs_awdfs_uayey}, which reports the lateral acceleration $a_y$ over the curvilinear abscissa $\al$, we can notice that during cornering the AIWD vehicle (blue line) maintains a higher mean absolute value of the lateral acceleration.

In particular, traveling the chicane composed by turn 3 and 4, starting at $\al = 0.27$, the lateral acceleration $a_y$ of the AIWD FS vehicle (blue line) is greater (in magnitude) than the lateral acceleration of the RWD FS vehicle (red line), except for the small range, in which $\al\in[0.27, 0.28]$.

Still referring to the second plot of Figure~\ref{fig:rwdfs_awdfs_uayey}, and  focusing on turn 4, starting approximately at $\al = 0.33$, the lateral acceleration $a_y$ of the AIWD vehicle has smaller variation than the lateral acceleration of the RWD vehicle, maintaining almost a constant value during cornering, denoting a smoother behavior.
%This behavior can be explained by considering that having four driving torques allows the vehicle to achieve a higher absolute yaw moment, and hence a higher yaw rate $r$, which results in higher lateral acceleration $a_y$. In fact if we consider, for simplicity, a 2D vehicle model for handling, as the one described in~\cite[Chapter~3.2]{Guiggiani:book:2023}, with the hypothesis listed in the introduction of the chapter, is possible to express the lateral acceleration as follows:
%\begin{equation}\label{eq:ayguiggiani}
%a_y = \dot{v} + ur,
%\end{equation}
%where $u$, $v$ and $r$ are respectively, longitudinal and lateral velocity, and yaw rate.

%Apart from the different trajectory, the AWD vehicle pushes the wheels close to the grip limit not only in the lateral direction but also in the longitudinal direction, thereby generating greater tangential forces.

%Qui non si capisce dove vuoi arrivare. Il ragionamento (o la spiegazione associata) non e' chiaro. Hai richiamato la formula della $a_y$ per cosa? Cosa aumenta diminuisce nei due termini? Non si capisce, prova a rileggerti cercando di desumere il messaggio solo da cio' che e' scritto. %Considering this evidence, we can affirm that the two vehicles have significantly different driving styles, and thus the effects of the added features could be misinterpreted.

\subsection{Rear Wheel Drive, All Steering Vehicle - RWD AS}
We can observe in Figure \ref{fig:RWD_AS_toes}, first plot, the road wheel angles $\de_{i,j}$ of the four wheels with respect to the curvilinear abscissa $\al$ for the RWD AS vehicle. It is worth recalling that, during cornering, this vehicle can exhibit different road wheel angle values for the wheels of the same axle.

In the first plot, the color code is such that wheels of the same axle share the color (red for the front axle, blue for the rear axle), while wheels of the same side share the same line style (solid line for the left side, dashed line for the right side). In the second plot, the forward speed $u$ with respect to the curvilinear abscissa $\al$ is depicted.
In both plots, the green vertical lines mark the start and end point of the turns. In the lower plot, the turns are labeled with a progressive number, according to Figure~\ref{fig:siena_track} and a letter indicating its direction (L for left-hand turns and R for right-hand turns).

Although a slight penalization is applied to the cost function to avoid discordant wheel angles on the same axle during turns, the independent behavior of the wheels allows for better control of ground forces and their direction, enhancing turning performance. This is especially noticeable in turn 5 (right turn), where the left wheels steer more than the right, and in turn 6 (left turn), where the right wheels steer more than the left. See also the accompanying video~\cite{video:RWDAS:2024}. Generally, the outer wheel, which bears a greater vertical load due to lateral transfer, has a larger road wheel angle thus generating more lateral force, while the inner wheel steers less since it contributes less to the total lateral force.
%\begin{figure}[h]
%	\centering
%	\includegraphics[scale = .65]{Images/DrivingStyles/rwdfs_rwdas_toeForces.pdf}
%	\caption{Toe angle $\de$, $F_x$ and $F_y$ forces for the front left wheel of a RWD FS and a RWD AS vehicle, in red and blue respectively. Dashed lines in the forces plots indicate the maximum and minimum force reachable.}
%	\label{fig:rwdfs_rwdas_toeForces}
%\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[scale = .8]{Images/DrivingStyles/RWD_AS_toes.pdf}
	\caption{Road wheel angle $\de$ of all wheels (above), and forward speed $u$ (below), evolution over curvilinear abscissa $\al$. As reported in the legend, wheels of the same axle share the same color, while wheels of the same side (left/right) share the same line style. The green dash-dot lines represent the starting and ending point of turns. The turns are indicated in the forward speed plot with the corresponding number (see Figure~\ref{fig:siena_track}) and a letter that identifies the direction (L for left turns and R for right turns).}
	\label{fig:RWD_AS_toes}
\end{figure}

Another aspect worth noting is the behavior of the rear wheels in relation to the front ones. In fact, the rear wheels steer in phase with the front wheels for most of the corners and for most of the corner entrances, except for some cases where, in the slower part of the turn, the rear wheels steer in counterphase with the front wheels.
In particular, referring again to Figure~\ref{fig:RWD_AS_toes}, it is possible to observe that in turn 2-5-7 the front and rear road wheel angles are concordant, while in the remaining turns (1-3-4-6), the road wheel angles are discordant for most part of the turn.
This behavior is generally consistent with Guiggiani's analysis of MAP's (Map of Achievable Performance) for a Single-Track model of vehicle. In fact, in~\cite[Section~6.8.1]{Guiggiani:book:2023} the author affirms that a desired behavior is to have a vehicle where the $\be$ angle spans a small range.
This can be achieved by having in-phase rear steering at high forward speeds and counter-phase rear steering at low forward speeds.

Although a general rule like this may not hold true for every corner on the track, it nonetheless represents the desired trend to optimize performance, which is here confirmed by quantitative lapsim outputs.
It also may be worthwhile to point out that a human driver can drive differently with respect to the optimizer, due to the fact that it lacks the perfect view of the track beforehand which, on the other hand, is profitably exploited by the optimizer.
%This in also confirmed if we observe that the amplitude of the beta angle, slightly reduced in many turns.
%\begin{figure}[h]
%	\centering
%	\includegraphics[scale = .65]{Images/DrivingStyles/rwdas_toe.pdf}
%	\caption{Toe angle $\de$ for the front and rear left wheel, and forward speed $u$ for a RWD AS vehicle. (For the toe angle, only the left side is plotted for clarity and readability, considering that the wheels on the same axle steer in-phase with each other)}
%	\label{fig:rwdas_toe}
%\end{figure}

%The last aspect we discuss here is the approach to the turns. The AS vehicle brakes less than the FS vehicle and approaches the corners with a higher forward speed. This behavior can be explained by considering that, with better control of the ground forces, the vehicle reaches the grip limit at higher velocities and lateral accelerations. Not only does the higher limit contribute to this, but having four steered wheels also generates a higher resistance force. This becomes clear when we decompose the four lateral forces of the wheels in the vehicle reference system.

\subsection{Rear Wheel Drive, Forward Steering, Dynamic Camber Control - RWD FS DCC}
The ability to adjust the camber of each wheel during cornering allows the vehicle to maximize the ground forces even more. In fact, in a traditional vehicle, the position and orientation of the wheels during a turn are determined by the passive motion of the suspension, in addition to the steering system for the front wheels. This means that, even if the suspension is well designed, the position of the wheels will not be always the best for performance and could be improved in each turn.

The first plot of Figure~\ref{fig:rwdfs_rwdfsdcc_gammaFyFR} shows the camber angles $\ga$ of the front right wheel (FR) over the curvilinear abscissa $\al$ for a RWD FS vehicle (red line) and for a RWD FS DCC vehicle (blue line), respectively.
The second plot of Figure~\ref{fig:rwdfs_rwdfsdcc_gammaFyFR} shows the lateral force $F_y$ of the front right wheel (FR) for the RWD FS vehicle (red solid line) and for the RWD FS DCC vehicle (blue solid line). Furthermore the maximum and minimum achievable lateral forces are shown in dashed lines and following the colors of the solid lines (RWD FS in red, RWD FS DCC in blue). Both plots show green dash-dot lines to delimit turns, while the first plot, in addition, reports labels with turn numbers (that follow the enumeration of Figure~\ref{fig:siena_track}), and an indication of the turn direction (L for left turns, R for right turns).

It is possible to note that, during cornering, e.g. turn 3, the camber tends to settle around a value of $-7.5$~degrees which contributes to increasing the lateral force by 128 N. This corresponds to a 7.6\% increase w.r.t. the baseline, which can be considered a significant achievement (see the zoomed area in the second plot at turn 3, where $\al\in [0.26, 0.33]$). See also the accompanying video~\cite{video:RWDFSDCC:2024}.

Similarly, in turn 6, starting at $\al = 0.69$, it can be observed that the maximum lateral force $F_y$ achievable is higher for the RWD FS DCC vehicle, implying a better wheel positioning relative to the road.
We remind that, according to the convention of positive lateral forces $F_y$ pointing to the left of the vehicle, ``the higher (more positive) the better'' is valid if we analyze a left turn, such as turn 3 or 6. Otherwise, in a right turn, an improvement on force limits would correspond to a lower minimum for achievable lateral force.
For this reason, considering here the front right (FR) wheel, we focus only on left turns.
%In the Figure \ref{fig:rwdfs_rwdfsdcc_gammaFyFR}, showing the camber angle $\ga$ evolution (above) and the lateral force $F_y$ (below) of the front right wheel over the curvilinear abscissa $\al$, the limits, indicated by dashed lines and coded with the same color of the actual forces, are higher than those of the traditional vehicle, implying a better wheel positioning relative to the road. This is more evident in the left turns, in which this wheel has a higher vertical load.

\begin{figure}[h]
	\centering
	\includegraphics[scale = .8]{Images/DrivingStyles/rwdfs_rwdfsdcc_gammaFyFR.pdf}
	\caption{Camber angle $\ga$ (first plot), and lateral force $F_y$ (second plot) over curvilinear abscissa $\al$, for the front right wheel (FR) of a RWD FS (red line) and a RWD FS DCC (blue line) vehicle. The dashed lines represent the upper and lower limit of the forces for each vehicle, coded by the same color. The green dash-dot lines represent the starting and ending point of turns. The turns are indicated in the forward speed plot with the number and the letter that identifies the direction (L for left turns and R for right turns). The zoomed area reports the section with $\al\in [0.26, 0.33]$ to highlight the improvement of $\simeq 130$\,N in the lateral force achievable by Dynamic Camber Control (DCC).}
	\label{fig:rwdfs_rwdfsdcc_gammaFyFR}
\end{figure}

%The vertical loads from the optimization, combined with the tyre model, are used to analyze the behavior of the camber angle during cornering. In particular, the goal is to verify that the value assumed by the camber angle guarantees the maximum lateral force. In Figure~\ref{fig:dcc_lateralforce_gammaspan}, a zoomed-view of the lateral force $F_y$ of the front right wheel (FR) in first part of the turn 3 ($\al\in[0.268, 0.286]$) over the curvilinear abscissa $\al$ is presented. The solid curve in black represents the optimal lateral force derived from the optimization, obtained with a value of the camber angle $\gamma^*(\al)$, varying with the curvilinear abscissa. The dashed red curve represents the potential force obtained with the minimum value of the added camber $q_{hk,\text{max}}=9$~degrees (see Section~\ref{sec:cambersystemmodeling}). The other colored lines represent the lateral forces obtained with the same vertical load $F_z(\al)$ derived by the optimization, and varying the total camber angle $\ga\in[\ga^*(\al)-\De\ga,\ga^*(\al)+\De\ga]$, with $\De\ga = 4$ degrees. As we can observe, by varying $\ga$ such that $\ga = \ga^*(\al) + \de\ga$, with the value of $\de\ga$ coded by colors in the bar on the right side, the lateral force presents a peak that, especially in the second part of the plot, doesn't correspond to $\de$ equal to the limit value $\pm\De\ga$. After the first part of the curve, which presents a delay due to the fact that we command the second derivative of $q_{hk}$, the solution, adjusting the additional camber, coasts the maximum value of the force.

%\begin{figure}[h]
%	\centering
%	\includegraphics[scale = .65]{Images/DrivingStyles/dcc_lateralforce_gammaspan}
%	\caption{Reachable lateral forces $F_y$ over curvilinear abscissa $\al$ when spanning the camber angle $\ga\in[\ga^*(\al)-\De\ga,\ga^*(\al)+\De\ga]$, equivalent to a window of $\pm 4$ degrees centered on the optimal value $\ga^*(\al)$, which varies with curvilinear abscissa $\al$. The black line indicates the optimal value of the force, while the red dashed line indicates the force obtained with the maximum, or minimum, depending on the sign, value of the additional camber $\De\ga$. The colored lines represents level curves with constant $\De\ga$, but, in general, different $\ga$.}
%	\label{fig:dcc_lateralforce_gammaspan}
%\end{figure}

For brevity, we only mention that the RWD AS DCC vehicle combines the advantages of the previous two, thereby guaranteeing the highest performance. Similarly to the RWD AS vehicle, it steers more with the outer wheels thus exploiting the lateral load transfers; at the same time, as registered for the RWD FS DCC vehicle, it adjusts the camber during cornering to settle around the value of maximum ground force available.

%In Figure~\ref{fig:rwdfs_rwdasdcc}, upper plot, the evolution of the road wheel angles $\de_{i,j}$ with respect to the curvilinear abscissa $\al$ is depicted using the same conventions of Figure~\ref{fig:RWD_AS_toes}.
%The lower plot shows the lateral force $F_y$ evolution over the curvilinear abscissa $\al$ and the force bounds, with the same layout of Figure~\ref{fig:rwdfs_rwdfsdcc_gammaFyFR}, for a RWD FS and RWD AS DCC vehicle, depicted in red and blue, respectively.
%
%With reference to the upper plot, we can observe the behavior noted for the RWD AS vehicle, which steers more with the outer wheels thus exploiting the lateral load transfers.
%Similarly to the RWD FS DCC, this vehicle adjusts the camber during cornering to settle around the value of maximum ground force available. Still referring to turn 3, as done in the zoomed area of Figure~\ref{fig:rwdfs_rwdfsdcc_gammaFyFR}, the maximum increase of the force grows now up to 12.1\%. These two effects combined also allow the vehicle to navigate corners with higher driving torque on the rear wheels or, in general, less braking torque on all four wheels, thus resulting in a higher forward speed.
%\begin{figure}[h]
%	\centering
%	\includegraphics[scale = .8]{Images/DrivingStyles/rwdfs_rwdasdcc.pdf}
%	\caption{Road wheel angles $\de_{i,j}$ of all wheels of a RWD AS DCC vehicle over curvilinear abscissa $\al$ (upper plot), and lateral force $F_y$ over curvilinear abscissa $\al$ (lower plot), for the front right wheel (FR) of a RWD FS (red line) and a RWD AS DCC (blue line) vehicle. The dashed lines represent the upper and lower limit of the forces for each vehicle, coded by the same color.}
%	\label{fig:rwdfs_rwdasdcc}
%\end{figure}

\section{AIWD 0S DCC: in depth analysis}
This section aims to showcase that on a track with characteristics like the N\"{u}rburgring section used or Siena, whose results we present here, a vehicle able of applying four independent acceleration torques (AIWD) and dynamically controlling the camber (DCC) has a very significant advantage in terms of lap time and cornering performance.

To stress the power of these features, a vehicle with such characteristics but \emph{without steering abilities} (0S, zero steering) has been considered. See also the accompanying video~\cite{video:AWDOSDCC:2024}.
Cornering is still possible if the vehicles applies different torques to the wheels on the same axle, as we can observe in Figure~\ref{fig:awd0sdcc_deltaT}.
Here, the evolution of the torques applied to the wheels $T_{i,j}$ over the curvilinear abscissa $\al$ for the front axle (upper plot) and the rear axle (lower plot) is illustrated.
The left wheels (FL and RL) are plotted with red lines, while the right wheels (FR and RR) are plotted with blue lines. Similarly to the previous figures, the green dash-dot lines in both plots delimit the turns, while the labels in the first plot denote the turn number, according to Figure~\ref{fig:siena_track}, and a letter which discriminates the direction (L for left-hand turns, R for right-hand turns).

During turns 2-5-6-7, the torque of the inner wheel settles to zero, as is possible to observe in the first panel of Figure~\ref{fig:awd0sdcc_deltaT}. Here, the inner wheel starts a free rolling motion, with no applied torque by the motor.

Otherwise a different behavior emerges in turn 4, starting around $\al=0.33$. Here, the two wheels experience torques that are approximately mirror-symmetric with respect to the horizontal line $T=100$\,Nm, with the inner torque (FR, blue line) that assumes negative values closely to $\al=0.33$, without losing contact to the ground.
\begin{figure}[h]
	\centering
	\includegraphics[scale = .8]{Images/DrivingStyles/awd0sdcc_deltaT.pdf}
	\caption{Wheel torques $T_{i,j}$ evolution over curvilinear abscissa $\al$, for the front wheels (above) and for the rear wheels (below) of a AIWD 0S DCC vehicle (0S indicates incapability of steering). The red lines refer to the left side while the blue lines refer to the right side. As for the previous plots, the green dash-dot lines indicates the turns, labeled in the first plot by the number and a letter L, for left turns, or R, for right turns.}
	\label{fig:awd0sdcc_deltaT}
\end{figure}

While the AIWD FS DCC vehicle completes a lap in 51.40 seconds, the AIWD 0S DCC vehicle does so in 51.64 seconds -- only 0.24 seconds slower but 3.66 seconds faster than a traditional RWD FS vehicle. This is achieved by properly adjusting camber and applying four independent torques, thus compensating for the lack of steering. The torque vectoring logic, used in both AIWD FS DCC and AIWD 0S DCC, plays a key role in optimizing turning maneuvers. By simply adjusting the intervention to avoid steering, the lap is completed with a 0.47\% increase in lap time compared to the AIWD FS DCC.
It is worth mentioning in passing that the model implements power limits as total values, neglecting potential motor saturation, and may lead to an overestimation of performance.
Therefore, this analysis should not be taken as a direct guideline for designing a AIWD 0S DCC vehicle but an informative case study highlighting factors relevant to cornering and steering.

%While an AWD FS DCC vehicle completes a lap in 51.40\,s, this vehicle (here named AWD 0S DCC) completes a lap in 51.64\,s, thus with only 0.24\,s more, but with 3.66\,s less than a traditional vehicle (RWD FS). This is possible because having the ability of adjust the camber and apply 4 independent torques compensate for the lack of steering. In particular the torque vectoring logic, used also for AWD FS DCC, provide the most for the turning maneuver, and hence, simply adjusting this intervention to avoid steering the vehicle completes the lap without significant problems. It's important to underline that, in the model, the power limits are implemented as total values, hence neglecting potential saturation of power for individual motors. This approach has been employed to compare different traction configurations (RWD/AWD); however, it may lead to an overestimation of achievable conditions. Nonetheless, this analysis should not be used as a guideline for designing a real AWD 0S DCC vehicle but rather serves as an informative case study that highlights various factors during cornering, including those related to steering.

### 2025-04-03 stochastic_vehicle_dynamics.tex
[MGa] basic structure

--- 2025-04-03 closed_loop_planning.tex
@@
-\section{Closed-loop uncertainty-aware planning over the full horizon}
-\label{sec:closed_loop_planning}
+\section{Stochastic vehicle dynamics}
+\label{sec:stochastic_vehicle_dynamics}
 
-Building upon our previous paper that presented a 3D multibody model suitable for embedding within an MLTP problem~\cite{Domenighini:Designs:2023} - treated here as the \emph{baseline} model - two models with increasing complexity are presented.
-The first represents a vehicle capable of independently steering all four wheels (AS) and applying four independent torques to each wheel (AIWD, aka, torque vectoring), whereas the second model additionally incorporates an active camber variation system for each wheel (DCC).
-%The kinematic tree of the two models is slightly different, but the procedure used to build the dynamic model is similar apart from a non-trivial aspect that will be discussed later on.
-In the remainder of this study, unless otherwise specified, the descriptions provided apply to both models.
-The figures related to Sections~\ref{sec:suspanal}-\ref{sec:tyremodel} are courtesy of the authors.
+Some text of a previous paper left for reference on equation and pseudo-code formatting.
 

### 2025-04-07 stochastic_vehicle_dynamics.tex
[MG] just started with notation

@@
 
-Some text of a previous paper left for reference on equation and pseudo-code formatting.
+We model the \emph{perturbed vehicle dynamics} as the following nonlinear continuous-time random dynamical system
+\begin{align}
+\dbx(t) = f(\bx(t), \bu(t)) + \bw(t),
+\end{align}
+where $\bu(t)$ are the deterministic control inputs, $\bw(t)$ is additive Gaussian white noise with zero mean and known covariance $\bQ(t)$, i.e.
+$\bw(t) \sim \calN(\bzero, \bQ(t))$, and the state $\bx(t)$ results in a Gaussian distribution with mean $\bmu(t)$ and covariance $\bP(t)$, so that $\bx(t) \sim \calN(\bmu(t), \bP(t))$.
 
-%% CHANGE THESE SECTIONS TO MERGE BOTH MODELS DESCRIPTIONS
-\subsection{Kinematic Tree Structure of the Vehicle}
-\begin{figure}[b]\centering
-	\begin{tikzpicture}[scale=1.2]\small
-		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw,gray](g)at(-0.1,-0.1){};
-		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](b)at(0,2){};
-		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](h11)at(-0.9,2.9){};
-		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](h12)at(0.9,2.95){};
-		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](h21)at(-1,1){};
-		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](h22)at(0.95,1){};
-		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](k11)at(-2.0,3){};
-		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](k12)at(2.0,3){};
-		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](k21)at(-2.1,1.1){};
-		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](k22)at(2.0,1.05){};
-		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](w11)at(-3.0,2.9){};
-		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](w12)at(3.0,2.9){};
-		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](w21)at(-3.1,1.){};
-		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](w22)at(3.0,.95){};
-		\draw[gray](g)to node(2){}(b);
-		\draw(b)to node(4){}(h11);
-		\draw(b)to(h12);
-		\draw(b)to(h21);
-		\draw(b)to(h22);
-		\draw(h11)to node(8){}(k11);
-		\draw(h12)to(k12);
-		\draw(h21)to(k21);
-		\draw(h22)to(k22);
-		\draw(k11)to node(6){}(w11);
-		\draw(k12)to(w12);
-		\draw(k21)to(w21);
-		\draw(k22)to(w22);
-		\draw(g)node[below,gray](1){$\{G\}$}(b)+(0,0.1)node[above](3){$\{B\}$}(h11)node[above](5){$\{H_1\}$}(h12)node[above]{$\{H_2\}$}(h21)node[below]{$\{H_3\}$}(h22)node[below]{$\{H_4\}$}
-		(k11)node[above](9){$\{K_1\}$}(k12)node[above]{$\{K_2\}$}(k21)node[below]{$\{K_3\}$}(k22)node[below]{$\{K_4\}$}
-		(w11)node[above](7){$\{W_1\}$}(w12)node[above]{$\{W_2\}$}(w21)node[below]{$\{W_3\}$}(w22)node[below]{$\{W_4\}$};
-		\node[gray,inner sep=1pt](world)at(1,-0.3){world};
-		\node[gray,inner sep=0pt,align=center](6DoF)at(-1.25,0.15){floating base\\(virtual 6-DoF)};
-		\node[inner sep=1pt](chassis)at(0.05,3){chassis};
-		\node[inner sep=0pt,align=center](susp)at(-2.2,1.6){suspension linkages\\(generalized 1-DoF)};
-		\node[inner sep=1pt](knuckle1)at(-.4,3.7){knuckles 1};
-		\node[inner sep=1pt](knuckle2)at(-2.05,3.8){knuckles 2};
-		\node[inner sep=0pt,align=center](hub)at(-4.5,2.25){hub bearings\\(revolute 1-DoF)};
-		\node[inner sep=0pt,align=center](camber)at(-2.,2.4){camber joint\\(revolute 1-DoF)};
-		\node[inner sep=1pt](rim)at(-3.4,3.65){rims};
-		\draw[gray](world.west)to(1.east);
-		\draw[gray](6DoF.north east)to(2);
-		\draw(chassis.south)to(3);
-		\draw(susp.north east)to(4);
-		\draw(knuckle1.south)to(5);
-		\draw(knuckle2.south)to(9);
-		\draw(hub.east)to(6);
-		\draw(camber.north)to(8);
-		\draw(rim.south east)to(7);
-	\end{tikzpicture}
-	\caption{Kinematic tree representing the interconnections between bodies. Bodies are referred to by their attached frame.}
-	\label{fig:tree}
-\end{figure}
-The vehicle is modeled as an articulated rigid-body system with chassis, knuckles, and wheels as rigid bodies.
-The chassis represents all sprung masses rigidly connected, while the knuckles (zero-mass bodies) decouple steering and vertical suspension motion from the wheel spindle.
-In the second model, knuckles are split into two zero-mass bodies, introducing an additional degree of freedom per wheel to simulate dynamic camber control (DCC).
-Wheels include all rotating unsprung masses, with non-rotating parts (e.g., brake calipers) incorporated into the wheel-tire mass.
+Due to the symmetry of the probability density function (pdf) with respect to $\bmu(t)$, it is possible to represent the time evolution of the pdf as: i) the deterministic evolution of the mean $\bmu(t)$
+\begin{align}
+\dbmu(t) = f(\bmu(t), \bu(t)),
+\end{align}
+and ii) the time evolution of the state covariance matrix $\bP(t)$ along $\bmu(t)$, which can be expressed as
+\begin{align}
+\dbP(t) = \bA(t) \bP(t) + \bP(t) \bA^T(t) + \bQ(t),\quad \bP(0) = \bP_0.
+\end{align}
+The \emph{initial} state covariance $\bP_0$ is assumed known and $\bA(t)$ is the classical shorthand for the Jacobian along the mean trajectory $\bmu(t)$, that is $\bA(t)=\bJ(\bmu(t),\bu(t))$, where $\bJ(\bx, \bu) = \frac{\pd f(\bx, \bu)}{\pd \bx}$.
 
-Each body has a right-handed barycentric frame: $\sref{B}$ (chassis), $\sref{H}$ and $\sref{K}$ (knuckles), and $\sref{W}$ (wheels).
-Additional frames $\sref{G}$ (ground) and $\sref{S}$ (centerline) are defined, with \(x\)-axis of $\sref{S}$ tangent to the centerline, pointing forward.
-Frame relationships are described by homogeneous transformation matrices $ g(d, \vPhi) \in SE(3) $, with $ d $ as position and \(\vPhi\) (Euler angles) parameterizing the rotation.
-
-As shown in Fig.~\ref{fig:tree}, bodies are connected by joints in a kinematic tree.
-The second model generalizes the first, where $\sref{H}$ and $\sref{K}$ coincide, so only the second model's kinematic tree is shown for brevity.
-%% START ORIGINAL
-%The vehicle is modeled as an articulated rigid-body system, with chassis, knuckles, and wheels treated as rigid bodies.
-The chassis includes all sprung masses, rigidly connected, while the knuckles (modeled as zero-mass bodies) decouple steering and vertical suspension motion from the wheel spindle.
-In the second model, the knuckles are split into two zero-mass bodies, adding one degree of freedom per wheel to simulate dynamic camber control (DCC).
-The wheels encompass all rotating unsprung masses, with non-rotating parts (e.g., brake calipers) included in the wheel-tire mass.
-%
-%Each body has a right-handed barycentric frame: $\sref{B}$ for the chassis, $\sref{H}$ (and $\sref{K}$ in the second model) for the knuckles, and $\sref{W}$ for the wheels.
-Two additional frames, $\sref{G}$ (ground) and $\sref{S}$ (attached to the vehicle's centerline), are also defined.
-The frame $\sref{S}$ lies in the tangent plane, with its $x$-axis tangent to the centerline, pointing forward.
-%
-%Relative positions and orientations between frames are described by homogeneous transformation matrices \( g(d, \vPhi) \in SE(3) \), where \( d \in \mathbb{R}^3 \) is the Cartesian position and \( \vPhi \in \mathbb{R}^3 \) are the Euler angles parameterizing the rotation matrix \( R \in SO(3) \).
-%
-%The bodies are connected by joints, forming the kinematic tree shown in Fig.~\ref{fig:tree}.
-The first model is a special case of the second, where $\sref{H}$ and $\sref{K}$ are always coincident, so only the kinematic tree for the second model is presented to avoid redundancy.
-
-%% END ORIGINAL
-
-%The vehicle is modeled as an articulated rigid-body system.
-In particular, chassis, knuckles and wheels are considered as rigid bodies.
-The chassis comprises all the sprung masses, including the driver, which are rigidly attached to one another.
-The knuckles are assumed here as zero-mass bodies, and their function is to decouple the motion of the suspension due to steering and vertical travel to the wheel spindle.
-In the second model, the knuckles are split into two zero-mass bodies, introducing one additional degree of freedom (DoF) for each wheel that allows to simulate the active camber variation system, here named dynamic camber control (DCC).
-The wheels include all the rotating unsprung masses, while all non-rotating parts of the unsprung masses, such as the brake calipers are included in the wheel-tire mass.
-%
-%Each body has a right-handed barycentral reference frame attached.
-The frames are denoted as $\sref{B}$, $\sref{H}$ (and $\sref{K}$) and $\sref{W}$ for the chassis, knuckles and wheels, respectively, where $\sref{K}$ is present only in the second model.
-Two additional frames are introduced: $\sref{G}$, fixed to the ground, and $\sref{S}$, that follows the vehicle on the ground.
-More precisely, $\sref{S}$ is attached to the centerline curve with its $x$ and $y$ axes lying in the tangent plane, and the $x$ axis also tangent to the centerline curve pointing forward.
-%
-%The relative positions and orientations between reference frames are described by homogeneous transformation matrices $g(d,\vPhi)\in SE(3)$, where $d\in \bbR^3$ is the relative Cartesian coordinates vector and $\vPhi \in \bbR^3$ are the three Euler angles that parameterize the rotation matrix $R\in SO(3)$.
-%
-%The bodies are connected by joints and the kinematic tree of the second model is shown in Figure \ref{fig:tree}.
-It is worth noting that the first model can be seen as a particular case of the second, in which the reference frames $\sref{H}$ and $\sref{K}$ are constrained to be always overlapping.
-For this reason, and to avoid cluttering the document, has been chosen to show the kinematic tree only for the second model.
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[scale = 0.8]{Images/vehicle_frames}
-	\caption{Reference frames: $\sref{G}$ is a fixed inertial reference frame with the $z$-axis antiparallel to the gravity vector; $\sref{S}$ is a reference frame following the vehicle on track; $\sref{B}$ and $\sref{K}$ are FLU (Forward Left Up) reference frames aligned with the principal axes of inertia of the chassis and wheel, respectively; $\sref{H}$ is an auxiliary reference frame necessary to add the camber DoF; and $\sref{W}$ rotates about the $y$-axis with respect to $\sref{K}$ by the wheel rotation angle.}
-	\label{fig:frames}
-\end{figure}
-
-The joints modeled are: hub bearings, connecting the rims $\sref{W}$ to the knuckles $\sref{H}$ (or $\sref{K}$ for the second model), and the suspension linkages, connecting the knuckles $\sref{H}$ to the chassis $\sref{B}$.
-For the second model a 1 DoF revolute joint connecting the two parts of the knuckles is also present.
-The chassis is a floating-base and its position and orientation with respect to the ground are unconstrained.
-This situations is conveniently encoded by a virtual six-DoF joint connecting $\{B\}$ to $\{G\}$.
-The resulting systems have 4+4+6=14 and 4+4+4+6=18 degrees of freedom, respectively.
-%After defining the system's topology, a set of suitable state variables needs to be defined to describe the configuration and motion of each body.
-
-%To fully describe the position and orientation of the frame $\sref{B}$ w.r.t. $\sref{G}$ we define vector $q_{gb}=(d_{gb};\varPhi_{gb})$.
-%Its first-order kinematics is encoded in twist $V_{gb}^b = (v_{gb};\omega_{gb})$.
-In general, a twist $V_{ij}^k$ represents the motion of the frame $\sref{J}$ w.r.t.
-the frame $\sref{I}$ in $\sref{K}$ components.
-Unless otherwise specified, we employ \emph{body} coordinates, therefore $k=j$; this choice aligns to ABA convention~\cite{Featherstone:book:2008}.
-For each of the 4 DoF's associated to the suspension motion, a variable $z$ (and its derivative $\dot{z}$) is defined representing spring's length variation with respect to the nominal value.
-With these definitions, along with variables $\delta$ and its derivatives, defined in Sec.\ref{sec:suspanal}, one can find the six-component vector $q_{bh}$ representing the pose of $\sref{H}$ w.r.t.
-$\sref{B}$ and the twist $V_{bh}$.
-Assuming the wheel is an axis-symmetrical rigid body, and given that the relative orientation of the frame $\sref{W}$ w.r.t.
-$\sref{H}$ (or $\sref{K}$) is not relevant for the scope, only the angular velocity $\omega$ is introduced.
-For the second model, it is also necessary to introduce the variable $q_{hk}$ to describe the relative position of the frame $\sref{K}$ w.r.t.
-$\sref{H}$.
-\subsection{Suspension Analysis}
-\label{sec:suspanal}
-The suspensions connect the chassis to the unsprung masses.
-This connection cannot be described by an elementary joint, such as prismatic or revolute.
-Therefore, a set of six variables $q_{bh}$, that parameterize a homogeneous transformation matrix, is necessary.
-The transformation can be realized by composing three translations and three rotations (the ZXY Euler sequence was selected) using the global Product of Exponentials (PoE) formula~\cite{Murray:book:1994}, which reads: \begin{equation}\label{eqn:PoE}
-	g_{bh}=e^{\hat{Y}_1q_{bh,1}}\cdots e^{\hat{Y}_6q_{bh,6}} g_{bh}(0),
-\end{equation}
-where the offset $g_{bh(0)}=I$ and $Y_i=e_i$, with $e_i\in\bbR^6$ are the six normalized screw vectors (canonical basis elements in $\bbR^6$) associated with the Cartesian coordinates and the Euler ZXY angles in $q_{bh}$.
-Clearly, the six components of $q_{bh}$ are not independent; the modeled suspension is, in fact, a double wishbone that has: 2 DoFs for wheels that can steer, and 1 DoF otherwise.
-In this case, both axles have wheels that can steer; hence, a pair of independent variables ($z;\delta $) is associated with each wheel.
-A detailed analysis is omitted here for the sake of brevity. The interested reader is referred to~\cite{Domenighini:Designs:2023} for a similar treatment using the PoE formalism.
-
-%As mentioned in the previous sections, $z$ represents the variation of the distance between the two attachment points of the spring-damper system, positive when the distance increases and negative when it decreases, while $\delta$ represents the motion of the steering rod attachment point to the chassis along its local $y$ axis (frame $\sref{B}$).
-Considering that the type of the suspension is known, six scalar equations of constraints can be expressed implicitly as $F(q_{bh};z,\delta)=0\in \bbR^6$.
-%
-%To find an explicit relation between inputs $(z,\delta)$ and outputs $q_{bh}$, the constraint equations are solved on a grid of pairs $(z^{(p)},\delta^{(l)})$, where $p = 1,\dots,N$ and $l = 1,\dots,M$.
-Then, the data are fitted (in the least-squares sense) with a 2D regression polynomial $q_{bh}(z,\delta)$ as follows:
-%\begin{equation}\label{eqn:poly}
-%	q_{bh,i}(z,\delta)=\sum_{r=0}^{3} \sum_{s=0}^{3-r} a_{{rs},i} z^r \delta^s.
-%\end{equation}
-%For the purpose a polynomial of degree three is sufficiently accurate.
-Now is possible to express the homogeneous transformation $g_{bh}\bigl(q_{bh}(z,\delta)\bigr)$ and to compute the twist $V_{bh}$ and $\dot{V}_{bh}$ as functions of $(z,\delta)$ and their derivatives.
-We start from:
-%\begin{equation}\label{eqn:Jbh}
-%	V_{bh}=\sum_{i=1}^6J_{bh,i}\dot{q}_{bh,i},
-%\end{equation}
-%where $J_{bh}$ is the \emph{geometric Jacobian} of the suspension joint relative to $q_{bh}$.
-Then, following the procedure thoroughly described in~\cite[Sec.
-2.2]{Domenighini:Designs:2023}, it is possible to obtain %
-%assumed with independent components.
-Each column of $J_{bh}$ can be computed based on the PoE formula \eqref{eqn:PoE}.
-For the $i$-th column, we have $J_{bh,i}=A_i^{-1}Y_i$, with $A_6 = I_{6\times 6}$ and $A_i=\Ad_{e^{\hat{Y}_{i}q_{bh,i}}\cdots e^{\hat{Y}_6q_{bh,6}}}$, $i = 1, \dots, 5$.
-The \emph{Adjoint operator} $\Ad_g$ has been defined as the $6\times6$ matrix satisfying $Y=\Ad_gX$ whenever $\hat{Y}=g\hat{X}g^{-1}$, so that, if $g=\left[\begin{smallmatrix}R&d\\0^T&1\end{smallmatrix}\right]$, then $\Ad_g=\left[\begin{smallmatrix}R&\hat{d}R\\0&R\end{smallmatrix}\right]$.\\
-%As a function of $(z,\delta)$, $\dot{q}_{bh,i}$ can be expressed as follows:
-%\begin{equation}\label{eqn:dt_qbh}
-%	\dot{q}_{bh,i}=\frac{\partial}{\partial z}q_{bh,i}\dot{z}+\frac{\partial}{\partial\delta}q_{bh,i}\dot{\delta},
-%\end{equation}
-%where the (geometric) partial derivatives $\frac{\partial}{\partial z}q_{bh,i}$ and $\frac{\partial}{\partial\delta}q_{bh,i}$ are obtained by differentiating Eq.~\eqref{eqn:poly}.
-%Then, by defining the Jacobians with respect to the true independent variables $z$ and $\delta$ as follows
-%\begin{align}\label{eqn:Jbh,z}
-%	J_{bh,z}=\sum_{i=1}^6J_{bh,i}\frac{\partial}{\partial z}q_{bh,i}\quad \textrm{and}\quad J_{bh,\delta}=\sum_{i=1}^6J_{bh,i}\frac{\partial}{\partial\delta}q_{bh,i},
-%\end{align}
-%it possible to obtain the rigid-body velocity $V_{bh}$ of the constrained motion from Eq.~\eqref{eqn:Jbh}, in terms of $\dot{z}$ and $\dot{\delta}$, as follows
-%\begin{equation}\label{eqn:Vbh}
-%	V_{bh}=J_{bh,z}\dot{z}+J_{bh,\delta}\dot{\delta}.
-%\end{equation}
-%
-%To write the dynamic equations the time derivative $\dot{V}_{bh}$ of the twist $V_{bh}$ is necessary. To this sake, it is sufficient to differentiate Eqs.~\eqref{eqn:Jbh}--\eqref{eqn:Vbh}.
-%\begin{equation}\label{eqn:dt_Jbh}
-%	\dot{V}_{bh}=\sum_{i=1}^6\dot{J}_{bh,i}\dot{q}_{bh,i}+\sum_{i=1}^6J_{bh,i}\ddot{q}_{bh,i}.
-%\end{equation}
-%
-%Here, following~\cite{Gabiccini:ICRA:2012}, the derivatives $\dot{J}_{bh,i}$ can computed using the formula
-%\begin{align}\label{eqn:dt_Jbhi}
-%	\dot{J}_{bh,i}=-\sum\limits_{i < j \leq 6}\ad_{J_{bh,j}\dot{q}_{bh,j}} J_{bh,i}
-%\end{align}
-%using Lie derivatives.
-In the matrix case, the Lie derivative $\hat{Z}$ of the vector field $\hat{X}$ with respect to vector field $\hat{Y}$ is calculated by as a commutator operation $\hat{Z}=\hat{Y}\hat{X}-\hat{X}\hat{Y}$; in the vector case, the $6\times6$ matrix $\ad_Y$ is used, defined such that $Z=\ad_Y X$.
-If $v$ and $\omega$ are the translational and rotational component of $V$, then $\ad_V=\left[\begin{smallmatrix}\hat{\omega}&\hat{v}\\0&\hat{\omega}\end{smallmatrix}\right]$.
-%
-%In Eq.~\eqref{eqn:dt_Jbh}, the accelerations $\ddot{q}_{bh,i}$ are obtained by differentiating the velocities in Eq.~\eqref{eqn:dt_qbh}:
-%\begin{equation}\label{eqn:dt_dt_qbh}
-%	\ddot{q}_{bh,i}=\frac{\partial^2}{\partial z^2}q_{bh,i}\dot{z}^2+2\frac{\partial^2}{\partial z\partial\delta}q_{bh,i}\dot{z}\dot{\delta}+\frac{\partial^2}{\partial\delta^2}q_{bh,i}\dot{\delta}^2+\frac{\partial}{\partial z}q_{bh,i}\ddot{z}+\frac{\partial}{\partial\delta}q_{bh,i}\ddot{\delta},
-%\end{equation}
-%where the partial derivatives are computed by differentiating the regression polynomials~in~\eqref{eqn:poly}.
-%After simple but tedious calculations, the explicit expression for the derivatives of the Jacobians in \eqref{eqn:Jbh,z} appear as follows:
-%\begin{align}
-%	\dot{J}_{bh,z}&=\sum_{i=1}^6\dot{J}_{bh,i}\frac{\partial}{\partial z}q_{bh,i}+\sum_{i=1}^6J_{bh,i}\Bigl(\frac{\partial^2}{\partial z^2}q_{bh,i}\dot{z}+\frac{\partial^2}{\partial z\partial\delta}q_{bh,i}\dot{\delta}\Bigr)\label{eqn:dt_Jbh,z}\quad\textrm{and}\\\dot{J}_{bh,\delta}&=\sum_{i=1}^6\dot{J}_{bh,i}\frac{\partial}{\partial\delta}q_{bh,i}+\sum_{i=1}^6J_{bh,i}\Bigl(\frac{\partial^2}{\partial z\partial\delta}q_{bh,i}\dot{\delta}+\frac{\partial^2}{\partial\delta^2}q_{bh,i}\dot{\delta}\Bigr)\label{eqn:dt_Jbh,d}.
-%\end{align}
-%Substituting Eqs.~\eqref{eqn:dt_Jbhi} and~\eqref{eqn:dt_dt_qbh} in Eqs.~\eqref{eqn:dt_Jbh}, and casting some intermediate results as~\eqref{eqn:dt_Jbh,z} and~\eqref{eqn:dt_Jbh,d},
-%the constrained rigid-body acceleration $\dot{V}_{bh}$ in the following form:
-%\begin{equation}\label{eqn:dt_Vbh}
-%	\dot{V}_{bh}=\dot{J}_{bh,z}\dot{z}+\dot{J}_{bh,\delta}\dot{\delta}+J_{bh,z}\ddot{z}+J_{bh,\delta}\ddot{\delta}.
-%\end{equation}
-%The explicit expressions for auxiliary Jacobians $J_{bh,z}$, $J_{bh,\delta}$, $\dot{J}_{bh,z}$ and $\dot{J}_{bh,\delta}$ are omitted here for brevity.
-The interested reader can find the detailed analytical derivation in~\cite[Eqs.
-(5)--(8)]{Domenighini:Designs:2023}.
-%
-%A generalized force $\tau\in\bbR$ is introduced that depends only on the independent variable $z\in\bbR$.
-$\tau$ is assumed to be generated only by the spring and the damper, considered aligned along the same axis.
-It is worth noting that, using the defined $z$, we are allowed to compute the force of the suspension without calculate the associated $q_{bh}$ nor using the Principle of Virtual Works.
-In this expression, also the contribution of the anti-roll bar is considered, which depends on the difference between the left and right trim, and the contribution of the bump-stops, which activate when the trim reaches its lower or upper limits.
-The intensity of $\tau$, for each wheel, is expressed as follow:
-%\begin{equation}
-%	\tau_{ij}=k_{s,i}(l_{c,i}-l_{0,i}+z_{ij})+c_{d,i}\dot{z}_{ij}+(-1)^j k_{a,i}(z_{i,1}-z_{i,2}) + \left(\frac{2z-z_{\text{max}}-z_{\text{min}}}{|z_{\text{max}}-z_{\text{min}}|}\right)^{b},
-%\end{equation}
-%where $l_{c,i}$ is the spring's length in the reference position, $l_{0,i}$ is the spring's length at rest, $k_{s,i}$ and $c_{d,i}$ are the stiffness and the damping coefficients, $k_{a,i}$ is the anti-roll bar stiffness, $b$ influences the stiffness of the bump-stops, and $z_{\text{min}}$ and $z_{\text{max}}$ are the knees of the bump-stops' force characteristics.
-\section{Tire Model}
-\label{sec:tyremodel}
-\begin{figure}[t]\centering
-	\tdplotsetmaincoords{0}{0}
-	\begin{tikzpicture}[tdplot_main_coords,scale=3.5]\small
-		\tdplotsetrotatedcoords{-166.5322}{-65.2292}{-166.5322}\draw[tdplot_rotated_coords,dashed,gray!75,fill=gray!10](-0.6,0,0)--(-0.6,1.1,0)--(0.6,1.1,0)--(0.6,0,0);\draw[tdplot_rotated_coords,very thick,fill=white](0,0)coordinate(C)..controls(0.25,0)and(0.5,0)..(0.5,0.45)arc(0:180:0.5)..controls(-0.5,0)and(-0.25,0)..(0,0);\draw[tdplot_rotated_coords,very thick,fill=gray!25](0,0.45)coordinate(O)circle(0.25);\draw[tdplot_rotated_coords,very thick,dashed](0.5,0.45)arc(0:-180:0.5);\draw[tdplot_rotated_coords,very thin,dashed](0,0,0)--(0,0.2,0)(0,0.7,0)--(0,0.95,0)(-0.5,0.45,0)--(-0.25,0.45,0)(0.25,0.45,0)--(0.5,0.45,0);\draw[tdplot_rotated_coords,very thin](0,0.2,0)--(0,0.7,0)(0,0.95,0)--(0,1.25,0)(-1,0.45,0)--(-0.5,0.45,0)(-0.25,0.45,0)--(0.25,0.45,0)(0.5,0.45,0)--(1,0.45,0);\draw[tdplot_rotated_coords,{Stealth[inset=0pt]}-{Stealth[inset=0pt]},very thin](0.75,0,0)--node[right]{$r$}(0.75,0.45,0);\tdplotsetrotatedcoordsorigin{(O)}\tdplotsetrotatedcoords{-166.5322}{-65.2292}{-146.5322}\draw[tdplot_rotated_coords,dashed](-0.25,0,0)--(-0.5,0,0);\draw[tdplot_rotated_coords,dashed](0,0.25,0)--(0,0.5,0);\draw[tdplot_rotated_coords,->](0,0,0)--(-0.25,0,0)(-0.5,0,0)--(-1,0,0)node[below left]{$x_h$};\draw[tdplot_rotated_coords,->](0,0,0)--(0,0.25,0)(0,0.5,0)--(0,0.8,0)node[above left]{$z_h$};\draw[tdplot_rotated_coords,->](0,0,0)--(0,0,0.8)node[right]{$y_h$};\draw[tdplot_rotated_coords,{Stealth[inset=0pt]}-{Stealth[inset=0pt]},very thin](-0.8,0)arc(180:160:0.8)node[midway,left]{$\sigma$};\tdplotsetrotatedcoordsorigin{(C)}\tdplotsetrotatedcoords{43.2192}{-27.9909}{-46.7808}\draw[tdplot_rotated_coords,dashed,gray!75,fill=gray!25](0,0,-0.6)--(0.6,0,-0.6)--(0.6,0,0.6)--(0,0,0.6);\draw[tdplot_rotated_coords,->](0,0,0)--(0.8,0,0)node[right]{$y_n$};\draw[tdplot_rotated_coords,->](0,0,0)node[circle,fill=black,inner sep=0pt,minimum size=3pt,draw]{}--(0,1.25,0)node[above]{$z_n$};\draw[tdplot_rotated_coords,->](0,0,-1)--(0,0,1)node[below left]{$x_n$};\draw[tdplot_rotated_coords,{Stealth[inset=0pt]}-{Stealth[inset=0pt]},very thin](0,1.1)arc(90:110:1.1)node[midway,above]{$\gamma$};\draw(0,-0.02)..controls(0,-0.3)and(0.3,-0.1)..(0.3,-0.3)node[below]{contact point};\draw(-0.5,0.7)..controls(-0.35,0.55)and(-0.6,0.7)..(-0.55,0.5)node[below,align=center]{wheel\\plane};\draw(0.4,0.15)..controls(0.3,0.25)and(0.55,0.2)..(0.5,0.3)node[above,align=center]{road\\plane};\draw(O)node[below right]{$O_h$}(C)node[below right]{$O_n$};\draw[tdplot_rotated_coords,thick,-{Stealth[inset=0pt]}](C)--+(0,0,0.8)node[below right]{$F_x$};\draw[tdplot_rotated_coords,thick,-{Stealth[inset=0pt]}](C)--+(0.5,0,0)node[above]{$F_y$};\draw[tdplot_rotated_coords,thick,-{Stealth[inset=0pt]}](C)--+(0,1,0)node[right]{$F_z$};
-	\end{tikzpicture}
-	\hfill
-	\begin{tikzpicture}[scale=4.25]\small
-		\coordinate(C)at(0,0);\coordinate(-)at(-0.375,0);\coordinate(+)at(0.325,0);\draw[->](-0.7,0)--(0.6,0)node[right]{$y_n$};\draw[->](0,0)node[circle,fill=black,inner sep=0pt,minimum size=3pt,draw]{}--(0,1.25)node[above]{$z_n$};\tdplotsetrotatedcoords{10}{0}{0}\draw[tdplot_rotated_coords,dash dot](0,-0.15)--(0,1.25)(-0.35,0.45)--(0.6,0.45);\draw[tdplot_rotated_coords,very thick](-0.25,0.2)--(-0.25,0.7)..controls(-0.25,0.95)and(-0.25,0.95)..(0,0.95)..controls(0.25,0.95)and(0.25,0.95)..(0.25,0.7)--(0.25,0.2)(-0.25,0.2)..controls(-0.25,0.025)and(-)..(C)..controls(+)and(0.25,-0.025)..(0.25,0.2);\draw[tdplot_rotated_coords,very thick,dashed](-0.25,0.2)..controls(-0.25,-0.05)and(-0.25,-0.05)..(0,-0.05)node(B){}..controls(0.25,-0.05)and(0.25,-0.05)..(0.25,0.2);\node[tdplot_rotated_coords](O)at(0,0.45){};\draw[very thin](O)--({(-0.55,0)}|-O);\draw[very thin,{Stealth[inset=0pt]}-{Stealth[inset=0pt]}](-0.475,0)--node[left]{$h$}({(-0.475,0)}|-O);\draw[very thin](B)--({(-0.7,0)}|-B);\draw[very thin,-{Stealth[inset=0pt]}](-0.625,0.3)--node[left]{$d$}(-0.625,0);\draw[very thin](-0.625,0)--({(-0.625,0)}|-B);\draw[very thin,-{Stealth[inset=0pt]}](-0.625,-0.15)--({(-0.625,0)}|-B);\draw[tdplot_rotated_coords,very thin](0.05,0)--(0.45,0);\draw[tdplot_rotated_coords,very thin](0.05,-0.05)--(0.6,-0.05);\draw[tdplot_rotated_coords,very thin,{Stealth[inset=0pt]}-{Stealth[inset=0pt]}](0.375,0)--node[right]{$r$}(0.375,0.45);\draw[tdplot_rotated_coords,very thin,{Stealth[inset=0pt]}-{Stealth[inset=0pt]}](0.525,-0.05)--node[right]{$r_0$}(0.525,0.45);\draw[very thin,{Stealth[inset=0pt]}-{Stealth[inset=0pt]}](0,1.125)arc(90:100:1.125)node[midway,above]{$\gamma$};\draw(0,-0.4);
-	\end{tikzpicture}
-	\caption{Decomposition of the tyre forces along the axes of the auxiliary frame $\{N\}$.
-The frame $\{N\}$ is such that: $x_n$ lies along the intersection between the wheel plane ($x_hz_h$-plane) and the road plane ($x_s y_s$-plane), pointing forward; $y_n$ points left (w.r.t.
-the driver) along the intersection line between the road plane ($x_s y_s$-plane) and the transverse plane ($y_h z_s$-plane) through the wheel center $O_h$; $z_n$ is normal to the road surface.
-The contact point is estimated in correspondence to the origin $O_n$.
-On the right, we report a detail of the tyre vertical deformation.}
-	\label{fig:tyre_wheel}
-\end{figure}
+\begin{align}
+\dbx = \begin{bmatrix}
+              \dot{\bq}\\
+              \dot{\bq}_v
+             \end{bmatrix} =
+             \begin{bmatrix}
+              \bq_v\\
+              \text{ABA}\text{dyn}(\bq, \bq_v, f_{xa}, f_{xb}, \delta, \bw),
+             \end{bmatrix} = F(\bx).
+	\label{eq:ABA_dyn}
+\end{align}
 

### 2025-04-07 MAIN.tex
[MG] just started with notation
[MGa] stochastic framework

@@
 \usepackage{algorithm,algpseudocode}
-\usepackage{tikz,tikz-3dplot}\usetikzlibrary{arrows.meta,angles}
+\usepackage{tikz,tikz-3dplot}
+\usetikzlibrary{arrows.meta,angles}
 
@@
 
-\author{Martino Gulisano\textsuperscript{a}, Matteo Masoni\textsuperscript{a}, and \name{Marco Gabiccini,\textsuperscript{a,*}
+\author{Martino Gulisano\textsuperscript{a}, Matteo Masoni\textsuperscript{a}, and \name{Marco Gabiccini\textsuperscript{a,*}
 \thanks{* Contact: Marco Gabiccini. Email: marco.gabiccini@unipi.it}

### 2025-04-07 stochastic_vehicle_dynamics.tex
[MGa] stochastic framework

@@
 Due to the symmetry of the probability density function (pdf) with respect to $\bmu(t)$, it is possible to represent the time evolution of the pdf as: i) the deterministic evolution of the mean $\bmu(t)$
-\begin{align}
+\begin{align}\label{eq:meandynamics}
 \dbmu(t) = f(\bmu(t), \bu(t)),
 \end{align}
-and ii) the time evolution of the state covariance matrix $\bP(t)$ along $\bmu(t)$, which can be expressed as
-\begin{align}
-\dbP(t) = \bA(t) \bP(t) + \bP(t) \bA^T(t) + \bQ(t),\quad \bP(0) = \bP_0.
+and ii) the time evolution of the state covariance matrix $\bP(t)$ along $\bmu(t)$, which can be expressed by the Lyapunov matrix differential equation
+\begin{align}\label{eq:dP}
+\dbP(t) = \bA(t) \bP(t) + \bP(t) \bA^T(t) + \bQ(t),\quad \bP(0) = \bP_0 = \bP_0^T.
 \end{align}
-The \emph{initial} state covariance $\bP_0$ is assumed known and $\bA(t)$ is the classical shorthand for the Jacobian along the mean trajectory $\bmu(t)$, that is $\bA(t)=\bJ(\bmu(t),\bu(t))$, where $\bJ(\bx, \bu) = \frac{\pd f(\bx, \bu)}{\pd \bx}$.
+In~\eqref{eq:dP}, $\bP_0$ is the \emph{initial} state covariance and $\bA(t)$ is the usual shorthand notation for the Jacobian along the mean trajectory $\bmu(t)$, that is $\bA(t)=\bJ(\bmu(t),\bu(t))$, where $\bJ(\bx, \bu) = \frac{\pd f(\bx, \bu)}{\pd \bx}$.
 
+As can be readily verified by differentiation~\cite{Gajic:LyapunovMatrixEquation:2010}, the analytical solution of~\eqref{eq:dP} has the form
+\begin{align}\label{eq:P_STM}
+\bP(t) = \bPhi(t,t_0) \bP_0 \bPhi(t,t_0)+\int_{t_0}^{t} \bPhi(t,\tau) \bQ(\tau) \bPhi^T(t,\tau) \dd \tau \quad \bP(0)=\bP_0,
+\end{align}
+where $\bPhi(t,t_0)$ is the \emph{state transition matrix}.
+This matrix encodes the evolution from $t_0$ to $t$ of a perturbation w.r.t.
+to a fiducial trajectory, the mean $\bmu(t)$.
+In symbols, $\de \bx(t) = \bPhi(t,t_0) \de \bx(t_0)$, with $\de \bx(\bullet) =\bx(\bullet) - \bmu(\bullet) $.
+In turn, the evolution of $\bPhi(t,t_k)$ from a generic $t_k$ to $t$ is driven by the following differential equation
+\begin{align}\label{eq:STM}
+\dbPhi(t,t_k) = \bA(t)\bPhi(t,t_k), \quad \bPhi(t_k, t_k) = \bI.
+\end{align}
+Even if in geometric mechanic and orbital mechanic books, see e.g.~\cite{Maruskin:DynamicalSystemsGeometric:2018} or~\cite{Tapley:StatisticalOrbitDetermination:2004}, the usual choice for a statistical trajectory determination is to employ~\eqref{eq:meandynamics} along with~\eqref{eq:P_STM} and~\eqref{eq:STM}, in view of using collocation integrators in stochastic trajectory planning, we proceed more directly using~\eqref{eq:meandynamics} and~\eqref{eq:dP}.
 \begin{align}

### 2025-04-08 stochastic_vehicle_dynamics.tex
[MGa] some rephrasing

@@
 \end{align}
-and ii) the time evolution of the state covariance matrix $\bP(t)$ along $\bmu(t)$, which can be expressed by the Lyapunov matrix differential equation
+and ii) the time evolution of the state covariance matrix $\bP(t)$ along $\bmu(t)$, which can be expressed by the \emph{Lyapunov matrix differential equation}
 \begin{align}\label{eq:dP}
@@
 This matrix encodes the evolution from $t_0$ to $t$ of a perturbation w.r.t.
-to a fiducial trajectory, the mean $\bmu(t)$.
-In symbols, $\de \bx(t) = \bPhi(t,t_0) \de \bx(t_0)$, with $\de \bx(\bullet) =\bx(\bullet) - \bmu(\bullet) $.
+to a fiducial trajectory -- the mean $\bmu(t)$.
+In symbols, $\barbx(t) = \bPhi(t,t_0) \barbx(t_0)$, with $\barbx(\bullet) =\bx(\bullet) - \bmu(\bullet) $.
 In turn, the evolution of $\bPhi(t,t_k)$ from a generic $t_k$ to $t$ is driven by the following differential equation
@@
 \end{align}
-Even if in geometric mechanic and orbital mechanic books, see e.g.~\cite{Maruskin:DynamicalSystemsGeometric:2018} or~\cite{Tapley:StatisticalOrbitDetermination:2004}, the usual choice for a statistical trajectory determination is to employ~\eqref{eq:meandynamics} along with~\eqref{eq:P_STM} and~\eqref{eq:STM}, in view of using collocation integrators in stochastic trajectory planning, we proceed more directly using~\eqref{eq:meandynamics} and~\eqref{eq:dP}.
+In geometric mechanics and orbital mechanics books, see e.g.~\cite{Maruskin:DynamicalSystemsGeometric:2018} or~\cite{Tapley:StatisticalOrbitDetermination:2004}, the usual choice for statistical trajectory determination is to employ~\eqref{eq:meandynamics} along with~\eqref{eq:P_STM} and~\eqref{eq:STM}.
+In our case, since we use collocation integrators for stochastic trajectory planning, we follow a more direct approach -- motivated by~\cite{Gillis:PracticalMethodsApproximate:2015} -- by directly employing~\eqref{eq:meandynamics} and~\eqref{eq:dP}.
 \begin{align}

### 2025-04-10 stochastic_vehicle_dynamics.tex
[MGa] stochastic OCP

@@
 to a fiducial trajectory -- the mean $\bmu(t)$.
-In symbols, $\barbx(t) = \bPhi(t,t_0) \barbx(t_0)$, with $\barbx(\bullet) =\bx(\bullet) - \bmu(\bullet) $.
+In symbols, $\barbx(t) = \bPhi(t,t_0) \barbx(t_0)$, with $\barbx(\cdot) =\bx(\cdot) - \bmu(\cdot) $.
 In turn, the evolution of $\bPhi(t,t_k)$ from a generic $t_k$ to $t$ is driven by the following differential equation
@@
 \end{align}
-In geometric mechanics and orbital mechanics books, see e.g.~\cite{Maruskin:DynamicalSystemsGeometric:2018} or~\cite{Tapley:StatisticalOrbitDetermination:2004}, the usual choice for statistical trajectory determination is to employ~\eqref{eq:meandynamics} along with~\eqref{eq:P_STM} and~\eqref{eq:STM}.
+In geometric mechanics and orbital mechanics, see e.g.~\cite{Maruskin:DynamicalSystemsGeometric:2018} or~\cite{Tapley:StatisticalOrbitDetermination:2004}, the usual choice for statistical trajectory determination is to employ~\eqref{eq:meandynamics} along with~\eqref{eq:P_STM} and~\eqref{eq:STM}.
 In our case, since we use collocation integrators for stochastic trajectory planning, we follow a more direct approach -- motivated by~\cite{Gillis:PracticalMethodsApproximate:2015} -- by directly employing~\eqref{eq:meandynamics} and~\eqref{eq:dP}.
+
+The continuous-time stochastic trajectory planning can be framed as the following nonlinear optimal control problem
 \begin{align}
-\dbx = \begin{bmatrix}
-              \dot{\bq}\\
-              \dot{\bq}_v
-             \end{bmatrix} =
-             \begin{bmatrix}
-              \bq_v\\
-              \text{ABA}\text{dyn}(\bq, \bq_v, f_{xa}, f_{xb}, \delta, \bw),
-             \end{bmatrix} = F(\bx).
-	\label{eq:ABA_dyn}
+& \underset{\bmu(t), \bu(t), \bP(t)}{\text{minimize}} \quad J(\bmu(t), \bu(t), \bP(t)) \label{eq:OCPcost} \\
+& \text{s.t.} \quad \dbmu(t)           = f(\bmu(t), \bu(t)) \label{eq:OCPdyn} \\
+& \phantom{\text{s.t.} \quad} \bmu(0)  = \bmu_0 \label{eq:OCPdynIC} \\
+& \phantom{\text{s.t.} \quad} \dbP(t) = \bA(t) \bP(t) + \bP(t) \bA^T(t) + \bQ(t) \label{eq:OCPdP} \\
+& \phantom{\text{s.t.} \quad} \bP(0)  = \bP_0 \succeq 0 \label{eq:OCPdPIC} \\ %& \phantom{\text{s.t.} \qquad} 0       \geq h_i(\bmu(t), \bu(t))
+%+ \overbrace{\gamma \bigg[\underbrace{\na^T_{\bx} h_i \bP(t) \na_{\bx} h_i}_{\text{variance of $h_i(\bx)$}}\bigg]^{\frac{1}{2}}}^{\text{safety margin}},
+&\phantom{\text{s.t.} \qquad} 0       \geq h_i(\bmu(t), \bu(t))
++ \be_i(\bmu(t), \bu(t), \bP(t)),
+\quad i = 1, \ldots, p \label{eq:OCPconstraints}
 \end{align}
-
-The tangential forces exchanged between the tire and road are computed using \mbox{Pacejka's} \emph{Magic Formula}:
-\begin{equation}\label{eqn_mf}
-	[F_x,F_y]=\mf(\kappa,\alpha,F_z,\gamma),
-\end{equation}
-where $ F_x, F_y $ are the longitudinal and lateral forces, dependent on tire slips $( \kappa, \alpha )$, vertical load $( F_z )$, and camber angle $( \gamma )$.
-As shown in Fig.~\ref{fig:tyre_wheel}, an auxiliary frame $\sref{N}$ is introduced to calculate $ F_z $.
-This depends upon the interpenetration $ d $ of the tire into the ground, following a penalty-based compliant tire model.
-This model treats the wheel as a radial spring-damper system with constant stiffness $ k_t $ and damping $ c_t $.
-The vertical force is expressed as:
-\begin{equation}\label{eqn_Fz}
-	F_z= F_0\log_2\left(1 + 2^{\frac{k_t d + c_t \dot{d}}{F_0}}\right)
-\end{equation}
-which is differentiable for $ d = 0 $ and $\dot{d} = 0$, improving numerical stability.
-The model ensures $ F_z $ approaches zero when the tire loses contact with the ground, while maintaining a non-zero gradient to guide optimization algorithms in scenarios where the tire detaches from the ground or encounters sudden changes.
+The cost function $J$ in~\eqref{eq:OCPcost} depends on the mean $\bmu(t)$, the controls $\bu(t)$ and the state covariance $\bP(t)$.
+The mean dynamics is expressed by~\eqref{eq:OCPdyn} with~\eqref{eq:OCPdynIC}, and the covariance dynamics is expressed by~\eqref{eq:OCPdP} with~\eqref{eq:OCPdPIC}.
+In eq.~\eqref{eq:OCPconstraints} the \emph{backoff terms} $\be_i$ account for the disturbances and serve the purpose of obtaining deterministic safety-margin types of constraint directly on $\bmu(t)$.
+This is equivalent to a linearization around the mean of the original \emph{chance constraint} on $\bx(t)$ expressed by $\Prob \{h_i(\bx) \leq 0\}\geq p$, where $p$ is the confidence level of constraint satisfaction.
+The backoff terms are given by
+$\be_i = \gamma \sig_i$.
+The coefficient $\gamma = \Phi^{-1}(p)$ is the quantile function, where $\Phi(z)=\Prob \{Z\leq z\}$ is the cdf of a standard normal distribution $Z \sim \calN(0,1)$, and acts as a \emph{tuning knob}: the greater the confidence level $p$ required, the higher the gain $\ga$\footnote{For example, with $p=0.84$ $\ga = 1.0$, with $p=0.97$ $\ga = 2.0$, with $p=0.99$ $\ga = 3.0$}.
+The term $\sig_i = \big[\na^T_{\bx} h_i(\bmu) \bP(t) \na_{\bx} h_i(\bmu)\big]^{\frac{1}{2}}$ represents the standard deviation of the constraint $h_i(\bx)$ linearized around the mean, i.e.
+of random variable $h_i(\bmu)+\na^T_{\bx} h_i(\bmu) (\bx - \bmu)$, and follows immediately from the propagation rule of covariance.
 

### 2025-04-14 stochastic_vehicle_dynamics.tex
[MG] towards discretized form of OCP, i.e. DOCP

@@
 
-The continuous-time stochastic trajectory planning can be framed as the following nonlinear optimal control problem
+Accordingly, the continuous-time stochastic trajectory planning can be framed as the following nonlinear optimal control problem
 \begin{align} 
@@
 + \be_i(\bmu(t), \bu(t), \bP(t)),
-\quad i = 1, \ldots, p \label{eq:OCPconstraints}
+\quad i \in \calI \label{eq:OCPconstraints}
 \end{align}
@@
 In eq.~\eqref{eq:OCPconstraints} the \emph{backoff terms} $\be_i$ account for the disturbances and serve the purpose of obtaining deterministic safety-margin types of constraint directly on $\bmu(t)$.
+$\calI$ is the set of indices defining the inequality constraints.
 This is equivalent to a linearization around the mean of the original \emph{chance constraint} on $\bx(t)$ expressed by $\Prob \{h_i(\bx) \leq 0\}\geq p$, where $p$ is the confidence level of constraint satisfaction.
@@
 
-\subsection{Camber System Modeling}
-\label{sec:cambersystemmodeling}
-The dynamic camber control (DCC) system is modeled as a simplified actuated revolute joint connecting the two virtual halves of the knuckle, though a real implementation might require a more complex 1-DoF joint.
-While this simplification may influence joint values from the optimizer, it results in nearly the same optimal wheel posture.
-The choice was driven by the limited literature on such systems for FSAE vehicles, which is the vehicle type employed in our analysis.
-A realistic model, if available, could be characterized using the same process applied to the suspension.
-During the design phase of the actual joint, some parameters could be treated as time-invariant optimization variables, constrained within dimensional bounds, to achieve the best performance configuration.
-%The dynamic camber control (DCC) system is modeled here as a simple actuated revolute joint connecting the two virtual halves of the knuckle.
-A real implementation could require a more complex generalized 1-DoF joint.
-Changing the joint characteristics might affect the joint values obtained from the optimizer but would result in approximately the same optimal wheel posture.
-This simplification was necessary due to the limited open literature on such systems for FSAE vehicles.
-If a realistic model of the system were available, the process used for the suspension could easily be repeated to characterize this joint.
-It is worth observing that, in the design phase of the actual controlled joint, some model parameters could be treated as time-invariant optimization variables, appropriately constrained to meet dimensional bounds, and optimized to find the best performance configuration.
+\subsection{Discretization via direct collocation}
+\label{sec:discretization}
 
-In Table~\ref{tab:camber_prop}, we report the range of variation of function $q_{hk}(t)$ and its first and second derivatives.
-Considering the scarcity of open literature available, reasonable values have been selected.
-In Figure~\ref{fig:camber_showing} the wheel is shown in a general posture with zero camber (shaded black) and with maximum positive camber (solid black).
-Due to the convention adopted for rotations, this posture corresponds to the minimum value of $q_{hk}$.
-\begin{table}[h]
-	\centering
-	\begin{tabular}{|c|c|c|}
-		\hline
-		$q_{hk}$ (rad) & $\dot{q}_{hk}$ (rad/s)& $\ddot{q}_{hk}$ (rad/s$^2$)\\\hline
-		$\pm \pi/20$ & $\pm \pi/20$ & $\pm \pi/10$ \\\hline
-	\end{tabular}
-	\caption{Upper and lower bounds for camber angle value $q_{hk}(t)$ and its derivatives.}
-	\label{tab:camber_prop}
-\end{table}
+The nonlinear optimal control problem~\eqref{eq:OCPcost} can be discretized by applying a suitable collocation integrator obtaining 
+\begin{alignat}{3}
+\underset{\bmu_k,\bxi_k, \bu_k, \bP_k,\bz_k}{\text{minimize}} \,
+& & & J_k(\bmu_k,\bxi_k, \bu_k, \bP_k) & & \label{eq:DOCPcost} \\
+\hspace*{-2.0 cm}\text{s.t.} \quad
+& \bzero      & = & \; \bPsimu_k(\bmu_{k-1},\bmu_k,\bxi_k, \bu_k,\bz_k),
+& \quad & k = 1,\ldots, N \label{eq:DOCPdyn} \\
+& \bmu_0      & = & \; \bar{\bmu}_0
+& & \label{eq:DOCPdynIC} \\
+& \bzero      & = & \; \bPsiP_k(\bmu_k,\bxi_k, \bu_k, \bP_{k-1},\bP_k,\bSi_k,\bz_k),
+& \quad & k = 1,\ldots, N \label{eq:DOCPdP} \\
+& \bP_0       & = & \; \bar{\bP}_0 \succeq 0
+& & \label{eq:DOCPdPIC} \\
+& \bzero      & = & \; \bOm_k(\bmu_k,\bxi_k, \bu_k,\bz_k),
+& \quad & k = 0,\ldots, N \label{eq:DOCPpath} \\
+& 0           & \geq & \; h_i(\bmu_k, \bu_k, \bz_k) + \be_i(\bmu_k, \bu_k, \bP_k, \bz_k),
+& \quad & k = 1,\ldots, N;\; i \in \calI \label{eq:DOCPconstraints}
+\end{alignat}
 
-\begin{figure}[h]
-	\centering
-	\includegraphics[scale = .6]{Images/camber_showing}
-	\caption{Real scale representation of the wheel in two configurations: $q_{hk}=0$ (shaded surface) and $q_{hk}=-\pi/20$ (solid surface).}
-	\label{fig:camber_showing}
-\end{figure}
 \subsection{External Wrenches}

### 2025-04-14 stochastic_vehicle_dynamics.tex
[MGa] better description for DOCP

@@
 \end{align}
-In geometric mechanics and orbital mechanics, see e.g.~\cite{Maruskin:DynamicalSystemsGeometric:2018} or~\cite{Tapley:StatisticalOrbitDetermination:2004}, the usual choice for statistical trajectory determination is to employ~\eqref{eq:meandynamics} along with~\eqref{eq:P_STM} and~\eqref{eq:STM}.
+In geometric and orbital mechanics, see e.g.~\cite{Maruskin:DynamicalSystemsGeometric:2018} or~\cite{Tapley:StatisticalOrbitDetermination:2004}, the usual choice for statistical trajectory determination is to employ~\eqref{eq:meandynamics} along with~\eqref{eq:P_STM} and~\eqref{eq:STM}.
 In our case, since we use collocation integrators for stochastic trajectory planning, we follow a more direct approach -- motivated by~\cite{Gillis:PracticalMethodsApproximate:2015} -- by directly employing~\eqref{eq:meandynamics} and~\eqref{eq:dP}.
@@
 
-The nonlinear optimal control problem~\eqref{eq:OCPcost} can be discretized by applying a suitable collocation integrator obtaining 
+The nonlinear optimal control problem~\eqref{eq:OCPcost} can be discretized by applying a suitable collocation integrator obtaining the following nonlinear program (NLP) 
 \begin{alignat}{3}
 \underset{\bmu_k,\bxi_k, \bu_k, \bP_k,\bz_k}{\text{minimize}} \,
-& & & J_k(\bmu_k,\bxi_k, \bu_k, \bP_k) & & \label{eq:DOCPcost} \\
+& & & J_k(\bmu_k,\bxi_k, \bu_k) & & \label{eq:DOCPcost} \\
 \hspace*{-2.0 cm}\text{s.t.} \quad
@@
 \end{alignat}
+To perform the discretization, the track centerline is parameterized using a curvilinear abscissa $\alpha \in [0,1]$, and uniformly sampled at $N+1$ points $\alpha_0, \ldots, \alpha_N$.
+Accordingly, $\bmu_k$ denotes the mean state at grid node $\alpha_k$, while $\bu_k$ and $\bz_k$ are assumed to be piecewise constant over each interval $[\al_k, \al_{k+1}]$.
+These represent the controls and the algebraic variables, respectively.
+The $\bz_k$'s are introduced as direct handles for physically meaningful quantities such contact forces.
+Similarly, $\bP_k$ represents the covariance matrix at the $k$-th node.
+Following the direct collocation approach, both the mean and covariance state trajectories are approximated, in the $k$-th interval, with polynomials $\pi_k(\tau)$, defined on the unit interval $\tau\in[0,1]$, and then scaled to match the width $h_k$ of the corresponding time step.
+On the unit interval, we select $d$ collocation points $\tau_1, \ldots, \tau_d$, associated with as many collocation states.
+Accordingly, we define $\bxi_k$ and $\bSi_k$ as the mean states and covariance matrices at the $d$ collocation points within each interval $[\al_k, \al_{k+1}]$, respectively.
+Therefore, $\bxi_k =
+(\bxi_{k,1}, \ldots, \bxi_{k,d})$, and similarly $\bSi_k=(\bSi_{k,1}, \ldots, \bSi_{k,d})$.
+Equations~\eqref{eq:DOCPdyn} and~\eqref{eq:DOCPdP} represent the collocation and continuity equations for mean and covariance, respectively, with initial conditions represented by~\eqref{eq:DOCPdynIC} and~\eqref{eq:DOCPdPIC}.
+Eqs.~\eqref{eq:DOCPpath} are path equality constraints and~\eqref{eq:DOCPconstraints} are the robustified inequality constraints.
 
+In certain cases~\cite{Gillis:PracticalMethodsApproximate:2015}, positive-definiteness-preserving Lyapunov discretization schemes are used in~\eqref{eq:DOCPdP}.
+However, for sufficiently fine discretization, we found that integrating only the lower triangular part of $\bP(t)$ in~\eqref{eq:OCPdPIC}  (and reconstructing its strictly upper part accordingly), ensures that all $\bP_k$ remain symmetric and positive definite when the initial $\bP_0$ is symmetric and positive definite.
+In all our tests we used direct collocation with a cubic polynomial and Gauss-Legendre collocation points.
 \subsection{External Wrenches}

### 2025-04-15 stochastic_vehicle_dynamics.tex
[MGa] small mod. No H steps ahead (open-loop) back-off

@@
 The mean dynamics is expressed by~\eqref{eq:OCPdyn} with~\eqref{eq:OCPdynIC}, and the covariance dynamics is expressed by~\eqref{eq:OCPdP} with~\eqref{eq:OCPdPIC}.
-In eq.~\eqref{eq:OCPconstraints} the \emph{backoff terms} $\be_i$ account for the disturbances and serve the purpose of obtaining deterministic safety-margin types of constraint directly on $\bmu(t)$.
+In eq.~\eqref{eq:OCPconstraints} the \emph{backoff terms} $\be_i$ account for the disturbances and serve the purpose of obtaining deterministic safety-margins directly on $\bmu(t)$.
 $\calI$ is the set of indices defining the inequality constraints.
@@
 However, for sufficiently fine discretization, we found that integrating only the lower triangular part of $\bP(t)$ in~\eqref{eq:OCPdPIC}  (and reconstructing its strictly upper part accordingly), ensures that all $\bP_k$ remain symmetric and positive definite when the initial $\bP_0$ is symmetric and positive definite.
-In all our tests we used direct collocation with a cubic polynomial and Gauss-Legendre collocation points.
+In all our tests direct collocation with cubic polynomial state representations and Gauss-Legendre collocation points were used.
 \subsection{External Wrenches}

### 2025-04-16 bis_open_loop_planning.tex
[MM] constraints formulation (first try)

--- 2025-04-03 open_loop_planning.tex
@@
 
-The model described in Sec.~\ref{chap:vehicle_model} is embedded into a minimum lap time problem which is briefly outlined below.
+The proposed approach requires the presence of $H+1$ versions of the matrix $\bP$ at each grid point.
+One of these versions is initialized with $\bar{\bP}_0$, while the others represent the evolution of matrices that were initialized at previous grid points, from 1 to $H$ steps earlier.
+This ensures that at each step it is possible to find a covariance matrix evolved for $H$ steps and use it to robustify the constraint.
+To adapt the problem, it is necessary to modify equations (\ref{eq:DOCPdP}) and (\ref{eq:DOCPconstraints}).
+In this case,  (\ref{eq:DOCPdP}) collects $H$ collocation and continuity equations and an initialization equation, while (\ref{eq:DOCPconstraints}) takes as input the most propagated version of the covariance matrix at the k-th step, here named $\bP^H_k$.
+The new formulation is reported:
+\begin{alignat}{3}
+\bzero       = & \; \bPsiP_k(\bmu_k,\bxi_k, \bu_k, \bP_{k-1}^{j-1},\bP_k^j,\bSi_k^j,\bz_k),
+& \quad & k = 1,\ldots, N; \; j=0,\ldots,H \label{eq:DOCPdPopenloop} \\
+0            \geq & \; h_i(\bmu_k, \bu_k, \bz_k) + \be_i(\bmu_k, \bu_k, \bP_k^j, \bz_k),
+& \quad & k = 1,\ldots, N;\; i \in \calI \label{eq:DOCPconstraintsopenloop}
+\end{alignat}
+This process is illustrated in Figure \ref{fig:DOCPgrid}, where the dashed rectangle represent the $k$-th step at which the constraints are being formulated.
+At each grid point, two versions of the covariance matrix are highlighted: $\bP^0_k$ (red node), which is the version to be initialized, and $\bP^H_k$ (green node), which is the most propagated version used to formulate the robust constraints in equation (\ref{eq:DOCPconstraints}).
+
+\begin{figure}
+	\centering
+	\begin{tikzpicture}[%
+		smallnode/.style={%
+			circle, draw, minimum size=3mm, inner sep=0pt
+		},
+		r_smallnode/.style={%
+			circle, draw, minimum size=3mm, inner sep=0pt, red
+		},
+		g_smallnode/.style={%
+			circle, draw, minimum size=3mm, inner sep=0pt, mygreen
+		},
+		every path/.style={->, thick},
+		x=1.5cm, y=1.5cm
+		]
+
+		\newcommand{\Plab}[2]{\bP^{#2}_{#1}}
+		\definecolor{mygreen}{RGB}{0 150 0}
+
+		% Nodes
+		% First row
+		\node[smallnode, label={[label distance=-2pt]above:\protect\(\Plab{k-2}{H-2}\)}] (Pk-2H-2) at (1,0) {};
+		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{H-1}\)}] (Pk-1H-1) at (2,0) {};
+		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k}{H}\)}] (PkH) at (3,0) {};
+		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{0}\)}] (Pk+10) at (4,0) {};
+		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{1}\)}] (Pk+21) at (5,0) {};
+
+		% Second row
+		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-2}{H-3}\)}] (Pk-2H-3) at (1,1) {};
+		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{H-2}\)}] (Pk-1H-2) at (2,1) {};
+		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k}{H-1}\)}] (PkH-1) at (3,1) {};
+		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{H}\)}] (Pk+1H) at (4,1) {};
+		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{0}\)}] (Pk+20) at (5,1) {};
+
+		% dots row 
+		\node[label={center,rotate=90:\(\dots\)}] (dots2) at (1, 1.7) {};
+		\node[label={center,rotate=90:\(\dots\)}] (dots3) at (2, 1.7) {};
+		\node[label={center,rotate=90:\(\dots\)}] (dots4) at (3, 1.7) {};
+		\node[label={center,rotate=90:\(\dots\)}] (dots5) at (4, 1.7) {};
+		\node[label={center,rotate=90:\(\dots\)}] (dots6) at (5, 1.7) {};
+
+		% Third row 
+		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k-2}{H}\)}] (Pk-2H) at (1,2) {};
+		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{0}\)}] (Pk-10) at (2,2) {};
+		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k}{1}\)}] (Pk1) at (3,2) {};
+		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{2}\)}] (Pk+12) at (4,2) {};
+		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{3}\)}] (Pk+23) at (5,2) {};
+
+		% Fourth row
+		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-2}{H-1}\)}] (Pk-2H-1) at (1,3) {};
+		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{H}\)}] (Pk-1H) at (2,3) {};
+		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k}{0}\)}] (Pk0) at (3,3) {};
+		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{1}\)}] (Pk+11) at (4,3) {};
+		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{2}\)}] (Pk+22) at (5,3) {};
+
+		% Curved arrows
+		\draw[bend left=20] (Pk-2H-2) to (Pk-1H-1);
+		\draw[bend left=20] (Pk-1H-1) to (PkH);
+		\draw[bend left=20] (Pk+10) to (Pk+21);
+
+		\draw[bend left=20] (Pk-2H-3) to (Pk-1H-2);
+		\draw[bend left=20] (Pk-1H-2) to (PkH-1);
+		\draw[bend left=20] (PkH-1) to (Pk+1H);
+
+		\draw[bend left=20] (Pk-10) to (Pk1);
+		\draw[bend left=20] (Pk1) to (Pk+12);
+		\draw[bend left=20] (Pk+12) to (Pk+23);
+
+		\draw[bend left=20] (Pk-10) to (Pk1);
+		\draw[bend left=20] (Pk1) to (Pk+12);
+		\draw[bend left=20] (Pk+12) to (Pk+23);
+
+		\draw[bend left=20] (Pk-2H-1) to (Pk-1H);
+		\draw[bend left=20] (Pk0) to (Pk+11);
+		\draw[bend left=20] (Pk+11) to (Pk+22);
+
+		% Dashed rectangle
+		\draw[dashed, thick, rounded corners] (2.6,-.25) rectangle (3.4,3.5);
+
+	\end{tikzpicture}
+	\caption{}
+	\label{fig:DOCPgrid}
+\end{figure}
 

### 2025-04-17 closed_loop_planning.tex
[MGa] textual description of closed-loop planning

@@
 
-Building upon our previous paper that presented a 3D multibody model suitable for embedding within an MLTP problem~\cite{Domenighini:Designs:2023} - treated here as the \emph{baseline} model - two models with increasing complexity are presented.
-The first represents a vehicle capable of independently steering all four wheels (AS) and applying four independent torques to each wheel (AIWD, aka, torque vectoring), whereas the second model additionally incorporates an active camber variation system for each wheel (DCC).
-%The kinematic tree of the two models is slightly different, but the procedure used to build the dynamic model is similar apart from a non-trivial aspect that will be discussed later on.
-In the remainder of this study, unless otherwise specified, the descriptions provided apply to both models.
-The figures related to Sections~\ref{sec:suspanal}-\ref{sec:tyremodel} are courtesy of the authors.
+The second approach plans over the full horizon while incorporating a feedback policy that mimics the driver's closed-loop behavior.
+This enables a more realistic propagation of the uncertainty encoded in the covariance matrix $\bP$, preventing its uncontrolled growth through the stabilizing effect of the driver's feedback.
 
-%% CHANGE THESE SECTIONS TO MERGE BOTH MODELS DESCRIPTIONS
-\subsection{Kinematic Tree Structure of the Vehicle}
-\begin{figure}[b]\centering
-	\begin{tikzpicture}[scale=1.2]\small
-		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw,gray](g)at(-0.1,-0.1){};
-		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](b)at(0,2){};
-		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](h11)at(-0.9,2.9){};
-		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](h12)at(0.9,2.95){};
-		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](h21)at(-1,1){};
-		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](h22)at(0.95,1){};
-		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](k11)at(-2.0,3){};
-		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](k12)at(2.0,3){};
-		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](k21)at(-2.1,1.1){};
-		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](k22)at(2.0,1.05){};
-		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](w11)at(-3.0,2.9){};
-		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](w12)at(3.0,2.9){};
-		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](w21)at(-3.1,1.){};
-		\node[fill,circle,minimum size=3pt,inner sep=0pt,draw](w22)at(3.0,.95){};
-		\draw[gray](g)to node(2){}(b);
-		\draw(b)to node(4){}(h11);
-		\draw(b)to(h12);
-		\draw(b)to(h21);
-		\draw(b)to(h22);
-		\draw(h11)to node(8){}(k11);
-		\draw(h12)to(k12);
-		\draw(h21)to(k21);
-		\draw(h22)to(k22);
-		\draw(k11)to node(6){}(w11);
-		\draw(k12)to(w12);
-		\draw(k21)to(w21);
-		\draw(k22)to(w22);
-		\draw(g)node[below,gray](1){$\{G\}$}(b)+(0,0.1)node[above](3){$\{B\}$}(h11)node[above](5){$\{H_1\}$}(h12)node[above]{$\{H_2\}$}(h21)node[below]{$\{H_3\}$}(h22)node[below]{$\{H_4\}$}
-		(k11)node[above](9){$\{K_1\}$}(k12)node[above]{$\{K_2\}$}(k21)node[below]{$\{K_3\}$}(k22)node[below]{$\{K_4\}$}
-		(w11)node[above](7){$\{W_1\}$}(w12)node[above]{$\{W_2\}$}(w21)node[below]{$\{W_3\}$}(w22)node[below]{$\{W_4\}$};
-		\node[gray,inner sep=1pt](world)at(1,-0.3){world};
-		\node[gray,inner sep=0pt,align=center](6DoF)at(-1.25,0.15){floating base\\(virtual 6-DoF)};
-		\node[inner sep=1pt](chassis)at(0.05,3){chassis};
-		\node[inner sep=0pt,align=center](susp)at(-2.2,1.6){suspension linkages\\(generalized 1-DoF)};
-		\node[inner sep=1pt](knuckle1)at(-.4,3.7){knuckles 1};
-		\node[inner sep=1pt](knuckle2)at(-2.05,3.8){knuckles 2};
-		\node[inner sep=0pt,align=center](hub)at(-4.5,2.25){hub bearings\\(revolute 1-DoF)};
-		\node[inner sep=0pt,align=center](camber)at(-2.,2.4){camber joint\\(revolute 1-DoF)};
-		\node[inner sep=1pt](rim)at(-3.4,3.65){rims};
-		\draw[gray](world.west)to(1.east);
-		\draw[gray](6DoF.north east)to(2);
-		\draw(chassis.south)to(3);
-		\draw(susp.north east)to(4);
-		\draw(knuckle1.south)to(5);
-		\draw(knuckle2.south)to(9);
-		\draw(hub.east)to(6);
-		\draw(camber.north)to(8);
-		\draw(rim.south east)to(7);
-	\end{tikzpicture}
-	\caption{Kinematic tree representing the interconnections between bodies. Bodies are referred to by their attached frame.}
-	\label{fig:tree}
-\end{figure}
-The vehicle is modeled as an articulated rigid-body system with chassis, knuckles, and wheels as rigid bodies.
-The chassis represents all sprung masses rigidly connected, while the knuckles (zero-mass bodies) decouple steering and vertical suspension motion from the wheel spindle.
-In the second model, knuckles are split into two zero-mass bodies, introducing an additional degree of freedom per wheel to simulate dynamic camber control (DCC).
-Wheels include all rotating unsprung masses, with non-rotating parts (e.g., brake calipers) incorporated into the wheel-tire mass.
-
-Each body has a right-handed barycentric frame: $\sref{B}$ (chassis), $\sref{H}$ and $\sref{K}$ (knuckles), and $\sref{W}$ (wheels).
-Additional frames $\sref{G}$ (ground) and $\sref{S}$ (centerline) are defined, with \(x\)-axis of $\sref{S}$ tangent to the centerline, pointing forward.
-Frame relationships are described by homogeneous transformation matrices $ g(d, \vPhi) \in SE(3) $, with $ d $ as position and \(\vPhi\) (Euler angles) parameterizing the rotation.
-
-As shown in Fig.~\ref{fig:tree}, bodies are connected by joints in a kinematic tree.
-The second model generalizes the first, where $\sref{H}$ and $\sref{K}$ coincide, so only the second model's kinematic tree is shown for brevity.
-%% START ORIGINAL
-%The vehicle is modeled as an articulated rigid-body system, with chassis, knuckles, and wheels treated as rigid bodies.
-The chassis includes all sprung masses, rigidly connected, while the knuckles (modeled as zero-mass bodies) decouple steering and vertical suspension motion from the wheel spindle.
-In the second model, the knuckles are split into two zero-mass bodies, adding one degree of freedom per wheel to simulate dynamic camber control (DCC).
-The wheels encompass all rotating unsprung masses, with non-rotating parts (e.g., brake calipers) included in the wheel-tire mass.
-%
-%Each body has a right-handed barycentric frame: $\sref{B}$ for the chassis, $\sref{H}$ (and $\sref{K}$ in the second model) for the knuckles, and $\sref{W}$ for the wheels.
-Two additional frames, $\sref{G}$ (ground) and $\sref{S}$ (attached to the vehicle's centerline), are also defined.
-The frame $\sref{S}$ lies in the tangent plane, with its $x$-axis tangent to the centerline, pointing forward.
-%
-%Relative positions and orientations between frames are described by homogeneous transformation matrices \( g(d, \vPhi) \in SE(3) \), where \( d \in \mathbb{R}^3 \) is the Cartesian position and \( \vPhi \in \mathbb{R}^3 \) are the Euler angles parameterizing the rotation matrix \( R \in SO(3) \).
-%
-%The bodies are connected by joints, forming the kinematic tree shown in Fig.~\ref{fig:tree}.
-The first model is a special case of the second, where $\sref{H}$ and $\sref{K}$ are always coincident, so only the kinematic tree for the second model is presented to avoid redundancy.
-
-%% END ORIGINAL
-
-%The vehicle is modeled as an articulated rigid-body system.
-In particular, chassis, knuckles and wheels are considered as rigid bodies.
-The chassis comprises all the sprung masses, including the driver, which are rigidly attached to one another.
-The knuckles are assumed here as zero-mass bodies, and their function is to decouple the motion of the suspension due to steering and vertical travel to the wheel spindle.
-In the second model, the knuckles are split into two zero-mass bodies, introducing one additional degree of freedom (DoF) for each wheel that allows to simulate the active camber variation system, here named dynamic camber control (DCC).
-The wheels include all the rotating unsprung masses, while all non-rotating parts of the unsprung masses, such as the brake calipers are included in the wheel-tire mass.
-%
-%Each body has a right-handed barycentral reference frame attached.
-The frames are denoted as $\sref{B}$, $\sref{H}$ (and $\sref{K}$) and $\sref{W}$ for the chassis, knuckles and wheels, respectively, where $\sref{K}$ is present only in the second model.
-Two additional frames are introduced: $\sref{G}$, fixed to the ground, and $\sref{S}$, that follows the vehicle on the ground.
-More precisely, $\sref{S}$ is attached to the centerline curve with its $x$ and $y$ axes lying in the tangent plane, and the $x$ axis also tangent to the centerline curve pointing forward.
-%
-%The relative positions and orientations between reference frames are described by homogeneous transformation matrices $g(d,\vPhi)\in SE(3)$, where $d\in \bbR^3$ is the relative Cartesian coordinates vector and $\vPhi \in \bbR^3$ are the three Euler angles that parameterize the rotation matrix $R\in SO(3)$.
-%
-%The bodies are connected by joints and the kinematic tree of the second model is shown in Figure \ref{fig:tree}.
-It is worth noting that the first model can be seen as a particular case of the second, in which the reference frames $\sref{H}$ and $\sref{K}$ are constrained to be always overlapping.
-For this reason, and to avoid cluttering the document, has been chosen to show the kinematic tree only for the second model.
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[scale = 0.8]{Images/vehicle_frames}
-	\caption{Reference frames: $\sref{G}$ is a fixed inertial reference frame with the $z$-axis antiparallel to the gravity vector; $\sref{S}$ is a reference frame following the vehicle on track; $\sref{B}$ and $\sref{K}$ are FLU (Forward Left Up) reference frames aligned with the principal axes of inertia of the chassis and wheel, respectively; $\sref{H}$ is an auxiliary reference frame necessary to add the camber DoF; and $\sref{W}$ rotates about the $y$-axis with respect to $\sref{K}$ by the wheel rotation angle.}
-	\label{fig:frames}
-\end{figure}
-
-The joints modeled are: hub bearings, connecting the rims $\sref{W}$ to the knuckles $\sref{H}$ (or $\sref{K}$ for the second model), and the suspension linkages, connecting the knuckles $\sref{H}$ to the chassis $\sref{B}$.
-For the second model a 1 DoF revolute joint connecting the two parts of the knuckles is also present.
-The chassis is a floating-base and its position and orientation with respect to the ground are unconstrained.
-This situations is conveniently encoded by a virtual six-DoF joint connecting $\{B\}$ to $\{G\}$.
-The resulting systems have 4+4+6=14 and 4+4+4+6=18 degrees of freedom, respectively.
-%After defining the system's topology, a set of suitable state variables needs to be defined to describe the configuration and motion of each body.
-
-%To fully describe the position and orientation of the frame $\sref{B}$ w.r.t. $\sref{G}$ we define vector $q_{gb}=(d_{gb};\varPhi_{gb})$.
-%Its first-order kinematics is encoded in twist $V_{gb}^b = (v_{gb};\omega_{gb})$.
-In general, a twist $V_{ij}^k$ represents the motion of the frame $\sref{J}$ w.r.t.
-the frame $\sref{I}$ in $\sref{K}$ components.
-Unless otherwise specified, we employ \emph{body} coordinates, therefore $k=j$; this choice aligns to ABA convention~\cite{Featherstone:book:2008}.
-For each of the 4 DoF's associated to the suspension motion, a variable $z$ (and its derivative $\dot{z}$) is defined representing spring's length variation with respect to the nominal value.
-With these definitions, along with variables $\delta$ and its derivatives, defined in Sec.\ref{sec:suspanal}, one can find the six-component vector $q_{bh}$ representing the pose of $\sref{H}$ w.r.t.
-$\sref{B}$ and the twist $V_{bh}$.
-Assuming the wheel is an axis-symmetrical rigid body, and given that the relative orientation of the frame $\sref{W}$ w.r.t.
-$\sref{H}$ (or $\sref{K}$) is not relevant for the scope, only the angular velocity $\omega$ is introduced.
-For the second model, it is also necessary to introduce the variable $q_{hk}$ to describe the relative position of the frame $\sref{K}$ w.r.t.
-$\sref{H}$.
+The approach is based on the three following steps: i) A nominal time-optimal trajectory is planned: this will be referred to as nominal feed-forward optimal trajectory.
+This is the trajectory a driver should follow in the ideal situation of a world without uncertainty.
+Therefore, in this step the mean trajectory is the actual trajectory a deterministic vehicle dynamics would trace.
+ii) A closed-loop controller mimicking the driver's action is computed which stabilizes the nominal feed-forward optimal trajectory computed at step i).
+We assume a discrete time-varying Linear Quadratic Regulator (LQR) controller is a good approximation of how an expert driver can track a prescribed trajectory.
+iii) In the last step the stochastic framework is reintroduced and the time-optimal planning problem is robustified using the Lyapunov framework.
+The main idea is to \emph{re-plan} a mean trajectory of a stochastic system such that, a time-varying LQR controller that aims at stabilizing it, is able to properly tame the propagation of the uncertainty.
+This amounts to ensuring that the robustified path constraints are satisfied.
+More specifically, this implies two key aspects: first, that the covariance matrix is propagated accounting explicitly for the controlled system dynamics; and second, that the path constraints are robustly satisfied by incorporating the covariance estimation propagated through the closed-loop controlled system.
+Some aspects that may initially appear technical are in fact essential.
+In this third step, not only are states and controls re-planned, but also the time-varying LQR gains.
+In fact, they need to explicitly account for the changed requirements associated with stabilizing a trajectory different from the nominal one, which must now satisfy the robustified constraints.
 \subsection{Suspension Analysis}

### 2025-04-17 bis_open_loop_planning.tex
[MM] DOCP for OpenLoop

@@
-\section{Open-loop planning via n-steps ahead predictions}
+\section{Open-loop planning via H-steps ahead predictions}
 \label{sec:open_loop_planning}
+The proposed approach requires the presence of $H+1$ versions of the matrix $\bP$ at each grid point of the discretized problem.
+One of these versions is initialized with $\bar{\bP}_0$, while the others represent the evolution of matrices that were initialized at previous grid points, from 1 to $H$ steps earlier.
+This ensures that at each step it is possible to find a covariance matrix evolved for $H$ steps and use it to robustify the constraints.
+The initialization of the covariance matrix is necessary due to its divergent dynamics.
+In fact, without a feedback control, the error on the position and orientation of the body reference frame can only increase along the track, compromising the conditioning of the matrix.
 
-The proposed approach requires the presence of $H+1$ versions of the matrix $\bP$ at each grid point.
-One of these versions is initialized with $\bar{\bP}_0$, while the others represent the evolution of matrices that were initialized at previous grid points, from 1 to $H$ steps earlier.
-This ensures that at each step it is possible to find a covariance matrix evolved for $H$ steps and use it to robustify the constraint.
-To adapt the problem, it is necessary to modify equations (\ref{eq:DOCPdP}) and (\ref{eq:DOCPconstraints}).
-In this case,  (\ref{eq:DOCPdP}) collects $H$ collocation and continuity equations and an initialization equation, while (\ref{eq:DOCPconstraints}) takes as input the most propagated version of the covariance matrix at the k-th step, here named $\bP^H_k$.
-The new formulation is reported:
+\subsection{Discretization via direct collocation}
+\label{sec:discretization}
+
+The nonlinear optimal control problem~\eqref{eq:OCPcost} can be discretized by applying a suitable collocation integrator obtaining the following nonlinear program (NLP) 
 \begin{alignat}{3}
-\bzero       = & \; \bPsiP_k(\bmu_k,\bxi_k, \bu_k, \bP_{k-1}^{j-1},\bP_k^j,\bSi_k^j,\bz_k),
-& \quad & k = 1,\ldots, N; \; j=0,\ldots,H \label{eq:DOCPdPopenloop} \\
-0            \geq & \; h_i(\bmu_k, \bu_k, \bz_k) + \be_i(\bmu_k, \bu_k, \bP_k^j, \bz_k),
+	\underset{\bmu_k,\bxi_k, \bu_k, \bP_k,\bz_k}{\text{minimize}} \,
+	& & & J_k(\bmu_k,\bxi_k, \bu_k) & & \label{eq:DOCPcost} \\
+	\hspace*{-2.0 cm}\text{s.t.} \quad
+	& \bzero      & = & \; \bPsimu_k(\bmu_{k-1},\bmu_k,\bxi_k, \bu_k,\bz_k),
+	& \quad & k = 1,\ldots, N \label{eq:DOCPdyn} \\
+	& \bmu_0      & = & \; \bar{\bmu}_0
+	& & \label{eq:DOCPdynIC} \\
+	& \bzero      & = & \; \bPsiP_k(\bmu_k,\bxi_k, \bu_k, \bP_{k-1}^{j-1},\bP_k^j,\bSi_k^j,\bz_k),
+	& \quad & \substack{k = 1,\ldots, N; \; \\ j=1,\ldots,H;\;} \label{eq:DOCPdPopenloop} \\
+	& \bP_0^j       & = & \; \bar{\bP}_0 \succeq 0
+	& \quad & j=0,\ldots,H; \; \label{eq:DOCPdPIC} \\
+	& \bP_k^0       & = & \; \bar{\bP}_0 \succeq 0
+	& \quad & k = 1,\ldots, N; \; \label{eq:DOCPdPinit} \\
+	& \bzero      & = & \; \bOm_k(\bmu_k,\bxi_k, \bu_k,\bz_k),
+	& \quad & k = 0,\ldots, N \label{eq:DOCPpath} \\
+	& 0           & \geq & \; h_i(\bmu_k, \bu_k, \bz_k) + \be_i(\bmu_k, \bu_k, \bP_k^H, \bz_k),
 	& \quad & k = 1,\ldots, N;\; i \in \calI \label{eq:DOCPconstraintsopenloop}
 \end{alignat}
-This process is illustrated in Figure \ref{fig:DOCPgrid}, where the dashed rectangle represent the $k$-th step at which the constraints are being formulated.
-At each grid point, two versions of the covariance matrix are highlighted: $\bP^0_k$ (red node), which is the version to be initialized, and $\bP^H_k$ (green node), which is the most propagated version used to formulate the robust constraints in equation (\ref{eq:DOCPconstraints}).
+To perform the discretization, the track centerline is parameterized using a curvilinear abscissa $\alpha \in [0,1]$, and uniformly sampled at $N+1$ points $\alpha_0, \ldots, \alpha_N$.
+Accordingly, $\bmu_k$ denotes the mean state at grid node $\alpha_k$, while $\bu_k$ and $\bz_k$ are assumed to be piecewise constant over each interval $[\al_k, \al_{k+1}]$.
+These represent the controls and the algebraic variables, respectively. The $\bz_k$'s are introduced as direct handles for physically meaningful quantities such contact forces. 
+Similarly, $\bP_k^j$ represents the $j$-th version of the covariance matrix at the $k$-th node.
+Following the direct collocation approach, both the mean and covariance state trajectories are approximated, in the $k$-th interval, with polynomials $\pi_k(\tau)$, defined on the unit interval $\tau\in[0,1]$, and then scaled to match the width $h_k$ of the corresponding time step.
+On the unit interval, we select $d$ collocation points $\tau_1, \ldots, \tau_d$, associated with as many collocation states.
+Accordingly, we define $\bxi_k$ and $\bSi_k^j$ as the mean states and covariance matrices at the $d$ collocation points within each interval $[\al_k, \al_{k+1}]$, respectively.
+Therefore, $\bxi_k =
+(\bxi_{k,1}, \ldots, \bxi_{k,d})$, and similarly $\bSi_k^j=(\bSi_{k,1}^j, \ldots, \bSi_{k,d}^j)$. 
+Equations~\eqref{eq:DOCPdyn} and~\eqref{eq:DOCPdPopenloop} represent the collocation and continuity equations for mean and covariance, respectively, with initial conditions represented by~\eqref{eq:DOCPdynIC} and~\eqref{eq:DOCPdPIC}.
+Eq.~\eqref{eq:DOCPdPinit} is needed to initialize the covariance matrices that evolved for $H$ step, as previously described. 
+Eqs.~\eqref{eq:DOCPpath} are path equality constraints and~\eqref{eq:DOCPconstraintsopenloop} are the robustified inequality constraints.
+
+A schematic representation of the approach used to manage the $H+1$ versions of the covariance matrix is shown in Figure \ref{fig:DOCPgrid}, where the dashed rectangle represent the $k$-th step at which the constraints are being formulated.
+At each grid point, two versions of the covariance matrix are highlighted: $\bP^0_k$ (red node), which is the version to be initialized, and $\bP^H_k$ (green node), which is the most propagated version used to formulate the robust constraints in Eq.~\eqref{eq:DOCPconstraintsopenloop}.
+
+In certain cases~\cite{Gillis:PracticalMethodsApproximate:2015}, positive-definiteness-preserving Lyapunov discretization schemes are used in~\eqref{eq:DOCPdPopenloop}.
+However, for sufficiently fine discretization, we found that integrating only the lower triangular part of $\bP(t)$ in~\eqref{eq:OCPdPIC}  (and reconstructing its strictly upper part accordingly), ensures that all $\bP_k^j$ remain symmetric and positive definite when the initial $\bP_0$ is symmetric and positive definite.
+In all our tests direct collocation with cubic polynomial state representations and Gauss-Legendre collocation points were used.
 
@@
 	\end{tikzpicture}
-	\caption{}
+	\caption{Schematic representation of the continuity and initialization Equations for the $H+1$ versions of the covariance matrix introduced in the discretized OCP.
+The dashed rectangle indicate the $k$-th step, in which the constraints are formulated.
+Each node represents a covariance matrix, labeled such that the subscript denotes the current grid step, while the superscript indicates the number of propagation steps it has undergone.
+At each step, two nodes are highlighted: the covariance matrix to be initialized (red node) and the matrix that has been propagated for $H$ steps (green node), which is used to formulate the robust constraints.}
 	\label{fig:DOCPgrid}
@@
 
-\subsection{Direct Collocation}
-The objective is to recast the MLTP into a Non Linear Program (NLP) using a direct orthogonal collocation approach, specifically based on Gauss-Legendre collocation points.
-The problem is formulated within the CasADi framework~\cite{Andersson:MPC:2019} and solved using the interior point method implemented in IPOPT~\cite{Wachter:MP:2006}.
-The resulting NLP can be expressed in the form:
-\begin{align}
-	\underset{x,v,u,z}{\textrm{minimize}}&\quad\sum_{k=1}^NL_k(x_k,v_k,u_k,z_k)\label{eqn:functional}\\
-	\textrm{subject to}&\quad g_k(x_{k-1},x_k,v_k,u_k,z_k)=0,\ k=1,\ldots,N,\label{eqn:equality}\\
-	&\quad h_k(x_k,u_k,z_k)\le0,\ k=1,\ldots,N.\label{eqn:inequality}
-\end{align}
-In Eqs.~\eqref{eqn:functional}, \eqref{eqn:equality} and \eqref{eqn:inequality}, in addition to the state and control vectors, $v_k$ and $z_k$ are included, representing the state vector at the collocation points and the algebraic quantities, respectively.
-The controls and the algebraic variables are approximated as piecewise constant, while the states are approximated as quadratic polynomials within each finite element.
-Specifically, the algebraic variables in the problem are the time steps $\nu_k$ and the lateral and vertical displacements of the Center of Mass relative to the origin of the $k$-th frame $\sref{S}$, attached to the track's centerline ($e_y$ and $e_z$).
-Their function is to serve as handle variables in the optimization process to promote problem sparsity.
-
-The track centerline is represented by a curvilinear abscissa $\alpha\in[0,1]$, sampled at $N+1$ points $\alpha_0,\ldots,\alpha_N$.
-Each grid point is associated with a reference frame $\sref{S_k}$, whose origin lies at the corresponding point on the track centerline, as illustrated in Figure~\ref{fig:track_frames}.
-A more detailed description of the 3D track modeling can be found in~\cite{Bartali:Meccanica:2023}.
-Specifically, this representation allows us to capture track slope and banking, but it does not support non-developable track surfaces.
-
-\begin{figure}[h]\centering
-	\includegraphics[width=0.8\textwidth]{Images/track_frames.pdf}
-	\caption{Track frames corresponding to adjacent nodes of the grid over $\alpha$.
-Note that the grid points need not be equispaced.
-If desired, they can be chosen irregularly, so as to capture more features of the track where it is more critical (e.g.
-in proximity to corners) and save space where it is more regular (e.g.
-on the straights).}
-	\label{fig:track_frames}
-\end{figure}
-
-%According the direct collocation method, the state evolution on the $k$-th element of the grid is approximated with a polynomial $p_k$.
-The number of collocation points depends on the degree $d$ of the polynomial chosen and their position inside the interval can be selected based on different criteria; here the Gauss-Legendre collocation points are used.
-The polynomial is built on the unit interval and then scaled to match the length $\nu_k$ of the relative time step.
-%
-%In Eq.~\eqref{eqn:equality}, besides specific constraints inserted to bias the solution towards desired behaviours, two equations are paramount. The first  one is:
-%\begin{equation}\label{eqn:collocation}
-%	\frac{1}{\nu_k}\frac{\partial}{\partial\tau}p_k(\tau_i)=f(v_{k,i},u_k),
-%\end{equation}
-%the collocation equation, that constrains the state at each collocation point to comply with the system dynamics. The second one, which is $p(1)=x_{k+1}$, ensures continuity across time intervals.
-
-In addition to the classical collocation and continuity constraints~\cite{Domenighini:Designs:2023}, the inequalities in Eq.~\eqref{eqn:inequality} for our MLTP ensure the vehicle remains within track bounds.
-These include the requirement for the CoM to lie on the $yz$ plane of the frame $\sref{S_k}$, as well as power and adherence limits.
-
-In our optimizations, the cost function at each step is expressed as follows:
-\begin{equation}\label{eqn:cost}
-	L_k=\nu_k^2+k_\delta(\delta_k-\delta_{k-1})^2+k_T(T_k-T_{k-1})^2+k_v x_{k,2}^2 + l_k(\ddot{q}_{hk,k},\delta_k,x_k),
-\end{equation}
-where the first term penalizes travel time, the second and third prevent abrupt variations of the inputs and avoid singular arcs.
-The quadratic penalization ensures a well-defined dependence of the Hamiltonian on the control variables, preventing indeterminate solutions.
-The last two serve to shape the driving behaviors according to the designer preferences.
-
-Their primary goal is to improve the optimization landscape, ensuring that the solver is less likely trapped by undesirable local minima.
-As an example, the fourth term, which includes the lateral velocity $x_{k,2}$, has been included to prevent undesirable behaviors which could emerge as a result of the increased system's DoFs in combination with the simplified aerodynamic modeling adopted in this work.
-Some local minima could automatically be avoided by incorporating more realistic aerodynamic maps and tyre wear models.
-These enhancements will be explored in future works.
-In this context, an interesting direction is presented in~\cite{Masouleh:IEEE:2016}, where the authors introduce an aerodynamic model that explicitly accounts for the influence of suspension and vehicle configuration.
-This approach offers a more comprehensive representation of the aero forces and could be effectively used to refine the current aerodynamic map.
-By incorporating such a model, we expect not only a more physically consistent representation of the aerodynamics but also a smoother and more meaningful optimization landscape, reducing the occurrence of non-physical local minima and improving overall solution robustness.%More in details, the all steering vehicle, having the capability to steer independently with all four wheels, and considering that the aerodynamic drag only depends on the forward velocity $x_1$, could run with excessively large $\beta$ angles, even on straights paths.
-Even if this behavior could be appropriate in certain track sectors, is undesirable from an engineering perspective.
-%In some cases, it could also be the outcome of the solver being stuck in local minima: in these events the above term helps discourage such behaviours.
-
-In the current context, we imposed coefficient $k_v$ to be a function of the (absolute value of) track curvature $c_k$ according to Eq.~\eqref{eqn:lateral_cost}.
-Here, the two constant $\bar{k}_v$ and $k_c$ are selected heuristically for each track; the first increases/decreases the weight of the cost term, while the second varies the hardness of the intervention in the range $[0,1]$.
-\begin{equation}\label{eqn:lateral_cost}
-	k_v = \bar{k}_v(1-\tanh(k_c c_k))
-\end{equation}
-
-To ensure reproducibility, we report the values of the main penalty weights for the cost function terms used in the Siena track simulations: $k_{\delta} = 0.05$, $k_T = 0.005$, $\bar{k}_v = 0.005$, and $k_c = 100$.
+\subsection{Robust adherence constraint}

### 2025-04-17 stochastic_vehicle_dynamics.tex
[MM] DOCP for OpenLoop

--- 2025-04-14 stochastic_vehicle_dynamics.tex
@@
 
-%To compute the tangential components of the force exchanged by tyre and road, a formulation of the Pacejka's \emph{Magic Formula} ~\cite[Section~4.3.2]{Pacejka:book:2012} is used, which reads:
-%\begin{equation}\label{eqn_mf}
-%	[F_x,F_y]=\mf(\kappa,\alpha,F_z,\gamma).
-%\end{equation}
-%The Magic Formula $\mf$ computes the longitudinal and lateral forces $F_x,F_y$ as a function of the \emph{tyre slips} $\kappa,\alpha$~\cite[Section~1.2.1]{Pacejka:book:2012}.
-The formula is also sensitive to variations of vertical load $F_z$ and camber angle $\gamma$.
+%\subsection{Discretization via direct collocation}
+%\label{sec:discretization}
 %
-%The auxiliary frame $\sref{N}$ is introduced to compute quantities used in Eq.~\eqref{eqn_mf} and the (attempted) interpenetration $d$ (see Figure.~\ref{fig:tyre_wheel}) of the tyre in the ground, necessary to compute the dynamic vertical load $F_z$.
-The construction of the frame $\sref{N}$ is reported in Figure~\ref{fig:tyre_wheel}.
-It is worth noting that the frame $\sref{N}$ is fully described if $\sref{H}$ and $\sref{S}$ are known.
-%The model used for the contact is a unilateral, penalty-based compliant tyre model, that considers the wheel as a radial spring with constant stiffness $k_t$ and constant damping coefficient $c_t$\footnote{The introduction of a more refined model with nonlinear spring and damping characteristics would not pose particular difficulties.}.
-The traditional role of compression in a spring is conducted by $d$, which is computed considering the road and the lowermost point of the non-deformed tyre.
-The resultant function that relate $d$, $\dot{d}$ and $F_z$ is:
-%\begin{equation}\label{eqn_Fz}
-%	F_z= F_0\log_2\left(1 + 2^{\frac{k_t d + c_t \dot{d}}{F_0}}\right)
-%\end{equation}
-%This function, with $F_0$ tuned to be negligible, is differentiable for $d=0$ and $\dot{d}=0$, enhancing numerical performance, and behaves as a spring-damper system, as requested, when the tyre is in contact with the ground.
-When the tyre detaches completely from the ground, the resultant vertical forces tends to zero.
-However, since its gradient is not zero, this formulation helps the optimization algorithm to figure out proper search directions.
-This allows to deal with jumps and other situations in which the tyre may lose contact with the ground.
-
-\subsection{Discretization via direct collocation}
-\label{sec:discretization}
-
-The nonlinear optimal control problem~\eqref{eq:OCPcost} can be discretized by applying a suitable collocation integrator obtaining 
-\begin{alignat}{3}
-\underset{\bmu_k,\bxi_k, \bu_k, \bP_k,\bz_k}{\text{minimize}} \,
-& & & J_k(\bmu_k,\bxi_k, \bu_k, \bP_k) & & \label{eq:DOCPcost} \\
-\hspace*{-2.0 cm}\text{s.t.} \quad
-& \bzero      & = & \; \bPsimu_k(\bmu_{k-1},\bmu_k,\bxi_k, \bu_k,\bz_k),
-& \quad & k = 1,\ldots, N \label{eq:DOCPdyn} \\
-& \bmu_0      & = & \; \bar{\bmu}_0
-& & \label{eq:DOCPdynIC} \\
-& \bzero      & = & \; \bPsiP_k(\bmu_k,\bxi_k, \bu_k, \bP_{k-1},\bP_k,\bSi_k,\bz_k),
-& \quad & k = 1,\ldots, N \label{eq:DOCPdP} \\
-& \bP_0       & = & \; \bar{\bP}_0 \succeq 0
-& & \label{eq:DOCPdPIC} \\
-& \bzero      & = & \; \bOm_k(\bmu_k,\bxi_k, \bu_k,\bz_k),
-& \quad & k = 0,\ldots, N \label{eq:DOCPpath} \\
-& 0           & \geq & \; h_i(\bmu_k, \bu_k, \bz_k) + \be_i(\bmu_k, \bu_k, \bP_k, \bz_k),
-& \quad & k = 1,\ldots, N;\; i \in \calI \label{eq:DOCPconstraints}
-\end{alignat}
-
-\subsection{External Wrenches}
-\label{sec:extwren}
-The external wrenches acting on each body, from wheels to chassis, are expressed in body-fixed components.
-Only gravity, aerodynamic forces, and road-tyre interactions are considered.
-Following~\cite{Guiggiani:book:2023}, the aerodynamic wrench in frame $\{B\}$ is $W_a=-\frac{1}{2}\rho S v_{gb,1}^2 [C_x, 0, C_z, 0, -h_0 C_x - a_f C_{zf} + a_r C_{zr}, 0]^T$, where $\rho$ is air density, $S$ is the frontal area, $v_{gb,1}$ the forward velocity (component of \emph{distal rigid-body twist} $V_{gb}$, see~\cite[p.~4]{Domenighini:Designs:2023}), and $C_x, C_z$ are drag/lift coefficients.
-$C_z$ is split into $C_{zf}=k_a C_z$ and $C_{zr}=(1-k_a)C_z$ by the aerodynamic balance coefficient $k_a$.
-
-The total external wrench on a wheel is $W_{we}=\Ad_{g_{hn}}^*W_t+W_{w_h}$,
-%\begin{equation}\label{eqn:Wwe}
-%W_{we}=\Ad_{g_{hn}}^*W_t+W_{w_h},
-%\end{equation}
-where $W_t=[F_x, F_y, F_z, 0, 0, 0]^T$ is the ground wrench transferred from $\sref{N}$ to $\sref{H}$ by the \emph{starred} Adjoint operator, and $W_{w_h}=[F_{w_h}^T R_{hg}^T, 0, 0, 0]^T$ is the weight contribution with $F_{w_h}=[0, 0, -m_h g]^T$ in frame $\{G\}$, where $m_h$ is the wheel mass.
-
-Knuckles experience no external forces, therefore:
-\begin{equation}\label{eqn:Whe}
-	W_{he}=[0,0,0,0,0,0]^T\quad\textrm{and}\quad W_{ke}=[0,0,0,0,0,0]^T.
-\end{equation}
-The chassis's external wrench is $W_{be}=W_a+W_{w_b}$, %\begin{equation}\label{eqn:Wbe}
-%	W_{be}=W_a+W_{w_b},
-%\end{equation}
-where $W_a$ and $W_{w_b}$ are aerodynamic and weight contributions, respectively.
-For $W_{w_b}=[F_{w_b}^T R_{bg}^T, 0, 0, 0]^T$, the weight force is $F_{w_b}=[0, 0, -m_b g]^T$ in frame $\{G\}$, with $m_b$ as the chassis mass.
-%The purpose of this section is to describe the external wrenches acting on each body, from the wheels to the chassis, each expressed in body-fixed components.
-%Besides gravity, aerodynamic forces and the interactions between road and tyres are the only external forces considered.
-%Adopting the notation in~\cite{Guiggiani:book:2023}, the aerodynamic contribution to the wrenches in frame $\{B\}$ is $W_a=-\frac{1}{2}\rho Sv_{gb,1}^2[C_x,0,C_z,0,-h_0C_x-a_fC_{zf}+a_rC_{zr},0]^T$, where $\rho$ is the air density, $S$ the frontal area of the vehicle, $v_{gb,1}$ the forward velocity, $h_0$ the nominal height of the CoM from ground and $a_f,a_r$ the distances of the CoM from the front and rear axle, respectively.
-The drag and lift coefficient $C_x$ and $C_z$ are dimensionless parameters that depend on the shape of the vehicle's body.
-$C_z$ can be further partitioned into $C_{zf}=k_a C_z$ and $C_{zr}=(1-k_a) C_z$ according to the \emph{aerodynamic balance coefficient} $k_a$, which is a parameter of the vehicle set-up.
+%The nonlinear optimal control problem~\eqref{eq:OCPcost} can be discretized by applying a suitable collocation integrator obtaining the following nonlinear program (NLP) 
+%\begin{alignat}{3}
+%\underset{\bmu_k,\bxi_k, \bu_k, \bP_k,\bz_k}{\text{minimize}} \,
+%& & & J_k(\bmu_k,\bxi_k, \bu_k) & & \label{eq:DOCPcost} \\
+%\hspace*{-2.0 cm}\text{s.t.} \quad
+%& \bzero      & = & \; \bPsimu_k(\bmu_{k-1},\bmu_k,\bxi_k, \bu_k,\bz_k),
+%& \quad & k = 1,\ldots, N \label{eq:DOCPdyn} \\
+%& \bmu_0      & = & \; \bar{\bmu}_0
+%& & \label{eq:DOCPdynIC} \\
+%& \bzero      & = & \; \bPsiP_k(\bmu_k,\bxi_k, \bu_k, \bP_{k-1},\bP_k,\bSi_k,\bz_k),
+%& \quad & k = 1,\ldots, N \label{eq:DOCPdP} \\
+%& \bP_0       & = & \; \bar{\bP}_0 \succeq 0
+%& & \label{eq:DOCPdPIC} \\
+%& \bzero      & = & \; \bOm_k(\bmu_k,\bxi_k, \bu_k,\bz_k),
+%& \quad & k = 0,\ldots, N \label{eq:DOCPpath} \\
+%& 0           & \geq & \; h_i(\bmu_k, \bu_k, \bz_k) + \be_i(\bmu_k, \bu_k, \bP_k, \bz_k),
+%& \quad & k = 1,\ldots, N;\; i \in \calI \label{eq:DOCPconstraints}
+%\end{alignat}
+%To perform the discretization, the track centerline is parameterized using a curvilinear abscissa $\alpha \in [0,1]$, and uniformly sampled at $N+1$ points $\alpha_0, \ldots, \alpha_N$.
+%Accordingly, $\bmu_k$ denotes the mean state at grid node $\alpha_k$, while $\bu_k$ and $\bz_k$ are assumed to be piecewise constant over each interval $[\al_k, \al_{k+1}]$.
+%These represent the controls and the algebraic variables, respectively.
+The $\bz_k$'s are introduced as direct handles for physically meaningful quantities such contact forces.
+Similarly, $\bP_k$ represents the covariance matrix at the $k$-th node.
+%Following the direct collocation approach, both the mean and covariance state trajectories are approximated, in the $k$-th interval, with polynomials $\pi_k(\tau)$, defined on the unit interval $\tau\in[0,1]$, and then scaled to match the width $h_k$ of the corresponding time step.
+On the unit interval, we select $d$ collocation points $\tau_1, \ldots, \tau_d$, associated with as many collocation states.
+Accordingly, we define $\bxi_k$ and $\bSi_k$ as the mean states and covariance matrices at the $d$ collocation points within each interval $[\al_k, \al_{k+1}]$, respectively.
+Therefore, $\bxi_k =
+%(\bxi_{k,1}, \ldots, \bxi_{k,d})$, and similarly $\bSi_k=(\bSi_{k,1}, \ldots, \bSi_{k,d})$.
+Equations~\eqref{eq:DOCPdyn} and~\eqref{eq:DOCPdP} represent the collocation and continuity equations for mean and covariance, respectively, with initial conditions represented by~\eqref{eq:DOCPdynIC} and~\eqref{eq:DOCPdPIC}.
+Eqs.~\eqref{eq:DOCPpath} are path equality constraints and~\eqref{eq:DOCPconstraints} are the robustified inequality constraints.
 %
-%The total external wrench acting on a (general) wheel is:
-%\begin{equation}\label{eqn:Wwe}
-%	W_{we}=\Ad_{g_{hn}}^*W_t+W_{w_h},
-%\end{equation}
-%where $W_t=[F_x,F_y,F_z,0,0,0]^T$ are the ground forces computed in $\sref{N}$ and properly transferred to $\sref{H}$ components by the \emph{starred} Adjoint operator, while $W_{w_h} = [ F_{w_h}^T R^T_{hg}, 0,0,0 ]^T$, with $F_{w_h}=[0,0,-m_h g]^T$ in frame $\{G\}$, represents the weight contribution and $m_{h}$ is the wheel mass.
-%
-%Under our hypotheses, the knuckles are not subjected to external forces, thus \begin{equation}\label{eqn:Whe}
-%	W_{he}=[0,0,0,0,0,0]^T\quad\textrm{and}\quad W_{ke}=[0,0,0,0,0,0]^T.
-%\end{equation}
-%
-%The resultant external wrench acting on the chassis is therefore:
-%\begin{equation}\label{eqn:Wbe}
-%	W_{be}=W_a+W_{w_b},
-%\end{equation}
-%where $W_a$ and $W_{w_b}$ are, respectively, the contributions of aerodynamic and chassis weight forces, whose general expression is of the form $W_{w_b} = [ F_{w_b}^T R^T_{bg}, 0,0,0 ]^T$.
-Explicitly we have $F_{w_b}=[0,0,-m_b g]^T$ in frame $\{G\}$, where $m_b$ is the mass of the chassis.
-
-\subsection{Recursive computation of the dynamics equations with ABA}
-\label{sec:aba}
-Thanks to the ABA algorithm, assuming known joint positions and velocities and joint forces and torques, it is possible to compute recursively and efficiently the accelerations of the joint variables.
-Following~\cite{Domenighini:Designs:2023}, twists are streamlined as follows: $V_{gb}^b = V_{gb} = V_b$, $V_{bh}^h=V_{bh}$, $V_{hw}^h=V_{hw}$.
-For the second model (DCC) the last expression is replaced with $V_{kw}^k=V_{kw}$ and $V_{hk}^k=V_{hk}$ is added.
-Given that the first model (without camber control) is a particular case of the DCC one, to avoid duplications only the algorithm for DCC is reported here.
-
-The traditional ABA requires knowledge of the active forces and torques, denoted as $\tau_{ij}$, exerted by the parent body ($i$) on the child body ($j$).
-Its purpose is to solve the \emph{Forward Dynamics} problem of an articulated systems of rigid bodies.
-In our floating-base system, given the twist $V_b$ of the parent body $\{B\}$, the values and time derivatives of all joint variables ($q$ and $\dot{q}$), and the torque vector ($\tau$), the objective is to compute the joint accelerations ($\dot{V}_{b}$ and $\ddot{q}$).
-%The fundamental concept is that of the inertia of an articulated body.
-This is represented by that of the root rigid body (original parent) plus the portion of the sub-trees, composed by the children bodies, whose inertia that can be structurally transferred backwards through the joints.
-%This depends on the architecture, the pose and the joints' types.
-The overall computation is performed in three sweeps defined recursively.
-One of the greatest computational benefits when it is employed to assemble symbolic dynamic equations, as in our case, is that \emph{no matrix inversion is required}, thus contributing to streamline the algebraic expressions.
-
-The first step of the algorithm, as described in \textbf{Step 1}, performs the forward propagation of rigid-body velocities: from the root to the leaves of the tree.
-%Twists are calculated as the sum of parent twist and the relative twist which is expressed consistently with the joint type.
-The interesting Jacobians are listed here: $J_{hk}=[0,0,0,1,0,0]^T$ and $J_{kw}=[0,0,0,0,1,0]^T$, and represent two revolute joints with axes $x$ and $y$ respectively.
-In Algorithm~\ref{alg:step1} the terms ``Knuckle 1'' and ``Knuckle 2'' are used to identify the two halves of the knuckle, where, using kinematic trees terminology, the first is the parent of the second.
-The Adjoint operator is defined as follows:
-\begin{equation}\label{eq:Adjoint} \Ad_{g}=\left[\begin{array}{cc}R&\hat{d}R\\0&R\end{array}\right].
-\end{equation}
-In the case $g_{bh}\in SE(3)$ is considered, $R_{bh}\in SO(3)$ and $d_{bh}\in \bbR^3$ are functions of $(z,\delta)$.
-In the case of $g_{hk}$, the expressions are simply $d_{hk}=[0,0,0]^T$ and $R = \rotX(q_{hk})$, the latter being an elementary rotation about the common $x$-axis of the angle $q_{hk}$.
-
-\begin{algorithm}[h]
-	\floatname{algorithm}{Step}
-	\caption{Forward Propagation of Velocity}\label{alg:step1}
-	\begin{algorithmic}[1]
-		\vspace{1mm}
-		\For{$h=h_1,h_2,h_3,h_4$}
-		\vspace{1mm}
-		\State{$V_{bh}=J_{bh,z}\dot{z}+J_{bh,\delta}\dot{\delta}$}
-		\vspace{1mm}
-		\State{$V_h=\Ad_{g_{bh}}^{-1}V_b+V_{bh}$}\Comment{Knuckle 1 Rigid-Body Velocity}
-		\vspace{1mm}
-		\State{$V_{hk} = J_{hk}\dot{q}_{hk}$}
-		\vspace{1mm}
-		\State{$V_k=\Ad_{g_{hk}}^{-1}V_h+V_{hk}$}\Comment{Knuckle 2 Rigid-Body Velocity}
-		\vspace{1mm}
-		\State{$V_{kw}=J_{kw}\omega$}
-		\vspace{1mm}
-		\State{$V_w=V_k+V_{kw}$}\Comment{Rim Rigid-Body Velocity}
-		\vspace{1mm}
-		\EndFor
-		\vspace{1mm}
-	\end{algorithmic}
-\end{algorithm}
-
-The second step of the algorithm, as detailed in \textbf{Step 2}, performs the backward propagation of inertia and bias terms.
-\begin{algorithm}[h]
-	\floatname{algorithm}{Step}
-	\caption{Backward Propagation of Articulated Inertia and Bias}\label{alg:step2}
-	\begin{algorithmic}[1]
-		\vspace{1mm}
-		\For{$h=h_1,h_2,h_3,h_4$}
-		\vspace{1mm}
-		\State{$\hat{M}_w=M_w$}\Comment{Rim Articulated Inertia}
-		\vspace{1mm}
-		\State{$\hat{B}_w=-W_{we}+\ad_{V_w}^*M_wV_w$}\Comment{Rim Articulated Bias}
-		\vspace{1mm}
-		\State{$\bar{M}_w=\hat{M}_w-\dfrac{\hat{M}_wJ_{hw}J_{hw}^T\hat{M}_w}{J_{hw}^T\hat{M}_wJ_{hw}}$}
-		\vspace{1mm}
-		\State{$\bar{B}_w=\hat{B}_w-\dfrac{\hat{M}_wJ_{hw}J_{hw}^T\hat{B}_w}{J_{hw}^T\hat{M}_wJ_{hw}}-\bar{M}_w\ad_{V_{hw}}V_w$}
-		\vspace{1mm}
-		\State{$\hat{M}_k= \bar{M}_w$
-		}\Comment{Knuckle 2 Articulated Inertia}
-		\vspace{1mm}
-		\State{$\hat{B}_k= \bar{B}_w$
-		}\Comment{Knuckle 2 Articulated Bias}
-		\vspace{1mm}
-		\State{$\hat{M}_h= \Ad_{g_{hk}}^*\hat{M}_k\Ad_{g_{hk}}^{-1}$
-		}\Comment{Knuckle 1 Articulated Inertia}
-		\vspace{1mm}
-		\State{$\hat{B}_h= \Ad_{g_{hk}}^*\bigl(\hat{M}_k(J_{hk}\ddot{q}_{hk}-\ad_{V_{hk}}\Ad_{g_{kh}}V_k)+\hat{B}_k\bigr)$
-		}\Comment{Knuckle 1 Articulated Bias}
-		\vspace{1mm}
-		\State{$\bar{M}_h=\hat{M}_h-\dfrac{\hat{M}_hJ_{bh,z}J_{bh,z}^T\hat{M}_h}{J_{bh,z}^T\hat{M}_hJ_{bh,z}}$}
-		\vspace{1mm}
-		\State{$\bar{B}_h=\hat{B}_h-\dfrac{\hat{M}_hJ_{bh,z}(J_{bh,z}^T\hat{B}_h-\tau)}{J_{bh,z}^T\hat{M}_hJ_{bh,z}}-\bar{M}_h(\ad_{V_{bh}}V_h-\dot{J}_{bh,z}\dot{z}-\dot{J}_{bh,\delta}\dot{\delta}-J_{bh,\delta}\ddot{\delta})$}
-		\vspace{1mm}
-		\EndFor
-		\vspace{1mm}
-		\State{$\hat{M}_b=M_b+\displaystyle\sum_h\Ad_{g_{bh}}^*\bar{M}_h\Ad_{g_{bh}}^{-1}$}\Comment{Chassis Articulated Inertia}
-		\vspace{1mm}
-		\State{$\hat{B}_b=-W_{be}+\ad_{V_b}^*M_bV_b+\displaystyle\sum_h\Ad_{g_{bh}}^*\bar{B}_h$}\Comment{Chassis Articulated Bias}
-		\vspace{1mm}
-	\end{algorithmic}
-\end{algorithm}
-
-%Both inertia and bias are expressed in the body-fixed components.
-In the case the DCC architecture is considered, the camber angle acceleration $\ddot{q}_{hk}$ is known, since $q_{hk(t)}$ is the assumed input, while the torque $\tau_{hk}$ has to be computed together with the rest of unknown accelerations, which are the outcome of the third sweep.
-This leads to a \emph{Hybrid Dynamics} problem that requires to carefully revisit the ABA's second and third steps.
-The approach used is inspired by~\cite[Section~2.3]{Kim:notes:2012} where the the notation and conventions here employed were adapted according to~\cite{Domenighini:Designs:2023}.
-%Specifically, apart from the labels assigned to the quantities, the primary difference between the two methods is the ordering of the twist vector.
-In this work, the twist vector is defined as $V = [v,\omega]^T$, whereas in the other method, it is defined as $V=[\omega,v]^T$.
-This difference might seem minor but leads to significant changes in the definitions of Adjoint operators, inertia matrices, bias and wrench vectors.
-
-%For a generic articulated body the Newton-Euler equations can be expressed in the form:
-%\begin{equation}\label{eqn:NE}
-%	W=\hat{M}\dot{V}+\hat{B}.
-%\end{equation}
-%For a leaf node of the kinematic tree, having no children, the articulated inertia $\hat{M}$ is simply the generalized inertia $M$ of the body itself. Consistently, the bias $\hat{B}=B$, where:
-%\begin{equation}
-%	B=-W_e+\ad_V^*MV,
-%\end{equation}
-%and where $W_e$ is the resultant external wrench applied to the body and $\ad_V^*MV$ accounts for generalized gyroscopic and centrifugal forces, which are bi-linear in $V$.
-According to~\cite[Section~4.3.3]{Murray:book:1994} $\ad^*=-\ad^T$ has been defined.
-(If $v$ and $\omega$ are the translational and rotational component of $V$, then $\ad^*_V=\left[\begin{smallmatrix}\hat{\omega}&0\\\hat{v}&\hat{\omega}\end{smallmatrix}\right]$).
-
-Starting from the leaf nodes, which in our case are represented by the wheels ($\sref{W}$), after providing information such as external wrenches, twists, joint geometry, body masses, joint forces / torques for traditional connections, and time evolution of variables defining kinematically imposed joints, the backward propagation of inertia and bias can be performed.
-
-Finally the last step, denoted as \textbf{Step 3}, starting from the root node, computes joint variable accelerations and consequently body twists derivatives.
-For the kinematically controlled camber joints, the torques $\tau_{hk}$ needed to maintain the desired state of motion (DCC architecture) are computed as a by-product of the local inverse dynamics.
-This information is crucial to prevent actuators from  saturating or operating outside their bandwidth in the planned trajectories.
-
-The root twist acceleration is determined by solving equation $W=\hat{M}\dot{V}+\hat{B}$, considering that the chassis is free-floating.
-The other unknown quantities, such as variable accelerations and torques, are computed projecting the Newton-Euler equations of each body along the respective Jacobian and then solving for the variable.
-See also~\cite{Domenighini:Designs:2023} for more details in the derivation of the equations.
-
-\begin{algorithm}[h]
-	\floatname{algorithm}{Step}
-	\caption{Forward Propagation of Acceleration}\label{alg:step3}
-	\begin{algorithmic}[1]
-		\State{$\dot{V}_b=-\hat{M}_b^{-1}\hat{B}_b$}\Comment{Chassis Rigid-Body Acceleration}
-		\vspace{1mm}
-		\For{$h=h_1,h_2,h_3,h_4$}
-		\vspace{1mm}
-		\State{$\ddot{z}=-\dfrac{J_{bh,z}^T\bigl(\hat{B}_h+\hat{M}_h(\Ad_{g_{bh}}^{-1}\dot{V}_b-\ad_{V_{bh}}V_h+\dot{J}_{bh,z}\dot{z}+\dot{J}_{bh,\delta}\dot{\delta}+J_{bh,\delta}\ddot{\delta})\bigr)-\tau}{J_{bh,z}^T\hat{M}_hJ_{bh,z}}$}
-		\vspace{1mm}
-		\State{$\dot{V}_{bh}=\dot{J}_{bh,z}\dot{z}+\dot{J}_{bh,\delta}\dot{\delta}+J_{bh,z}\ddot{z}+J_{bh,\delta}\ddot{\delta}$}
-		\vspace{1mm}
-		\State{$\dot{V}_h=\Ad_{g_{bh}}^{-1}\dot{V}_b-\ad_{V_{bh}}V_h+\dot{V}_{bh}$}\Comment{Knuckle 1 Rigid-Body Acceleration}
-		\vspace{1mm}
-		\State{$\tau_{hk} = J_{hk}^T\bigl(\hat{B}_k+\hat{M}_k(\Ad_{g_{kh}}\dot{V}_h-\ad_{V_{hk}}V_k + J_{hk}\ddot{q}_{hk})\bigr)$}
-		\vspace{1mm}
-		\State{$\dot{V}_{bh} = J_{hk}\ddot{q}_{hk}$}
-		\vspace{1mm}
-		\State{$\dot{V}_k=\Ad_{g_{hk}}^{-1}\dot{V}_h-\ad_{V_{hk}}V_k+\dot{V}_{bh}$}\Comment{Knuckle 2 Rigid-Body Acceleration}
-		\State{$\dot{\omega}=-\dfrac{J_{kw}^T\bigl(\hat{B}_w+\hat{M}_w(\dot{V}_k-\ad_{V_{kw}}V_w)\bigr)}{J_{kw}^T\hat{M}_wJ_{kw}}$}
-		\vspace{1mm}
-		\EndFor
-		\vspace{1mm}
-	\end{algorithmic}
-\end{algorithm}
-
-\subsection{Dynamic Model}
-The dynamics equations, assembled via the procedure described  in Sec.~\ref{sec:aba}, are implemented in the software code by using a state-space representation $\dot{x}=f(x,u)$.
-%\begin{equation}\label{eqn:f}
-%	\dot{x}=f(x,u).
-%\end{equation}
-The state vector is $x = (q_{gb};V_{gb};z;\dot{z};\omega;q_{hk};\dot{q}_{hk})\in\bbR^{32}$, with components defined in previous sections.
-As already mentioned, in the model the camber joint is kinematically controlled as per the DCC architecture.
-Hence, the associated control vector is $u=(T_a;T_b;\delta;\ddot{q}_{hk})\in\bbR^{16}$, where each element is in $\bbR^{4}$ and collects the controls of the branch to each wheel.
-The torque is split into acceleration and braking components, allowing for minor modifications to model on-board motors coupled with traditional brakes and/or in-wheel motors as well.
-These components have their reaction torques in two different bodies in case of on-board motors and traditional brakes: the chassis and the knuckles, respectively.
-The $\delta$ vector collects the four translations of the steering rod attachment points, as described in Section~\ref{sec:suspanal}, while for the vector $\ddot{q}_{hk}$ no further explanation should be necessary.
-Clearly, this representation is valid for the model with active camber.
-The simpler model (no DCC) shares with the former the first 24 components of the state vector and the first 12 of the control vector.
-
-The time derivative of the state vector has also been expressed previously except for the first element, $q_{gb}$.
-Its derivative is a function of the state and can be recovered using a standard procedure~\cite{Domenighini:Designs:2023} as follows:
-\begin{equation}\label{eqn:dotqgb}
-	\dot{q}_{gb}=J_{gb}^{-1}(q_{gb})V_{gb},
-\end{equation}
-where $J_{gb}(q_{gb})\in\bbR^{6\times6}$ is the Jacobian of the virtual joint connecting the ground to the chassis, and it is computed via the PoE parameterization of the transformation $g_{gb}$.
-%, similarly to $J_{bh}$ in Eq.~\eqref{eqn:Jbh}.
+%In certain cases~\cite{Gillis:PracticalMethodsApproximate:2015}, positive-definiteness-preserving Lyapunov discretization schemes are used in~\eqref{eq:DOCPdP}.
+However, for sufficiently fine discretization, we found that integrating only the lower triangular part of $\bP(t)$ in~\eqref{eq:OCPdPIC}  (and reconstructing its strictly upper part accordingly), ensures that all $\bP_k$ remain symmetric and positive definite when the initial $\bP_0$ is symmetric and positive definite.
+In all our tests direct collocation with cubic polynomial state representations and Gauss-Legendre collocation points were used.

### 2025-04-17 closed_loop_planning.tex
[MGa] TV-LQR, first version

@@
 In fact, they need to explicitly account for the changed requirements associated with stabilizing a trajectory different from the nominal one, which must now satisfy the robustified constraints.
-\subsection{Suspension Analysis}
-\label{sec:suspanal}
-The suspensions connect the chassis to the unsprung masses.
-This connection cannot be described by an elementary joint, such as prismatic or revolute.
-Therefore, a set of six variables $q_{bh}$, that parameterize a homogeneous transformation matrix, is necessary.
-The transformation can be realized by composing three translations and three rotations (the ZXY Euler sequence was selected) using the global Product of Exponentials (PoE) formula~\cite{Murray:book:1994}, which reads: \begin{equation}\label{eqn:PoE}
-	g_{bh}=e^{\hat{Y}_1q_{bh,1}}\cdots e^{\hat{Y}_6q_{bh,6}} g_{bh}(0),
-\end{equation}
-where the offset $g_{bh(0)}=I$ and $Y_i=e_i$, with $e_i\in\bbR^6$ are the six normalized screw vectors (canonical basis elements in $\bbR^6$) associated with the Cartesian coordinates and the Euler ZXY angles in $q_{bh}$.
-Clearly, the six components of $q_{bh}$ are not independent; the modeled suspension is, in fact, a double wishbone that has: 2 DoFs for wheels that can steer, and 1 DoF otherwise.
-In this case, both axles have wheels that can steer; hence, a pair of independent variables ($z;\delta $) is associated with each wheel.
-A detailed analysis is omitted here for the sake of brevity. The interested reader is referred to~\cite{Domenighini:Designs:2023} for a similar treatment using the PoE formalism.
+\subsection{Step 1 -- Nominal feed-forward time-optimal trajectory planning}
+\label{sec:nominalFF}
 
-%As mentioned in the previous sections, $z$ represents the variation of the distance between the two attachment points of the spring-damper system, positive when the distance increases and negative when it decreases, while $\delta$ represents the motion of the steering rod attachment point to the chassis along its local $y$ axis (frame $\sref{B}$).
-Considering that the type of the suspension is known, six scalar equations of constraints can be expressed implicitly as $F(q_{bh};z,\delta)=0\in \bbR^6$.
-%
-%To find an explicit relation between inputs $(z,\delta)$ and outputs $q_{bh}$, the constraint equations are solved on a grid of pairs $(z^{(p)},\delta^{(l)})$, where $p = 1,\dots,N$ and $l = 1,\dots,M$.
-Then, the data are fitted (in the least-squares sense) with a 2D regression polynomial $q_{bh}(z,\delta)$ as follows:
-%\begin{equation}\label{eqn:poly}
-%	q_{bh,i}(z,\delta)=\sum_{r=0}^{3} \sum_{s=0}^{3-r} a_{{rs},i} z^r \delta^s.
-%\end{equation}
-%For the purpose a polynomial of degree three is sufficiently accurate.
-Now is possible to express the homogeneous transformation $g_{bh}\bigl(q_{bh}(z,\delta)\bigr)$ and to compute the twist $V_{bh}$ and $\dot{V}_{bh}$ as functions of $(z,\delta)$ and their derivatives.
-We start from:
-%\begin{equation}\label{eqn:Jbh}
-%	V_{bh}=\sum_{i=1}^6J_{bh,i}\dot{q}_{bh,i},
-%\end{equation}
-%where $J_{bh}$ is the \emph{geometric Jacobian} of the suspension joint relative to $q_{bh}$.
-Then, following the procedure thoroughly described in~\cite[Sec.
-2.2]{Domenighini:Designs:2023}, it is possible to obtain %
-%assumed with independent components.
-Each column of $J_{bh}$ can be computed based on the PoE formula \eqref{eqn:PoE}.
-For the $i$-th column, we have $J_{bh,i}=A_i^{-1}Y_i$, with $A_6 = I_{6\times 6}$ and $A_i=\Ad_{e^{\hat{Y}_{i}q_{bh,i}}\cdots e^{\hat{Y}_6q_{bh,6}}}$, $i = 1, \dots, 5$.
-The \emph{Adjoint operator} $\Ad_g$ has been defined as the $6\times6$ matrix satisfying $Y=\Ad_gX$ whenever $\hat{Y}=g\hat{X}g^{-1}$, so that, if $g=\left[\begin{smallmatrix}R&d\\0^T&1\end{smallmatrix}\right]$, then $\Ad_g=\left[\begin{smallmatrix}R&\hat{d}R\\0&R\end{smallmatrix}\right]$.\\
-%As a function of $(z,\delta)$, $\dot{q}_{bh,i}$ can be expressed as follows:
-%\begin{equation}\label{eqn:dt_qbh}
-%	\dot{q}_{bh,i}=\frac{\partial}{\partial z}q_{bh,i}\dot{z}+\frac{\partial}{\partial\delta}q_{bh,i}\dot{\delta},
-%\end{equation}
-%where the (geometric) partial derivatives $\frac{\partial}{\partial z}q_{bh,i}$ and $\frac{\partial}{\partial\delta}q_{bh,i}$ are obtained by differentiating Eq.~\eqref{eqn:poly}.
-%Then, by defining the Jacobians with respect to the true independent variables $z$ and $\delta$ as follows
-%\begin{align}\label{eqn:Jbh,z}
-%	J_{bh,z}=\sum_{i=1}^6J_{bh,i}\frac{\partial}{\partial z}q_{bh,i}\quad \textrm{and}\quad J_{bh,\delta}=\sum_{i=1}^6J_{bh,i}\frac{\partial}{\partial\delta}q_{bh,i},
-%\end{align}
-%it possible to obtain the rigid-body velocity $V_{bh}$ of the constrained motion from Eq.~\eqref{eqn:Jbh}, in terms of $\dot{z}$ and $\dot{\delta}$, as follows
-%\begin{equation}\label{eqn:Vbh}
-%	V_{bh}=J_{bh,z}\dot{z}+J_{bh,\delta}\dot{\delta}.
-%\end{equation}
-%
-%To write the dynamic equations the time derivative $\dot{V}_{bh}$ of the twist $V_{bh}$ is necessary. To this sake, it is sufficient to differentiate Eqs.~\eqref{eqn:Jbh}--\eqref{eqn:Vbh}.
-%\begin{equation}\label{eqn:dt_Jbh}
-%	\dot{V}_{bh}=\sum_{i=1}^6\dot{J}_{bh,i}\dot{q}_{bh,i}+\sum_{i=1}^6J_{bh,i}\ddot{q}_{bh,i}.
-%\end{equation}
-%
-%Here, following~\cite{Gabiccini:ICRA:2012}, the derivatives $\dot{J}_{bh,i}$ can computed using the formula
-%\begin{align}\label{eqn:dt_Jbhi}
-%	\dot{J}_{bh,i}=-\sum\limits_{i < j \leq 6}\ad_{J_{bh,j}\dot{q}_{bh,j}} J_{bh,i}
-%\end{align}
-%using Lie derivatives.
-In the matrix case, the Lie derivative $\hat{Z}$ of the vector field $\hat{X}$ with respect to vector field $\hat{Y}$ is calculated by as a commutator operation $\hat{Z}=\hat{Y}\hat{X}-\hat{X}\hat{Y}$; in the vector case, the $6\times6$ matrix $\ad_Y$ is used, defined such that $Z=\ad_Y X$.
-If $v$ and $\omega$ are the translational and rotational component of $V$, then $\ad_V=\left[\begin{smallmatrix}\hat{\omega}&\hat{v}\\0&\hat{\omega}\end{smallmatrix}\right]$.
-%
-%In Eq.~\eqref{eqn:dt_Jbh}, the accelerations $\ddot{q}_{bh,i}$ are obtained by differentiating the velocities in Eq.~\eqref{eqn:dt_qbh}:
-%\begin{equation}\label{eqn:dt_dt_qbh}
-%	\ddot{q}_{bh,i}=\frac{\partial^2}{\partial z^2}q_{bh,i}\dot{z}^2+2\frac{\partial^2}{\partial z\partial\delta}q_{bh,i}\dot{z}\dot{\delta}+\frac{\partial^2}{\partial\delta^2}q_{bh,i}\dot{\delta}^2+\frac{\partial}{\partial z}q_{bh,i}\ddot{z}+\frac{\partial}{\partial\delta}q_{bh,i}\ddot{\delta},
-%\end{equation}
-%where the partial derivatives are computed by differentiating the regression polynomials~in~\eqref{eqn:poly}.
-%After simple but tedious calculations, the explicit expression for the derivatives of the Jacobians in \eqref{eqn:Jbh,z} appear as follows:
-%\begin{align}
-%	\dot{J}_{bh,z}&=\sum_{i=1}^6\dot{J}_{bh,i}\frac{\partial}{\partial z}q_{bh,i}+\sum_{i=1}^6J_{bh,i}\Bigl(\frac{\partial^2}{\partial z^2}q_{bh,i}\dot{z}+\frac{\partial^2}{\partial z\partial\delta}q_{bh,i}\dot{\delta}\Bigr)\label{eqn:dt_Jbh,z}\quad\textrm{and}\\\dot{J}_{bh,\delta}&=\sum_{i=1}^6\dot{J}_{bh,i}\frac{\partial}{\partial\delta}q_{bh,i}+\sum_{i=1}^6J_{bh,i}\Bigl(\frac{\partial^2}{\partial z\partial\delta}q_{bh,i}\dot{\delta}+\frac{\partial^2}{\partial\delta^2}q_{bh,i}\dot{\delta}\Bigr)\label{eqn:dt_Jbh,d}.
-%\end{align}
-%Substituting Eqs.~\eqref{eqn:dt_Jbhi} and~\eqref{eqn:dt_dt_qbh} in Eqs.~\eqref{eqn:dt_Jbh}, and casting some intermediate results as~\eqref{eqn:dt_Jbh,z} and~\eqref{eqn:dt_Jbh,d},
-%the constrained rigid-body acceleration $\dot{V}_{bh}$ in the following form:
-%\begin{equation}\label{eqn:dt_Vbh}
-%	\dot{V}_{bh}=\dot{J}_{bh,z}\dot{z}+\dot{J}_{bh,\delta}\dot{\delta}+J_{bh,z}\ddot{z}+J_{bh,\delta}\ddot{\delta}.
-%\end{equation}
-%The explicit expressions for auxiliary Jacobians $J_{bh,z}$, $J_{bh,\delta}$, $\dot{J}_{bh,z}$ and $\dot{J}_{bh,\delta}$ are omitted here for brevity.
-The interested reader can find the detailed analytical derivation in~\cite[Eqs.
-(5)--(8)]{Domenighini:Designs:2023}.
-%
-%A generalized force $\tau\in\bbR$ is introduced that depends only on the independent variable $z\in\bbR$.
-$\tau$ is assumed to be generated only by the spring and the damper, considered aligned along the same axis.
-It is worth noting that, using the defined $z$, we are allowed to compute the force of the suspension without calculate the associated $q_{bh}$ nor using the Principle of Virtual Works.
-In this expression, also the contribution of the anti-roll bar is considered, which depends on the difference between the left and right trim, and the contribution of the bump-stops, which activate when the trim reaches its lower or upper limits.
-The intensity of $\tau$, for each wheel, is expressed as follow:
-%\begin{equation}
-%	\tau_{ij}=k_{s,i}(l_{c,i}-l_{0,i}+z_{ij})+c_{d,i}\dot{z}_{ij}+(-1)^j k_{a,i}(z_{i,1}-z_{i,2}) + \left(\frac{2z-z_{\text{max}}-z_{\text{min}}}{|z_{\text{max}}-z_{\text{min}}|}\right)^{b},
-%\end{equation}
-%where $l_{c,i}$ is the spring's length in the reference position, $l_{0,i}$ is the spring's length at rest, $k_{s,i}$ and $c_{d,i}$ are the stiffness and the damping coefficients, $k_{a,i}$ is the anti-roll bar stiffness, $b$ influences the stiffness of the bump-stops, and $z_{\text{min}}$ and $z_{\text{max}}$ are the knees of the bump-stops' force characteristics.
+\subsection{Step 2 -- Time-varying LQR stabilizing controller}
+\label{sec:LQR}
+
+Given the nominal trajectory $(\hbmu_k, \hbu_k)$ over $k=0,\ldots, N$, and introduced the deviation variables $\bbmu_k=\bmu_k - \hbmu_k$ and $\bbu = \bu_k - \hbu_k$, we design a linear state-feedback law $\bbu_k= -\hbK_k \bbmu_k$, or equivalently $\bu_k = \hbu_k - \hbK_k (\bmu_k - \hbmu_k)$, to stabilize the deviations along the nominal trajectory.
+To compute the state-feedback matrices $\hbK_k$, we linearize the system around $(\hbmu_k, \hbu_k)$, such that $\bbmu_{k+1} \approx \hbA_{k}\bbmu_k + \hbB_k \bbu_k$, where $\hbA_k = f_{,\bmu}(\hbmu_k,\hbu_k)$ and $\hbB_k = f_{,\bu}(\hbmu_k,\hbu_k)$, and define the quadratic regulator (tracking) cost function
+\begin{align}
+J = \sum_{k=0}^{N-1} \big( \bbmu_k^{T} \bW_k \bbmu_k + \bbu_k^T \bR_k \bbu_k \big) + \bbmu_N^T \bW_N \bbmu_N,
+\end{align}
+where $\bW_N=\bW_N^T \succ 0$, $\bW_k=\bW_k^T \succeq 0$, $\bR_k=\bR_k^T \succ 0$.
+It can be shown that the optimal cost-to-go is given by $J_k^*(\bbmu_k) = \bbmu_k^T \bS_k \bbmu_k$, with $\bS_k = \bS_k^T \succ 0$, where $\bS_k$ can be computed through the (backward) Riccati recursion for $k=N-1,\ldots, 0$
+\begin{align}
+\bS_N &= \bW_N,\\ 
+\bS_{k} &= \bW_k + \hbA_k^T \bS_{k+1} \hbA_k - \hbA_k^T \bS_{k+1}\hbB_k \hbK_k,\quad k={N-1, \ldots, 0}
+\end{align}
+where  $\hbK_k =\big( \bR_k+\hbB_k^T \bS_{k+1} \hbB_k \big)^{-1} \hbB_k^T \bS_{k+1} \hbA_k$.
+The optimal (stabilizing) feedback policy is given by $\bbu_k=-\hbK_k \bbmu_k$.
+We assume that the gain sequence $\{\hbK_k\}$, $k=0,\ldots, N$ represents the control strategy that an expert driver would employ to optimally stabilize the system along the nominal trajectory.
+
+\subsection{Step 3 -- Closed-loop robustified planning}
+\label{sec:CLrobustified}
+
 \section{Tire Model}

### 2025-04-17 stochastic_vehicle_dynamics.tex
[MM] DOCP re-formulation

@@
 
-%\subsection{Discretization via direct collocation}
-%\label{sec:discretization}
-%
-%The nonlinear optimal control problem~\eqref{eq:OCPcost} can be discretized by applying a suitable collocation integrator obtaining the following nonlinear program (NLP) 
-%\begin{alignat}{3}
-%\underset{\bmu_k,\bxi_k, \bu_k, \bP_k,\bz_k}{\text{minimize}} \,
-%& & & J_k(\bmu_k,\bxi_k, \bu_k) & & \label{eq:DOCPcost} \\
-%\hspace*{-2.0 cm}\text{s.t.} \quad
-%& \bzero      & = & \; \bPsimu_k(\bmu_{k-1},\bmu_k,\bxi_k, \bu_k,\bz_k),
-%& \quad & k = 1,\ldots, N \label{eq:DOCPdyn} \\
-%& \bmu_0      & = & \; \bar{\bmu}_0
-%& & \label{eq:DOCPdynIC} \\
-%& \bzero      & = & \; \bPsiP_k(\bmu_k,\bxi_k, \bu_k, \bP_{k-1},\bP_k,\bSi_k,\bz_k),
-%& \quad & k = 1,\ldots, N \label{eq:DOCPdP} \\
-%& \bP_0       & = & \; \bar{\bP}_0 \succeq 0
-%& & \label{eq:DOCPdPIC} \\
-%& \bzero      & = & \; \bOm_k(\bmu_k,\bxi_k, \bu_k,\bz_k),
-%& \quad & k = 0,\ldots, N \label{eq:DOCPpath} \\
-%& 0           & \geq & \; h_i(\bmu_k, \bu_k, \bz_k) + \be_i(\bmu_k, \bu_k, \bP_k, \bz_k),
-%& \quad & k = 1,\ldots, N;\; i \in \calI \label{eq:DOCPconstraints}
-%\end{alignat}
-%To perform the discretization, the track centerline is parameterized using a curvilinear abscissa $\alpha \in [0,1]$, and uniformly sampled at $N+1$ points $\alpha_0, \ldots, \alpha_N$.
-%Accordingly, $\bmu_k$ denotes the mean state at grid node $\alpha_k$, while $\bu_k$ and $\bz_k$ are assumed to be piecewise constant over each interval $[\al_k, \al_{k+1}]$.
-%These represent the controls and the algebraic variables, respectively.
+\subsection{Discretization via direct collocation}
+\label{sec:discretization}
+
+The nonlinear optimal control problem~\eqref{eq:OCPcost} can be discretized by applying a suitable collocation integrator obtaining the following nonlinear program (NLP) 
+\begin{alignat}{3}
+\underset{\bmu_k,\bxi_k, \bu_k, \bP_k,\bz_k}{\text{minimize}} \,
+& & & J_k(\bmu_k,\bxi_k, \bu_k) & & \label{eq:DOCPcost} \\
+\hspace*{-2.0 cm}\text{s.t.} \quad
+& \bzero      & = & \; \bPsimu_k(\bmu_{k-1},\bmu_k,\bxi_k, \bu_k,\bz_k),
+& \quad & k = 1,\ldots, N \label{eq:DOCPdyn} \\
+& \bmu_0      & = & \; \bar{\bmu}_0
+& & \label{eq:DOCPdynIC} \\
+& \bzero      & = & \; \bPsiP_k(\bmu_k,\bxi_k, \bu_k, \bP_{k-1},\bP_k,\bSi_k,\bz_k),
+& \quad & k = 1,\ldots, N \label{eq:DOCPdP} \\
+& \bP_0       & = & \; \bar{\bP}_0 \succeq 0
+& & \label{eq:DOCPdPIC} \\
+& \bzero      & = & \; \bOm_k(\bmu_k,\bxi_k, \bu_k,\bz_k),
+& \quad & k = 0,\ldots, N \label{eq:DOCPpath} \\
+& 0           & \geq & \; h_i(\bmu_k, \bu_k, \bz_k) + \be_i(\bmu_k, \bu_k, \bP_k, \bz_k),
+& \quad & k = 1,\ldots, N;\; i \in \calI \label{eq:DOCPconstraints}
+\end{alignat}
+To perform the discretization, the track centerline is parameterized using a curvilinear abscissa $\alpha \in [0,1]$, and uniformly sampled at $N+1$ points $\alpha_0, \ldots, \alpha_N$.
+Accordingly, $\bmu_k$ denotes the mean state at grid node $\alpha_k$, while $\bu_k$ and $\bz_k$ are assumed to be piecewise constant over each interval $[\al_k, \al_{k+1}]$.
+These represent the controls and the algebraic variables, respectively.
 The $\bz_k$'s are introduced as direct handles for physically meaningful quantities such contact forces.
 Similarly, $\bP_k$ represents the covariance matrix at the $k$-th node.
-%Following the direct collocation approach, both the mean and covariance state trajectories are approximated, in the $k$-th interval, with polynomials $\pi_k(\tau)$, defined on the unit interval $\tau\in[0,1]$, and then scaled to match the width $h_k$ of the corresponding time step.
+Following the direct collocation approach, both the mean and covariance state trajectories are approximated, in the $k$-th interval, with polynomials $\pi_k(\tau)$, defined on the unit interval $\tau\in[0,1]$, and then scaled to match the width $h_k$ of the corresponding time step.
 On the unit interval, we select $d$ collocation points $\tau_1, \ldots, \tau_d$, associated with as many collocation states.
@@
 Therefore, $\bxi_k =
-%(\bxi_{k,1}, \ldots, \bxi_{k,d})$, and similarly $\bSi_k=(\bSi_{k,1}, \ldots, \bSi_{k,d})$.
+(\bxi_{k,1}, \ldots, \bxi_{k,d})$, and similarly $\bSi_k=(\bSi_{k,1}, \ldots, \bSi_{k,d})$.
 Equations~\eqref{eq:DOCPdyn} and~\eqref{eq:DOCPdP} represent the collocation and continuity equations for mean and covariance, respectively, with initial conditions represented by~\eqref{eq:DOCPdynIC} and~\eqref{eq:DOCPdPIC}.
 Eqs.~\eqref{eq:DOCPpath} are path equality constraints and~\eqref{eq:DOCPconstraints} are the robustified inequality constraints.
-%
-%In certain cases~\cite{Gillis:PracticalMethodsApproximate:2015}, positive-definiteness-preserving Lyapunov discretization schemes are used in~\eqref{eq:DOCPdP}.
+
+In certain cases~\cite{Gillis:PracticalMethodsApproximate:2015}, positive-definiteness-preserving Lyapunov discretization schemes are used in~\eqref{eq:DOCPdP}.
 However, for sufficiently fine discretization, we found that integrating only the lower triangular part of $\bP(t)$ in~\eqref{eq:OCPdPIC}  (and reconstructing its strictly upper part accordingly), ensures that all $\bP_k$ remain symmetric and positive definite when the initial $\bP_0$ is symmetric and positive definite.

### 2025-04-17 closed_loop_planning.tex
[MGa] draft for closed loop

@@
 \begin{align}
-J = \sum_{k=0}^{N-1} \big( \bbmu_k^{T} \bW_k \bbmu_k + \bbu_k^T \bR_k \bbu_k \big) + \bbmu_N^T \bW_N \bbmu_N,
+J = \bbmu_N^T \bW_N \bbmu_N + \sum_{k=0}^{N-1} \big( \bbmu_k^{T} \bW_k \bbmu_k + \bbu_k^T \bR_k \bbu_k \big),
 \end{align}
@@
 
+%\begin{align}
+%    \underset{\bmu_k,\bxi_k, \bu_k, \bP_k,\bz_k, \bK_k}{\text{minimize}} \quad & J_k(\bmu_k,\bxi_k, \bu_k) \label{eq:ROCPcost}\\
+%       \text{s.t.}\notag\\
+%          & f_0(x) - t \le 0 \quad \phantom{i=1, \dots, m}\\
+%          & f_i(x) \le 0, \quad i=1, \dots, m \quad i=1, \dots, m\\
+%          & h_i(x) = 0,   \quad i=1, \dots, p,
+%\end{align}
+
+\begin{alignat}{3}
+\underset{\bmu_k,\bxi_k, \bu_k, \bP_k,\bz_k, \bK_k}{\text{minimize}}
+    & \quad J_k(\bmu_k,\bxi_k, \bu_k) & & \label{eq:ROCPcost} \\
+\text{s.t.}\notag\\
+\bzero      & = \; \bPsimu_k(\bmu_{k-1},\bmu_k,\bxi_k, \bu_k,\bz_k),
+\quad k = 1,\ldots, N \label{eq:ROCPdyn} \\
+\bmu_0      & = \; \bar{\bmu}_0 \label{eq:ROCPdynIC} \\
+\bzero      & = \; \bPsiPCL_k(\bmu_k,\bxi_k, \bu_k, \tbP_{k-1},\tbP_k,\bSi_k,\bz_k, \bK_k),
+\quad k = 1,\ldots, N \label{eq:ROCPdP} \\
+\tbP_0       & = \; \bar{\bP}_0 \succeq 0 \label{eq:ROCPdPIC} \\
+\bzero      & = \; \bOm_k(\bmu_k,\bxi_k, \bu_k,\bz_k), \quad k = 0,\ldots, N \label{eq:ROCPpath} \\
+0    & \geq  h_i(\bmu_k, \bu_k, \bz_k) + \be_i(\bmu_k, \bu_k, \tbP_k, \bz_k), \quad k = 1,\ldots, N;\; \quad i \in \calI \label{eq:ROCPconstraints}\\
+-\de \bK   & \leq \bK_k - \hbK_k \leq  \de \bK \label{eq:ROCPdeltaK}, \quad k = 0,\ldots, N
+\end{alignat} 
+where~\eqref{eq:ROCPdP} denotes the discretized form for time evolution of the covariance matrix under a stabilizing controller of the form $\bbu_k=-\bK_k \bbmu_k$.
+As apposed to~\eqref{eq:OCPdP}, $\tbP$ would evolve, in continuous time, according to $\dot{\tbP}(t) = (\bA(t)-\bB(t) \bK(t)) \tbP(t) + \tbP(t) (\bA(t)-\bB(t) \bK(t))^T + \bQ(t)$.
+To be continued...
 \section{Tire Model}

### 2025-04-17 open_loop_planning.tex

@@
-\section{Open-loop planning via n-steps ahead predictions}
+\section{Open-loop planning via H-steps ahead predictions}
 \label{sec:open_loop_planning}
+The proposed approach requires the presence of $H+1$ versions of the matrix $\bP$ at each grid point of the discretized problem.
+One of these versions is initialized with $\bar{\bP}_0$, while the others represent the evolution of matrices that were initialized at previous grid points, from 1 to $H$ steps earlier.
+This ensures that at each step it is possible to find a covariance matrix evolved for $H$ steps and use it to robustify the constraints.
+The initialization of the covariance matrix is necessary due to its divergent dynamics.
+In fact, without a feedback control, the error on the position and orientation of the body reference frame can only increase along the track, compromising the conditioning of the matrix.
 
-The model described in Sec.~\ref{chap:vehicle_model} is embedded into a minimum lap time problem which is briefly outlined below.
+%\subsection{Discretization via direct collocation}
+%\label{sec:openloopdiscretization}
+Introducing the versioning of the covariance matrix approach requires a modification of the Equations described in Section~\ref{sec:discretization}.
+In particular Eqs.~\eqref{eq:DOCPdP}, \eqref{eq:DOCPdPIC}, and \eqref{eq:DOCPconstraints} are replaced by Eqs.~\eqref{eq:DOCPdPopenloop}, \eqref{eq:DOCPdPinitopenloop}, and \eqref{eq:DOCPconstraintsopenloop}.
+Eq.~\eqref{eq:DOCPdPopenloop} collects the continuity and collocation Equations for all the versions of the covariance matrix that have propagated from the previous step.
+To account for the evolution of $H$ matrices, it is necessary to introduce $H\times d$ matrices $\bSi_k^j=(\bSi_{k,1}^j, \ldots, \bSi_{k,d}^j)$, which represent the values of each matrix version at the collocation points.
 
-\subsection{Direct Collocation}
-The objective is to recast the MLTP into a Non Linear Program (NLP) using a direct orthogonal collocation approach, specifically based on Gauss-Legendre collocation points.
-The problem is formulated within the CasADi framework~\cite{Andersson:MPC:2019} and solved using the interior point method implemented in IPOPT~\cite{Wachter:MP:2006}.
-The resulting NLP can be expressed in the form:
-\begin{align}
-	\underset{x,v,u,z}{\textrm{minimize}}&\quad\sum_{k=1}^NL_k(x_k,v_k,u_k,z_k)\label{eqn:functional}\\
-	\textrm{subject to}&\quad g_k(x_{k-1},x_k,v_k,u_k,z_k)=0,\ k=1,\ldots,N,\label{eqn:equality}\\
-	&\quad h_k(x_k,u_k,z_k)\le0,\ k=1,\ldots,N.\label{eqn:inequality}
-\end{align}
-In Eqs.~\eqref{eqn:functional}, \eqref{eqn:equality} and \eqref{eqn:inequality}, in addition to the state and control vectors, $v_k$ and $z_k$ are included, representing the state vector at the collocation points and the algebraic quantities, respectively.
-The controls and the algebraic variables are approximated as piecewise constant, while the states are approximated as quadratic polynomials within each finite element.
-Specifically, the algebraic variables in the problem are the time steps $\nu_k$ and the lateral and vertical displacements of the Center of Mass relative to the origin of the $k$-th frame $\sref{S}$, attached to the track's centerline ($e_y$ and $e_z$).
-Their function is to serve as handle variables in the optimization process to promote problem sparsity.
+The specific formulation is expressed as follows:
+\begin{alignat}{3}
+	\underset{\bmu_k,\bxi_k, \bu_k, \bP_k,\bz_k}{\text{minimize}} \,
+	& & & J_k(\bmu_k,\bxi_k, \bu_k) & & \label{eq:DOCPcost} \\
+	\hspace*{-2.0 cm}\text{s.t.} \quad
+	& \bzero      & = & \; \bPsimu_k(\bmu_{k-1},\bmu_k,\bxi_k, \bu_k,\bz_k),
+	& \quad & k = 1,\ldots, N \label{eq:DOCPdyn} \\
+	& \bmu_0      & = & \; \bar{\bmu}_0
+	& & \label{eq:DOCPdynIC} \\
+	& \bzero      & = & \; \bPsiP_k(\bmu_k,\bxi_k, \bu_k, \bP_{k-1}^{j-1},\bP_k^j,\bSi_k^j,\bz_k),
+	& \quad & \substack{k = 1,\ldots, N; \; \\ j=1,\ldots,H;\;} \label{eq:DOCPdPopenloop} \\
+	& \bP_k^0       & = & \; \bar{\bP}_0 \succeq 0
+	& \quad & k = 1,\ldots, N; \; \label{eq:DOCPdPinitopenloop} \\
+	& \bzero      & = & \; \bOm_k(\bmu_k,\bxi_k, \bu_k,\bz_k),
+	& \quad & k = 0,\ldots, N \label{eq:DOCPpath} \\
+	& 0           & \geq & \; h_i(\bmu_k, \bu_k, \bz_k) + \be_i(\bmu_k, \bu_k, \bP_k^H, \bz_k),
+	& \quad & k = 1,\ldots, N;\; i \in \calI \label{eq:DOCPconstraintsopenloop}
+\end{alignat}
 
-The track centerline is represented by a curvilinear abscissa $\alpha\in[0,1]$, sampled at $N+1$ points $\alpha_0,\ldots,\alpha_N$.
-Each grid point is associated with a reference frame $\sref{S_k}$, whose origin lies at the corresponding point on the track centerline, as illustrated in Figure~\ref{fig:track_frames}.
-A more detailed description of the 3D track modeling can be found in~\cite{Bartali:Meccanica:2023}.
-Specifically, this representation allows us to capture track slope and banking, but it does not support non-developable track surfaces.
+Eq.~\eqref{eq:DOCPdPinitopenloop} is used to initialize the correct version at the $k$-th step while Eq.~\eqref{eq:DOCPconstraintsopenloop} specify that the back-off term of the constraints is evaluated with the most propagated covariance matrix version, $\bP^H_k$.
+A schematic representation of the approach used to manage the $H+1$ versions of the covariance matrix is shown in Figure \ref{fig:DOCPgrid}, where the dashed rectangle represent the $k$-th step at which the constraints are being formulated.
+At each grid point, two versions of the covariance matrix are highlighted: $\bP^0_k$ (red node), which is the version to be initialized, and $\bP^H_k$ (green node), which is the most propagated version used to formulate the robust constraints in Eq.~\eqref{eq:DOCPconstraintsopenloop}.
 
-\begin{figure}[h]\centering
-	\includegraphics[width=0.8\textwidth]{Images/track_frames.pdf}
-	\caption{Track frames corresponding to adjacent nodes of the grid over $\alpha$.
-Note that the grid points need not be equispaced.
-If desired, they can be chosen irregularly, so as to capture more features of the track where it is more critical (e.g.
-in proximity to corners) and save space where it is more regular (e.g.
-on the straights).}
-	\label{fig:track_frames}
+\begin{figure}
+	\centering
+	\begin{tikzpicture}[%
+		smallnode/.style={%
+			circle, draw, minimum size=3mm, inner sep=0pt
+		},
+		r_smallnode/.style={%
+			circle, draw, minimum size=3mm, inner sep=0pt, red
+		},
+		g_smallnode/.style={%
+			circle, draw, minimum size=3mm, inner sep=0pt, mygreen
+		},
+		every path/.style={->, thick},
+		x=1.5cm, y=1.5cm
+		]
+
+		\newcommand{\Plab}[2]{\bP^{#2}_{#1}}
+		\definecolor{mygreen}{RGB}{0 150 0}
+
+		% Nodes
+		% First row
+		\node[smallnode, label={[label distance=-2pt]above:\protect\(\Plab{k-2}{H-2}\)}] (Pk-2H-2) at (1,0) {};
+		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{H-1}\)}] (Pk-1H-1) at (2,0) {};
+		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k}{H}\)}] (PkH) at (3,0) {};
+		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{0}\)}] (Pk+10) at (4,0) {};
+		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{1}\)}] (Pk+21) at (5,0) {};
+
+		% Second row
+		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-2}{H-3}\)}] (Pk-2H-3) at (1,1) {};
+		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{H-2}\)}] (Pk-1H-2) at (2,1) {};
+		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k}{H-1}\)}] (PkH-1) at (3,1) {};
+		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{H}\)}] (Pk+1H) at (4,1) {};
+		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{0}\)}] (Pk+20) at (5,1) {};
+
+		% dots row 
+		\node[label={center,rotate=90:\(\dots\)}] (dots2) at (1, 1.7) {};
+		\node[label={center,rotate=90:\(\dots\)}] (dots3) at (2, 1.7) {};
+		\node[label={center,rotate=90:\(\dots\)}] (dots4) at (3, 1.7) {};
+		\node[label={center,rotate=90:\(\dots\)}] (dots5) at (4, 1.7) {};
+		\node[label={center,rotate=90:\(\dots\)}] (dots6) at (5, 1.7) {};
+
+		% Third row 
+		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k-2}{H}\)}] (Pk-2H) at (1,2) {};
+		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{0}\)}] (Pk-10) at (2,2) {};
+		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k}{1}\)}] (Pk1) at (3,2) {};
+		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{2}\)}] (Pk+12) at (4,2) {};
+		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{3}\)}] (Pk+23) at (5,2) {};
+
+		% Fourth row
+		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-2}{H-1}\)}] (Pk-2H-1) at (1,3) {};
+		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{H}\)}] (Pk-1H) at (2,3) {};
+		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k}{0}\)}] (Pk0) at (3,3) {};
+		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{1}\)}] (Pk+11) at (4,3) {};
+		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{2}\)}] (Pk+22) at (5,3) {};
+
+		% Curved arrows
+		\draw[bend left=20] (Pk-2H-2) to (Pk-1H-1);
+		\draw[bend left=20] (Pk-1H-1) to (PkH);
+		\draw[bend left=20] (Pk+10) to (Pk+21);
+
+		\draw[bend left=20] (Pk-2H-3) to (Pk-1H-2);
+		\draw[bend left=20] (Pk-1H-2) to (PkH-1);
+		\draw[bend left=20] (PkH-1) to (Pk+1H);
+
+		\draw[bend left=20] (Pk-10) to (Pk1);
+		\draw[bend left=20] (Pk1) to (Pk+12);
+		\draw[bend left=20] (Pk+12) to (Pk+23);
+
+		\draw[bend left=20] (Pk-10) to (Pk1);
+		\draw[bend left=20] (Pk1) to (Pk+12);
+		\draw[bend left=20] (Pk+12) to (Pk+23);
+
+		\draw[bend left=20] (Pk-2H-1) to (Pk-1H);
+		\draw[bend left=20] (Pk0) to (Pk+11);
+		\draw[bend left=20] (Pk+11) to (Pk+22);
+
+		% Dashed rectangle
+		\draw[dashed, thick, rounded corners] (2.6,-.25) rectangle (3.4,3.5);
+
+	\end{tikzpicture}
+	\caption{Schematic representation of the continuity and initialization Equations for the $H+1$ versions of the covariance matrix introduced in the discretized OCP.
+The dashed rectangle indicate the $k$-th step, in which the constraints are formulated.
+Each node represents a covariance matrix, labeled such that the subscript denotes the current grid step, while the superscript indicates the number of propagation steps it has undergone.
+At each step, two nodes are highlighted: the covariance matrix to be initialized (red node) and the matrix that has been propagated for $H$ steps (green node), which is used to formulate the robust constraints.}
+	\label{fig:DOCPgrid}
 \end{figure}
 
-%According the direct collocation method, the state evolution on the $k$-th element of the grid is approximated with a polynomial $p_k$.
-The number of collocation points depends on the degree $d$ of the polynomial chosen and their position inside the interval can be selected based on different criteria; here the Gauss-Legendre collocation points are used.
-The polynomial is built on the unit interval and then scaled to match the length $\nu_k$ of the relative time step.
-%
-%In Eq.~\eqref{eqn:equality}, besides specific constraints inserted to bias the solution towards desired behaviours, two equations are paramount. The first  one is:
-%\begin{equation}\label{eqn:collocation}
-%	\frac{1}{\nu_k}\frac{\partial}{\partial\tau}p_k(\tau_i)=f(v_{k,i},u_k),
-%\end{equation}
-%the collocation equation, that constrains the state at each collocation point to comply with the system dynamics. The second one, which is $p(1)=x_{k+1}$, ensures continuity across time intervals.
-
-In addition to the classical collocation and continuity constraints~\cite{Domenighini:Designs:2023}, the inequalities in Eq.~\eqref{eqn:inequality} for our MLTP ensure the vehicle remains within track bounds.
-These include the requirement for the CoM to lie on the $yz$ plane of the frame $\sref{S_k}$, as well as power and adherence limits.
-
-In our optimizations, the cost function at each step is expressed as follows:
-\begin{equation}\label{eqn:cost}
-	L_k=\nu_k^2+k_\delta(\delta_k-\delta_{k-1})^2+k_T(T_k-T_{k-1})^2+k_v x_{k,2}^2 + l_k(\ddot{q}_{hk,k},\delta_k,x_k),
-\end{equation}
-where the first term penalizes travel time, the second and third prevent abrupt variations of the inputs and avoid singular arcs.
-The quadratic penalization ensures a well-defined dependence of the Hamiltonian on the control variables, preventing indeterminate solutions.
-The last two serve to shape the driving behaviors according to the designer preferences.
-
-Their primary goal is to improve the optimization landscape, ensuring that the solver is less likely trapped by undesirable local minima.
-As an example, the fourth term, which includes the lateral velocity $x_{k,2}$, has been included to prevent undesirable behaviors which could emerge as a result of the increased system's DoFs in combination with the simplified aerodynamic modeling adopted in this work.
-Some local minima could automatically be avoided by incorporating more realistic aerodynamic maps and tyre wear models.
-These enhancements will be explored in future works.
-In this context, an interesting direction is presented in~\cite{Masouleh:IEEE:2016}, where the authors introduce an aerodynamic model that explicitly accounts for the influence of suspension and vehicle configuration.
-This approach offers a more comprehensive representation of the aero forces and could be effectively used to refine the current aerodynamic map.
-By incorporating such a model, we expect not only a more physically consistent representation of the aerodynamics but also a smoother and more meaningful optimization landscape, reducing the occurrence of non-physical local minima and improving overall solution robustness.%More in details, the all steering vehicle, having the capability to steer independently with all four wheels, and considering that the aerodynamic drag only depends on the forward velocity $x_1$, could run with excessively large $\beta$ angles, even on straights paths.
-Even if this behavior could be appropriate in certain track sectors, is undesirable from an engineering perspective.
-%In some cases, it could also be the outcome of the solver being stuck in local minima: in these events the above term helps discourage such behaviours.
-
-In the current context, we imposed coefficient $k_v$ to be a function of the (absolute value of) track curvature $c_k$ according to Eq.~\eqref{eqn:lateral_cost}.
-Here, the two constant $\bar{k}_v$ and $k_c$ are selected heuristically for each track; the first increases/decreases the weight of the cost term, while the second varies the hardness of the intervention in the range $[0,1]$.
-\begin{equation}\label{eqn:lateral_cost}
-	k_v = \bar{k}_v(1-\tanh(k_c c_k))
-\end{equation}
-
-To ensure reproducibility, we report the values of the main penalty weights for the cost function terms used in the Siena track simulations: $k_{\delta} = 0.05$, $k_T = 0.005$, $\bar{k}_v = 0.005$, and $k_c = 100$.
+\subsection{Robust adherence constraint}

### 2025-04-18 closed_loop_planning.tex
[MGa] updated closed loop and mod open loop
Restructured eqs in \subequations, ...

@@
 This is the trajectory a driver should follow in the ideal situation of a world without uncertainty.
-Therefore, in this step the mean trajectory is the actual trajectory a deterministic vehicle dynamics would trace.
+In this step the mean trajectory is the actual trajectory a deterministic vehicle dynamics would trace.
+
 ii) A closed-loop controller mimicking the driver's action is computed which stabilizes the nominal feed-forward optimal trajectory computed at step i).
 We assume a discrete time-varying Linear Quadratic Regulator (LQR) controller is a good approximation of how an expert driver can track a prescribed trajectory.
+
 iii) In the last step the stochastic framework is reintroduced and the time-optimal planning problem is robustified using the Lyapunov framework.
@@
 This amounts to ensuring that the robustified path constraints are satisfied.
-More specifically, this implies two key aspects: first, that the covariance matrix is propagated accounting explicitly for the controlled system dynamics; and second, that the path constraints are robustly satisfied by incorporating the covariance estimation propagated through the closed-loop controlled system.
-Some aspects that may initially appear technical are in fact essential.
+More specifically, this accounts for two key aspects: firstly, that the covariance matrix is propagated accounting explicitly for the controlled system dynamics; and secondly, that the path constraints are robustly satisfied by incorporating the covariance estimation propagated through the closed-loop controlled system.
+Some aspects that may appear a bit technical are in fact essential.
 In this third step, not only are states and controls re-planned, but also the time-varying LQR gains.
-In fact, they need to explicitly account for the changed requirements associated with stabilizing a trajectory different from the nominal one, which must now satisfy the robustified constraints.
+As a matter of fact, they need to explicitly account for the changed requirements associated with stabilizing a trajectory different from the nominal one, which must now satisfy the robustified constraints.
 \subsection{Step 1 -- Nominal feed-forward time-optimal trajectory planning}
 \label{sec:nominalFF}
+The first step consists in planning a nominal feed-forward trajectory.
+This assumes deterministic system dynamics, with \( (\hbmu_k, \hbu_k) \) denoting the differential states and controls, and \( \hbz_k \) the algebraic states, for \( k = 0, \ldots, N \).
+To avoid redundancy, we note that the NLP to be solved is a special case of~\eqref{eq:DOCP}, where~\eqref{eq:DOCPdP},~\eqref{eq:DOCPdPIC}, and the backoff terms are omitted, since the goal is to compute a nominal trajectory.
 
@@
 
-Given the nominal trajectory $(\hbmu_k, \hbu_k)$ over $k=0,\ldots, N$, and introduced the deviation variables $\bbmu_k=\bmu_k - \hbmu_k$ and $\bbu = \bu_k - \hbu_k$, we design a linear state-feedback law $\bbu_k= -\hbK_k \bbmu_k$, or equivalently $\bu_k = \hbu_k - \hbK_k (\bmu_k - \hbmu_k)$, to stabilize the deviations along the nominal trajectory.
+Given the nominal trajectory $(\hbmu_k, \hbu_k)$ over $k=0,\ldots, N$, and introduced the deviation variables $\bbmu_k=\bmu_k - \hbmu_k$ and $\bbu = \bu_k - \hbu_k$, we design a linear state-feedback law $\bbu_k= -\hbK_k \bbmu_k$, or equivalently $\bu_k = \hbu_k - \hbK_k (\bmu_k - \hbmu_k)$, to mitigate disturbances while tracking the nominal trajectory.
 To compute the state-feedback matrices $\hbK_k$, we linearize the system around $(\hbmu_k, \hbu_k)$, such that $\bbmu_{k+1} \approx \hbA_{k}\bbmu_k + \hbB_k \bbu_k$, where $\hbA_k = f_{,\bmu}(\hbmu_k,\hbu_k)$ and $\hbB_k = f_{,\bu}(\hbmu_k,\hbu_k)$, and define the quadratic regulator (tracking) cost function
@@
 \label{sec:CLrobustified}
-
 %\begin{align}
@@
 %\end{align}
-
+As the third and last step, a closed-loop robustified planning problem is set up.
+Its aim is the definition of a mean trajectory of a stochastic system such that, working in tandem with state-feedback controller that aims at stabilizing it, it is able to mitigate the perturbations and reduce the propagation of uncertainty.
+This helps to guarantee that the robustified path constraints can be satisfied over the entire planning horizon.
+The discretized version of this problem takes the form of the following NLP:
+\begin{subequations}\label{eq:ROCP}
 \begin{alignat}{3}
@@
 \end{alignat}
-where~\eqref{eq:ROCPdP} denotes the discretized form for time evolution of the covariance matrix under a stabilizing controller of the form $\bbu_k=-\bK_k \bbmu_k$.
-As apposed to~\eqref{eq:OCPdP}, $\tbP$ would evolve, in continuous time, according to $\dot{\tbP}(t) = (\bA(t)-\bB(t) \bK(t)) \tbP(t) + \tbP(t) (\bA(t)-\bB(t) \bK(t))^T + \bQ(t)$.
+\end{subequations} 
+Here, \eqref{eq:ROCPdP} denotes the discretized form for time evolution of the covariance matrix under a stabilizing controller of the form $\bbu_k=-\bK_k \bbmu_k$.
+To mark the difference with~\eqref{eq:OCPdP}, a new symbol $\tbP$ is introduced to the denote the covariance matrix in this case.
+In continuous time, $\tbP$ evolves according to $\dot{\tbP}(t) = (\bA(t)-\bB(t) \bK(t)) \tbP(t) + \tbP(t) (\bA(t)-\bB(t) \bK(t))^T + \bQ(t)$.
 To be continued...

### 2025-04-18 open_loop_planning.tex
[MGa] updated closed loop and mod open loop
Restructured eqs in \subequations, ...

@@
 \label{sec:open_loop_planning}
-The proposed approach requires the presence of $H+1$ versions of the matrix $\bP$ at each grid point of the discretized problem.
+In this section, we address the problem of planning under uncertainty for a stochastic system, where the covariance matrix is propagated in a feed-forward fashion for a fixed prediction horizon of $H$ steps at each grid point $k$ of the discretized trajectory.
+In the planning problem, we conservatively evaluate the maximum propagation of the covariance, that is, the matrix evolved over $H$ steps, to ensure the tightest possible enforcement of constraints.
+This maximizes the robustness of the resulting solution, as the constraint backoff is evaluated in the worst-case scenario.
+
+To enable this, we associate multiple versions of the covariance matrix with each grid point.
+Specifically, $H+1$ instances of the matrix $\bP$ are maintained at each step.
+To this sake we introduce the notation $\bP_k^j$, where $j = 0, \ldots, H$, and $\bP_k^j$ represents the version of the covariance matrix at step $k$ that was initialized $j$ steps earlier.
+Hence, $\bP_k^0$ originates at node $k$ itself and will be propagated forward and employed in subsequent steps (exactly when evolved as $\bP_{k+H}^H$ ), while $\bP_k^H$ is the instance that originated $H$ steps earlier and is the most propagated one at step $k$.
+
+The repeated initialization of the covariance matrix at each step is necessary to prevent its unbounded growth, as we assume that the system evolves without feedback control.
+In the absence of feedback, the uncertainty on the position and orientation of the body frame accumulates along the track, which may severely and unrealistically degrade the conditioning of the covariance matrix.
+For this reason, the propagation horizon in this open-loop setting is a critical design parameter that must be bounded to realistically capture the evolution of disturbances occurring prior to any corrective driver response.
+
+This approach requires modifications to the formulation~\eqref{eq:DOCP} introduced in Section~\ref{sec:discretization}.
+In particular, Eqs.~\eqref{eq:DOCPdP}, \eqref{eq:DOCPdPIC}, and \eqref{eq:DOCPconstraints} are replaced by Eqs.~\eqref{eq:DOCPdPopenloop}, \eqref{eq:DOCPdPinitopenloop}, and \eqref{eq:DOCPconstraintsopenloop}.
+Equation~\eqref{eq:DOCPdPopenloop} collects the continuity and collocation equations for all versions of the covariance matrix propagated from previous steps.
+To support the evolution of $H$ versions, we introduce $H \times d$ matrices $\bSi_k^j = (\bSi_{k,1}^j, \ldots, \bSi_{k,d}^j)$, representing the collocation values of each version $j$ of the covariance matrix within interval $k$.
+
+%The proposed approach requires the presence of $H+1$ versions of the matrix $\bP$ at each grid point of the discretized problem.
 One of these versions is initialized with $\bar{\bP}_0$, while the others represent the evolution of matrices that were initialized at previous grid points, from 1 to $H$ steps earlier.
@@
 In fact, without a feedback control, the error on the position and orientation of the body reference frame can only increase along the track, compromising the conditioning of the matrix.
-
-%\subsection{Discretization via direct collocation}
-%\label{sec:openloopdiscretization}
-Introducing the versioning of the covariance matrix approach requires a modification of the Equations described in Section~\ref{sec:discretization}.
+%
+%Introducing the versioning of the covariance matrix approach requires a modification of the Equations described in Section~\ref{sec:discretization}.
 In particular Eqs.~\eqref{eq:DOCPdP}, \eqref{eq:DOCPdPIC}, and \eqref{eq:DOCPconstraints} are replaced by Eqs.~\eqref{eq:DOCPdPopenloop}, \eqref{eq:DOCPdPinitopenloop}, and \eqref{eq:DOCPconstraintsopenloop}.
@@
 The specific formulation is expressed as follows:
+\begin{subequations}\label{eq:DOCPopenloop}
 \begin{alignat}{3}
 	\underset{\bmu_k,\bxi_k, \bu_k, \bP_k,\bz_k}{\text{minimize}} \,
-	& & & J_k(\bmu_k,\bxi_k, \bu_k) & & \label{eq:DOCPcost} \\
+	& & & J_k(\bmu_k,\bxi_k, \bu_k) & & \label{eq:DOCPcostopenloop} \\
 	\hspace*{-2.0 cm}\text{s.t.} \quad
 	& \bzero      & = & \; \bPsimu_k(\bmu_{k-1},\bmu_k,\bxi_k, \bu_k,\bz_k),
-	& \quad & k = 1,\ldots, N \label{eq:DOCPdyn} \\
+	& \quad & k = 1,\ldots, N \label{eq:DOCPdynopenloop} \\
 	& \bmu_0      & = & \; \bar{\bmu}_0
-	& & \label{eq:DOCPdynIC} \\
+	& & \label{eq:DOCPdynICopenloop} \\
 	& \bzero      & = & \; \bPsiP_k(\bmu_k,\bxi_k, \bu_k, \bP_{k-1}^{j-1},\bP_k^j,\bSi_k^j,\bz_k),
@@
 	& \bzero      & = & \; \bOm_k(\bmu_k,\bxi_k, \bu_k,\bz_k),
-	& \quad & k = 0,\ldots, N \label{eq:DOCPpath} \\
+	& \quad & k = 0,\ldots, N \label{eq:DOCPpathopenloop} \\
 	& 0           & \geq & \; h_i(\bmu_k, \bu_k, \bz_k) + \be_i(\bmu_k, \bu_k, \bP_k^H, \bz_k),
@@
 \end{alignat}
+\end{subequations}
+Equation~\eqref{eq:DOCPdPinitopenloop} defines the initialization of the appropriate covariance matrix version at step $k$, while Eq.~\eqref{eq:DOCPconstraintsopenloop} specifies that the back-off term in the constraints is computed using the most propagated version, $\bP^H_k$.
+A schematic illustration of the management of the $H+1$ covariance matrix instances is provided in Figure~\ref{fig:DOCPgrid}.
+The dashed rectangle highlights the $k$-th discretization step, where the constraints are evaluated.
+At each grid point, two particular versions of the covariance matrix are emphasized: $\bP^0_k$ (red node), representing the instance to be initialized at that step, and $\bP^H_k$ (green node), corresponding to the version that has been propagated over the full horizon of $H$ steps and is used to robustify the constraints in Eq.~\eqref{eq:DOCPconstraintsopenloop}.
 
-Eq.~\eqref{eq:DOCPdPinitopenloop} is used to initialize the correct version at the $k$-th step while Eq.~\eqref{eq:DOCPconstraintsopenloop} specify that the back-off term of the constraints is evaluated with the most propagated covariance matrix version, $\bP^H_k$.
-A schematic representation of the approach used to manage the $H+1$ versions of the covariance matrix is shown in Figure \ref{fig:DOCPgrid}, where the dashed rectangle represent the $k$-th step at which the constraints are being formulated.
+%Eq.~\eqref{eq:DOCPdPinitopenloop} is used to initialize the correct version at the $k$-th step while Eq.~\eqref{eq:DOCPconstraintsopenloop} specify that the back-off term of the constraints is evaluated with the most propagated covariance matrix version, $\bP^H_k$.
+%A schematic representation of the approach used to manage the $H+1$ versions of the covariance matrix is shown in Figure \ref{fig:DOCPgrid}, where the dashed rectangle represent the $k$-th step at which the constraints are being formulated.
 At each grid point, two versions of the covariance matrix are highlighted: $\bP^0_k$ (red node), which is the version to be initialized, and $\bP^H_k$ (green node), which is the most propagated version used to formulate the robust constraints in Eq.~\eqref{eq:DOCPconstraintsopenloop}.
@@
 	\end{tikzpicture}
-	\caption{Schematic representation of the continuity and initialization Equations for the $H+1$ versions of the covariance matrix introduced in the discretized OCP.
+\caption{Schematic representation of the continuity and initialization equations associated with the $H+1$ propagated instances of the covariance matrix introduced in problem~\eqref{eq:DOCPopenloop}.
+The dashed rectangle highlights the $k$-th discretization step, at which the robust constraints are enforced.
+Each node corresponds to a specific covariance matrix, where the subscript denotes the current grid index and the superscript indicates the number of propagation steps since initialization.
+At each step, two key instances are emphasized: the matrix to be initialized at that step (red node), and the matrix that has been propagated over $H$ steps (green node), which is used in the evaluation of the constraint backoff.}
+	%\caption{Schematic representation of the continuity and initialization Equations for the $H+1$ versions of the covariance matrix introduced in the discretized OCP.
 The dashed rectangle indicate the $k$-th step, in which the constraints are formulated.
@@
 \end{figure}
-
-\subsection{Robust adherence constraint}

### 2025-04-18 stochastic_vehicle_dynamics.tex
[MGa] updated closed loop and mod open loop
Restructured eqs in \subequations, ...

@@
 Accordingly, the continuous-time stochastic trajectory planning can be framed as the following nonlinear optimal control problem
+\begin{subequations}\label{eq:OCP}
 \begin{align}
@@
 \end{align}
+\end{subequations}
 The cost function $J$ in~\eqref{eq:OCPcost} depends on the mean $\bmu(t)$, the controls $\bu(t)$ and the state covariance $\bP(t)$.
 The mean dynamics is expressed by~\eqref{eq:OCPdyn} with~\eqref{eq:OCPdynIC}, and the covariance dynamics is expressed by~\eqref{eq:OCPdP} with~\eqref{eq:OCPdPIC}.
-In eq.~\eqref{eq:OCPconstraints} the \emph{backoff terms} $\be_i$ account for the disturbances and serve the purpose of obtaining deterministic safety-margins directly on $\bmu(t)$.
+In eq.~\eqref{eq:OCPconstraints} the \emph{backoff terms} $\be_i$ account for the disturbances and serve the purpose of obtaining deterministic safety margins directly on $\bmu(t)$.
 $\calI$ is the set of indices defining the inequality constraints.
-This is equivalent to a linearization around the mean of the original \emph{chance constraint} on $\bx(t)$ expressed by $\Prob \{h_i(\bx) \leq 0\}\geq p$, where $p$ is the confidence level of constraint satisfaction.
-The backoff terms are given by
+The backoff terms stem for a linearization around the mean of the original \emph{chance constraint} on $\bx(t)$ expressed by $\Prob \{h_i(\bx) \leq 0\}\geq p$, where $p$ is the confidence level of constraint satisfaction.
+Explicitly the backoff terms are given by
 $\be_i = \gamma \sig_i$.
@@
 
-The nonlinear optimal control problem~\eqref{eq:OCPcost} can be discretized by applying a suitable collocation integrator obtaining the following nonlinear program (NLP) 
+The nonlinear optimal control problem~\eqref{eq:OCP} can be discretized by applying a suitable collocation integrator obtaining the following nonlinear program (NLP)
+\begin{subequations}\label{eq:DOCP}
 \begin{alignat}{3}
@@
 \end{alignat}
+\end{subequations}
 To perform the discretization, the track centerline is parameterized using a curvilinear abscissa $\alpha \in [0,1]$, and uniformly sampled at $N+1$ points $\alpha_0, \ldots, \alpha_N$.
-Accordingly, $\bmu_k$ denotes the mean state at grid node $\alpha_k$, while $\bu_k$ and $\bz_k$ are assumed to be piecewise constant over each interval $[\al_k, \al_{k+1}]$.
-These represent the controls and the algebraic variables, respectively.
-The $\bz_k$'s are introduced as direct handles for physically meaningful quantities such contact forces.
-Similarly, $\bP_k$ represents the covariance matrix at the $k$-th node.
+Accordingly, $\bmu_k$ denotes the mean state at grid node $\alpha_k$, while the controls $\bu_k$ and the algebraic variables $\bz_k$ are assumed to be piecewise constant over each interval $[\al_k, \al_{k+1}]$.
+The $\bz_k$'s are introduced as direct handles for physically meaningful quantities such contact forces. Similarly, $\bP_k$ represents the covariance matrix at the $k$-th node.
 Following the direct collocation approach, both the mean and covariance state trajectories are approximated, in the $k$-th interval, with polynomials $\pi_k(\tau)$, defined on the unit interval $\tau\in[0,1]$, and then scaled to match the width $h_k$ of the corresponding time step.
@@
 In certain cases~\cite{Gillis:PracticalMethodsApproximate:2015}, positive-definiteness-preserving Lyapunov discretization schemes are used in~\eqref{eq:DOCPdP}.
-However, for sufficiently fine discretization, we found that integrating only the lower triangular part of $\bP(t)$ in~\eqref{eq:OCPdPIC}  (and reconstructing its strictly upper part accordingly), ensures that all $\bP_k$ remain symmetric and positive definite when the initial $\bP_0$ is symmetric and positive definite.
-In all our tests direct collocation with cubic polynomial state representations and Gauss-Legendre collocation points were used.
+However, for sufficiently fine discretization, we found that integrating only the lower triangular part of $\bP(t)$ in~\eqref{eq:OCPdPIC}  (and reconstructing its strictly upper part accordingly), ensures that $\bP_k$ remain symmetric and positive definite when the initial $\bP_0$ is symmetric and positive definite.
+In all our tests direct collocation with cubic polynomial state representations and Gauss-Legendre collocation points provided accurate results.

### 2025-04-18 MAIN.tex
[MGa] some rephrasing
[MGa] stochastic OCP
[MM] constraints formulation (first try)
[MGa] TV-LQR, first version
[MM] DOCP re-formulation
[MGa] also my open loop figure
Let's see...

@@
 \usepackage{tikz,tikz-3dplot}
-\usetikzlibrary{arrows.meta,angles}
+\usetikzlibrary{arrows.meta,angles,decorations.pathmorphing,patterns}
+
+\usepackage{bm}
 
@@
 
-\author{Martino Gulisano\textsuperscript{a}, Matteo Masoni\textsuperscript{a}, and \name{Marco Gabiccini\textsuperscript{a,*}
+\author{Martino Gulisano\textsuperscript{a}, Matteo Masoni\textsuperscript{a}, \name{Marco Gabiccini\textsuperscript{a,*} and Massimo Guiggiani\textsuperscript{a}
 \thanks{* Contact: Marco Gabiccini. Email: marco.gabiccini@unipi.it}
@@
 \input{stochastic_vehicle_dynamics}
-\input{open_loop_planning}
+\input{marco_open_loop_planning}
 %for back-off calculations

### 2025-04-18 marco_open_loop_planning.tex
[MGa] also my open loop figure
Let's see...

--- 2025-04-18 open_loop_planning.tex
@@
 	\centering
-	\begin{tikzpicture}[%
-		smallnode/.style={%
-			circle, draw, minimum size=3mm, inner sep=0pt
-		},
-		r_smallnode/.style={%
-			circle, draw, minimum size=3mm, inner sep=0pt, red
-		},
-		g_smallnode/.style={%
-			circle, draw, minimum size=3mm, inner sep=0pt, mygreen
-		},
+%\begin{tikzpicture}[%
+%		smallnode/.style={%
+%			circle, draw, minimum size=3mm, inner sep=0pt
+%		},
+%		r_smallnode/.style={%
+%			circle, draw, minimum size=3mm, inner sep=0pt, red
+%		},
+%		g_smallnode/.style={%
+%			circle, draw, minimum size=3mm, inner sep=0pt, mygreen
+%		},
+%		every path/.style={->, thick},
+%		x=1.5cm, y=1.5cm
+%		]
+%		
+%		\newcommand{\Plab}[2]{\bP^{#2}_{#1}}
+%		\definecolor{mygreen}{RGB}{0 150 0}
+%		
+%		% Nodes
+%		% First row
+%		\node[smallnode, label={[label distance=-2pt]above:\protect\(\Plab{k-2}{H-2}\)}] (Pk-2H-2) at (1,0) {};
+%		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{H-1}\)}] (Pk-1H-1) at (2,0) {};
+%		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k}{H}\)}] (PkH) at (3,0) {};
+%		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{0}\)}] (Pk+10) at (4,0) {};
+%		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{1}\)}] (Pk+21) at (5,0) {};
+%		
+%		% Second row
+%		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-2}{H-3}\)}] (Pk-2H-3) at (1,1) {};
+%		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{H-2}\)}] (Pk-1H-2) at (2,1) {};
+%		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k}{H-1}\)}] (PkH-1) at (3,1) {};
+%		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{H}\)}] (Pk+1H) at (4,1) {};
+%		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{0}\)}] (Pk+20) at (5,1) {};
+%		
+%		% dots row
+%		\node[label={center,rotate=90:\(\dots\)}] (dots2) at (1, 1.7) {};
+%		\node[label={center,rotate=90:\(\dots\)}] (dots3) at (2, 1.7) {};
+%		\node[label={center,rotate=90:\(\dots\)}] (dots4) at (3, 1.7) {};
+%		\node[label={center,rotate=90:\(\dots\)}] (dots5) at (4, 1.7) {};
+%		\node[label={center,rotate=90:\(\dots\)}] (dots6) at (5, 1.7) {};
+%		
+%		% Third row
+%		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k-2}{H}\)}] (Pk-2H) at (1,2) {};
+%		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{0}\)}] (Pk-10) at (2,2) {};
+%		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k}{1}\)}] (Pk1) at (3,2) {};
+%		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{2}\)}] (Pk+12) at (4,2) {};
+%		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{3}\)}] (Pk+23) at (5,2) {};
+%		
+%		% Fourth row
+%		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-2}{H-1}\)}] (Pk-2H-1) at (1,3) {};
+%		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{H}\)}] (Pk-1H) at (2,3) {};
+%		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k}{0}\)}] (Pk0) at (3,3) {};
+%		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{1}\)}] (Pk+11) at (4,3) {};
+%		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{2}\)}] (Pk+22) at (5,3) {};
+%		
+%		% Curved arrows
+%		\draw[bend left=20] (Pk-2H-2) to (Pk-1H-1);
+%		\draw[bend left=20] (Pk-1H-1) to (PkH);
+%		\draw[bend left=20] (Pk+10) to (Pk+21);
+%		
+%		\draw[bend left=20] (Pk-2H-3) to (Pk-1H-2);
+%		\draw[bend left=20] (Pk-1H-2) to (PkH-1);
+%		\draw[bend left=20] (PkH-1) to (Pk+1H);
+%		
+%		\draw[bend left=20] (Pk-10) to (Pk1);
+%		\draw[bend left=20] (Pk1) to (Pk+12);
+%		\draw[bend left=20] (Pk+12) to (Pk+23);
+%		
+%		\draw[bend left=20] (Pk-10) to (Pk1);
+%		\draw[bend left=20] (Pk1) to (Pk+12);
+%		\draw[bend left=20] (Pk+12) to (Pk+23);
+%		
+%		\draw[bend left=20] (Pk-2H-1) to (Pk-1H);
+%		\draw[bend left=20] (Pk0) to (Pk+11);
+%		\draw[bend left=20] (Pk+11) to (Pk+22);
+%		
+%		% Dashed rectangle
+%		\draw[dashed, thick, rounded corners] (2.6,-.25) rectangle (3.4,3.5);
+%		
+%	\end{tikzpicture}
+\definecolor{mygreen}{RGB}{0,150,0}
+%\begin{tikzpicture}[
+%    smallnode/.style={circle, draw, minimum size=3mm, inner sep=0pt},
+%    r_smallnode/.style={circle, draw=red, minimum size=3mm, inner sep=0pt},
+%    g_smallnode/.style={circle, draw=mygreen, minimum size=3mm, inner sep=0pt},
+%    every path/.style={->, thick},
+%    x=1.4cm, y=1.2cm
+%]
+%
+%% Row j=0 (y = 3)
+%\node[r_smallnode, label={[label distance=-2pt]above:\(\bP_0^0\)}] (P00) at (1,3) {};
+%\node[smallnode, label={[label distance=-2pt]above:\(\bP_1^1\)}] (P11) at (2,3) {};
+%\node[smallnode, label={[label distance=-2pt]above:\(\bP_2^2\)}] (P22) at (3,3) {};
+%\node[g_smallnode, label={[label distance=-2pt]above:\(\bP_3^3\)}] (P33) at (4,3) {};
+%\node[smallnode, label={[label distance=-2pt]above:\(\bP_4^4\)}] (P44) at (5,3) {};
+%\node at (6,3) {\(\cdots\)};
+%
+%% Row j=1 (y = 2)
+%\node[r_smallnode, label={[label distance=-2pt]above:\(\bP_1^0\)}] (P10) at (2,2) {};
+%\node[smallnode, label={[label distance=-2pt]above:\(\bP_2^1\)}] (P21) at (3,2) {};
+%\node[smallnode, label={[label distance=-2pt]above:\(\bP_3^2\)}] (P32) at (4,2) {};
+%\node[g_smallnode, label={[label distance=-2pt]above:\(\bP_4^3\)}] (P43) at (5,2) {};
+%\node at (6,2) {\(\cdots\)};
+%
+%% Row j=2 (y = 1)
+%\node[r_smallnode, label={[label distance=-2pt]above:\(\bP_2^0\)}] (P20) at (3,1) {};
+%\node[smallnode, label={[label distance=-2pt]above:\(\bP_3^1\)}] (P31) at (4,1) {};
+%\node[smallnode, label={[label distance=-2pt]above:\(\bP_4^2\)}] (P42) at (5,1) {};
+%\node[g_smallnode, label={[label distance=-2pt]above:\(\bP_5^3\)}] (P53) at (6,1) {};
+%\node at (7,1) {\(\cdots\)};
+%
+%% Row j=3 (y = 0)
+%\node[r_smallnode, label={[label distance=-2pt]above:\(\bP_3^0\)}] (P30) at (4,0) {};
+%\node[smallnode, label={[label distance=-2pt]above:\(\bP_4^1\)}] (P41) at (5,0) {};
+%\node[smallnode, label={[label distance=-2pt]above:\(\bP_5^2\)}] (P52) at (6,0) {};
+%\node[g_smallnode, label={[label distance=-2pt]above:\(\bP_6^3\)}] (P63) at (7,0) {};
+%\node at (8,0) {\(\cdots\)};
+%
+%% Arrows: P_k^j --> P_{k+1}^{j+1}
+%\draw (P00) -- (P11);
+%\draw (P11) -- (P22);
+%\draw (P22) -- (P33);
+%\draw (P33) -- (P44);
+%
+%\draw (P10) -- (P21);
+%\draw (P21) -- (P32);
+%\draw (P32) -- (P43);
+%
+%\draw (P20) -- (P31);
+%\draw (P31) -- (P42);
+%\draw (P42) -- (P53);
+%
+%\draw (P30) -- (P41);
+%\draw (P41) -- (P52);
+%\draw (P52) -- (P63);
+%
+%% Dashed box around column k = 3 (x=4)
+%\draw[dashed, thick, rounded corners] (3.6,-0.5) rectangle (4.4,3.5);
+%
+%% Optional top axis labels
+%\node at (1, 4.1) {\scriptsize \(k=0\)};
+%\node at (2, 4.1) {\scriptsize \(k=1\)};
+%\node at (3, 4.1) {\scriptsize \(k=2\)};
+%\node at (4, 4.1) {\scriptsize \(k=3\)};
+%\node at (5, 4.1) {\scriptsize \(k=4\)};
+%\node at (6, 4.1) {\scriptsize \(k=5\)};
+%\node at (7, 4.1) {\scriptsize \(k=6\)};
+%
+%\end{tikzpicture}
+
+\begin{tikzpicture}[
+    smallnode/.style={circle, draw, minimum size=3mm, inner sep=0pt},
+    r_smallnode/.style={circle, draw=red, minimum size=3mm, inner sep=0pt},
+    g_smallnode/.style={circle, draw=mygreen, minimum size=3mm, inner sep=0pt},
+    r_hatnode/.style={
+    circle, draw=red, pattern=north east lines, pattern color=red,
+    minimum size=3mm, inner sep=0pt},
+    g_hatnode/.style={
+    circle, draw=mygreen, pattern=north east lines, pattern color=mygreen,
+    minimum size=3mm, inner sep=0pt},
     every path/.style={->, thick},
-		x=1.5cm, y=1.5cm
+    x=1.4cm, y=1.2cm
 ]
 
-		\newcommand{\Plab}[2]{\bP^{#2}_{#1}}
-		\definecolor{mygreen}{RGB}{0 150 0}
+% ROW j = 0 (y=3)
+\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k-3}^0\)}] (P00) at (1,3) {};
+\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k-2}^1\)}] (P11) at (2,3) {};
+\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k-1}^2\)}] (P22) at (3,3) {};
+\node[g_hatnode, label={[label distance=-2pt]above:\(\bP_k^3\)}] (P33) at (4,3) {};
+\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k+1}^0\)}] (P44) at (5,3) {};
+\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+2}^1\)}] (P55) at (6,3) {};
 
-		% Nodes
-		% First row
-		\node[smallnode, label={[label distance=-2pt]above:\protect\(\Plab{k-2}{H-2}\)}] (Pk-2H-2) at (1,0) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{H-1}\)}] (Pk-1H-1) at (2,0) {};
-		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k}{H}\)}] (PkH) at (3,0) {};
-		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{0}\)}] (Pk+10) at (4,0) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{1}\)}] (Pk+21) at (5,0) {};
+% ROW j = 1 (y=2)
+\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k-2}^0\)}] (P10) at (2,2) {};
+\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k-1}^1\)}] (P21) at (3,2) {};
+\node[smallnode, label={[label distance=-2pt]above:\(\bP_k^2\)}] (P32) at (4,2) {};
+\node[g_hatnode, label={[label distance=-2pt]above:\(\bP_{k+1}^3\)}] (P43) at (5,2) {};
+\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k+2}^0\)}] (P54) at (6,2) {};
+\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+3}^1\)}] (P65) at (7,2) {};
 
-		% Second row
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-2}{H-3}\)}] (Pk-2H-3) at (1,1) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{H-2}\)}] (Pk-1H-2) at (2,1) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k}{H-1}\)}] (PkH-1) at (3,1) {};
-		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{H}\)}] (Pk+1H) at (4,1) {};
-		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{0}\)}] (Pk+20) at (5,1) {};
+% ROW j = 2 (y=1)
+\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k-1}^0\)}] (P20) at (3,1) {};
+\node[smallnode, label={[label distance=-2pt]above:\(\bP_k^1\)}] (P31) at (4,1) {};
+\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+1}^2\)}] (P42) at (5,1) {};
+\node[g_hatnode, label={[label distance=-2pt]above:\(\bP_{k+2}^3\)}] (P53) at (6,1) {};
+\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k+3}^0\)}] (P64) at (7,1) {};
+\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+4}^1\)}] (P75) at (8,1) {};
 
-		% dots row
-		\node[label={center,rotate=90:\(\dots\)}] (dots2) at (1, 1.7) {};
-		\node[label={center,rotate=90:\(\dots\)}] (dots3) at (2, 1.7) {};
-		\node[label={center,rotate=90:\(\dots\)}] (dots4) at (3, 1.7) {};
-		\node[label={center,rotate=90:\(\dots\)}] (dots5) at (4, 1.7) {};
-		\node[label={center,rotate=90:\(\dots\)}] (dots6) at (5, 1.7) {};
+% ROW j = 3 (y=0)
+\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_k^0\)}] (P30) at (4,0) {};
+\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+1}^1\)}] (P41) at (5,0) {};
+\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+2}^2\)}] (P52) at (6,0) {};
+\node[g_hatnode, label={[label distance=-2pt]above:\(\bP_{k+3}^3\)}] (P63) at (7,0) {};
+\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k+4}^0\)}] (P74) at (8,0) {};
+\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+5}^1\)}] (P85) at (9,0) {};
 
-		% Third row
-		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k-2}{H}\)}] (Pk-2H) at (1,2) {};
-		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{0}\)}] (Pk-10) at (2,2) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k}{1}\)}] (Pk1) at (3,2) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{2}\)}] (Pk+12) at (4,2) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{3}\)}] (Pk+23) at (5,2) {};
+% ARROWS
+\draw (P00) -- (P11);
+\draw (P11) -- (P22);
+\draw (P22) -- (P33);
+\draw (P22) -- (P33);
 
-		% Fourth row
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-2}{H-1}\)}] (Pk-2H-1) at (1,3) {};
-		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{H}\)}] (Pk-1H) at (2,3) {};
-		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k}{0}\)}] (Pk0) at (3,3) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{1}\)}] (Pk+11) at (4,3) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{2}\)}] (Pk+22) at (5,3) {};
+\draw (P44) -- (P55);
+\draw[dashed,->] (P55) -- ++(1,0);
 
-		% Curved arrows
-		\draw[bend left=20] (Pk-2H-2) to (Pk-1H-1);
-		\draw[bend left=20] (Pk-1H-1) to (PkH);
-		\draw[bend left=20] (Pk+10) to (Pk+21);
+\draw (P10) -- (P21);
+\draw (P21) -- (P32);
+\draw (P32) -- (P43);
+\draw (P54) -- (P65);
+\draw[dashed,->] (P65) -- ++(1,0);
 
-		\draw[bend left=20] (Pk-2H-3) to (Pk-1H-2);
-		\draw[bend left=20] (Pk-1H-2) to (PkH-1);
-		\draw[bend left=20] (PkH-1) to (Pk+1H);
+\draw (P20) -- (P31);
+\draw (P31) -- (P42);
+\draw (P42) -- (P53);
+\draw (P64) -- (P75);
+\draw[dashed,->] (P75) -- ++(1,0);
 
-		\draw[bend left=20] (Pk-10) to (Pk1);
-		\draw[bend left=20] (Pk1) to (Pk+12);
-		\draw[bend left=20] (Pk+12) to (Pk+23);
+\draw (P30) -- (P41);
+\draw (P41) -- (P52);
+\draw (P52) -- (P63);
 
-		\draw[bend left=20] (Pk-10) to (Pk1);
-		\draw[bend left=20] (Pk1) to (Pk+12);
-		\draw[bend left=20] (Pk+12) to (Pk+23);
+\draw (P74) -- (P85);
+\draw[dashed,->] (P85) -- ++(1,0);
 
-		\draw[bend left=20] (Pk-2H-1) to (Pk-1H);
-		\draw[bend left=20] (Pk0) to (Pk+11);
-		\draw[bend left=20] (Pk+11) to (Pk+22);
+% Dashed box around column k (x=4)
+\draw[dashed, thick, rounded corners] (3.6,-0.5) rectangle (4.4,3.7);
 
-		% Dashed rectangle
-		\draw[dashed, thick, rounded corners] (2.6,-.25) rectangle (3.4,3.5);
+% Top axis labels
+\node at (0, 4.3) {\scriptsize \(\cdots\)};
+\node at (1, 4.3) {\scriptsize \(k-3\)};
+\node at (2, 4.3) {\scriptsize \(k-2\)};
+\node at (3, 4.3) {\scriptsize \(k-1\)};
+\node (knode) at (4, 4.3) {\scriptsize \(k\)};
+\node at (5, 4.3) {\scriptsize \(k+1\)};
+\node at (6, 4.3) {\scriptsize \(k+2\)};
+\node at (7, 4.3) {\scriptsize \(k+3\)};
+\node at (8, 4.3) {\scriptsize \(k+4\)};
+\node at (9, 4.3) {\scriptsize \(\cdots\)};
+
+% Wiggly arrow from label to top edge of dashed box
+\draw[->, thick, decorate, decoration={snake, amplitude=0.25mm, segment length=2.5mm}]
+  (knode.south) -- (4, 3.8);
+
+% Legend box (top right)
+%\begin{scope}[shift={(6.5,2.5)}]
+%    \node[r_smallnode] (r) at (0,0) {};
+%    \node[anchor=west] at (0.4,0) {\scriptsize Initialized matrix \(\bP_k^0\)};
+%
+%    \node[g_smallnode] (g) at (0,-0.8) {};
+%    \node[anchor=west] at (0.4,-0.8) {\scriptsize Most propagated matrix \(\bP_k^3\)};
+%\end{scope}
 
 \end{tikzpicture}
-\caption{Schematic representation of the continuity and initialization equations associated with the $H+1$ propagated instances of the covariance matrix introduced in problem~\eqref{eq:DOCPopenloop}.
+
+\caption{Cartoon representation (depicted for $H=3$) of the continuity and initialization equations associated with the $H+1$ propagated instances of the covariance matrix introduced in problem~\eqref{eq:DOCPopenloop}.
 The dashed rectangle highlights the $k$-th discretization step, at which the robust constraints are enforced.

### 2025-04-18 open_loop_planning.tex

@@
 		% First row
-		\node[smallnode, label={[label distance=-2pt]above:\protect\(\Plab{k-2}{H-2}\)}] (Pk-2H-2) at (1,0) {};
+		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-2}{H-2}\)}] (Pk-2H-2) at (1,0) {};
 		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{H-1}\)}] (Pk-1H-1) at (2,0) {};
@@
 
+		\draw[bend left=20] (Pk-2H-2) to (Pk-1H-1);
+		\draw[bend left=20] (Pk-1H-1) to (PkH);
+		\draw[bend left=20] (Pk+10) to (Pk+21);
+
+		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k-H}{0}\)}] (Pk-H0) at (0,0) {};
+		\node[label={center:\(\dots\)}] (dots10) at (-.5, 0) {};
+		\node[label={center:\(\dots\)}] (dots11) at (.5, 0) {};
+		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+H}{H-1}\)}] (Pk+HH-1) at (6,0) {};
+		\node[label={center:\(\dots\)}] (dots12) at (5.5, 0) {};
+		\node[label={center:\(\dots\)}] (dots13) at (6.5, 0) {};
 		% Second row
@@
 
+		\draw[bend left=20] (Pk-2H-3) to (Pk-1H-2);
+		\draw[bend left=20] (Pk-1H-2) to (PkH-1);
+		\draw[bend left=20] (PkH-1) to (Pk+1H);
+
+		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k-H}{H}\)}] (Pk-HH) at (0,1) {};
+		\node[label={center:\(\dots\)}] (dots20) at (-.5, 1) {};
+		\node[label={center:\(\dots\)}] (dots21) at (.5, 1) {};
+		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+H}{H-2}\)}] (Pk+HH-2) at (6,1) {};
+		\node[label={center:\(\dots\)}] (dots22) at (5.5, 1) {};
+		\node[label={center:\(\dots\)}] (dots23) at (6.5, 1) {};
+
 		% dots row
+		\node[label={center,rotate=90:\(\dots\)}] (dots1) at (0, 1.7) {};
 		\node[label={center,rotate=90:\(\dots\)}] (dots2) at (1, 1.7) {};
@@
 		\node[label={center,rotate=90:\(\dots\)}] (dots6) at (5, 1.7) {};
+		\node[label={center,rotate=90:\(\dots\)}] (dots7) at (6, 1.7) {};
 
@@
 
+		\draw[bend left=20] (Pk-10) to (Pk1);
+		\draw[bend left=20] (Pk1) to (Pk+12);
+		\draw[bend left=20] (Pk+12) to (Pk+23);
+
+		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-H}{2}\)}] (Pk-H2) at (0,2) {};
+		\node[label={center:\(\dots\)}] (dots30) at (-.5, 2) {};
+		\node[label={center:\(\dots\)}] (dots31) at (.5, 2) {};
+		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k+H}{0}\)}] (Pk+H0) at (6,2) {};
+		\node[label={center:\(\dots\)}] (dots32) at (5.5, 2) {};
+		\node[label={center:\(\dots\)}] (dots33) at (6.5, 2) {};
+
 		% Fourth row
@@
 
-		% Curved arrows
-		\draw[bend left=20] (Pk-2H-2) to (Pk-1H-1);
-		\draw[bend left=20] (Pk-1H-1) to (PkH);
-		\draw[bend left=20] (Pk+10) to (Pk+21);
-
-		\draw[bend left=20] (Pk-2H-3) to (Pk-1H-2);
-		\draw[bend left=20] (Pk-1H-2) to (PkH-1);
-		\draw[bend left=20] (PkH-1) to (Pk+1H);
-
-		\draw[bend left=20] (Pk-10) to (Pk1);
-		\draw[bend left=20] (Pk1) to (Pk+12);
-		\draw[bend left=20] (Pk+12) to (Pk+23);
-
-		\draw[bend left=20] (Pk-10) to (Pk1);
-		\draw[bend left=20] (Pk1) to (Pk+12);
-		\draw[bend left=20] (Pk+12) to (Pk+23);
-
 		\draw[bend left=20] (Pk-2H-1) to (Pk-1H);
@@
 		\draw[bend left=20] (Pk+11) to (Pk+22);
+
+		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-H}{1}\)}] (Pk-H1) at (0,3) {};
+		\node[label={center:\(\dots\)}] (dots40) at (-.5, 3) {};
+		\node[label={center:\(\dots\)}] (dots41) at (.5, 3) {};
+		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k+H}{H}\)}] (Pk+HH) at (6,3) {};
+		\node[label={center:\(\dots\)}] (dots42) at (5.5, 3) {};
+		\node[label={center:\(\dots\)}] (dots43) at (6.5, 3) {};
 

### 2025-04-18 stochastic_vehicle_dynamics.tex

@@
-\section{Stochastic vehicle dynamics}
+\section{TMP NAME}
+\label{sec:tmp}
+
+\subsection{Vehicle model}
+\label{sec:vehicle_model}
+The vehicle model employed in this study is a Single Track model featuring nonlinear axle characteristics.
+The lateral forces $Y_1$ and $Y_2$ are computed using a Pacejka's Magic Formula, which depends on the total vertical load acting on the axle and the theoretical lateral slip evaluated at the midpoint of the axle.
+The longitudinal forces $X_1$ and $X_2$ are provided as inputs to the system.
+Specifically, the first two components of the input vector $\bu$ correspond to acceleration and braking forces, while the third component represents the steering angle of the front wheels.
+This input formulation facilitates the distribution of braking forces between the front and rear axles, while assigning acceleration exclusively to the rear axle, consistent with the rear-wheel-drive configuration of the modeled vehicle.
+
+By treating the longitudinal forces as inputs, the longitudinal dynamics can be explicitly solved, allowing the computation of vertical load transfers prior to their use in the expression of the lateral forces.
+Consequently, the time derivative of the state becomes an explicit function of the state vector $\bx$ and the input vector $\bu$, as follows:
+\begin{equation}
+	\dbx(t) = f(\bx(t), \bu(t))
+\end{equation}
+
+The state vector $\bx$ has six components and includes the three kinematic quantities $u$, $v$, and $r$, along with the position and orientation of the barycentrical reference frame.
+As illustrated in Figure~\ref{fig:vehicle_model} $u$ and $v$ represent the longitudinal and lateral velocity of the CoM w.r.t its reference frame, while $r$ represents vehicle's yaw rate.
+
+\begin{figure}
+	\centering
+	\begin{tikzpicture}[scale=2, x = {(0,1cm)}, y = {(-1cm,0)},
+		force/.style={->, thick, red!60!black, >={Latex[length=10pt, width=4pt]}},
+		vec/.style={->, thick, mgreen, >={Latex[length=10pt, width=4pt]}},
+		aux/.style={thin},
+		frame/.style={->,thin,>={Latex}},
+		every node/.style={font=\small, text=black}
+		]
+		\definecolor{mgreen}{RGB}{0,100,0}
+		\begin{scope}[rotate = 25,shift = {(-.3,-2)}]
+			% Coordinates
+			\coordinate (rear) at (0,0);
+			\coordinate (G) at (1.2,0);
+			\coordinate (front) at (2.8,0);
+
+			% Main body line
+			\draw[blue!50!black, thick] (rear) -- (front);
+
+			% Rear wheel
+			\draw[fill=black!60!white, draw=black, rounded corners=2pt]
+			($(rear)+(-.2,.1)$) rectangle ($(rear)+(+.2,-.1)$);
+
+			% Front wheel (rotated around its center)
+			\begin{scope}[shift={(front)}, rotate=25]
+				\draw[fill=black!60!white, draw=black, rounded corners=2pt]
+				(-0.2,.1) rectangle (0.2,-.1);
+			\end{scope}
+
+			% Local axes at rear
+			\draw[force] (rear) -- ++(.5,0) node[above,right] {$X_2$};
+			\draw[force] (rear) -- ++(0,1.3) node[above] {$Y_2$};
+
+			% Local axes at front
+			\draw[force] (front) -- ++(-.3,0) node[right] {$X_1$};
+			\draw[force] (front) -- ++(0,1.5) node[above] {$Y_1$};
+
+			% Velocity vectors at G
+			\draw[vec] (G) -- ++(.6,0) node[left, shift={(-.2,-.1)}] {$u\,\mathbf{i}$};
+			\draw[vec] (G) -- ++(0,.3) node[above] {$v\,\mathbf{j}$};
+
+			% Angular velocity
+			\begin{scope}[rotate = -115]
+				\draw[vec] 
+				(-.1,.8) arc[start angle=0, end angle=50, radius=1]
+				node[midway, right=2pt] {$r\,\mathbf{k}$};
+			\end{scope}
+
+			% G point
+			\fill (G) circle (1pt) node[right] {$G$};
+
+			% Body frame
+			\draw[frame] (G) --++ (1,0) node[right] {$x_b$};
+			\draw[frame] (G) --++ (0,1) node[above] {$y_b$};
+
+			% Distances
+%			\draw[<->,>={Latex}] (3,-1) -- node[left] {$a_1$} (1.5,-1);
+%			\draw[<->,>={Latex}] (1.5,-1) -- node[left] {$a_2$} (0,-1);
+
+			% Steering angle (auxiliary vertical line)
+			\draw[aux] (front) -- ++(.4,0);
+			\draw[aux] (front) -- ++({0.4*cos(25)}, {0.4*sin(25)});
+
+			% Steering angle arc
+			\draw[->,thin,>={Latex}] ([shift={(0.35,0)}]front) arc[start angle=0, end angle=25, radius=0.35];
+			\node[left,shift={(.15,0)}] at ([shift={(0.35,0)}]front) {$\delta$};
+
+			% Psi angle aux
+			\draw[aux] (G) -- ++({0.4*cos(25)}, {-0.4*sin(25)});
+
+			% Psi angle
+			\draw[->,thin,>={Latex}] ([shift={({0.35*cos(25)},{-.35*sin(25)})}]G) arc[start angle=-25, end angle=0, radius=0.35];
+			\node[left,shift={(.2,-.05)}] at ([shift={({0.35*cos(25)},{-.35*sin(25)})}]G) {$\psi$};
+
+		\end{scope}
+		\coordinate (ground) at (0,0);
+		\draw[frame] (ground) --++ (1,0) node[right] {$x$};
+		\draw[frame] (ground) --++ (0,1) node[above] {$y$};
+		\draw[->,thick,>=Latex] (ground) --++ (G);
+%		\draw[->,thick,>=Latex] (ground) --++ (G) node[midway, shift={(0,-.3)}] {$\boldsymbol{p}$};
+
+	\end{tikzpicture}
+	\caption{Single Track model.
+The ground force acting on each axle is decomposed into longitudinal and lateral components, indicated by red arrows.
+The longitudinal components are part of the input vector $\bu$, encompassing both acceleration and braking forces, while the lateral components are computed using Pacejka's Magic Formula based on the vertical load and slip angle at each axle.
+The figure also depicts the fixed reference frame and the body reference frame, with the origin located at the vehicle's center of mass (CoM).
+The state variables $u$ and $v$ represent the longitudinal and lateral velocities of the CoM w.r.t.
+the body frame, while $r$ denotes the yaw rate.
+The angle $\psi$ is used to describe the orientation of the body frame w.r.t.
+the fixed frame.}
+	\label{fig:vehicle_model}
+\end{figure}
+
+\subsection{Stochastic vehicle dynamics}
 \label{sec:stochastic_vehicle_dynamics}

### 2025-04-22 closed_loop_planning.tex
[MM] Old paper sections deleted from closed loop

@@
 To be continued...
-\section{Tire Model}
-\label{sec:tyremodel}
-\begin{figure}[t]\centering
-	\tdplotsetmaincoords{0}{0}
-	\begin{tikzpicture}[tdplot_main_coords,scale=3.5]\small
-		\tdplotsetrotatedcoords{-166.5322}{-65.2292}{-166.5322}\draw[tdplot_rotated_coords,dashed,gray!75,fill=gray!10](-0.6,0,0)--(-0.6,1.1,0)--(0.6,1.1,0)--(0.6,0,0);\draw[tdplot_rotated_coords,very thick,fill=white](0,0)coordinate(C)..controls(0.25,0)and(0.5,0)..(0.5,0.45)arc(0:180:0.5)..controls(-0.5,0)and(-0.25,0)..(0,0);\draw[tdplot_rotated_coords,very thick,fill=gray!25](0,0.45)coordinate(O)circle(0.25);\draw[tdplot_rotated_coords,very thick,dashed](0.5,0.45)arc(0:-180:0.5);\draw[tdplot_rotated_coords,very thin,dashed](0,0,0)--(0,0.2,0)(0,0.7,0)--(0,0.95,0)(-0.5,0.45,0)--(-0.25,0.45,0)(0.25,0.45,0)--(0.5,0.45,0);\draw[tdplot_rotated_coords,very thin](0,0.2,0)--(0,0.7,0)(0,0.95,0)--(0,1.25,0)(-1,0.45,0)--(-0.5,0.45,0)(-0.25,0.45,0)--(0.25,0.45,0)(0.5,0.45,0)--(1,0.45,0);\draw[tdplot_rotated_coords,{Stealth[inset=0pt]}-{Stealth[inset=0pt]},very thin](0.75,0,0)--node[right]{$r$}(0.75,0.45,0);\tdplotsetrotatedcoordsorigin{(O)}\tdplotsetrotatedcoords{-166.5322}{-65.2292}{-146.5322}\draw[tdplot_rotated_coords,dashed](-0.25,0,0)--(-0.5,0,0);\draw[tdplot_rotated_coords,dashed](0,0.25,0)--(0,0.5,0);\draw[tdplot_rotated_coords,->](0,0,0)--(-0.25,0,0)(-0.5,0,0)--(-1,0,0)node[below left]{$x_h$};\draw[tdplot_rotated_coords,->](0,0,0)--(0,0.25,0)(0,0.5,0)--(0,0.8,0)node[above left]{$z_h$};\draw[tdplot_rotated_coords,->](0,0,0)--(0,0,0.8)node[right]{$y_h$};\draw[tdplot_rotated_coords,{Stealth[inset=0pt]}-{Stealth[inset=0pt]},very thin](-0.8,0)arc(180:160:0.8)node[midway,left]{$\sigma$};\tdplotsetrotatedcoordsorigin{(C)}\tdplotsetrotatedcoords{43.2192}{-27.9909}{-46.7808}\draw[tdplot_rotated_coords,dashed,gray!75,fill=gray!25](0,0,-0.6)--(0.6,0,-0.6)--(0.6,0,0.6)--(0,0,0.6);\draw[tdplot_rotated_coords,->](0,0,0)--(0.8,0,0)node[right]{$y_n$};\draw[tdplot_rotated_coords,->](0,0,0)node[circle,fill=black,inner sep=0pt,minimum size=3pt,draw]{}--(0,1.25,0)node[above]{$z_n$};\draw[tdplot_rotated_coords,->](0,0,-1)--(0,0,1)node[below left]{$x_n$};\draw[tdplot_rotated_coords,{Stealth[inset=0pt]}-{Stealth[inset=0pt]},very thin](0,1.1)arc(90:110:1.1)node[midway,above]{$\gamma$};\draw(0,-0.02)..controls(0,-0.3)and(0.3,-0.1)..(0.3,-0.3)node[below]{contact point};\draw(-0.5,0.7)..controls(-0.35,0.55)and(-0.6,0.7)..(-0.55,0.5)node[below,align=center]{wheel\\plane};\draw(0.4,0.15)..controls(0.3,0.25)and(0.55,0.2)..(0.5,0.3)node[above,align=center]{road\\plane};\draw(O)node[below right]{$O_h$}(C)node[below right]{$O_n$};\draw[tdplot_rotated_coords,thick,-{Stealth[inset=0pt]}](C)--+(0,0,0.8)node[below right]{$F_x$};\draw[tdplot_rotated_coords,thick,-{Stealth[inset=0pt]}](C)--+(0.5,0,0)node[above]{$F_y$};\draw[tdplot_rotated_coords,thick,-{Stealth[inset=0pt]}](C)--+(0,1,0)node[right]{$F_z$};
-	\end{tikzpicture}
-	\hfill
-	\begin{tikzpicture}[scale=4.25]\small
-		\coordinate(C)at(0,0);\coordinate(-)at(-0.375,0);\coordinate(+)at(0.325,0);\draw[->](-0.7,0)--(0.6,0)node[right]{$y_n$};\draw[->](0,0)node[circle,fill=black,inner sep=0pt,minimum size=3pt,draw]{}--(0,1.25)node[above]{$z_n$};\tdplotsetrotatedcoords{10}{0}{0}\draw[tdplot_rotated_coords,dash dot](0,-0.15)--(0,1.25)(-0.35,0.45)--(0.6,0.45);\draw[tdplot_rotated_coords,very thick](-0.25,0.2)--(-0.25,0.7)..controls(-0.25,0.95)and(-0.25,0.95)..(0,0.95)..controls(0.25,0.95)and(0.25,0.95)..(0.25,0.7)--(0.25,0.2)(-0.25,0.2)..controls(-0.25,0.025)and(-)..(C)..controls(+)and(0.25,-0.025)..(0.25,0.2);\draw[tdplot_rotated_coords,very thick,dashed](-0.25,0.2)..controls(-0.25,-0.05)and(-0.25,-0.05)..(0,-0.05)node(B){}..controls(0.25,-0.05)and(0.25,-0.05)..(0.25,0.2);\node[tdplot_rotated_coords](O)at(0,0.45){};\draw[very thin](O)--({(-0.55,0)}|-O);\draw[very thin,{Stealth[inset=0pt]}-{Stealth[inset=0pt]}](-0.475,0)--node[left]{$h$}({(-0.475,0)}|-O);\draw[very thin](B)--({(-0.7,0)}|-B);\draw[very thin,-{Stealth[inset=0pt]}](-0.625,0.3)--node[left]{$d$}(-0.625,0);\draw[very thin](-0.625,0)--({(-0.625,0)}|-B);\draw[very thin,-{Stealth[inset=0pt]}](-0.625,-0.15)--({(-0.625,0)}|-B);\draw[tdplot_rotated_coords,very thin](0.05,0)--(0.45,0);\draw[tdplot_rotated_coords,very thin](0.05,-0.05)--(0.6,-0.05);\draw[tdplot_rotated_coords,very thin,{Stealth[inset=0pt]}-{Stealth[inset=0pt]}](0.375,0)--node[right]{$r$}(0.375,0.45);\draw[tdplot_rotated_coords,very thin,{Stealth[inset=0pt]}-{Stealth[inset=0pt]}](0.525,-0.05)--node[right]{$r_0$}(0.525,0.45);\draw[very thin,{Stealth[inset=0pt]}-{Stealth[inset=0pt]}](0,1.125)arc(90:100:1.125)node[midway,above]{$\gamma$};\draw(0,-0.4);
-	\end{tikzpicture}
-	\caption{Decomposition of the tyre forces along the axes of the auxiliary frame $\{N\}$.
-The frame $\{N\}$ is such that: $x_n$ lies along the intersection between the wheel plane ($x_hz_h$-plane) and the road plane ($x_s y_s$-plane), pointing forward; $y_n$ points left (w.r.t.
-the driver) along the intersection line between the road plane ($x_s y_s$-plane) and the transverse plane ($y_h z_s$-plane) through the wheel center $O_h$; $z_n$ is normal to the road surface.
-The contact point is estimated in correspondence to the origin $O_n$.
-On the right, we report a detail of the tyre vertical deformation.}
-	\label{fig:tyre_wheel}
-\end{figure}
-
-The tangential forces exchanged between the tire and road are computed using \mbox{Pacejka's} \emph{Magic Formula}:
-\begin{equation}\label{eqn_mf}
-	[F_x,F_y]=\mf(\kappa,\alpha,F_z,\gamma),
-\end{equation}
-where $ F_x, F_y $ are the longitudinal and lateral forces, dependent on tire slips $( \kappa, \alpha )$, vertical load $( F_z )$, and camber angle $( \gamma )$.
-As shown in Fig.~\ref{fig:tyre_wheel}, an auxiliary frame $\sref{N}$ is introduced to calculate $ F_z $.
-This depends upon the interpenetration $ d $ of the tire into the ground, following a penalty-based compliant tire model.
-This model treats the wheel as a radial spring-damper system with constant stiffness $ k_t $ and damping $ c_t $.
-The vertical force is expressed as:
-\begin{equation}\label{eqn_Fz}
-	F_z= F_0\log_2\left(1 + 2^{\frac{k_t d + c_t \dot{d}}{F_0}}\right)
-\end{equation}
-which is differentiable for $ d = 0 $ and $\dot{d} = 0$, improving numerical stability.
-The model ensures $ F_z $ approaches zero when the tire loses contact with the ground, while maintaining a non-zero gradient to guide optimization algorithms in scenarios where the tire detaches from the ground or encounters sudden changes.
-
-%To compute the tangential components of the force exchanged by tyre and road, a formulation of the Pacejka's \emph{Magic Formula} ~\cite[Section~4.3.2]{Pacejka:book:2012} is used, which reads:
-%\begin{equation}\label{eqn_mf}
-%	[F_x,F_y]=\mf(\kappa,\alpha,F_z,\gamma).
-%\end{equation}
-%The Magic Formula $\mf$ computes the longitudinal and lateral forces $F_x,F_y$ as a function of the \emph{tyre slips} $\kappa,\alpha$~\cite[Section~1.2.1]{Pacejka:book:2012}.
-The formula is also sensitive to variations of vertical load $F_z$ and camber angle $\gamma$.
-%
-%The auxiliary frame $\sref{N}$ is introduced to compute quantities used in Eq.~\eqref{eqn_mf} and the (attempted) interpenetration $d$ (see Figure.~\ref{fig:tyre_wheel}) of the tyre in the ground, necessary to compute the dynamic vertical load $F_z$.
-The construction of the frame $\sref{N}$ is reported in Figure~\ref{fig:tyre_wheel}.
-It is worth noting that the frame $\sref{N}$ is fully described if $\sref{H}$ and $\sref{S}$ are known.
-%The model used for the contact is a unilateral, penalty-based compliant tyre model, that considers the wheel as a radial spring with constant stiffness $k_t$ and constant damping coefficient $c_t$\footnote{The introduction of a more refined model with nonlinear spring and damping characteristics would not pose particular difficulties.}.
-The traditional role of compression in a spring is conducted by $d$, which is computed considering the road and the lowermost point of the non-deformed tyre.
-The resultant function that relate $d$, $\dot{d}$ and $F_z$ is:
-%\begin{equation}\label{eqn_Fz}
-%	F_z= F_0\log_2\left(1 + 2^{\frac{k_t d + c_t \dot{d}}{F_0}}\right)
-%\end{equation}
-%This function, with $F_0$ tuned to be negligible, is differentiable for $d=0$ and $\dot{d}=0$, enhancing numerical performance, and behaves as a spring-damper system, as requested, when the tyre is in contact with the ground.
-When the tyre detaches completely from the ground, the resultant vertical forces tends to zero.
-However, since its gradient is not zero, this formulation helps the optimization algorithm to figure out proper search directions.
-This allows to deal with jumps and other situations in which the tyre may lose contact with the ground.
-
-\subsection{Camber System Modeling}
-\label{sec:cambersystemmodeling}
-The dynamic camber control (DCC) system is modeled as a simplified actuated revolute joint connecting the two virtual halves of the knuckle, though a real implementation might require a more complex 1-DoF joint.
-While this simplification may influence joint values from the optimizer, it results in nearly the same optimal wheel posture.
-The choice was driven by the limited literature on such systems for FSAE vehicles, which is the vehicle type employed in our analysis.
-A realistic model, if available, could be characterized using the same process applied to the suspension.
-During the design phase of the actual joint, some parameters could be treated as time-invariant optimization variables, constrained within dimensional bounds, to achieve the best performance configuration.
-%The dynamic camber control (DCC) system is modeled here as a simple actuated revolute joint connecting the two virtual halves of the knuckle.
-A real implementation could require a more complex generalized 1-DoF joint.
-Changing the joint characteristics might affect the joint values obtained from the optimizer but would result in approximately the same optimal wheel posture.
-This simplification was necessary due to the limited open literature on such systems for FSAE vehicles.
-If a realistic model of the system were available, the process used for the suspension could easily be repeated to characterize this joint.
-It is worth observing that, in the design phase of the actual controlled joint, some model parameters could be treated as time-invariant optimization variables, appropriately constrained to meet dimensional bounds, and optimized to find the best performance configuration.
-
-In Table~\ref{tab:camber_prop}, we report the range of variation of function $q_{hk}(t)$ and its first and second derivatives.
-Considering the scarcity of open literature available, reasonable values have been selected.
-In Figure~\ref{fig:camber_showing} the wheel is shown in a general posture with zero camber (shaded black) and with maximum positive camber (solid black).
-Due to the convention adopted for rotations, this posture corresponds to the minimum value of $q_{hk}$.
-\begin{table}[h]
-	\centering
-	\begin{tabular}{|c|c|c|}
-		\hline
-		$q_{hk}$ (rad) & $\dot{q}_{hk}$ (rad/s)& $\ddot{q}_{hk}$ (rad/s$^2$)\\\hline
-		$\pm \pi/20$ & $\pm \pi/20$ & $\pm \pi/10$ \\\hline
-	\end{tabular}
-	\caption{Upper and lower bounds for camber angle value $q_{hk}(t)$ and its derivatives.}
-	\label{tab:camber_prop}
-\end{table}
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[scale = .6]{Images/camber_showing}
-	\caption{Real scale representation of the wheel in two configurations: $q_{hk}=0$ (shaded surface) and $q_{hk}=-\pi/20$ (solid surface).}
-	\label{fig:camber_showing}
-\end{figure}
-\subsection{External Wrenches}
-\label{sec:extwren}
-The external wrenches acting on each body, from wheels to chassis, are expressed in body-fixed components.
-Only gravity, aerodynamic forces, and road-tyre interactions are considered.
-Following~\cite{Guiggiani:book:2023}, the aerodynamic wrench in frame $\{B\}$ is $W_a=-\frac{1}{2}\rho S v_{gb,1}^2 [C_x, 0, C_z, 0, -h_0 C_x - a_f C_{zf} + a_r C_{zr}, 0]^T$, where $\rho$ is air density, $S$ is the frontal area, $v_{gb,1}$ the forward velocity (component of \emph{distal rigid-body twist} $V_{gb}$, see~\cite[p.~4]{Domenighini:Designs:2023}), and $C_x, C_z$ are drag/lift coefficients.
-$C_z$ is split into $C_{zf}=k_a C_z$ and $C_{zr}=(1-k_a)C_z$ by the aerodynamic balance coefficient $k_a$.
-
-The total external wrench on a wheel is $W_{we}=\Ad_{g_{hn}}^*W_t+W_{w_h}$,
-%\begin{equation}\label{eqn:Wwe}
-%W_{we}=\Ad_{g_{hn}}^*W_t+W_{w_h},
-%\end{equation}
-where $W_t=[F_x, F_y, F_z, 0, 0, 0]^T$ is the ground wrench transferred from $\sref{N}$ to $\sref{H}$ by the \emph{starred} Adjoint operator, and $W_{w_h}=[F_{w_h}^T R_{hg}^T, 0, 0, 0]^T$ is the weight contribution with $F_{w_h}=[0, 0, -m_h g]^T$ in frame $\{G\}$, where $m_h$ is the wheel mass.
-
-Knuckles experience no external forces, therefore:
-\begin{equation}\label{eqn:Whe}
-	W_{he}=[0,0,0,0,0,0]^T\quad\textrm{and}\quad W_{ke}=[0,0,0,0,0,0]^T.
-\end{equation}
-The chassis's external wrench is $W_{be}=W_a+W_{w_b}$, %\begin{equation}\label{eqn:Wbe}
-%	W_{be}=W_a+W_{w_b},
-%\end{equation}
-where $W_a$ and $W_{w_b}$ are aerodynamic and weight contributions, respectively.
-For $W_{w_b}=[F_{w_b}^T R_{bg}^T, 0, 0, 0]^T$, the weight force is $F_{w_b}=[0, 0, -m_b g]^T$ in frame $\{G\}$, with $m_b$ as the chassis mass.
-%The purpose of this section is to describe the external wrenches acting on each body, from the wheels to the chassis, each expressed in body-fixed components.
-%Besides gravity, aerodynamic forces and the interactions between road and tyres are the only external forces considered.
-%Adopting the notation in~\cite{Guiggiani:book:2023}, the aerodynamic contribution to the wrenches in frame $\{B\}$ is $W_a=-\frac{1}{2}\rho Sv_{gb,1}^2[C_x,0,C_z,0,-h_0C_x-a_fC_{zf}+a_rC_{zr},0]^T$, where $\rho$ is the air density, $S$ the frontal area of the vehicle, $v_{gb,1}$ the forward velocity, $h_0$ the nominal height of the CoM from ground and $a_f,a_r$ the distances of the CoM from the front and rear axle, respectively.
-The drag and lift coefficient $C_x$ and $C_z$ are dimensionless parameters that depend on the shape of the vehicle's body.
-$C_z$ can be further partitioned into $C_{zf}=k_a C_z$ and $C_{zr}=(1-k_a) C_z$ according to the \emph{aerodynamic balance coefficient} $k_a$, which is a parameter of the vehicle set-up.
-%
-%The total external wrench acting on a (general) wheel is:
-%\begin{equation}\label{eqn:Wwe}
-%	W_{we}=\Ad_{g_{hn}}^*W_t+W_{w_h},
-%\end{equation}
-%where $W_t=[F_x,F_y,F_z,0,0,0]^T$ are the ground forces computed in $\sref{N}$ and properly transferred to $\sref{H}$ components by the \emph{starred} Adjoint operator, while $W_{w_h} = [ F_{w_h}^T R^T_{hg}, 0,0,0 ]^T$, with $F_{w_h}=[0,0,-m_h g]^T$ in frame $\{G\}$, represents the weight contribution and $m_{h}$ is the wheel mass.
-%
-%Under our hypotheses, the knuckles are not subjected to external forces, thus \begin{equation}\label{eqn:Whe}
-%	W_{he}=[0,0,0,0,0,0]^T\quad\textrm{and}\quad W_{ke}=[0,0,0,0,0,0]^T.
-%\end{equation}
-%
-%The resultant external wrench acting on the chassis is therefore:
-%\begin{equation}\label{eqn:Wbe}
-%	W_{be}=W_a+W_{w_b},
-%\end{equation}
-%where $W_a$ and $W_{w_b}$ are, respectively, the contributions of aerodynamic and chassis weight forces, whose general expression is of the form $W_{w_b} = [ F_{w_b}^T R^T_{bg}, 0,0,0 ]^T$.
-Explicitly we have $F_{w_b}=[0,0,-m_b g]^T$ in frame $\{G\}$, where $m_b$ is the mass of the chassis.
-
-\subsection{Recursive computation of the dynamics equations with ABA}
-\label{sec:aba}
-Thanks to the ABA algorithm, assuming known joint positions and velocities and joint forces and torques, it is possible to compute recursively and efficiently the accelerations of the joint variables.
-Following~\cite{Domenighini:Designs:2023}, twists are streamlined as follows: $V_{gb}^b = V_{gb} = V_b$, $V_{bh}^h=V_{bh}$, $V_{hw}^h=V_{hw}$.
-For the second model (DCC) the last expression is replaced with $V_{kw}^k=V_{kw}$ and $V_{hk}^k=V_{hk}$ is added.
-Given that the first model (without camber control) is a particular case of the DCC one, to avoid duplications only the algorithm for DCC is reported here.
-
-The traditional ABA requires knowledge of the active forces and torques, denoted as $\tau_{ij}$, exerted by the parent body ($i$) on the child body ($j$).
-Its purpose is to solve the \emph{Forward Dynamics} problem of an articulated systems of rigid bodies.
-In our floating-base system, given the twist $V_b$ of the parent body $\{B\}$, the values and time derivatives of all joint variables ($q$ and $\dot{q}$), and the torque vector ($\tau$), the objective is to compute the joint accelerations ($\dot{V}_{b}$ and $\ddot{q}$).
-%The fundamental concept is that of the inertia of an articulated body.
-This is represented by that of the root rigid body (original parent) plus the portion of the sub-trees, composed by the children bodies, whose inertia that can be structurally transferred backwards through the joints.
-%This depends on the architecture, the pose and the joints' types.
-The overall computation is performed in three sweeps defined recursively.
-One of the greatest computational benefits when it is employed to assemble symbolic dynamic equations, as in our case, is that \emph{no matrix inversion is required}, thus contributing to streamline the algebraic expressions.
-
-The first step of the algorithm, as described in \textbf{Step 1}, performs the forward propagation of rigid-body velocities: from the root to the leaves of the tree.
-%Twists are calculated as the sum of parent twist and the relative twist which is expressed consistently with the joint type.
-The interesting Jacobians are listed here: $J_{hk}=[0,0,0,1,0,0]^T$ and $J_{kw}=[0,0,0,0,1,0]^T$, and represent two revolute joints with axes $x$ and $y$ respectively.
-In Algorithm~\ref{alg:step1} the terms ``Knuckle 1'' and ``Knuckle 2'' are used to identify the two halves of the knuckle, where, using kinematic trees terminology, the first is the parent of the second.
-The Adjoint operator is defined as follows:
-\begin{equation}\label{eq:Adjoint} \Ad_{g}=\left[\begin{array}{cc}R&\hat{d}R\\0&R\end{array}\right].
-\end{equation}
-In the case $g_{bh}\in SE(3)$ is considered, $R_{bh}\in SO(3)$ and $d_{bh}\in \bbR^3$ are functions of $(z,\delta)$.
-In the case of $g_{hk}$, the expressions are simply $d_{hk}=[0,0,0]^T$ and $R = \rotX(q_{hk})$, the latter being an elementary rotation about the common $x$-axis of the angle $q_{hk}$.
-
-\begin{algorithm}[h]
-	\floatname{algorithm}{Step}
-	\caption{Forward Propagation of Velocity}\label{alg:step1}
-	\begin{algorithmic}[1]
-		\vspace{1mm}
-		\For{$h=h_1,h_2,h_3,h_4$}
-		\vspace{1mm}
-		\State{$V_{bh}=J_{bh,z}\dot{z}+J_{bh,\delta}\dot{\delta}$}
-		\vspace{1mm}
-		\State{$V_h=\Ad_{g_{bh}}^{-1}V_b+V_{bh}$}\Comment{Knuckle 1 Rigid-Body Velocity}
-		\vspace{1mm}
-		\State{$V_{hk} = J_{hk}\dot{q}_{hk}$}
-		\vspace{1mm}
-		\State{$V_k=\Ad_{g_{hk}}^{-1}V_h+V_{hk}$}\Comment{Knuckle 2 Rigid-Body Velocity}
-		\vspace{1mm}
-		\State{$V_{kw}=J_{kw}\omega$}
-		\vspace{1mm}
-		\State{$V_w=V_k+V_{kw}$}\Comment{Rim Rigid-Body Velocity}
-		\vspace{1mm}
-		\EndFor
-		\vspace{1mm}
-	\end{algorithmic}
-\end{algorithm}
-
-The second step of the algorithm, as detailed in \textbf{Step 2}, performs the backward propagation of inertia and bias terms.
-\begin{algorithm}[h]
-	\floatname{algorithm}{Step}
-	\caption{Backward Propagation of Articulated Inertia and Bias}\label{alg:step2}
-	\begin{algorithmic}[1]
-		\vspace{1mm}
-		\For{$h=h_1,h_2,h_3,h_4$}
-		\vspace{1mm}
-		\State{$\hat{M}_w=M_w$}\Comment{Rim Articulated Inertia}
-		\vspace{1mm}
-		\State{$\hat{B}_w=-W_{we}+\ad_{V_w}^*M_wV_w$}\Comment{Rim Articulated Bias}
-		\vspace{1mm}
-		\State{$\bar{M}_w=\hat{M}_w-\dfrac{\hat{M}_wJ_{hw}J_{hw}^T\hat{M}_w}{J_{hw}^T\hat{M}_wJ_{hw}}$}
-		\vspace{1mm}
-		\State{$\bar{B}_w=\hat{B}_w-\dfrac{\hat{M}_wJ_{hw}J_{hw}^T\hat{B}_w}{J_{hw}^T\hat{M}_wJ_{hw}}-\bar{M}_w\ad_{V_{hw}}V_w$}
-		\vspace{1mm}
-		\State{$\hat{M}_k= \bar{M}_w$
-		}\Comment{Knuckle 2 Articulated Inertia}
-		\vspace{1mm}
-		\State{$\hat{B}_k= \bar{B}_w$
-		}\Comment{Knuckle 2 Articulated Bias}
-		\vspace{1mm}
-		\State{$\hat{M}_h= \Ad_{g_{hk}}^*\hat{M}_k\Ad_{g_{hk}}^{-1}$
-		}\Comment{Knuckle 1 Articulated Inertia}
-		\vspace{1mm}
-		\State{$\hat{B}_h= \Ad_{g_{hk}}^*\bigl(\hat{M}_k(J_{hk}\ddot{q}_{hk}-\ad_{V_{hk}}\Ad_{g_{kh}}V_k)+\hat{B}_k\bigr)$
-		}\Comment{Knuckle 1 Articulated Bias}
-		\vspace{1mm}
-		\State{$\bar{M}_h=\hat{M}_h-\dfrac{\hat{M}_hJ_{bh,z}J_{bh,z}^T\hat{M}_h}{J_{bh,z}^T\hat{M}_hJ_{bh,z}}$}
-		\vspace{1mm}
-		\State{$\bar{B}_h=\hat{B}_h-\dfrac{\hat{M}_hJ_{bh,z}(J_{bh,z}^T\hat{B}_h-\tau)}{J_{bh,z}^T\hat{M}_hJ_{bh,z}}-\bar{M}_h(\ad_{V_{bh}}V_h-\dot{J}_{bh,z}\dot{z}-\dot{J}_{bh,\delta}\dot{\delta}-J_{bh,\delta}\ddot{\delta})$}
-		\vspace{1mm}
-		\EndFor
-		\vspace{1mm}
-		\State{$\hat{M}_b=M_b+\displaystyle\sum_h\Ad_{g_{bh}}^*\bar{M}_h\Ad_{g_{bh}}^{-1}$}\Comment{Chassis Articulated Inertia}
-		\vspace{1mm}
-		\State{$\hat{B}_b=-W_{be}+\ad_{V_b}^*M_bV_b+\displaystyle\sum_h\Ad_{g_{bh}}^*\bar{B}_h$}\Comment{Chassis Articulated Bias}
-		\vspace{1mm}
-	\end{algorithmic}
-\end{algorithm}
-
-%Both inertia and bias are expressed in the body-fixed components.
-In the case the DCC architecture is considered, the camber angle acceleration $\ddot{q}_{hk}$ is known, since $q_{hk(t)}$ is the assumed input, while the torque $\tau_{hk}$ has to be computed together with the rest of unknown accelerations, which are the outcome of the third sweep.
-This leads to a \emph{Hybrid Dynamics} problem that requires to carefully revisit the ABA's second and third steps.
-The approach used is inspired by~\cite[Section~2.3]{Kim:notes:2012} where the the notation and conventions here employed were adapted according to~\cite{Domenighini:Designs:2023}.
-%Specifically, apart from the labels assigned to the quantities, the primary difference between the two methods is the ordering of the twist vector.
-In this work, the twist vector is defined as $V = [v,\omega]^T$, whereas in the other method, it is defined as $V=[\omega,v]^T$.
-This difference might seem minor but leads to significant changes in the definitions of Adjoint operators, inertia matrices, bias and wrench vectors.
-
-%For a generic articulated body the Newton-Euler equations can be expressed in the form:
-%\begin{equation}\label{eqn:NE}
-%	W=\hat{M}\dot{V}+\hat{B}.
-%\end{equation}
-%For a leaf node of the kinematic tree, having no children, the articulated inertia $\hat{M}$ is simply the generalized inertia $M$ of the body itself. Consistently, the bias $\hat{B}=B$, where:
-%\begin{equation}
-%	B=-W_e+\ad_V^*MV,
-%\end{equation}
-%and where $W_e$ is the resultant external wrench applied to the body and $\ad_V^*MV$ accounts for generalized gyroscopic and centrifugal forces, which are bi-linear in $V$.
-According to~\cite[Section~4.3.3]{Murray:book:1994} $\ad^*=-\ad^T$ has been defined.
-(If $v$ and $\omega$ are the translational and rotational component of $V$, then $\ad^*_V=\left[\begin{smallmatrix}\hat{\omega}&0\\\hat{v}&\hat{\omega}\end{smallmatrix}\right]$).
-
-Starting from the leaf nodes, which in our case are represented by the wheels ($\sref{W}$), after providing information such as external wrenches, twists, joint geometry, body masses, joint forces / torques for traditional connections, and time evolution of variables defining kinematically imposed joints, the backward propagation of inertia and bias can be performed.
-
-Finally the last step, denoted as \textbf{Step 3}, starting from the root node, computes joint variable accelerations and consequently body twists derivatives.
-For the kinematically controlled camber joints, the torques $\tau_{hk}$ needed to maintain the desired state of motion (DCC architecture) are computed as a by-product of the local inverse dynamics.
-This information is crucial to prevent actuators from  saturating or operating outside their bandwidth in the planned trajectories.
-
-The root twist acceleration is determined by solving equation $W=\hat{M}\dot{V}+\hat{B}$, considering that the chassis is free-floating.
-The other unknown quantities, such as variable accelerations and torques, are computed projecting the Newton-Euler equations of each body along the respective Jacobian and then solving for the variable.
-See also~\cite{Domenighini:Designs:2023} for more details in the derivation of the equations.
-
-\begin{algorithm}[h]
-	\floatname{algorithm}{Step}
-	\caption{Forward Propagation of Acceleration}\label{alg:step3}
-	\begin{algorithmic}[1]
-		\State{$\dot{V}_b=-\hat{M}_b^{-1}\hat{B}_b$}\Comment{Chassis Rigid-Body Acceleration}
-		\vspace{1mm}
-		\For{$h=h_1,h_2,h_3,h_4$}
-		\vspace{1mm}
-		\State{$\ddot{z}=-\dfrac{J_{bh,z}^T\bigl(\hat{B}_h+\hat{M}_h(\Ad_{g_{bh}}^{-1}\dot{V}_b-\ad_{V_{bh}}V_h+\dot{J}_{bh,z}\dot{z}+\dot{J}_{bh,\delta}\dot{\delta}+J_{bh,\delta}\ddot{\delta})\bigr)-\tau}{J_{bh,z}^T\hat{M}_hJ_{bh,z}}$}
-		\vspace{1mm}
-		\State{$\dot{V}_{bh}=\dot{J}_{bh,z}\dot{z}+\dot{J}_{bh,\delta}\dot{\delta}+J_{bh,z}\ddot{z}+J_{bh,\delta}\ddot{\delta}$}
-		\vspace{1mm}
-		\State{$\dot{V}_h=\Ad_{g_{bh}}^{-1}\dot{V}_b-\ad_{V_{bh}}V_h+\dot{V}_{bh}$}\Comment{Knuckle 1 Rigid-Body Acceleration}
-		\vspace{1mm}
-		\State{$\tau_{hk} = J_{hk}^T\bigl(\hat{B}_k+\hat{M}_k(\Ad_{g_{kh}}\dot{V}_h-\ad_{V_{hk}}V_k + J_{hk}\ddot{q}_{hk})\bigr)$}
-		\vspace{1mm}
-		\State{$\dot{V}_{bh} = J_{hk}\ddot{q}_{hk}$}
-		\vspace{1mm}
-		\State{$\dot{V}_k=\Ad_{g_{hk}}^{-1}\dot{V}_h-\ad_{V_{hk}}V_k+\dot{V}_{bh}$}\Comment{Knuckle 2 Rigid-Body Acceleration}
-		\State{$\dot{\omega}=-\dfrac{J_{kw}^T\bigl(\hat{B}_w+\hat{M}_w(\dot{V}_k-\ad_{V_{kw}}V_w)\bigr)}{J_{kw}^T\hat{M}_wJ_{kw}}$}
-		\vspace{1mm}
-		\EndFor
-		\vspace{1mm}
-	\end{algorithmic}
-\end{algorithm}
-
-\subsection{Dynamic Model}
-The dynamics equations, assembled via the procedure described  in Sec.~\ref{sec:aba}, are implemented in the software code by using a state-space representation $\dot{x}=f(x,u)$.
-%\begin{equation}\label{eqn:f}
-%	\dot{x}=f(x,u).
-%\end{equation}
-The state vector is $x = (q_{gb};V_{gb};z;\dot{z};\omega;q_{hk};\dot{q}_{hk})\in\bbR^{32}$, with components defined in previous sections.
-As already mentioned, in the model the camber joint is kinematically controlled as per the DCC architecture.
-Hence, the associated control vector is $u=(T_a;T_b;\delta;\ddot{q}_{hk})\in\bbR^{16}$, where each element is in $\bbR^{4}$ and collects the controls of the branch to each wheel.
-The torque is split into acceleration and braking components, allowing for minor modifications to model on-board motors coupled with traditional brakes and/or in-wheel motors as well.
-These components have their reaction torques in two different bodies in case of on-board motors and traditional brakes: the chassis and the knuckles, respectively.
-The $\delta$ vector collects the four translations of the steering rod attachment points, as described in Section~\ref{sec:suspanal}, while for the vector $\ddot{q}_{hk}$ no further explanation should be necessary.
-Clearly, this representation is valid for the model with active camber.
-The simpler model (no DCC) shares with the former the first 24 components of the state vector and the first 12 of the control vector.
-
-The time derivative of the state vector has also been expressed previously except for the first element, $q_{gb}$.
-Its derivative is a function of the state and can be recovered using a standard procedure~\cite{Domenighini:Designs:2023} as follows:
-\begin{equation}\label{eqn:dotqgb}
-	\dot{q}_{gb}=J_{gb}^{-1}(q_{gb})V_{gb},
-\end{equation}
-where $J_{gb}(q_{gb})\in\bbR^{6\times6}$ is the Jacobian of the virtual joint connecting the ground to the chassis, and it is computed via the PoE parameterization of the transformation $g_{gb}$.
-%, similarly to $J_{bh}$ in Eq.~\eqref{eqn:Jbh}.

### 2025-04-22 open_loop_planning.tex
[MM] merged figure versions

@@
 At each grid point, two versions of the covariance matrix are highlighted: $\bP^0_k$ (red node), which is the version to be initialized, and $\bP^H_k$ (green node), which is the most propagated version used to formulate the robust constraints in Eq.~\eqref{eq:DOCPconstraintsopenloop}.
-
+\def\version{marco}
+\ifdefstring{\version}{matteo}{
 \begin{figure}
@@
 \end{figure}
+}{false}
+
+\ifdefstring{\version}{marco}{
+\begin{figure}
+	\centering
+	%\begin{tikzpicture}[%
+	%		smallnode/.style={%
+		%			circle, draw, minimum size=3mm, inner sep=0pt
+		%		},
+	%		r_smallnode/.style={%
+		%			circle, draw, minimum size=3mm, inner sep=0pt, red
+		%		},
+	%		g_smallnode/.style={%
+		%			circle, draw, minimum size=3mm, inner sep=0pt, mygreen
+		%		},
+	%		every path/.style={->, thick},
+	%		x=1.5cm, y=1.5cm
+	%		]
+	%		
+	%		\newcommand{\Plab}[2]{\bP^{#2}_{#1}}
+	%		\definecolor{mygreen}{RGB}{0 150 0}
+	%		
+	%		% Nodes
+	%		% First row
+	%		\node[smallnode, label={[label distance=-2pt]above:\protect\(\Plab{k-2}{H-2}\)}] (Pk-2H-2) at (1,0) {};
+	%		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{H-1}\)}] (Pk-1H-1) at (2,0) {};
+	%		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k}{H}\)}] (PkH) at (3,0) {};
+	%		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{0}\)}] (Pk+10) at (4,0) {};
+	%		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{1}\)}] (Pk+21) at (5,0) {};
+	%		
+	%		% Second row
+	%		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-2}{H-3}\)}] (Pk-2H-3) at (1,1) {};
+	%		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{H-2}\)}] (Pk-1H-2) at (2,1) {};
+	%		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k}{H-1}\)}] (PkH-1) at (3,1) {};
+	%		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{H}\)}] (Pk+1H) at (4,1) {};
+	%		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{0}\)}] (Pk+20) at (5,1) {};
+	%		
+	%		% dots row
+	%		\node[label={center,rotate=90:\(\dots\)}] (dots2) at (1, 1.7) {};
+	%		\node[label={center,rotate=90:\(\dots\)}] (dots3) at (2, 1.7) {};
+	%		\node[label={center,rotate=90:\(\dots\)}] (dots4) at (3, 1.7) {};
+	%		\node[label={center,rotate=90:\(\dots\)}] (dots5) at (4, 1.7) {};
+	%		\node[label={center,rotate=90:\(\dots\)}] (dots6) at (5, 1.7) {};
+	%		
+	%		% Third row
+	%		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k-2}{H}\)}] (Pk-2H) at (1,2) {};
+	%		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{0}\)}] (Pk-10) at (2,2) {};
+	%		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k}{1}\)}] (Pk1) at (3,2) {};
+	%		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{2}\)}] (Pk+12) at (4,2) {};
+	%		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{3}\)}] (Pk+23) at (5,2) {};
+	%		
+	%		% Fourth row
+	%		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-2}{H-1}\)}] (Pk-2H-1) at (1,3) {};
+	%		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{H}\)}] (Pk-1H) at (2,3) {};
+	%		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k}{0}\)}] (Pk0) at (3,3) {};
+	%		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{1}\)}] (Pk+11) at (4,3) {};
+	%		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{2}\)}] (Pk+22) at (5,3) {};
+	%		
+	%		% Curved arrows
+	%		\draw[bend left=20] (Pk-2H-2) to (Pk-1H-1);
+	%		\draw[bend left=20] (Pk-1H-1) to (PkH);
+	%		\draw[bend left=20] (Pk+10) to (Pk+21);
+	%		
+	%		\draw[bend left=20] (Pk-2H-3) to (Pk-1H-2);
+	%		\draw[bend left=20] (Pk-1H-2) to (PkH-1);
+	%		\draw[bend left=20] (PkH-1) to (Pk+1H);
+	%		
+	%		\draw[bend left=20] (Pk-10) to (Pk1);
+	%		\draw[bend left=20] (Pk1) to (Pk+12);
+	%		\draw[bend left=20] (Pk+12) to (Pk+23);
+	%		
+	%		\draw[bend left=20] (Pk-10) to (Pk1);
+	%		\draw[bend left=20] (Pk1) to (Pk+12);
+	%		\draw[bend left=20] (Pk+12) to (Pk+23);
+	%		
+	%		\draw[bend left=20] (Pk-2H-1) to (Pk-1H);
+	%		\draw[bend left=20] (Pk0) to (Pk+11);
+	%		\draw[bend left=20] (Pk+11) to (Pk+22);
+	%		
+	%		% Dashed rectangle
+	%		\draw[dashed, thick, rounded corners] (2.6,-.25) rectangle (3.4,3.5);
+	%		
+	%	\end{tikzpicture}
+\definecolor{mygreen}{RGB}{0,150,0}
+%\begin{tikzpicture}[
+%    smallnode/.style={circle, draw, minimum size=3mm, inner sep=0pt},
+%    r_smallnode/.style={circle, draw=red, minimum size=3mm, inner sep=0pt},
+%    g_smallnode/.style={circle, draw=mygreen, minimum size=3mm, inner sep=0pt},
+%    every path/.style={->, thick},
+%    x=1.4cm, y=1.2cm
+%]
+%
+%% Row j=0 (y = 3)
+%\node[r_smallnode, label={[label distance=-2pt]above:\(\bP_0^0\)}] (P00) at (1,3) {};
+%\node[smallnode, label={[label distance=-2pt]above:\(\bP_1^1\)}] (P11) at (2,3) {};
+%\node[smallnode, label={[label distance=-2pt]above:\(\bP_2^2\)}] (P22) at (3,3) {};
+%\node[g_smallnode, label={[label distance=-2pt]above:\(\bP_3^3\)}] (P33) at (4,3) {};
+%\node[smallnode, label={[label distance=-2pt]above:\(\bP_4^4\)}] (P44) at (5,3) {};
+%\node at (6,3) {\(\cdots\)};
+%
+%% Row j=1 (y = 2)
+%\node[r_smallnode, label={[label distance=-2pt]above:\(\bP_1^0\)}] (P10) at (2,2) {};
+%\node[smallnode, label={[label distance=-2pt]above:\(\bP_2^1\)}] (P21) at (3,2) {};
+%\node[smallnode, label={[label distance=-2pt]above:\(\bP_3^2\)}] (P32) at (4,2) {};
+%\node[g_smallnode, label={[label distance=-2pt]above:\(\bP_4^3\)}] (P43) at (5,2) {};
+%\node at (6,2) {\(\cdots\)};
+%
+%% Row j=2 (y = 1)
+%\node[r_smallnode, label={[label distance=-2pt]above:\(\bP_2^0\)}] (P20) at (3,1) {};
+%\node[smallnode, label={[label distance=-2pt]above:\(\bP_3^1\)}] (P31) at (4,1) {};
+%\node[smallnode, label={[label distance=-2pt]above:\(\bP_4^2\)}] (P42) at (5,1) {};
+%\node[g_smallnode, label={[label distance=-2pt]above:\(\bP_5^3\)}] (P53) at (6,1) {};
+%\node at (7,1) {\(\cdots\)};
+%
+%% Row j=3 (y = 0)
+%\node[r_smallnode, label={[label distance=-2pt]above:\(\bP_3^0\)}] (P30) at (4,0) {};
+%\node[smallnode, label={[label distance=-2pt]above:\(\bP_4^1\)}] (P41) at (5,0) {};
+%\node[smallnode, label={[label distance=-2pt]above:\(\bP_5^2\)}] (P52) at (6,0) {};
+%\node[g_smallnode, label={[label distance=-2pt]above:\(\bP_6^3\)}] (P63) at (7,0) {};
+%\node at (8,0) {\(\cdots\)};
+%
+%% Arrows: P_k^j --> P_{k+1}^{j+1}
+%\draw (P00) -- (P11);
+%\draw (P11) -- (P22);
+%\draw (P22) -- (P33);
+%\draw (P33) -- (P44);
+%
+%\draw (P10) -- (P21);
+%\draw (P21) -- (P32);
+%\draw (P32) -- (P43);
+%
+%\draw (P20) -- (P31);
+%\draw (P31) -- (P42);
+%\draw (P42) -- (P53);
+%
+%\draw (P30) -- (P41);
+%\draw (P41) -- (P52);
+%\draw (P52) -- (P63);
+%
+%% Dashed box around column k = 3 (x=4)
+%\draw[dashed, thick, rounded corners] (3.6,-0.5) rectangle (4.4,3.5);
+%
+%% Optional top axis labels
+%\node at (1, 4.1) {\scriptsize \(k=0\)};
+%\node at (2, 4.1) {\scriptsize \(k=1\)};
+%\node at (3, 4.1) {\scriptsize \(k=2\)};
+%\node at (4, 4.1) {\scriptsize \(k=3\)};
+%\node at (5, 4.1) {\scriptsize \(k=4\)};
+%\node at (6, 4.1) {\scriptsize \(k=5\)};
+%\node at (7, 4.1) {\scriptsize \(k=6\)};
+%
+%\end{tikzpicture}
+
+\begin{tikzpicture}[
+	smallnode/.style={circle, draw, minimum size=3mm, inner sep=0pt},
+	r_smallnode/.style={circle, draw=red, minimum size=3mm, inner sep=0pt},
+	g_smallnode/.style={circle, draw=mygreen, minimum size=3mm, inner sep=0pt},
+	r_hatnode/.style={
+		circle, draw=red, pattern=north east lines, pattern color=red,
+		minimum size=3mm, inner sep=0pt},
+	g_hatnode/.style={
+		circle, draw=mygreen, pattern=north east lines, pattern color=mygreen,
+		minimum size=3mm, inner sep=0pt},
+	every path/.style={->, thick},
+	x=1.4cm, y=1.2cm
+	]
+
+	% ROW j = 0 (y=3)
+	\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k-3}^0\)}] (P00) at (1,3) {};
+	\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k-2}^1\)}] (P11) at (2,3) {};
+	\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k-1}^2\)}] (P22) at (3,3) {};
+	\node[g_hatnode, label={[label distance=-2pt]above:\(\bP_k^3\)}] (P33) at (4,3) {};
+	\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k+1}^0\)}] (P44) at (5,3) {};
+	\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+2}^1\)}] (P55) at (6,3) {};
+
+	% ROW j = 1 (y=2)
+	\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k-2}^0\)}] (P10) at (2,2) {};
+	\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k-1}^1\)}] (P21) at (3,2) {};
+	\node[smallnode, label={[label distance=-2pt]above:\(\bP_k^2\)}] (P32) at (4,2) {};
+	\node[g_hatnode, label={[label distance=-2pt]above:\(\bP_{k+1}^3\)}] (P43) at (5,2) {};
+	\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k+2}^0\)}] (P54) at (6,2) {};
+	\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+3}^1\)}] (P65) at (7,2) {};
+
+	% ROW j = 2 (y=1)
+	\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k-1}^0\)}] (P20) at (3,1) {};
+	\node[smallnode, label={[label distance=-2pt]above:\(\bP_k^1\)}] (P31) at (4,1) {};
+	\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+1}^2\)}] (P42) at (5,1) {};
+	\node[g_hatnode, label={[label distance=-2pt]above:\(\bP_{k+2}^3\)}] (P53) at (6,1) {};
+	\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k+3}^0\)}] (P64) at (7,1) {};
+	\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+4}^1\)}] (P75) at (8,1) {};
+
+	% ROW j = 3 (y=0)
+	\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_k^0\)}] (P30) at (4,0) {};
+	\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+1}^1\)}] (P41) at (5,0) {};
+	\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+2}^2\)}] (P52) at (6,0) {};
+	\node[g_hatnode, label={[label distance=-2pt]above:\(\bP_{k+3}^3\)}] (P63) at (7,0) {};
+	\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k+4}^0\)}] (P74) at (8,0) {};
+	\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+5}^1\)}] (P85) at (9,0) {};
+
+	% ARROWS
+	\draw (P00) -- (P11);
+	\draw (P11) -- (P22);
+	\draw (P22) -- (P33);
+	\draw (P22) -- (P33);
+
+	\draw (P44) -- (P55);
+	\draw[dashed,->] (P55) -- ++(1,0);
+
+	\draw (P10) -- (P21);
+	\draw (P21) -- (P32);
+	\draw (P32) -- (P43);
+	\draw (P54) -- (P65);
+	\draw[dashed,->] (P65) -- ++(1,0);
+
+	\draw (P20) -- (P31);
+	\draw (P31) -- (P42);
+	\draw (P42) -- (P53);
+	\draw (P64) -- (P75);
+	\draw[dashed,->] (P75) -- ++(1,0);
+
+	\draw (P30) -- (P41);
+	\draw (P41) -- (P52);
+	\draw (P52) -- (P63);
+
+	\draw (P74) -- (P85);
+	\draw[dashed,->] (P85) -- ++(1,0);
+
+	% Dashed box around column k (x=4)
+	\draw[dashed, thick, rounded corners] (3.6,-0.5) rectangle (4.4,3.7);
+
+	% Top axis labels
+	\node at (0, 4.3) {\scriptsize \(\cdots\)};
+	\node at (1, 4.3) {\scriptsize \(k-3\)};
+	\node at (2, 4.3) {\scriptsize \(k-2\)};
+	\node at (3, 4.3) {\scriptsize \(k-1\)};
+	\node (knode) at (4, 4.3) {\scriptsize \(k\)};
+	\node at (5, 4.3) {\scriptsize \(k+1\)};
+	\node at (6, 4.3) {\scriptsize \(k+2\)};
+	\node at (7, 4.3) {\scriptsize \(k+3\)};
+	\node at (8, 4.3) {\scriptsize \(k+4\)};
+	\node at (9, 4.3) {\scriptsize \(\cdots\)};
+
+	% Wiggly arrow from label to top edge of dashed box
+	\draw[->, thick, decorate, decoration={snake, amplitude=0.25mm, segment length=2.5mm}]
+	(knode.south) -- (4, 3.8);
+
+	% Legend box (top right)
+	%\begin{scope}[shift={(6.5,2.5)}]
+	%    \node[r_smallnode] (r) at (0,0) {};
+	%    \node[anchor=west] at (0.4,0) {\scriptsize Initialized matrix \(\bP_k^0\)};
+	%
+	%    \node[g_smallnode] (g) at (0,-0.8) {};
+	%    \node[anchor=west] at (0.4,-0.8) {\scriptsize Most propagated matrix \(\bP_k^3\)};
+	%\end{scope}
+
+\end{tikzpicture}
+
+\caption{Cartoon representation (depicted for $H=3$) of the continuity and initialization equations associated with the $H+1$ propagated instances of the covariance matrix introduced in problem~\eqref{eq:DOCPopenloop}.
+The dashed rectangle highlights the $k$-th discretization step, at which the robust constraints are enforced.
+Each node corresponds to a specific covariance matrix, where the subscript denotes the current grid index and the superscript indicates the number of propagation steps since initialization.
+At each step, two key instances are emphasized: the matrix to be initialized at that step (red node), and the matrix that has been propagated over $H$ steps (green node), which is used in the evaluation of the constraint backoff.}
+%\caption{Schematic representation of the continuity and initialization Equations for the $H+1$ versions of the covariance matrix introduced in the discretized OCP.
+The dashed rectangle indicate the $k$-th step, in which the constraints are formulated.
+Each node represents a covariance matrix, labeled such that the subscript denotes the current grid step, while the superscript indicates the number of propagation steps it has undergone.
+At each step, two nodes are highlighted: the covariance matrix to be initialized (red node) and the matrix that has been propagated for $H$ steps (green node), which is used to formulate the robust constraints.}
+\label{fig:DOCPgrid}
+\end{figure}
+}{}

### 2025-04-22 closed_loop_planning.tex
[MGa] more on closed-loop

@@
 
-The second approach plans over the full horizon while incorporating a feedback policy that mimics the driver's closed-loop behavior.
+The second approach we propose plans over the full horizon while incorporating a feedback policy that mimics the driver's closed-loop behavior.
 This enables a more realistic propagation of the uncertainty encoded in the covariance matrix $\bP$, preventing its uncontrolled growth through the stabilizing effect of the driver's feedback.
 
-The approach is based on the three following steps: i) A nominal time-optimal trajectory is planned: this will be referred to as nominal feed-forward optimal trajectory.
-This is the trajectory a driver should follow in the ideal situation of a world without uncertainty.
-In this step the mean trajectory is the actual trajectory a deterministic vehicle dynamics would trace.
+The approach is based on the three following steps: (i) A nominal time-optimal trajectory is planned: this will be referred to as nominal feed-forward optimal trajectory.
+This represents the trajectory that an expert driver would follow in an ideal scenario without uncertainty.
+In this step, the mean trajectory coincides with the actual path that the vehicle would follow under deterministic dynamics.
 
-ii) A closed-loop controller mimicking the driver's action is computed which stabilizes the nominal feed-forward optimal trajectory computed at step i).
+(ii) A closed-loop controller mimicking the driver's action is computed which stabilizes the nominal feed-forward optimal trajectory computed at step (i).
 We assume a discrete time-varying Linear Quadratic Regulator (LQR) controller is a good approximation of how an expert driver can track a prescribed trajectory.
 
-iii) In the last step the stochastic framework is reintroduced and the time-optimal planning problem is robustified using the Lyapunov framework.
+(iii) In the last step the stochastic framework is reintroduced and the time-optimal planning problem is robustified using the Lyapunov framework.
 The main idea is to \emph{re-plan} a mean trajectory of a stochastic system such that, a time-varying LQR controller that aims at stabilizing it, is able to properly tame the propagation of the uncertainty.
 This amounts to ensuring that the robustified path constraints are satisfied.
-More specifically, this accounts for two key aspects: firstly, that the covariance matrix is propagated accounting explicitly for the controlled system dynamics; and secondly, that the path constraints are robustly satisfied by incorporating the covariance estimation propagated through the closed-loop controlled system.
+More specifically, this considers two key aspects: firstly, that the covariance matrix is propagated accounting explicitly for the controlled system dynamics; secondly, that the path constraints are robustly satisfied by incorporating the covariance estimation propagated through the closed-loop controlled system.
 Some aspects that may appear a bit technical are in fact essential.
@@
 \label{sec:nominalFF}
-The first step consists in planning a nominal feed-forward trajectory.
-This assumes deterministic system dynamics, with \( (\hbmu_k, \hbu_k) \) denoting the differential states and controls, and \( \hbz_k \) the algebraic states, for \( k = 0, \ldots, N \).
+Step (i) consists in planning a nominal feed-forward trajectory.
+This assumes deterministic system dynamics, where $\hbmu_k$, $\hbu_k$, and $\hbz_k$ denote differential states, controls, and algebraic states, respectively, for \( k = 0, \ldots, N \).
 To avoid redundancy, we note that the NLP to be solved is a special case of~\eqref{eq:DOCP}, where~\eqref{eq:DOCPdP},~\eqref{eq:DOCPdPIC}, and the backoff terms are omitted, since the goal is to compute a nominal trajectory.
@@
 
-Given the nominal trajectory $(\hbmu_k, \hbu_k)$ over $k=0,\ldots, N$, and introduced the deviation variables $\bbmu_k=\bmu_k - \hbmu_k$ and $\bbu = \bu_k - \hbu_k$, we design a linear state-feedback law $\bbu_k= -\hbK_k \bbmu_k$, or equivalently $\bu_k = \hbu_k - \hbK_k (\bmu_k - \hbmu_k)$, to mitigate disturbances while tracking the nominal trajectory.
+Given the nominal trajectory $(\hbmu_k, \hbu_k)$ over $k=0,\ldots, N$, planned at step (i), and introduced the deviation variables $\bbmu_k=\bmu_k - \hbmu_k$ and $\bbu = \bu_k - \hbu_k$, we design a linear state-feedback law $\bbu_k= -\hbK_k \bbmu_k$, or equivalently $\bu_k = \hbu_k - \hbK_k (\bmu_k - \hbmu_k)$, to mitigate disturbances while tracking the nominal trajectory.
+
 To compute the state-feedback matrices $\hbK_k$, we linearize the system around $(\hbmu_k, \hbu_k)$, such that $\bbmu_{k+1} \approx \hbA_{k}\bbmu_k + \hbB_k \bbu_k$, where $\hbA_k = f_{,\bmu}(\hbmu_k,\hbu_k)$ and $\hbB_k = f_{,\bu}(\hbmu_k,\hbu_k)$, and define the quadratic regulator (tracking) cost function
@@
 The optimal (stabilizing) feedback policy is given by $\bbu_k=-\hbK_k \bbmu_k$.
-We assume that the gain sequence $\{\hbK_k\}$, $k=0,\ldots, N$ represents the control strategy that an expert driver would employ to optimally stabilize the system along the nominal trajectory.
+We assume that the gain sequence $\{\hbK_k\}$, $k=0,\ldots, N$ represents a reasonable approximation of the control strategy that an expert driver would adopt to optimally stabilize the system along the nominal trajectory.
 
@@
 As the third and last step, a closed-loop robustified planning problem is set up.
-Its aim is the definition of a mean trajectory of a stochastic system such that, working in tandem with state-feedback controller that aims at stabilizing it, it is able to mitigate the perturbations and reduce the propagation of uncertainty.
+Its aim is the definition of a mean trajectory of a stochastic system such that, working in tandem with state-feedback controller that aims at stabilizing it, is able to mitigate the perturbations and reduce the propagation of uncertainty.
 This helps to guarantee that the robustified path constraints can be satisfied over the entire planning horizon.
@@
 0    & \geq  h_i(\bmu_k, \bu_k, \bz_k) + \be_i(\bmu_k, \bu_k, \tbP_k, \bz_k), \quad k = 1,\ldots, N;\; \quad i \in \calI \label{eq:ROCPconstraints}\\
--\de \bK   & \leq \bK_k - \hbK_k \leq  \de \bK \label{eq:ROCPdeltaK}, \quad k = 0,\ldots, N
+-\de \bK_k   & \leq \bK_k - \hbK_k \leq  \de \bK_k \label{eq:ROCPdeltaK}, \quad k = 0,\ldots, N
 \end{alignat}
 \end{subequations}
-Here, \eqref{eq:ROCPdP} denotes the discretized form for time evolution of the covariance matrix under a stabilizing controller of the form $\bbu_k=-\bK_k \bbmu_k$.
-To mark the difference with~\eqref{eq:OCPdP}, a new symbol $\tbP$ is introduced to the denote the covariance matrix in this case.
+Here,~\eqref{eq:ROCPdP} denotes the discrete-time covariance dynamics under the stabilizing state-feedback controller \( \bbu_k = -\bK_k \bbmu_k \).
+To distinguish this case from~\eqref{eq:DOCP}, we use the symbol \( \tbP \) to denote the closed-loop (CL) covariance matrix, whose discrete evolution corresponds to a collocation-based approximation of the continuous-time dynamics $\dot{\tbP}(t) = \tbA(t) \tbP(t) + \tbP(t)\tbA(t)^T + \bQ(t)$, with $\tbA(t)=\bA(t)-\bB(t)\bK(t)$.
+It is worth noting that, accordingly, in equation~\eqref{eq:ROCPconstraints} the backoff term $\be_i$ is a function of closed-loop uncertainty $\tbP_k$.
+Another key aspect is that the state-feedback matrix $\bK_k$ is also treated as a decision variable.
+This reflects the fact that the driver may adjust his action to stabilize the new (to-be-planned) nominal trajectory using a policy that differs from the previous $\hbK_k$, computed in step (ii) via LQR.
+To promote convergence, however, $\bK_k$ is restrained within prescribed bounds ($\de \bK_k$) relative to $\hbK_k$.
+In our tests $\de \bK_k = 0.1 \hbK_k$.
+
+%To mark the difference with~\eqref{eq:OCPdP}, besides using P,CL to denote closed-loop dynamics, a new symbol $\tbP$ is introduced to the denote the covariance matrix in this case.
 In continuous time, $\tbP$ evolves according to $\dot{\tbP}(t) = (\bA(t)-\bB(t) \bK(t)) \tbP(t) + \tbP(t) (\bA(t)-\bB(t) \bK(t))^T + \bQ(t)$.

### 2025-04-22 open_loop_planning.tex
[MM] robust adherence constraint

--- 2025-04-18 open_loop_planning.tex
@@
 \end{figure}
+}{}
+
+\ifdefstring{\version}{marco}{
+\begin{figure}
+	\centering
+	\definecolor{mygreen}{RGB}{0,150,0}
+	\begin{tikzpicture}[
+		smallnode/.style={circle, draw, minimum size=3mm, inner sep=0pt},
+		r_smallnode/.style={circle, draw=red, minimum size=3mm, inner sep=0pt},
+		g_smallnode/.style={circle, draw=mygreen, minimum size=3mm, inner sep=0pt},
+		r_hatnode/.style={
+			circle, draw=red, pattern=north east lines, pattern color=red,
+			minimum size=3mm, inner sep=0pt},
+		g_hatnode/.style={
+			circle, draw=mygreen, pattern=north east lines, pattern color=mygreen,
+			minimum size=3mm, inner sep=0pt},
+		every path/.style={->, thick},
+		x=1.4cm, y=1.2cm
+		]
+
+		% ROW j = 0 (y=3)
+		\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k-3}^0\)}] (P00) at (1,3) {};
+		\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k-2}^1\)}] (P11) at (2,3) {};
+		\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k-1}^2\)}] (P22) at (3,3) {};
+		\node[g_hatnode, label={[label distance=-2pt]above:\(\bP_k^3\)}] (P33) at (4,3) {};
+		\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k+1}^0\)}] (P44) at (5,3) {};
+		\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+2}^1\)}] (P55) at (6,3) {};
+
+		% ROW j = 1 (y=2)
+		\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k-2}^0\)}] (P10) at (2,2) {};
+		\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k-1}^1\)}] (P21) at (3,2) {};
+		\node[smallnode, label={[label distance=-2pt]above:\(\bP_k^2\)}] (P32) at (4,2) {};
+		\node[g_hatnode, label={[label distance=-2pt]above:\(\bP_{k+1}^3\)}] (P43) at (5,2) {};
+		\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k+2}^0\)}] (P54) at (6,2) {};
+		\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+3}^1\)}] (P65) at (7,2) {};
+
+		% ROW j = 2 (y=1)
+		\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k-1}^0\)}] (P20) at (3,1) {};
+		\node[smallnode, label={[label distance=-2pt]above:\(\bP_k^1\)}] (P31) at (4,1) {};
+		\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+1}^2\)}] (P42) at (5,1) {};
+		\node[g_hatnode, label={[label distance=-2pt]above:\(\bP_{k+2}^3\)}] (P53) at (6,1) {};
+		\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k+3}^0\)}] (P64) at (7,1) {};
+		\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+4}^1\)}] (P75) at (8,1) {};
+
+		% ROW j = 3 (y=0)
+		\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_k^0\)}] (P30) at (4,0) {};
+		\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+1}^1\)}] (P41) at (5,0) {};
+		\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+2}^2\)}] (P52) at (6,0) {};
+		\node[g_hatnode, label={[label distance=-2pt]above:\(\bP_{k+3}^3\)}] (P63) at (7,0) {};
+		\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k+4}^0\)}] (P74) at (8,0) {};
+		\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+5}^1\)}] (P85) at (9,0) {};
+
+		% ARROWS
+		\draw (P00) -- (P11);
+		\draw (P11) -- (P22);
+		\draw (P22) -- (P33);
+		\draw (P22) -- (P33);
+
+		\draw (P44) -- (P55);
+		\draw[dashed,->] (P55) -- ++(1,0);
+
+		\draw (P10) -- (P21);
+		\draw (P21) -- (P32);
+		\draw (P32) -- (P43);
+		\draw (P54) -- (P65);
+		\draw[dashed,->] (P65) -- ++(1,0);
+
+		\draw (P20) -- (P31);
+		\draw (P31) -- (P42);
+		\draw (P42) -- (P53);
+		\draw (P64) -- (P75);
+		\draw[dashed,->] (P75) -- ++(1,0);
+
+		\draw (P30) -- (P41);
+		\draw (P41) -- (P52);
+		\draw (P52) -- (P63);
+
+		\draw (P74) -- (P85);
+		\draw[dashed,->] (P85) -- ++(1,0);
+
+		% Dashed box around column k (x=4)
+		\draw[dashed, thick, rounded corners] (3.6,-0.5) rectangle (4.4,3.7);
+
+		% Top axis labels
+		\node at (0, 4.3) {\scriptsize \(\cdots\)};
+		\node at (1, 4.3) {\scriptsize \(k-3\)};
+		\node at (2, 4.3) {\scriptsize \(k-2\)};
+		\node at (3, 4.3) {\scriptsize \(k-1\)};
+		\node (knode) at (4, 4.3) {\scriptsize \(k\)};
+		\node at (5, 4.3) {\scriptsize \(k+1\)};
+		\node at (6, 4.3) {\scriptsize \(k+2\)};
+		\node at (7, 4.3) {\scriptsize \(k+3\)};
+		\node at (8, 4.3) {\scriptsize \(k+4\)};
+		\node at (9, 4.3) {\scriptsize \(\cdots\)};
+
+		% Wiggly arrow from label to top edge of dashed box
+		\draw[->, thick, decorate, decoration={snake, amplitude=0.25mm, segment length=2.5mm}]
+		(knode.south) -- (4, 3.8);
+
+		% Legend box (top right)
+		%\begin{scope}[shift={(6.5,2.5)}]
+		%    \node[r_smallnode] (r) at (0,0) {};
+		%    \node[anchor=west] at (0.4,0) {\scriptsize Initialized matrix \(\bP_k^0\)};
+		%
+		%    \node[g_smallnode] (g) at (0,-0.8) {};
+		%    \node[anchor=west] at (0.4,-0.8) {\scriptsize Most propagated matrix \(\bP_k^3\)};
+		%\end{scope}
+
+	\end{tikzpicture}
+
+	\caption{Cartoon representation (depicted for $H=3$) of the continuity and initialization equations associated with the $H+1$ propagated instances of the covariance matrix introduced in problem~\eqref{eq:DOCPopenloop}.
+The dashed rectangle highlights the $k$-th discretization step, at which the robust constraints are enforced.
+Each node corresponds to a specific covariance matrix, where the subscript denotes the current grid index and the superscript indicates the number of propagation steps since initialization.
+At each step, two key instances are emphasized: the matrix to be initialized at that step (red node), and the matrix that has been propagated over $H$ steps (green node), which is used in the evaluation of the constraint backoff.}
+	%\caption{Schematic representation of the continuity and initialization Equations for the $H+1$ versions of the covariance matrix introduced in the discretized OCP.
+The dashed rectangle indicate the $k$-th step, in which the constraints are formulated.
+Each node represents a covariance matrix, labeled such that the subscript denotes the current grid step, while the superscript indicates the number of propagation steps it has undergone.
+At each step, two nodes are highlighted: the covariance matrix to be initialized (red node) and the matrix that has been propagated for $H$ steps (green node), which is used to formulate the robust constraints.}
+	\label{fig:DOCPgrid}
+\end{figure}
+}{}
+
+\subsection{Robust adherence limit constraint formulation}
+The model employs a tire formulation that neglects combined slip effects, i.e., it does not account for the simultaneous utilization of longitudinal and lateral tyre forces.
+As a result, the optimal control problem must include an additional constraint to ensure that the total ground reaction forces remain within the bounds of the tyre's adherence ellipse.
+
+The base version of the constraint for each axle can be expressed as a function of $\bx$ and $\bu$, as follows:
+\begin{equation}
+	h(\bx,\bu) = \left[ \left( \frac{X(\bx,\bu)}{\mu_x} \right)^2+ \left( \frac{Y(\bx,\bu)}{\mu_y}\right)^2\right]\frac{1}{Z^2(\bx,\bu)}-1\leq0
+\end{equation}
+Here, $X$ and $Y$ denote the longitudinal and lateral components of the in-plane ground forces, respectively, as illustrated in Figure~\ref{fig:vehicle_model}, while $Z$ represents the total vertical load acting on the axle.
+
+The gradient of the constraint w.r.t.
+the state vector $\bx$, denoted by $\na_{\bx}h(\bmu)$, is used to evaluate the standard deviation $\sigma$ associated with the constraint.
+This gradient can be computed using the chain rule, as follows:
+\begin{equation}
+	\na_{\bx}h(\bmu,\bu) = 
+	\frac{\pd h}{\pd X}\at_{\bmu,\bu}\na_{\bx}X(\bmu,\bu) + 
+	\frac{\pd h}{\pd Y}\at_{\bmu,\bu}\na_{\bx}Y(\bmu,\bu) + 
+	\frac{\pd h}{\pd Z}\at_{\bmu,\bu}\na_{\bx}Z(\bmu,\bu)
+\end{equation}
+where the first element is zero because $X$ is a function of the input vector $\bu$ only.
+
+For example, introducing a back-off term $\be=0.1$ in the constraint implies that the axle's adherence capacity can be exploited up to 90\% of its limit.
+This allows the optimizer to determine the most appropriate trade-off between the longitudinal force $X$ and the lateral force $Y$ --- often by reducing both components to some extent --- based on the specific requirements of the manoeuvre.

### 2025-04-28 open_loop_planning.tex
[MM] position change for adherence constraint description

@@
 }{}
-
-\subsection{Robust adherence limit constraint formulation}
-The model employs a tire formulation that neglects combined slip effects, i.e., it does not account for the simultaneous utilization of longitudinal and lateral tyre forces.
-As a result, the optimal control problem must include an additional constraint to ensure that the total ground reaction forces remain within the bounds of the tyre's adherence ellipse.
-
-The base version of the constraint for each axle can be expressed as a function of $\bx$ and $\bu$, as follows:
-\begin{equation}
-	h(\bx,\bu) = \left[ \left( \frac{X(\bx,\bu)}{\mu_x} \right)^2+ \left( \frac{Y(\bx,\bu)}{\mu_y}\right)^2\right]\frac{1}{Z^2(\bx,\bu)}-1\leq0
-\end{equation}
-Here, $X$ and $Y$ denote the longitudinal and lateral components of the in-plane ground forces, respectively, as illustrated in Figure~\ref{fig:vehicle_model}, while $Z$ represents the total vertical load acting on the axle.
-
-The gradient of the constraint w.r.t.
-the state vector $\bx$, denoted by $\na_{\bx}h(\bmu)$, is used to evaluate the standard deviation $\sigma$ associated with the constraint.
-This gradient can be computed using the chain rule, as follows:
-\begin{equation}
-	\na_{\bx}h(\bmu,\bu) = 
-	\frac{\pd h}{\pd X}\at_{\bmu,\bu}\na_{\bx}X(\bmu,\bu) + 
-	\frac{\pd h}{\pd Y}\at_{\bmu,\bu}\na_{\bx}Y(\bmu,\bu) + 
-	\frac{\pd h}{\pd Z}\at_{\bmu,\bu}\na_{\bx}Z(\bmu,\bu)
-\end{equation}
-where the first element is zero because $X$ is a function of the input vector $\bu$ only.
-
-For example, introducing a back-off term $\be=0.1$ in the constraint implies that the axle's adherence capacity can be exploited up to 90\% of its limit.
-This allows the optimizer to determine the most appropriate trade-off between the longitudinal force $X$ and the lateral force $Y$ --- often by reducing both components to some extent --- based on the specific requirements of the manoeuvre.

### 2025-04-28 stochastic_vehicle_dynamics.tex
[MM] position change for adherence constraint description

@@
 In all our tests direct collocation with cubic polynomial state representations and Gauss-Legendre collocation points provided accurate results.
+
+\subsection{Robust adherence limit constraint formulation}
+The model employs a tire formulation that neglects combined slip effects, i.e., it does not account for the simultaneous utilization of longitudinal and lateral tyre forces.
+As a result, the optimal control problem must include an additional constraint to ensure that the total ground reaction forces remain within the bounds of the tyre's adherence ellipse.
+
+The base version of the constraint for each axle can be expressed as a function of $\bx$ and $\bu$, as follows:
+\begin{equation}
+	h^\textrm{A}_j(\bx,\bu) = \left[ \left( \frac{X_j(\bx,\bu)}{\mu_{x,j}} \right)^2+ \left( \frac{Y_j(\bx,\bu)}{\mu_{y,j}}\right)^2\right]\frac{1}{Z_j^2(\bx,\bu)}-1\leq0\qquad j=1,2.\label{eq:adherence_constraint}
+\end{equation}
+where the superscript A denotes adherence, and the subscript $j$ indicates the axle index, with $j=1$ referring to the front axle and $j=2$ to the rear axle.
+Here, $X_j$ and $Y_j$ denote the longitudinal and lateral components of the in-plane ground forces, respectively, as illustrated in Figure~\ref{fig:vehicle_model}, while $Z_j$ represents the total vertical load acting on the axle.
+As can be seen from~\eqref{eq:adherence_constraint}, the points $\left(X_j,Y_j\right)$ are constrained to lie within an ellipse whose semi-axes are given by $\mu_{x,j}Z_j$ and $\mu_{y,j}Z_j$.
+
+The gradient of the constraint w.r.t.
+the state vector $\bx$, denoted by $\na_{\bx}h^\textrm{A}_j(\bmu)$, is used to evaluate the standard deviation $\sigma^\textrm{A}_j$ associated with the constraint.
+This gradient can be computed using the chain rule, as follows:
+\begin{equation}
+	\na_{\bx}h^\textrm{A}_j(\bmu,\bu) = 
+	\frac{\pd h^\textrm{A}_j}{\pd X_j}\at_{\bmu,\bu}\na_{\bx}X_j(\bmu,\bu) + 
+	\frac{\pd h^\textrm{A}_j}{\pd Y_j}\at_{\bmu,\bu}\na_{\bx}Y_j(\bmu,\bu) + 
+	\frac{\pd h^\textrm{A}_j}{\pd Z_j}\at_{\bmu,\bu}\na_{\bx}Z_j(\bmu,\bu)
+\end{equation}
+where the first element is zero because $X_j$ is a function of the input vector $\bu$ only.
+
+For example, introducing a back-off term $\be=0.1$ in the constraint implies that the axle's adherence capacity can be exploited up to 90\% of its limit.
+This allows the optimizer to determine the most appropriate trade-off between the longitudinal force $X$ and the lateral force $Y$ --- often by reducing both components to some extent --- based on the specific requirements of the manoeuvre.

### 2025-04-28 stochastic_vehicle_dynamics.tex
[MM] adherence constraint adjustments

@@
 As can be seen from~\eqref{eq:adherence_constraint}, the points $\left(X_j,Y_j\right)$ are constrained to lie within an ellipse whose semi-axes are given by $\mu_{x,j}Z_j$ and $\mu_{y,j}Z_j$.
+The constraint, without the back-off term, is represented in the left panel of Figure~\ref{fig:robust_constraints} by the solid line ellipse in the plane $X_j-Y_j$.
 
-The gradient of the constraint w.r.t.
+To better clarify the effect of the back-off term on the constraint we can suppose to have a combination of $\bx$, $\bu$, $P$ and $\ga^\textrm{A}_j$ so that $\be^\textrm{A}_j=0.2$.
+This implies that the semi-axes of the ellipse are reduced by a factor $\sqrt{1-\be^\textrm{A}_j} = \sqrt{0.8}$, as shown in the left panel Figure~\ref{fig:robust_constraints}, where the ellipse in dashed line delimits the available region with the back-off applied.
+This allows the optimizer to determine the most appropriate trade-off between the longitudinal force $X_j$ and the lateral force $Y_j$ --- often by reducing both components to some extent --- based on the specific requirements of the manoeuvre.
+
+%The gradient of the constraint w.r.t.
 the state vector $\bx$, denoted by $\na_{\bx}h^\textrm{A}_j(\bmu)$, is used to evaluate the standard deviation $\sigma^\textrm{A}_j$ associated with the constraint.
 This gradient can be computed using the chain rule, as follows:
-\begin{equation}
-	\na_{\bx}h^\textrm{A}_j(\bmu,\bu) = 
-	\frac{\pd h^\textrm{A}_j}{\pd X_j}\at_{\bmu,\bu}\na_{\bx}X_j(\bmu,\bu) + 
-	\frac{\pd h^\textrm{A}_j}{\pd Y_j}\at_{\bmu,\bu}\na_{\bx}Y_j(\bmu,\bu) + 
-	\frac{\pd h^\textrm{A}_j}{\pd Z_j}\at_{\bmu,\bu}\na_{\bx}Z_j(\bmu,\bu)
-\end{equation}
-where the first element is zero because $X_j$ is a function of the input vector $\bu$ only.
+%\begin{equation}
+%	\na_{\bx}h^\textrm{A}_j(\bmu,\bu) = 
+%	\frac{\pd h^\textrm{A}_j}{\pd X_j}\at_{\bmu,\bu}\na_{\bx}X_j(\bmu,\bu) + 
+%	\frac{\pd h^\textrm{A}_j}{\pd Y_j}\at_{\bmu,\bu}\na_{\bx}Y_j(\bmu,\bu) + 
+%	\frac{\pd h^\textrm{A}_j}{\pd Z_j}\at_{\bmu,\bu}\na_{\bx}Z_j(\bmu,\bu)
+%\end{equation}
+%where the first element is zero because $X_j$ is a function of the input vector $\bu$ only.
 
-For example, introducing a back-off term $\be=0.1$ in the constraint implies that the axle's adherence capacity can be exploited up to 90\% of its limit.
-This allows the optimizer to determine the most appropriate trade-off between the longitudinal force $X$ and the lateral force $Y$ --- often by reducing both components to some extent --- based on the specific requirements of the manoeuvre.
+\begin{figure}
+	\centering
+	\begin{tikzpicture}[scale = .7]
+		% Parameters
+		\def\outerX{3}   % x radius of outer ellipse
+		\def\outerY{2}   % y radius of outer ellipse
+		\def\gap{0.2}    % distance between outer and inner ellipse
+		\def\innerX{2.68}
+		\def\innerY{1.79}
+
+		% Draw outer ellipse
+		\draw[] (0,0) ellipse [x radius=\outerX, y radius=\outerY];
+
+		% Draw inner ellipse filled with color
+		\fill[black!30] (0,0) ellipse [
+		x radius={\innerX},
+		y radius={\innerY}
+		];
+		\draw[dashed] (0,0) ellipse [
+		x radius={\innerX},
+		y radius={\innerY}
+		];
+
+		% Draw axes
+		\draw[->] (-\outerX-1,0) -- (\outerX+1,0) node[right] {$X_j$};
+		\draw[->] (0,-\outerY-1) -- (0,\outerY+1) node[above] {$Y_j$};
+
+		% Arrows pointing to the ellipses
+		\draw[->,thin] (1.7,2) -- ({\outerX*cos(60)},{\outerY*sin(60)}) node[midway, above right] {$\beta=0$};
+		\draw[->,thin] (2.3,1.7) -- ({(\innerX)*cos(45)},{(\innerY)*sin(45)}) node[midway, right, shift={(0.05,.2)}] {$\beta=\gap$};
+
+		% Arrows pointing to the ellipses
+		\draw[->,thin] ({\outerX+.3},.3) -- (\outerX,0) node[midway, above right,shift={(.05,0)}] {$\mu_x Z_j$};
+		\draw[->,thin] (.3,{\outerY+.3}) -- (0,\outerY) node[midway, above right,shift={(.05,0)}] {$\mu_y Z_j$};
+
+	\end{tikzpicture}
+	\hfill
+	\begin{tikzpicture}[scale = .7]
+		% Parameters
+		\def\outerX{3}   % x radius of outer ellipse
+		\def\outerY{2}   % y radius of outer ellipse
+		\def\gap{0.2}    % distance between outer and inner ellipse
+		\def\innerX{2.68}
+		\def\innerY{1.79}
+
+		% Draw outer ellipse
+		\draw[] (0,0) ellipse [x radius=\outerX, y radius=\outerY];
+
+		% Draw inner ellipse filled with color
+		\fill[black!30] (0,0) ellipse [
+		x radius={\innerX},
+		y radius={\innerY}
+		];
+		\node[above] at (0,0) {PATH CONSTRAINT FIG};
+		\draw[dashed] (0,0) ellipse [
+		x radius={\innerX},
+		y radius={\innerY}
+		];
+
+		% Draw axes
+		\draw[->] (-\outerX-1,0) -- (\outerX+1,0) node[right] {$X_i$};
+		\draw[->] (0,-\outerY-1) -- (0,\outerY+1) node[above] {$Y_i$};
+
+		% Arrows pointing to the ellipses
+		\draw[->,thin] (1.7,2) -- ({\outerX*cos(60)},{\outerY*sin(60)}) node[midway, above right] {$\beta=0$};
+		\draw[->,thin] (2.3,1.7) -- ({(\innerX)*cos(45)},{(\innerY)*sin(45)}) node[midway, right, shift={(0.05,.2)}] {$\beta=\gap$};
+
+		% Arrows pointing to the ellipses
+		\draw[->,thin] ({\outerX+.3},.3) -- (\outerX,0) node[midway, above right,shift={(.05,0)}] {$\mu_x Z_i$};
+		\draw[->,thin] (.3,{\outerY+.3}) -- (0,\outerY) node[midway, above right,shift={(.05,0)}] {$\mu_y Z_i$};
+
+	\end{tikzpicture}
+	\caption{Graphical representation of the two constraints analyzed in the section.
+The left panel shows the adherence constraint in the plane $X_j-Y_j$.
+The ellipse in solid line, whose semi-axes are given by $\mu_{x,j}Z_j$ and $\mu_{y,j}Z_j$, represents the border of the available region for the base constraint, without back-off.
+The ellipse in dashed line delimits the available region when a back-off $\be^\textrm{A}_j=0.2$ is present.
+This ellipse has its semi-axes reduced by a factor $\sqrt{1-\be^\textrm{A}_j}$ w.r.t.
+the one in solid line.
+	}
+	\label{fig:robust_constraints}
+\end{figure}

### 2025-04-28 Fig/TikZ_Sources/tikz_adherence_constraint.tex
[MM] tikz sources folder created

--- 2025-04-18 stochastic_vehicle_dynamics.tex
@@
-\section{TMP NAME}
-\label{sec:tmp}
+\begin{tikzpicture}[scale = .7]
+	% Parameters
+	\def\outerX{3}   % x radius of outer ellipse
+	\def\outerY{2}   % y radius of outer ellipse
+	\def\gap{0.2}    % distance between outer and inner ellipse
+	\def\innerX{2.68}
+	\def\innerY{1.79}
 
-\subsection{Vehicle model}
-\label{sec:vehicle_model}
-The vehicle model employed in this study is a Single Track model featuring nonlinear axle characteristics.
-The lateral forces $Y_1$ and $Y_2$ are computed using a Pacejka's Magic Formula, which depends on the total vertical load acting on the axle and the theoretical lateral slip evaluated at the midpoint of the axle.
-The longitudinal forces $X_1$ and $X_2$ are provided as inputs to the system.
-Specifically, the first two components of the input vector $\bu$ correspond to acceleration and braking forces, while the third component represents the steering angle of the front wheels.
-This input formulation facilitates the distribution of braking forces between the front and rear axles, while assigning acceleration exclusively to the rear axle, consistent with the rear-wheel-drive configuration of the modeled vehicle.
+	% Draw outer ellipse
+	\draw[] (0,0) ellipse [x radius=\outerX, y radius=\outerY];
 
-By treating the longitudinal forces as inputs, the longitudinal dynamics can be explicitly solved, allowing the computation of vertical load transfers prior to their use in the expression of the lateral forces.
-Consequently, the time derivative of the state becomes an explicit function of the state vector $\bx$ and the input vector $\bu$, as follows:
-\begin{equation}
-	\dbx(t) = f(\bx(t), \bu(t))
-\end{equation}
+	% Draw inner ellipse filled with color
+	\fill[black!30] (0,0) ellipse [
+	x radius={\innerX},
+	y radius={\innerY}
+	];
+	\draw[dashed] (0,0) ellipse [
+	x radius={\innerX},
+	y radius={\innerY}
+	];
 
-The state vector $\bx$ has six components and includes the three kinematic quantities $u$, $v$, and $r$, along with the position and orientation of the barycentrical reference frame.
-As illustrated in Figure~\ref{fig:vehicle_model} $u$ and $v$ represent the longitudinal and lateral velocity of the CoM w.r.t its reference frame, while $r$ represents vehicle's yaw rate.
+	% Draw axes
+	\draw[->] (-\outerX-1,0) -- (\outerX+1,0) node[right] {$X_j$};
+	\draw[->] (0,-\outerY-1) -- (0,\outerY+1) node[above] {$Y_j$};
 
-\begin{figure}
-	\centering
-	\begin{tikzpicture}[scale=2, x = {(0,1cm)}, y = {(-1cm,0)},
-		force/.style={->, thick, red!60!black, >={Latex[length=10pt, width=4pt]}},
-		vec/.style={->, thick, mgreen, >={Latex[length=10pt, width=4pt]}},
-		aux/.style={thin},
-		frame/.style={->,thin,>={Latex}},
-		every node/.style={font=\small, text=black}
-		]
-		\definecolor{mgreen}{RGB}{0,100,0}
-		\begin{scope}[rotate = 25,shift = {(-.3,-2)}]
-			% Coordinates
-			\coordinate (rear) at (0,0);
-			\coordinate (G) at (1.2,0);
-			\coordinate (front) at (2.8,0);
+	% Arrows pointing to the ellipses
+	\draw[->,thin] (1.7,2) -- ({\outerX*cos(60)},{\outerY*sin(60)}) node[midway, above right] {$\beta=0$};
+	\draw[->,thin] (2.3,1.7) -- ({(\innerX)*cos(45)},{(\innerY)*sin(45)}) node[midway, right, shift={(0.05,.2)}] {$\beta=\gap$};
 
-			% Main body line
-			\draw[blue!50!black, thick] (rear) -- (front);
-
-			% Rear wheel
-			\draw[fill=black!60!white, draw=black, rounded corners=2pt]
-			($(rear)+(-.2,.1)$) rectangle ($(rear)+(+.2,-.1)$);
-
-			% Front wheel (rotated around its center)
-			\begin{scope}[shift={(front)}, rotate=25]
-				\draw[fill=black!60!white, draw=black, rounded corners=2pt]
-				(-0.2,.1) rectangle (0.2,-.1);
-			\end{scope}
-
-			% Local axes at rear
-			\draw[force] (rear) -- ++(.5,0) node[above,right] {$X_2$};
-			\draw[force] (rear) -- ++(0,1.3) node[above] {$Y_2$};
-
-			% Local axes at front
-			\draw[force] (front) -- ++(-.3,0) node[right] {$X_1$};
-			\draw[force] (front) -- ++(0,1.5) node[above] {$Y_1$};
-
-			% Velocity vectors at G
-			\draw[vec] (G) -- ++(.6,0) node[left, shift={(-.2,-.1)}] {$u\,\mathbf{i}$};
-			\draw[vec] (G) -- ++(0,.3) node[above] {$v\,\mathbf{j}$};
-
-			% Angular velocity
-			\begin{scope}[rotate = -115]
-				\draw[vec] 
-				(-.1,.8) arc[start angle=0, end angle=50, radius=1]
-				node[midway, right=2pt] {$r\,\mathbf{k}$};
-			\end{scope}
-
-			% G point
-			\fill (G) circle (1pt) node[right] {$G$};
-
-			% Body frame
-			\draw[frame] (G) --++ (1,0) node[right] {$x_b$};
-			\draw[frame] (G) --++ (0,1) node[above] {$y_b$};
-
-			% Distances
-%			\draw[<->,>={Latex}] (3,-1) -- node[left] {$a_1$} (1.5,-1);
-%			\draw[<->,>={Latex}] (1.5,-1) -- node[left] {$a_2$} (0,-1);
-
-			% Steering angle (auxiliary vertical line)
-			\draw[aux] (front) -- ++(.4,0);
-			\draw[aux] (front) -- ++({0.4*cos(25)}, {0.4*sin(25)});
-
-			% Steering angle arc
-			\draw[->,thin,>={Latex}] ([shift={(0.35,0)}]front) arc[start angle=0, end angle=25, radius=0.35];
-			\node[left,shift={(.15,0)}] at ([shift={(0.35,0)}]front) {$\delta$};
-
-			% Psi angle aux
-			\draw[aux] (G) -- ++({0.4*cos(25)}, {-0.4*sin(25)});
-
-			% Psi angle
-			\draw[->,thin,>={Latex}] ([shift={({0.35*cos(25)},{-.35*sin(25)})}]G) arc[start angle=-25, end angle=0, radius=0.35];
-			\node[left,shift={(.2,-.05)}] at ([shift={({0.35*cos(25)},{-.35*sin(25)})}]G) {$\psi$};
-
-		\end{scope}
-		\coordinate (ground) at (0,0);
-		\draw[frame] (ground) --++ (1,0) node[right] {$x$};
-		\draw[frame] (ground) --++ (0,1) node[above] {$y$};
-		\draw[->,thick,>=Latex] (ground) --++ (G);
-%		\draw[->,thick,>=Latex] (ground) --++ (G) node[midway, shift={(0,-.3)}] {$\boldsymbol{p}$};
+	% Arrows pointing to the ellipses
+	\draw[->,thin] ({\outerX+.3},.3) -- (\outerX,0) node[midway, above right,shift={(.05,0)}] {$\mu_x Z_j$};
+	\draw[->,thin] (.3,{\outerY+.3}) -- (0,\outerY) node[midway, above right,shift={(.05,0)}] {$\mu_y Z_j$};
 
 \end{tikzpicture}
-	\caption{Single Track model.
-The ground force acting on each axle is decomposed into longitudinal and lateral components, indicated by red arrows.
-The longitudinal components are part of the input vector $\bu$, encompassing both acceleration and braking forces, while the lateral components are computed using Pacejka's Magic Formula based on the vertical load and slip angle at each axle.
-The figure also depicts the fixed reference frame and the body reference frame, with the origin located at the vehicle's center of mass (CoM).
-The state variables $u$ and $v$ represent the longitudinal and lateral velocities of the CoM w.r.t.
-the body frame, while $r$ denotes the yaw rate.
-The angle $\psi$ is used to describe the orientation of the body frame w.r.t.
-the fixed frame.}
-	\label{fig:vehicle_model}
-\end{figure}
-
-\subsection{Stochastic vehicle dynamics}
-\label{sec:stochastic_vehicle_dynamics}
-
-We model the \emph{perturbed vehicle dynamics} as the following nonlinear continuous-time random dynamical system
-\begin{align}
-\dbx(t) = f(\bx(t), \bu(t)) + \bw(t),
-\end{align}
-where $\bu(t)$ are the deterministic control inputs, $\bw(t)$ is additive Gaussian white noise with zero mean and known covariance $\bQ(t)$, i.e.
-$\bw(t) \sim \calN(\bzero, \bQ(t))$, and the state $\bx(t)$ results in a Gaussian distribution with mean $\bmu(t)$ and covariance $\bP(t)$, so that $\bx(t) \sim \calN(\bmu(t), \bP(t))$.
-
-Due to the symmetry of the probability density function (pdf) with respect to $\bmu(t)$, it is possible to represent the time evolution of the pdf as: i) the deterministic evolution of the mean $\bmu(t)$
-\begin{align}\label{eq:meandynamics}
-\dbmu(t) = f(\bmu(t), \bu(t)),
-\end{align}
-and ii) the time evolution of the state covariance matrix $\bP(t)$ along $\bmu(t)$, which can be expressed by the \emph{Lyapunov matrix differential equation}
-\begin{align}\label{eq:dP}
-\dbP(t) = \bA(t) \bP(t) + \bP(t) \bA^T(t) + \bQ(t),\quad \bP(0) = \bP_0 = \bP_0^T.
-\end{align}
-In~\eqref{eq:dP}, $\bP_0$ is the \emph{initial} state covariance and $\bA(t)$ is the usual shorthand notation for the Jacobian along the mean trajectory $\bmu(t)$, that is $\bA(t)=\bJ(\bmu(t),\bu(t))$, where $\bJ(\bx, \bu) = \frac{\pd f(\bx, \bu)}{\pd \bx}$.
-
-As can be readily verified by differentiation~\cite{Gajic:LyapunovMatrixEquation:2010}, the analytical solution of~\eqref{eq:dP} has the form
-\begin{align}\label{eq:P_STM}
-\bP(t) = \bPhi(t,t_0) \bP_0 \bPhi(t,t_0)+\int_{t_0}^{t} \bPhi(t,\tau) \bQ(\tau) \bPhi^T(t,\tau) \dd \tau \quad \bP(0)=\bP_0,
-\end{align}
-where $\bPhi(t,t_0)$ is the \emph{state transition matrix}.
-This matrix encodes the evolution from $t_0$ to $t$ of a perturbation w.r.t.
-to a fiducial trajectory -- the mean $\bmu(t)$.
-In symbols, $\barbx(t) = \bPhi(t,t_0) \barbx(t_0)$, with $\barbx(\cdot) =\bx(\cdot) - \bmu(\cdot) $.
-In turn, the evolution of $\bPhi(t,t_k)$ from a generic $t_k$ to $t$ is driven by the following differential equation
-\begin{align}\label{eq:STM}
-\dbPhi(t,t_k) = \bA(t)\bPhi(t,t_k), \quad \bPhi(t_k, t_k) = \bI.
-\end{align}
-In geometric and orbital mechanics, see e.g.~\cite{Maruskin:DynamicalSystemsGeometric:2018} or~\cite{Tapley:StatisticalOrbitDetermination:2004}, the usual choice for statistical trajectory determination is to employ~\eqref{eq:meandynamics} along with~\eqref{eq:P_STM} and~\eqref{eq:STM}.
-In our case, since we use collocation integrators for stochastic trajectory planning, we follow a more direct approach -- motivated by~\cite{Gillis:PracticalMethodsApproximate:2015} -- by directly employing~\eqref{eq:meandynamics} and~\eqref{eq:dP}.
-
-Accordingly, the continuous-time stochastic trajectory planning can be framed as the following nonlinear optimal control problem
-\begin{subequations}\label{eq:OCP}
-\begin{align}
-& \underset{\bmu(t), \bu(t), \bP(t)}{\text{minimize}} \quad J(\bmu(t), \bu(t), \bP(t)) \label{eq:OCPcost} \\
-& \text{s.t.} \quad \dbmu(t)           = f(\bmu(t), \bu(t)) \label{eq:OCPdyn} \\
-& \phantom{\text{s.t.} \quad} \bmu(0)  = \bmu_0 \label{eq:OCPdynIC} \\
-& \phantom{\text{s.t.} \quad} \dbP(t) = \bA(t) \bP(t) + \bP(t) \bA^T(t) + \bQ(t) \label{eq:OCPdP} \\
-& \phantom{\text{s.t.} \quad} \bP(0)  = \bP_0 \succeq 0 \label{eq:OCPdPIC} \\ %& \phantom{\text{s.t.} \qquad} 0       \geq h_i(\bmu(t), \bu(t))
-%+ \overbrace{\gamma \bigg[\underbrace{\na^T_{\bx} h_i \bP(t) \na_{\bx} h_i}_{\text{variance of $h_i(\bx)$}}\bigg]^{\frac{1}{2}}}^{\text{safety margin}},
-&\phantom{\text{s.t.} \qquad} 0       \geq h_i(\bmu(t), \bu(t))
-+ \be_i(\bmu(t), \bu(t), \bP(t)),
-\quad i \in \calI \label{eq:OCPconstraints}
-\end{align}
-\end{subequations}
-The cost function $J$ in~\eqref{eq:OCPcost} depends on the mean $\bmu(t)$, the controls $\bu(t)$ and the state covariance $\bP(t)$.
-The mean dynamics is expressed by~\eqref{eq:OCPdyn} with~\eqref{eq:OCPdynIC}, and the covariance dynamics is expressed by~\eqref{eq:OCPdP} with~\eqref{eq:OCPdPIC}.
-In eq.~\eqref{eq:OCPconstraints} the \emph{backoff terms} $\be_i$ account for the disturbances and serve the purpose of obtaining deterministic safety margins directly on $\bmu(t)$.
-$\calI$ is the set of indices defining the inequality constraints.
-The backoff terms stem for a linearization around the mean of the original \emph{chance constraint} on $\bx(t)$ expressed by $\Prob \{h_i(\bx) \leq 0\}\geq p$, where $p$ is the confidence level of constraint satisfaction.
-Explicitly the backoff terms are given by
-$\be_i = \gamma \sig_i$.
-The coefficient $\gamma = \Phi^{-1}(p)$ is the quantile function, where $\Phi(z)=\Prob \{Z\leq z\}$ is the cdf of a standard normal distribution $Z \sim \calN(0,1)$, and acts as a \emph{tuning knob}: the greater the confidence level $p$ required, the higher the gain $\ga$\footnote{For example, with $p=0.84$ $\ga = 1.0$, with $p=0.97$ $\ga = 2.0$, with $p=0.99$ $\ga = 3.0$}.
-The term $\sig_i = \big[\na^T_{\bx} h_i(\bmu) \bP(t) \na_{\bx} h_i(\bmu)\big]^{\frac{1}{2}}$ represents the standard deviation of the constraint $h_i(\bx)$ linearized around the mean, i.e.
-of random variable $h_i(\bmu)+\na^T_{\bx} h_i(\bmu) (\bx - \bmu)$, and follows immediately from the propagation rule of covariance.
-
-\subsection{Discretization via direct collocation}
-\label{sec:discretization}
-
-The nonlinear optimal control problem~\eqref{eq:OCP} can be discretized by applying a suitable collocation integrator obtaining the following nonlinear program (NLP)
-\begin{subequations}\label{eq:DOCP}
-\begin{alignat}{3}
-\underset{\bmu_k,\bxi_k, \bu_k, \bP_k,\bz_k}{\text{minimize}} \,
-& & & J_k(\bmu_k,\bxi_k, \bu_k) & & \label{eq:DOCPcost} \\
-\hspace*{-2.0 cm}\text{s.t.} \quad
-& \bzero      & = & \; \bPsimu_k(\bmu_{k-1},\bmu_k,\bxi_k, \bu_k,\bz_k),
-& \quad & k = 1,\ldots, N \label{eq:DOCPdyn} \\
-& \bmu_0      & = & \; \bar{\bmu}_0
-& & \label{eq:DOCPdynIC} \\
-& \bzero      & = & \; \bPsiP_k(\bmu_k,\bxi_k, \bu_k, \bP_{k-1},\bP_k,\bSi_k,\bz_k),
-& \quad & k = 1,\ldots, N \label{eq:DOCPdP} \\
-& \bP_0       & = & \; \bar{\bP}_0 \succeq 0
-& & \label{eq:DOCPdPIC} \\
-& \bzero      & = & \; \bOm_k(\bmu_k,\bxi_k, \bu_k,\bz_k),
-& \quad & k = 0,\ldots, N \label{eq:DOCPpath} \\
-& 0           & \geq & \; h_i(\bmu_k, \bu_k, \bz_k) + \be_i(\bmu_k, \bu_k, \bP_k, \bz_k),
-& \quad & k = 1,\ldots, N;\; i \in \calI \label{eq:DOCPconstraints}
-\end{alignat}
-\end{subequations}
-To perform the discretization, the track centerline is parameterized using a curvilinear abscissa $\alpha \in [0,1]$, and uniformly sampled at $N+1$ points $\alpha_0, \ldots, \alpha_N$.
-Accordingly, $\bmu_k$ denotes the mean state at grid node $\alpha_k$, while the controls $\bu_k$ and the algebraic variables $\bz_k$ are assumed to be piecewise constant over each interval $[\al_k, \al_{k+1}]$.
-The $\bz_k$'s are introduced as direct handles for physically meaningful quantities such contact forces. Similarly, $\bP_k$ represents the covariance matrix at the $k$-th node.
-Following the direct collocation approach, both the mean and covariance state trajectories are approximated, in the $k$-th interval, with polynomials $\pi_k(\tau)$, defined on the unit interval $\tau\in[0,1]$, and then scaled to match the width $h_k$ of the corresponding time step.
-On the unit interval, we select $d$ collocation points $\tau_1, \ldots, \tau_d$, associated with as many collocation states.
-Accordingly, we define $\bxi_k$ and $\bSi_k$ as the mean states and covariance matrices at the $d$ collocation points within each interval $[\al_k, \al_{k+1}]$, respectively.
-Therefore, $\bxi_k =
-(\bxi_{k,1}, \ldots, \bxi_{k,d})$, and similarly $\bSi_k=(\bSi_{k,1}, \ldots, \bSi_{k,d})$.
-Equations~\eqref{eq:DOCPdyn} and~\eqref{eq:DOCPdP} represent the collocation and continuity equations for mean and covariance, respectively, with initial conditions represented by~\eqref{eq:DOCPdynIC} and~\eqref{eq:DOCPdPIC}.
-Eqs.~\eqref{eq:DOCPpath} are path equality constraints and~\eqref{eq:DOCPconstraints} are the robustified inequality constraints.
-
-In certain cases~\cite{Gillis:PracticalMethodsApproximate:2015}, positive-definiteness-preserving Lyapunov discretization schemes are used in~\eqref{eq:DOCPdP}.
-However, for sufficiently fine discretization, we found that integrating only the lower triangular part of $\bP(t)$ in~\eqref{eq:OCPdPIC}  (and reconstructing its strictly upper part accordingly), ensures that $\bP_k$ remain symmetric and positive definite when the initial $\bP_0$ is symmetric and positive definite.
-In all our tests direct collocation with cubic polynomial state representations and Gauss-Legendre collocation points provided accurate results.

### 2025-04-28 Fig/TikZ_Sources/tikz_covariance_evolution.tex
[MM] tikz sources folder created

--- 2025-04-17 open_loop_planning.tex
@@
-\section{Open-loop planning via H-steps ahead predictions}
-\label{sec:open_loop_planning}
-The proposed approach requires the presence of $H+1$ versions of the matrix $\bP$ at each grid point of the discretized problem.
-One of these versions is initialized with $\bar{\bP}_0$, while the others represent the evolution of matrices that were initialized at previous grid points, from 1 to $H$ steps earlier.
-This ensures that at each step it is possible to find a covariance matrix evolved for $H$ steps and use it to robustify the constraints.
-The initialization of the covariance matrix is necessary due to its divergent dynamics.
-In fact, without a feedback control, the error on the position and orientation of the body reference frame can only increase along the track, compromising the conditioning of the matrix.
-
-%\subsection{Discretization via direct collocation}
-%\label{sec:openloopdiscretization}
-Introducing the versioning of the covariance matrix approach requires a modification of the Equations described in Section~\ref{sec:discretization}.
-In particular Eqs.~\eqref{eq:DOCPdP}, \eqref{eq:DOCPdPIC}, and \eqref{eq:DOCPconstraints} are replaced by Eqs.~\eqref{eq:DOCPdPopenloop}, \eqref{eq:DOCPdPinitopenloop}, and \eqref{eq:DOCPconstraintsopenloop}.
-Eq.~\eqref{eq:DOCPdPopenloop} collects the continuity and collocation Equations for all the versions of the covariance matrix that have propagated from the previous step.
-To account for the evolution of $H$ matrices, it is necessary to introduce $H\times d$ matrices $\bSi_k^j=(\bSi_{k,1}^j, \ldots, \bSi_{k,d}^j)$, which represent the values of each matrix version at the collocation points.
-
-The specific formulation is expressed as follows:
-\begin{alignat}{3}
-	\underset{\bmu_k,\bxi_k, \bu_k, \bP_k,\bz_k}{\text{minimize}} \,
-	& & & J_k(\bmu_k,\bxi_k, \bu_k) & & \label{eq:DOCPcost} \\
-	\hspace*{-2.0 cm}\text{s.t.} \quad
-	& \bzero      & = & \; \bPsimu_k(\bmu_{k-1},\bmu_k,\bxi_k, \bu_k,\bz_k),
-	& \quad & k = 1,\ldots, N \label{eq:DOCPdyn} \\
-	& \bmu_0      & = & \; \bar{\bmu}_0
-	& & \label{eq:DOCPdynIC} \\
-	& \bzero      & = & \; \bPsiP_k(\bmu_k,\bxi_k, \bu_k, \bP_{k-1}^{j-1},\bP_k^j,\bSi_k^j,\bz_k),
-	& \quad & \substack{k = 1,\ldots, N; \; \\ j=1,\ldots,H;\;} \label{eq:DOCPdPopenloop} \\
-	& \bP_k^0       & = & \; \bar{\bP}_0 \succeq 0
-	& \quad & k = 1,\ldots, N; \; \label{eq:DOCPdPinitopenloop} \\
-	& \bzero      & = & \; \bOm_k(\bmu_k,\bxi_k, \bu_k,\bz_k),
-	& \quad & k = 0,\ldots, N \label{eq:DOCPpath} \\
-	& 0           & \geq & \; h_i(\bmu_k, \bu_k, \bz_k) + \be_i(\bmu_k, \bu_k, \bP_k^H, \bz_k),
-	& \quad & k = 1,\ldots, N;\; i \in \calI \label{eq:DOCPconstraintsopenloop}
-\end{alignat}
-
-Eq.~\eqref{eq:DOCPdPinitopenloop} is used to initialize the correct version at the $k$-th step while Eq.~\eqref{eq:DOCPconstraintsopenloop} specify that the back-off term of the constraints is evaluated with the most propagated covariance matrix version, $\bP^H_k$.
-A schematic representation of the approach used to manage the $H+1$ versions of the covariance matrix is shown in Figure \ref{fig:DOCPgrid}, where the dashed rectangle represent the $k$-th step at which the constraints are being formulated.
-At each grid point, two versions of the covariance matrix are highlighted: $\bP^0_k$ (red node), which is the version to be initialized, and $\bP^H_k$ (green node), which is the most propagated version used to formulate the robust constraints in Eq.~\eqref{eq:DOCPconstraintsopenloop}.
-
-\begin{figure}
-	\centering
-	\begin{tikzpicture}[%
-		smallnode/.style={%
-			circle, draw, minimum size=3mm, inner sep=0pt
-		},
-		r_smallnode/.style={%
-			circle, draw, minimum size=3mm, inner sep=0pt, red
-		},
-		g_smallnode/.style={%
-			circle, draw, minimum size=3mm, inner sep=0pt, mygreen
-		},
+\begin{tikzpicture}[
+	smallnode/.style={circle, draw, minimum size=3mm, inner sep=0pt},
+	r_smallnode/.style={circle, draw=red, minimum size=3mm, inner sep=0pt},
+	g_smallnode/.style={circle, draw=mygreen, minimum size=3mm, inner sep=0pt},
+	r_hatnode/.style={
+		circle, draw=red, pattern=north east lines, pattern color=red,
+		minimum size=3mm, inner sep=0pt},
+	g_hatnode/.style={
+		circle, draw=mygreen, pattern=north east lines, pattern color=mygreen,
+		minimum size=3mm, inner sep=0pt},
 	every path/.style={->, thick},
-		x=1.5cm, y=1.5cm
+	x=1.4cm, y=1.2cm
 	]
 
-		\newcommand{\Plab}[2]{\bP^{#2}_{#1}}
-		\definecolor{mygreen}{RGB}{0 150 0}
+	% ROW j = 0 (y=3)
+	\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k-3}^0\)}] (P00) at (1,3) {};
+	\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k-2}^1\)}] (P11) at (2,3) {};
+	\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k-1}^2\)}] (P22) at (3,3) {};
+	\node[g_hatnode, label={[label distance=-2pt]above:\(\bP_k^3\)}] (P33) at (4,3) {};
+	\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k+1}^0\)}] (P44) at (5,3) {};
+	\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+2}^1\)}] (P55) at (6,3) {};
 
-		% Nodes
-		% First row
-		\node[smallnode, label={[label distance=-2pt]above:\protect\(\Plab{k-2}{H-2}\)}] (Pk-2H-2) at (1,0) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{H-1}\)}] (Pk-1H-1) at (2,0) {};
-		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k}{H}\)}] (PkH) at (3,0) {};
-		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{0}\)}] (Pk+10) at (4,0) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{1}\)}] (Pk+21) at (5,0) {};
+	% ROW j = 1 (y=2)
+	\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k-2}^0\)}] (P10) at (2,2) {};
+	\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k-1}^1\)}] (P21) at (3,2) {};
+	\node[smallnode, label={[label distance=-2pt]above:\(\bP_k^2\)}] (P32) at (4,2) {};
+	\node[g_hatnode, label={[label distance=-2pt]above:\(\bP_{k+1}^3\)}] (P43) at (5,2) {};
+	\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k+2}^0\)}] (P54) at (6,2) {};
+	\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+3}^1\)}] (P65) at (7,2) {};
 
-		% Second row
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-2}{H-3}\)}] (Pk-2H-3) at (1,1) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{H-2}\)}] (Pk-1H-2) at (2,1) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k}{H-1}\)}] (PkH-1) at (3,1) {};
-		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{H}\)}] (Pk+1H) at (4,1) {};
-		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{0}\)}] (Pk+20) at (5,1) {};
+	% ROW j = 2 (y=1)
+	\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k-1}^0\)}] (P20) at (3,1) {};
+	\node[smallnode, label={[label distance=-2pt]above:\(\bP_k^1\)}] (P31) at (4,1) {};
+	\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+1}^2\)}] (P42) at (5,1) {};
+	\node[g_hatnode, label={[label distance=-2pt]above:\(\bP_{k+2}^3\)}] (P53) at (6,1) {};
+	\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k+3}^0\)}] (P64) at (7,1) {};
+	\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+4}^1\)}] (P75) at (8,1) {};
 
-		% dots row 
-		\node[label={center,rotate=90:\(\dots\)}] (dots2) at (1, 1.7) {};
-		\node[label={center,rotate=90:\(\dots\)}] (dots3) at (2, 1.7) {};
-		\node[label={center,rotate=90:\(\dots\)}] (dots4) at (3, 1.7) {};
-		\node[label={center,rotate=90:\(\dots\)}] (dots5) at (4, 1.7) {};
-		\node[label={center,rotate=90:\(\dots\)}] (dots6) at (5, 1.7) {};
+	% ROW j = 3 (y=0)
+	\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_k^0\)}] (P30) at (4,0) {};
+	\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+1}^1\)}] (P41) at (5,0) {};
+	\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+2}^2\)}] (P52) at (6,0) {};
+	\node[g_hatnode, label={[label distance=-2pt]above:\(\bP_{k+3}^3\)}] (P63) at (7,0) {};
+	\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k+4}^0\)}] (P74) at (8,0) {};
+	\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+5}^1\)}] (P85) at (9,0) {};
 
-		% Third row 
-		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k-2}{H}\)}] (Pk-2H) at (1,2) {};
-		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{0}\)}] (Pk-10) at (2,2) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k}{1}\)}] (Pk1) at (3,2) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{2}\)}] (Pk+12) at (4,2) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{3}\)}] (Pk+23) at (5,2) {};
+	% ARROWS
+	\draw (P00) -- (P11);
+	\draw (P11) -- (P22);
+	\draw (P22) -- (P33);
+	\draw (P22) -- (P33);
 
-		% Fourth row
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-2}{H-1}\)}] (Pk-2H-1) at (1,3) {};
-		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{H}\)}] (Pk-1H) at (2,3) {};
-		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k}{0}\)}] (Pk0) at (3,3) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{1}\)}] (Pk+11) at (4,3) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{2}\)}] (Pk+22) at (5,3) {};
+	\draw (P44) -- (P55);
+	\draw[dashed,->] (P55) -- ++(1,0);
 
-		% Curved arrows
-		\draw[bend left=20] (Pk-2H-2) to (Pk-1H-1);
-		\draw[bend left=20] (Pk-1H-1) to (PkH);
-		\draw[bend left=20] (Pk+10) to (Pk+21);
+	\draw (P10) -- (P21);
+	\draw (P21) -- (P32);
+	\draw (P32) -- (P43);
+	\draw (P54) -- (P65);
+	\draw[dashed,->] (P65) -- ++(1,0);
 
-		\draw[bend left=20] (Pk-2H-3) to (Pk-1H-2);
-		\draw[bend left=20] (Pk-1H-2) to (PkH-1);
-		\draw[bend left=20] (PkH-1) to (Pk+1H);
+	\draw (P20) -- (P31);
+	\draw (P31) -- (P42);
+	\draw (P42) -- (P53);
+	\draw (P64) -- (P75);
+	\draw[dashed,->] (P75) -- ++(1,0);
 
-		\draw[bend left=20] (Pk-10) to (Pk1);
-		\draw[bend left=20] (Pk1) to (Pk+12);
-		\draw[bend left=20] (Pk+12) to (Pk+23);
+	\draw (P30) -- (P41);
+	\draw (P41) -- (P52);
+	\draw (P52) -- (P63);
 
-		\draw[bend left=20] (Pk-10) to (Pk1);
-		\draw[bend left=20] (Pk1) to (Pk+12);
-		\draw[bend left=20] (Pk+12) to (Pk+23);
+	\draw (P74) -- (P85);
+	\draw[dashed,->] (P85) -- ++(1,0);
 
-		\draw[bend left=20] (Pk-2H-1) to (Pk-1H);
-		\draw[bend left=20] (Pk0) to (Pk+11);
-		\draw[bend left=20] (Pk+11) to (Pk+22);
+	% Dashed box around column k (x=4)
+	\draw[dashed, thick, rounded corners] (3.6,-0.5) rectangle (4.4,3.7);
 
-		% Dashed rectangle
-		\draw[dashed, thick, rounded corners] (2.6,-.25) rectangle (3.4,3.5);
+	% Top axis labels
+	\node at (0, 4.3) {\scriptsize \(\cdots\)};
+	\node at (1, 4.3) {\scriptsize \(k-3\)};
+	\node at (2, 4.3) {\scriptsize \(k-2\)};
+	\node at (3, 4.3) {\scriptsize \(k-1\)};
+	\node (knode) at (4, 4.3) {\scriptsize \(k\)};
+	\node at (5, 4.3) {\scriptsize \(k+1\)};
+	\node at (6, 4.3) {\scriptsize \(k+2\)};
+	\node at (7, 4.3) {\scriptsize \(k+3\)};
+	\node at (8, 4.3) {\scriptsize \(k+4\)};
+	\node at (9, 4.3) {\scriptsize \(\cdots\)};
+
+	% Wiggly arrow from label to top edge of dashed box
+	\draw[->, thick, decorate, decoration={snake, amplitude=0.25mm, segment length=2.5mm}]
+	(knode.south) -- (4, 3.8);
+
+	% Legend box (top right)
+	%\begin{scope}[shift={(6.5,2.5)}]
+	%    \node[r_smallnode] (r) at (0,0) {};
+	%    \node[anchor=west] at (0.4,0) {\scriptsize Initialized matrix \(\bP_k^0\)};
+	%
+	%    \node[g_smallnode] (g) at (0,-0.8) {};
+	%    \node[anchor=west] at (0.4,-0.8) {\scriptsize Most propagated matrix \(\bP_k^3\)};
+	%\end{scope}
 
 \end{tikzpicture}
-	\caption{Schematic representation of the continuity and initialization Equations for the $H+1$ versions of the covariance matrix introduced in the discretized OCP.
-The dashed rectangle indicate the $k$-th step, in which the constraints are formulated.
-Each node represents a covariance matrix, labeled such that the subscript denotes the current grid step, while the superscript indicates the number of propagation steps it has undergone.
-At each step, two nodes are highlighted: the covariance matrix to be initialized (red node) and the matrix that has been propagated for $H$ steps (green node), which is used to formulate the robust constraints.}
-	\label{fig:DOCPgrid}
-\end{figure}
-
-\subsection{Robust adherence constraint}

### 2025-04-28 Fig/TikZ_Sources/tikz_covariance_evolution_matteo.tex
[MM] tikz sources folder created

--- 2025-04-28 open_loop_planning.tex
@@
-\section{Open-loop planning via H-steps ahead predictions}
-\label{sec:open_loop_planning}
-In this section, we address the problem of planning under uncertainty for a stochastic system, where the covariance matrix is propagated in a feed-forward fashion for a fixed prediction horizon of $H$ steps at each grid point $k$ of the discretized trajectory.
-In the planning problem, we conservatively evaluate the maximum propagation of the covariance, that is, the matrix evolved over $H$ steps, to ensure the tightest possible enforcement of constraints.
-This maximizes the robustness of the resulting solution, as the constraint backoff is evaluated in the worst-case scenario.
-
-To enable this, we associate multiple versions of the covariance matrix with each grid point.
-Specifically, $H+1$ instances of the matrix $\bP$ are maintained at each step.
-To this sake we introduce the notation $\bP_k^j$, where $j = 0, \ldots, H$, and $\bP_k^j$ represents the version of the covariance matrix at step $k$ that was initialized $j$ steps earlier.
-Hence, $\bP_k^0$ originates at node $k$ itself and will be propagated forward and employed in subsequent steps (exactly when evolved as $\bP_{k+H}^H$ ), while $\bP_k^H$ is the instance that originated $H$ steps earlier and is the most propagated one at step $k$.
-
-The repeated initialization of the covariance matrix at each step is necessary to prevent its unbounded growth, as we assume that the system evolves without feedback control.
-In the absence of feedback, the uncertainty on the position and orientation of the body frame accumulates along the track, which may severely and unrealistically degrade the conditioning of the covariance matrix.
-For this reason, the propagation horizon in this open-loop setting is a critical design parameter that must be bounded to realistically capture the evolution of disturbances occurring prior to any corrective driver response.
-
-This approach requires modifications to the formulation~\eqref{eq:DOCP} introduced in Section~\ref{sec:discretization}.
-In particular, Eqs.~\eqref{eq:DOCPdP}, \eqref{eq:DOCPdPIC}, and \eqref{eq:DOCPconstraints} are replaced by Eqs.~\eqref{eq:DOCPdPopenloop}, \eqref{eq:DOCPdPinitopenloop}, and \eqref{eq:DOCPconstraintsopenloop}.
-Equation~\eqref{eq:DOCPdPopenloop} collects the continuity and collocation equations for all versions of the covariance matrix propagated from previous steps.
-To support the evolution of $H$ versions, we introduce $H \times d$ matrices $\bSi_k^j = (\bSi_{k,1}^j, \ldots, \bSi_{k,d}^j)$, representing the collocation values of each version $j$ of the covariance matrix within interval $k$.
-
-%The proposed approach requires the presence of $H+1$ versions of the matrix $\bP$ at each grid point of the discretized problem.
-One of these versions is initialized with $\bar{\bP}_0$, while the others represent the evolution of matrices that were initialized at previous grid points, from 1 to $H$ steps earlier.
-This ensures that at each step it is possible to find a covariance matrix evolved for $H$ steps and use it to robustify the constraints.
-The initialization of the covariance matrix is necessary due to its divergent dynamics.
-In fact, without a feedback control, the error on the position and orientation of the body reference frame can only increase along the track, compromising the conditioning of the matrix.
-%
-%Introducing the versioning of the covariance matrix approach requires a modification of the Equations described in Section~\ref{sec:discretization}.
-In particular Eqs.~\eqref{eq:DOCPdP}, \eqref{eq:DOCPdPIC}, and \eqref{eq:DOCPconstraints} are replaced by Eqs.~\eqref{eq:DOCPdPopenloop}, \eqref{eq:DOCPdPinitopenloop}, and \eqref{eq:DOCPconstraintsopenloop}.
-Eq.~\eqref{eq:DOCPdPopenloop} collects the continuity and collocation Equations for all the versions of the covariance matrix that have propagated from the previous step.
-To account for the evolution of $H$ matrices, it is necessary to introduce $H\times d$ matrices $\bSi_k^j=(\bSi_{k,1}^j, \ldots, \bSi_{k,d}^j)$, which represent the values of each matrix version at the collocation points.
-
-The specific formulation is expressed as follows:
-\begin{subequations}\label{eq:DOCPopenloop}
-\begin{alignat}{3}
-	\underset{\bmu_k,\bxi_k, \bu_k, \bP_k,\bz_k}{\text{minimize}} \,
-	& & & J_k(\bmu_k,\bxi_k, \bu_k) & & \label{eq:DOCPcostopenloop} \\
-	\hspace*{-2.0 cm}\text{s.t.} \quad
-	& \bzero      & = & \; \bPsimu_k(\bmu_{k-1},\bmu_k,\bxi_k, \bu_k,\bz_k),
-	& \quad & k = 1,\ldots, N \label{eq:DOCPdynopenloop} \\
-	& \bmu_0      & = & \; \bar{\bmu}_0
-	& & \label{eq:DOCPdynICopenloop} \\
-	& \bzero      & = & \; \bPsiP_k(\bmu_k,\bxi_k, \bu_k, \bP_{k-1}^{j-1},\bP_k^j,\bSi_k^j,\bz_k),
-	& \quad & \substack{k = 1,\ldots, N; \; \\ j=1,\ldots,H;\;} \label{eq:DOCPdPopenloop} \\
-	& \bP_k^0       & = & \; \bar{\bP}_0 \succeq 0
-	& \quad & k = 1,\ldots, N; \; \label{eq:DOCPdPinitopenloop} \\
-	& \bzero      & = & \; \bOm_k(\bmu_k,\bxi_k, \bu_k,\bz_k),
-	& \quad & k = 0,\ldots, N \label{eq:DOCPpathopenloop} \\
-	& 0           & \geq & \; h_i(\bmu_k, \bu_k, \bz_k) + \be_i(\bmu_k, \bu_k, \bP_k^H, \bz_k),
-	& \quad & k = 1,\ldots, N;\; i \in \calI \label{eq:DOCPconstraintsopenloop}
-\end{alignat}
-\end{subequations}
-Equation~\eqref{eq:DOCPdPinitopenloop} defines the initialization of the appropriate covariance matrix version at step $k$, while Eq.~\eqref{eq:DOCPconstraintsopenloop} specifies that the back-off term in the constraints is computed using the most propagated version, $\bP^H_k$.
-A schematic illustration of the management of the $H+1$ covariance matrix instances is provided in Figure~\ref{fig:DOCPgrid}.
-The dashed rectangle highlights the $k$-th discretization step, where the constraints are evaluated.
-At each grid point, two particular versions of the covariance matrix are emphasized: $\bP^0_k$ (red node), representing the instance to be initialized at that step, and $\bP^H_k$ (green node), corresponding to the version that has been propagated over the full horizon of $H$ steps and is used to robustify the constraints in Eq.~\eqref{eq:DOCPconstraintsopenloop}.
-
-%Eq.~\eqref{eq:DOCPdPinitopenloop} is used to initialize the correct version at the $k$-th step while Eq.~\eqref{eq:DOCPconstraintsopenloop} specify that the back-off term of the constraints is evaluated with the most propagated covariance matrix version, $\bP^H_k$.
-%A schematic representation of the approach used to manage the $H+1$ versions of the covariance matrix is shown in Figure \ref{fig:DOCPgrid}, where the dashed rectangle represent the $k$-th step at which the constraints are being formulated.
-At each grid point, two versions of the covariance matrix are highlighted: $\bP^0_k$ (red node), which is the version to be initialized, and $\bP^H_k$ (green node), which is the most propagated version used to formulate the robust constraints in Eq.~\eqref{eq:DOCPconstraintsopenloop}.
-\def\version{marco}
-\ifdefstring{\version}{matteo}{
-\begin{figure}
-	\centering
 \begin{tikzpicture}[%
@@
 \end{tikzpicture}
-	\caption{Schematic representation of the continuity and initialization equations associated with the $H+1$ propagated instances of the covariance matrix introduced in problem~\eqref{eq:DOCPopenloop}.
-The dashed rectangle highlights the $k$-th discretization step, at which the robust constraints are enforced.
-Each node corresponds to a specific covariance matrix, where the subscript denotes the current grid index and the superscript indicates the number of propagation steps since initialization.
-At each step, two key instances are emphasized: the matrix to be initialized at that step (red node), and the matrix that has been propagated over $H$ steps (green node), which is used in the evaluation of the constraint backoff.}
-	%\caption{Schematic representation of the continuity and initialization Equations for the $H+1$ versions of the covariance matrix introduced in the discretized OCP.
-The dashed rectangle indicate the $k$-th step, in which the constraints are formulated.
-Each node represents a covariance matrix, labeled such that the subscript denotes the current grid step, while the superscript indicates the number of propagation steps it has undergone.
-At each step, two nodes are highlighted: the covariance matrix to be initialized (red node) and the matrix that has been propagated for $H$ steps (green node), which is used to formulate the robust constraints.}
-	\label{fig:DOCPgrid}
-\end{figure}
-}{}
-
-\ifdefstring{\version}{marco}{
-\begin{figure}
-	\centering
-	\definecolor{mygreen}{RGB}{0,150,0}
-	\begin{tikzpicture}[
-		smallnode/.style={circle, draw, minimum size=3mm, inner sep=0pt},
-		r_smallnode/.style={circle, draw=red, minimum size=3mm, inner sep=0pt},
-		g_smallnode/.style={circle, draw=mygreen, minimum size=3mm, inner sep=0pt},
-		r_hatnode/.style={
-			circle, draw=red, pattern=north east lines, pattern color=red,
-			minimum size=3mm, inner sep=0pt},
-		g_hatnode/.style={
-			circle, draw=mygreen, pattern=north east lines, pattern color=mygreen,
-			minimum size=3mm, inner sep=0pt},
-		every path/.style={->, thick},
-		x=1.4cm, y=1.2cm
-		]
-
-		% ROW j = 0 (y=3)
-		\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k-3}^0\)}] (P00) at (1,3) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k-2}^1\)}] (P11) at (2,3) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k-1}^2\)}] (P22) at (3,3) {};
-		\node[g_hatnode, label={[label distance=-2pt]above:\(\bP_k^3\)}] (P33) at (4,3) {};
-		\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k+1}^0\)}] (P44) at (5,3) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+2}^1\)}] (P55) at (6,3) {};
-
-		% ROW j = 1 (y=2)
-		\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k-2}^0\)}] (P10) at (2,2) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k-1}^1\)}] (P21) at (3,2) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\bP_k^2\)}] (P32) at (4,2) {};
-		\node[g_hatnode, label={[label distance=-2pt]above:\(\bP_{k+1}^3\)}] (P43) at (5,2) {};
-		\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k+2}^0\)}] (P54) at (6,2) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+3}^1\)}] (P65) at (7,2) {};
-
-		% ROW j = 2 (y=1)
-		\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k-1}^0\)}] (P20) at (3,1) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\bP_k^1\)}] (P31) at (4,1) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+1}^2\)}] (P42) at (5,1) {};
-		\node[g_hatnode, label={[label distance=-2pt]above:\(\bP_{k+2}^3\)}] (P53) at (6,1) {};
-		\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k+3}^0\)}] (P64) at (7,1) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+4}^1\)}] (P75) at (8,1) {};
-
-		% ROW j = 3 (y=0)
-		\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_k^0\)}] (P30) at (4,0) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+1}^1\)}] (P41) at (5,0) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+2}^2\)}] (P52) at (6,0) {};
-		\node[g_hatnode, label={[label distance=-2pt]above:\(\bP_{k+3}^3\)}] (P63) at (7,0) {};
-		\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k+4}^0\)}] (P74) at (8,0) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+5}^1\)}] (P85) at (9,0) {};
-
-		% ARROWS
-		\draw (P00) -- (P11);
-		\draw (P11) -- (P22);
-		\draw (P22) -- (P33);
-		\draw (P22) -- (P33);
-
-		\draw (P44) -- (P55);
-		\draw[dashed,->] (P55) -- ++(1,0);
-
-		\draw (P10) -- (P21);
-		\draw (P21) -- (P32);
-		\draw (P32) -- (P43);
-		\draw (P54) -- (P65);
-		\draw[dashed,->] (P65) -- ++(1,0);
-
-		\draw (P20) -- (P31);
-		\draw (P31) -- (P42);
-		\draw (P42) -- (P53);
-		\draw (P64) -- (P75);
-		\draw[dashed,->] (P75) -- ++(1,0);
-
-		\draw (P30) -- (P41);
-		\draw (P41) -- (P52);
-		\draw (P52) -- (P63);
-
-		\draw (P74) -- (P85);
-		\draw[dashed,->] (P85) -- ++(1,0);
-
-		% Dashed box around column k (x=4)
-		\draw[dashed, thick, rounded corners] (3.6,-0.5) rectangle (4.4,3.7);
-
-		% Top axis labels
-		\node at (0, 4.3) {\scriptsize \(\cdots\)};
-		\node at (1, 4.3) {\scriptsize \(k-3\)};
-		\node at (2, 4.3) {\scriptsize \(k-2\)};
-		\node at (3, 4.3) {\scriptsize \(k-1\)};
-		\node (knode) at (4, 4.3) {\scriptsize \(k\)};
-		\node at (5, 4.3) {\scriptsize \(k+1\)};
-		\node at (6, 4.3) {\scriptsize \(k+2\)};
-		\node at (7, 4.3) {\scriptsize \(k+3\)};
-		\node at (8, 4.3) {\scriptsize \(k+4\)};
-		\node at (9, 4.3) {\scriptsize \(\cdots\)};
-
-		% Wiggly arrow from label to top edge of dashed box
-		\draw[->, thick, decorate, decoration={snake, amplitude=0.25mm, segment length=2.5mm}]
-		(knode.south) -- (4, 3.8);
-
-		% Legend box (top right)
-		%\begin{scope}[shift={(6.5,2.5)}]
-		%    \node[r_smallnode] (r) at (0,0) {};
-		%    \node[anchor=west] at (0.4,0) {\scriptsize Initialized matrix \(\bP_k^0\)};
-		%
-		%    \node[g_smallnode] (g) at (0,-0.8) {};
-		%    \node[anchor=west] at (0.4,-0.8) {\scriptsize Most propagated matrix \(\bP_k^3\)};
-		%\end{scope}
-
-	\end{tikzpicture}
-
-	\caption{Cartoon representation (depicted for $H=3$) of the continuity and initialization equations associated with the $H+1$ propagated instances of the covariance matrix introduced in problem~\eqref{eq:DOCPopenloop}.
-The dashed rectangle highlights the $k$-th discretization step, at which the robust constraints are enforced.
-Each node corresponds to a specific covariance matrix, where the subscript denotes the current grid index and the superscript indicates the number of propagation steps since initialization.
-At each step, two key instances are emphasized: the matrix to be initialized at that step (red node), and the matrix that has been propagated over $H$ steps (green node), which is used in the evaluation of the constraint backoff.}
-	%\caption{Schematic representation of the continuity and initialization Equations for the $H+1$ versions of the covariance matrix introduced in the discretized OCP.
-The dashed rectangle indicate the $k$-th step, in which the constraints are formulated.
-Each node represents a covariance matrix, labeled such that the subscript denotes the current grid step, while the superscript indicates the number of propagation steps it has undergone.
-At each step, two nodes are highlighted: the covariance matrix to be initialized (red node) and the matrix that has been propagated for $H$ steps (green node), which is used to formulate the robust constraints.}
-	\label{fig:DOCPgrid}
-\end{figure}
-}{}

### 2025-04-28 Fig/TikZ_Sources/tikz_stayontrack_constraint.tex
[MM] tikz sources folder created

--- 2025-04-18 stochastic_vehicle_dynamics.tex
@@
-\section{TMP NAME}
-\label{sec:tmp}
+\begin{tikzpicture}[scale = .7]
+	% Parameters
+	\def\outerX{3}   % x radius of outer ellipse
+	\def\outerY{2}   % y radius of outer ellipse
+	\def\gap{0.2}    % distance between outer and inner ellipse
+	\def\innerX{2.68}
+	\def\innerY{1.79}
 
-\subsection{Vehicle model}
-\label{sec:vehicle_model}
-The vehicle model employed in this study is a Single Track model featuring nonlinear axle characteristics.
-The lateral forces $Y_1$ and $Y_2$ are computed using a Pacejka's Magic Formula, which depends on the total vertical load acting on the axle and the theoretical lateral slip evaluated at the midpoint of the axle.
-The longitudinal forces $X_1$ and $X_2$ are provided as inputs to the system.
-Specifically, the first two components of the input vector $\bu$ correspond to acceleration and braking forces, while the third component represents the steering angle of the front wheels.
-This input formulation facilitates the distribution of braking forces between the front and rear axles, while assigning acceleration exclusively to the rear axle, consistent with the rear-wheel-drive configuration of the modeled vehicle.
+	% Draw outer ellipse
+	\draw[] (0,0) ellipse [x radius=\outerX, y radius=\outerY];
 
-By treating the longitudinal forces as inputs, the longitudinal dynamics can be explicitly solved, allowing the computation of vertical load transfers prior to their use in the expression of the lateral forces.
-Consequently, the time derivative of the state becomes an explicit function of the state vector $\bx$ and the input vector $\bu$, as follows:
-\begin{equation}
-	\dbx(t) = f(\bx(t), \bu(t))
-\end{equation}
+	% Draw inner ellipse filled with color
+	\fill[black!30] (0,0) ellipse [
+	x radius={\innerX},
+	y radius={\innerY}
+	];
+	\node[above] at (0,0) {PATH CONSTRAINT FIG};
+	\draw[dashed] (0,0) ellipse [
+	x radius={\innerX},
+	y radius={\innerY}
+	];
 
-The state vector $\bx$ has six components and includes the three kinematic quantities $u$, $v$, and $r$, along with the position and orientation of the barycentrical reference frame.
-As illustrated in Figure~\ref{fig:vehicle_model} $u$ and $v$ represent the longitudinal and lateral velocity of the CoM w.r.t its reference frame, while $r$ represents vehicle's yaw rate.
+	% Draw axes
+	\draw[->] (-\outerX-1,0) -- (\outerX+1,0) node[right] {$X_i$};
+	\draw[->] (0,-\outerY-1) -- (0,\outerY+1) node[above] {$Y_i$};
 
-\begin{figure}
-	\centering
-	\begin{tikzpicture}[scale=2, x = {(0,1cm)}, y = {(-1cm,0)},
-		force/.style={->, thick, red!60!black, >={Latex[length=10pt, width=4pt]}},
-		vec/.style={->, thick, mgreen, >={Latex[length=10pt, width=4pt]}},
-		aux/.style={thin},
-		frame/.style={->,thin,>={Latex}},
-		every node/.style={font=\small, text=black}
-		]
-		\definecolor{mgreen}{RGB}{0,100,0}
-		\begin{scope}[rotate = 25,shift = {(-.3,-2)}]
-			% Coordinates
-			\coordinate (rear) at (0,0);
-			\coordinate (G) at (1.2,0);
-			\coordinate (front) at (2.8,0);
+	% Arrows pointing to the ellipses
+	\draw[->,thin] (1.7,2) -- ({\outerX*cos(60)},{\outerY*sin(60)}) node[midway, above right] {$\beta=0$};
+	\draw[->,thin] (2.3,1.7) -- ({(\innerX)*cos(45)},{(\innerY)*sin(45)}) node[midway, right, shift={(0.05,.2)}] {$\beta=\gap$};
 
-			% Main body line
-			\draw[blue!50!black, thick] (rear) -- (front);
-
-			% Rear wheel
-			\draw[fill=black!60!white, draw=black, rounded corners=2pt]
-			($(rear)+(-.2,.1)$) rectangle ($(rear)+(+.2,-.1)$);
-
-			% Front wheel (rotated around its center)
-			\begin{scope}[shift={(front)}, rotate=25]
-				\draw[fill=black!60!white, draw=black, rounded corners=2pt]
-				(-0.2,.1) rectangle (0.2,-.1);
-			\end{scope}
-
-			% Local axes at rear
-			\draw[force] (rear) -- ++(.5,0) node[above,right] {$X_2$};
-			\draw[force] (rear) -- ++(0,1.3) node[above] {$Y_2$};
-
-			% Local axes at front
-			\draw[force] (front) -- ++(-.3,0) node[right] {$X_1$};
-			\draw[force] (front) -- ++(0,1.5) node[above] {$Y_1$};
-
-			% Velocity vectors at G
-			\draw[vec] (G) -- ++(.6,0) node[left, shift={(-.2,-.1)}] {$u\,\mathbf{i}$};
-			\draw[vec] (G) -- ++(0,.3) node[above] {$v\,\mathbf{j}$};
-
-			% Angular velocity
-			\begin{scope}[rotate = -115]
-				\draw[vec] 
-				(-.1,.8) arc[start angle=0, end angle=50, radius=1]
-				node[midway, right=2pt] {$r\,\mathbf{k}$};
-			\end{scope}
-
-			% G point
-			\fill (G) circle (1pt) node[right] {$G$};
-
-			% Body frame
-			\draw[frame] (G) --++ (1,0) node[right] {$x_b$};
-			\draw[frame] (G) --++ (0,1) node[above] {$y_b$};
-
-			% Distances
-%			\draw[<->,>={Latex}] (3,-1) -- node[left] {$a_1$} (1.5,-1);
-%			\draw[<->,>={Latex}] (1.5,-1) -- node[left] {$a_2$} (0,-1);
-
-			% Steering angle (auxiliary vertical line)
-			\draw[aux] (front) -- ++(.4,0);
-			\draw[aux] (front) -- ++({0.4*cos(25)}, {0.4*sin(25)});
-
-			% Steering angle arc
-			\draw[->,thin,>={Latex}] ([shift={(0.35,0)}]front) arc[start angle=0, end angle=25, radius=0.35];
-			\node[left,shift={(.15,0)}] at ([shift={(0.35,0)}]front) {$\delta$};
-
-			% Psi angle aux
-			\draw[aux] (G) -- ++({0.4*cos(25)}, {-0.4*sin(25)});
-
-			% Psi angle
-			\draw[->,thin,>={Latex}] ([shift={({0.35*cos(25)},{-.35*sin(25)})}]G) arc[start angle=-25, end angle=0, radius=0.35];
-			\node[left,shift={(.2,-.05)}] at ([shift={({0.35*cos(25)},{-.35*sin(25)})}]G) {$\psi$};
-
-		\end{scope}
-		\coordinate (ground) at (0,0);
-		\draw[frame] (ground) --++ (1,0) node[right] {$x$};
-		\draw[frame] (ground) --++ (0,1) node[above] {$y$};
-		\draw[->,thick,>=Latex] (ground) --++ (G);
-%		\draw[->,thick,>=Latex] (ground) --++ (G) node[midway, shift={(0,-.3)}] {$\boldsymbol{p}$};
+	% Arrows pointing to the ellipses
+	\draw[->,thin] ({\outerX+.3},.3) -- (\outerX,0) node[midway, above right,shift={(.05,0)}] {$\mu_x Z_i$};
+	\draw[->,thin] (.3,{\outerY+.3}) -- (0,\outerY) node[midway, above right,shift={(.05,0)}] {$\mu_y Z_i$};
 
 \end{tikzpicture}
-	\caption{Single Track model.
-The ground force acting on each axle is decomposed into longitudinal and lateral components, indicated by red arrows.
-The longitudinal components are part of the input vector $\bu$, encompassing both acceleration and braking forces, while the lateral components are computed using Pacejka's Magic Formula based on the vertical load and slip angle at each axle.
-The figure also depicts the fixed reference frame and the body reference frame, with the origin located at the vehicle's center of mass (CoM).
-The state variables $u$ and $v$ represent the longitudinal and lateral velocities of the CoM w.r.t.
-the body frame, while $r$ denotes the yaw rate.
-The angle $\psi$ is used to describe the orientation of the body frame w.r.t.
-the fixed frame.}
-	\label{fig:vehicle_model}
-\end{figure}
-
-\subsection{Stochastic vehicle dynamics}
-\label{sec:stochastic_vehicle_dynamics}
-
-We model the \emph{perturbed vehicle dynamics} as the following nonlinear continuous-time random dynamical system
-\begin{align}
-\dbx(t) = f(\bx(t), \bu(t)) + \bw(t),
-\end{align}
-where $\bu(t)$ are the deterministic control inputs, $\bw(t)$ is additive Gaussian white noise with zero mean and known covariance $\bQ(t)$, i.e.
-$\bw(t) \sim \calN(\bzero, \bQ(t))$, and the state $\bx(t)$ results in a Gaussian distribution with mean $\bmu(t)$ and covariance $\bP(t)$, so that $\bx(t) \sim \calN(\bmu(t), \bP(t))$.
-
-Due to the symmetry of the probability density function (pdf) with respect to $\bmu(t)$, it is possible to represent the time evolution of the pdf as: i) the deterministic evolution of the mean $\bmu(t)$
-\begin{align}\label{eq:meandynamics}
-\dbmu(t) = f(\bmu(t), \bu(t)),
-\end{align}
-and ii) the time evolution of the state covariance matrix $\bP(t)$ along $\bmu(t)$, which can be expressed by the \emph{Lyapunov matrix differential equation}
-\begin{align}\label{eq:dP}
-\dbP(t) = \bA(t) \bP(t) + \bP(t) \bA^T(t) + \bQ(t),\quad \bP(0) = \bP_0 = \bP_0^T.
-\end{align}
-In~\eqref{eq:dP}, $\bP_0$ is the \emph{initial} state covariance and $\bA(t)$ is the usual shorthand notation for the Jacobian along the mean trajectory $\bmu(t)$, that is $\bA(t)=\bJ(\bmu(t),\bu(t))$, where $\bJ(\bx, \bu) = \frac{\pd f(\bx, \bu)}{\pd \bx}$.
-
-As can be readily verified by differentiation~\cite{Gajic:LyapunovMatrixEquation:2010}, the analytical solution of~\eqref{eq:dP} has the form
-\begin{align}\label{eq:P_STM}
-\bP(t) = \bPhi(t,t_0) \bP_0 \bPhi(t,t_0)+\int_{t_0}^{t} \bPhi(t,\tau) \bQ(\tau) \bPhi^T(t,\tau) \dd \tau \quad \bP(0)=\bP_0,
-\end{align}
-where $\bPhi(t,t_0)$ is the \emph{state transition matrix}.
-This matrix encodes the evolution from $t_0$ to $t$ of a perturbation w.r.t.
-to a fiducial trajectory -- the mean $\bmu(t)$.
-In symbols, $\barbx(t) = \bPhi(t,t_0) \barbx(t_0)$, with $\barbx(\cdot) =\bx(\cdot) - \bmu(\cdot) $.
-In turn, the evolution of $\bPhi(t,t_k)$ from a generic $t_k$ to $t$ is driven by the following differential equation
-\begin{align}\label{eq:STM}
-\dbPhi(t,t_k) = \bA(t)\bPhi(t,t_k), \quad \bPhi(t_k, t_k) = \bI.
-\end{align}
-In geometric and orbital mechanics, see e.g.~\cite{Maruskin:DynamicalSystemsGeometric:2018} or~\cite{Tapley:StatisticalOrbitDetermination:2004}, the usual choice for statistical trajectory determination is to employ~\eqref{eq:meandynamics} along with~\eqref{eq:P_STM} and~\eqref{eq:STM}.
-In our case, since we use collocation integrators for stochastic trajectory planning, we follow a more direct approach -- motivated by~\cite{Gillis:PracticalMethodsApproximate:2015} -- by directly employing~\eqref{eq:meandynamics} and~\eqref{eq:dP}.
-
-Accordingly, the continuous-time stochastic trajectory planning can be framed as the following nonlinear optimal control problem
-\begin{subequations}\label{eq:OCP}
-\begin{align}
-& \underset{\bmu(t), \bu(t), \bP(t)}{\text{minimize}} \quad J(\bmu(t), \bu(t), \bP(t)) \label{eq:OCPcost} \\
-& \text{s.t.} \quad \dbmu(t)           = f(\bmu(t), \bu(t)) \label{eq:OCPdyn} \\
-& \phantom{\text{s.t.} \quad} \bmu(0)  = \bmu_0 \label{eq:OCPdynIC} \\
-& \phantom{\text{s.t.} \quad} \dbP(t) = \bA(t) \bP(t) + \bP(t) \bA^T(t) + \bQ(t) \label{eq:OCPdP} \\
-& \phantom{\text{s.t.} \quad} \bP(0)  = \bP_0 \succeq 0 \label{eq:OCPdPIC} \\ %& \phantom{\text{s.t.} \qquad} 0       \geq h_i(\bmu(t), \bu(t))
-%+ \overbrace{\gamma \bigg[\underbrace{\na^T_{\bx} h_i \bP(t) \na_{\bx} h_i}_{\text{variance of $h_i(\bx)$}}\bigg]^{\frac{1}{2}}}^{\text{safety margin}},
-&\phantom{\text{s.t.} \qquad} 0       \geq h_i(\bmu(t), \bu(t))
-+ \be_i(\bmu(t), \bu(t), \bP(t)),
-\quad i \in \calI \label{eq:OCPconstraints}
-\end{align}
-\end{subequations}
-The cost function $J$ in~\eqref{eq:OCPcost} depends on the mean $\bmu(t)$, the controls $\bu(t)$ and the state covariance $\bP(t)$.
-The mean dynamics is expressed by~\eqref{eq:OCPdyn} with~\eqref{eq:OCPdynIC}, and the covariance dynamics is expressed by~\eqref{eq:OCPdP} with~\eqref{eq:OCPdPIC}.
-In eq.~\eqref{eq:OCPconstraints} the \emph{backoff terms} $\be_i$ account for the disturbances and serve the purpose of obtaining deterministic safety margins directly on $\bmu(t)$.
-$\calI$ is the set of indices defining the inequality constraints.
-The backoff terms stem for a linearization around the mean of the original \emph{chance constraint} on $\bx(t)$ expressed by $\Prob \{h_i(\bx) \leq 0\}\geq p$, where $p$ is the confidence level of constraint satisfaction.
-Explicitly the backoff terms are given by
-$\be_i = \gamma \sig_i$.
-The coefficient $\gamma = \Phi^{-1}(p)$ is the quantile function, where $\Phi(z)=\Prob \{Z\leq z\}$ is the cdf of a standard normal distribution $Z \sim \calN(0,1)$, and acts as a \emph{tuning knob}: the greater the confidence level $p$ required, the higher the gain $\ga$\footnote{For example, with $p=0.84$ $\ga = 1.0$, with $p=0.97$ $\ga = 2.0$, with $p=0.99$ $\ga = 3.0$}.
-The term $\sig_i = \big[\na^T_{\bx} h_i(\bmu) \bP(t) \na_{\bx} h_i(\bmu)\big]^{\frac{1}{2}}$ represents the standard deviation of the constraint $h_i(\bx)$ linearized around the mean, i.e.
-of random variable $h_i(\bmu)+\na^T_{\bx} h_i(\bmu) (\bx - \bmu)$, and follows immediately from the propagation rule of covariance.
-
-\subsection{Discretization via direct collocation}
-\label{sec:discretization}
-
-The nonlinear optimal control problem~\eqref{eq:OCP} can be discretized by applying a suitable collocation integrator obtaining the following nonlinear program (NLP)
-\begin{subequations}\label{eq:DOCP}
-\begin{alignat}{3}
-\underset{\bmu_k,\bxi_k, \bu_k, \bP_k,\bz_k}{\text{minimize}} \,
-& & & J_k(\bmu_k,\bxi_k, \bu_k) & & \label{eq:DOCPcost} \\
-\hspace*{-2.0 cm}\text{s.t.} \quad
-& \bzero      & = & \; \bPsimu_k(\bmu_{k-1},\bmu_k,\bxi_k, \bu_k,\bz_k),
-& \quad & k = 1,\ldots, N \label{eq:DOCPdyn} \\
-& \bmu_0      & = & \; \bar{\bmu}_0
-& & \label{eq:DOCPdynIC} \\
-& \bzero      & = & \; \bPsiP_k(\bmu_k,\bxi_k, \bu_k, \bP_{k-1},\bP_k,\bSi_k,\bz_k),
-& \quad & k = 1,\ldots, N \label{eq:DOCPdP} \\
-& \bP_0       & = & \; \bar{\bP}_0 \succeq 0
-& & \label{eq:DOCPdPIC} \\
-& \bzero      & = & \; \bOm_k(\bmu_k,\bxi_k, \bu_k,\bz_k),
-& \quad & k = 0,\ldots, N \label{eq:DOCPpath} \\
-& 0           & \geq & \; h_i(\bmu_k, \bu_k, \bz_k) + \be_i(\bmu_k, \bu_k, \bP_k, \bz_k),
-& \quad & k = 1,\ldots, N;\; i \in \calI \label{eq:DOCPconstraints}
-\end{alignat}
-\end{subequations}
-To perform the discretization, the track centerline is parameterized using a curvilinear abscissa $\alpha \in [0,1]$, and uniformly sampled at $N+1$ points $\alpha_0, \ldots, \alpha_N$.
-Accordingly, $\bmu_k$ denotes the mean state at grid node $\alpha_k$, while the controls $\bu_k$ and the algebraic variables $\bz_k$ are assumed to be piecewise constant over each interval $[\al_k, \al_{k+1}]$.
-The $\bz_k$'s are introduced as direct handles for physically meaningful quantities such contact forces. Similarly, $\bP_k$ represents the covariance matrix at the $k$-th node.
-Following the direct collocation approach, both the mean and covariance state trajectories are approximated, in the $k$-th interval, with polynomials $\pi_k(\tau)$, defined on the unit interval $\tau\in[0,1]$, and then scaled to match the width $h_k$ of the corresponding time step.
-On the unit interval, we select $d$ collocation points $\tau_1, \ldots, \tau_d$, associated with as many collocation states.
-Accordingly, we define $\bxi_k$ and $\bSi_k$ as the mean states and covariance matrices at the $d$ collocation points within each interval $[\al_k, \al_{k+1}]$, respectively.
-Therefore, $\bxi_k =
-(\bxi_{k,1}, \ldots, \bxi_{k,d})$, and similarly $\bSi_k=(\bSi_{k,1}, \ldots, \bSi_{k,d})$.
-Equations~\eqref{eq:DOCPdyn} and~\eqref{eq:DOCPdP} represent the collocation and continuity equations for mean and covariance, respectively, with initial conditions represented by~\eqref{eq:DOCPdynIC} and~\eqref{eq:DOCPdPIC}.
-Eqs.~\eqref{eq:DOCPpath} are path equality constraints and~\eqref{eq:DOCPconstraints} are the robustified inequality constraints.
-
-In certain cases~\cite{Gillis:PracticalMethodsApproximate:2015}, positive-definiteness-preserving Lyapunov discretization schemes are used in~\eqref{eq:DOCPdP}.
-However, for sufficiently fine discretization, we found that integrating only the lower triangular part of $\bP(t)$ in~\eqref{eq:OCPdPIC}  (and reconstructing its strictly upper part accordingly), ensures that $\bP_k$ remain symmetric and positive definite when the initial $\bP_0$ is symmetric and positive definite.
-In all our tests direct collocation with cubic polynomial state representations and Gauss-Legendre collocation points provided accurate results.

### 2025-04-28 open_loop_planning.tex
[MM] tikz sources folder created

@@
 	\centering
-	\begin{tikzpicture}[%
-		smallnode/.style={%
-			circle, draw, minimum size=3mm, inner sep=0pt
-		},
-		r_smallnode/.style={%
-			circle, draw, minimum size=3mm, inner sep=0pt, red
-		},
-		g_smallnode/.style={%
-			circle, draw, minimum size=3mm, inner sep=0pt, mygreen
-		},
-		every path/.style={->, thick},
-		x=1.5cm, y=1.5cm
-		]
-
-		\newcommand{\Plab}[2]{\bP^{#2}_{#1}}
-		\definecolor{mygreen}{RGB}{0 150 0}
-
-		% Nodes
-		% First row
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-2}{H-2}\)}] (Pk-2H-2) at (1,0) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{H-1}\)}] (Pk-1H-1) at (2,0) {};
-		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k}{H}\)}] (PkH) at (3,0) {};
-		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{0}\)}] (Pk+10) at (4,0) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{1}\)}] (Pk+21) at (5,0) {};
-
-		\draw[bend left=20] (Pk-2H-2) to (Pk-1H-1);
-		\draw[bend left=20] (Pk-1H-1) to (PkH);
-		\draw[bend left=20] (Pk+10) to (Pk+21);
-
-		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k-H}{0}\)}] (Pk-H0) at (0,0) {};
-		\node[label={center:\(\dots\)}] (dots10) at (-.5, 0) {};
-		\node[label={center:\(\dots\)}] (dots11) at (.5, 0) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+H}{H-1}\)}] (Pk+HH-1) at (6,0) {};
-		\node[label={center:\(\dots\)}] (dots12) at (5.5, 0) {};
-		\node[label={center:\(\dots\)}] (dots13) at (6.5, 0) {};
-		% Second row
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-2}{H-3}\)}] (Pk-2H-3) at (1,1) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{H-2}\)}] (Pk-1H-2) at (2,1) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k}{H-1}\)}] (PkH-1) at (3,1) {};
-		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{H}\)}] (Pk+1H) at (4,1) {};
-		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{0}\)}] (Pk+20) at (5,1) {};
-
-		\draw[bend left=20] (Pk-2H-3) to (Pk-1H-2);
-		\draw[bend left=20] (Pk-1H-2) to (PkH-1);
-		\draw[bend left=20] (PkH-1) to (Pk+1H);
-
-		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k-H}{H}\)}] (Pk-HH) at (0,1) {};
-		\node[label={center:\(\dots\)}] (dots20) at (-.5, 1) {};
-		\node[label={center:\(\dots\)}] (dots21) at (.5, 1) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+H}{H-2}\)}] (Pk+HH-2) at (6,1) {};
-		\node[label={center:\(\dots\)}] (dots22) at (5.5, 1) {};
-		\node[label={center:\(\dots\)}] (dots23) at (6.5, 1) {};
-
-		% dots row
-		\node[label={center,rotate=90:\(\dots\)}] (dots1) at (0, 1.7) {};
-		\node[label={center,rotate=90:\(\dots\)}] (dots2) at (1, 1.7) {};
-		\node[label={center,rotate=90:\(\dots\)}] (dots3) at (2, 1.7) {};
-		\node[label={center,rotate=90:\(\dots\)}] (dots4) at (3, 1.7) {};
-		\node[label={center,rotate=90:\(\dots\)}] (dots5) at (4, 1.7) {};
-		\node[label={center,rotate=90:\(\dots\)}] (dots6) at (5, 1.7) {};
-		\node[label={center,rotate=90:\(\dots\)}] (dots7) at (6, 1.7) {};
-
-		% Third row
-		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k-2}{H}\)}] (Pk-2H) at (1,2) {};
-		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{0}\)}] (Pk-10) at (2,2) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k}{1}\)}] (Pk1) at (3,2) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{2}\)}] (Pk+12) at (4,2) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{3}\)}] (Pk+23) at (5,2) {};
-
-		\draw[bend left=20] (Pk-10) to (Pk1);
-		\draw[bend left=20] (Pk1) to (Pk+12);
-		\draw[bend left=20] (Pk+12) to (Pk+23);
-
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-H}{2}\)}] (Pk-H2) at (0,2) {};
-		\node[label={center:\(\dots\)}] (dots30) at (-.5, 2) {};
-		\node[label={center:\(\dots\)}] (dots31) at (.5, 2) {};
-		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k+H}{0}\)}] (Pk+H0) at (6,2) {};
-		\node[label={center:\(\dots\)}] (dots32) at (5.5, 2) {};
-		\node[label={center:\(\dots\)}] (dots33) at (6.5, 2) {};
-
-		% Fourth row
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-2}{H-1}\)}] (Pk-2H-1) at (1,3) {};
-		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{H}\)}] (Pk-1H) at (2,3) {};
-		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k}{0}\)}] (Pk0) at (3,3) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{1}\)}] (Pk+11) at (4,3) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{2}\)}] (Pk+22) at (5,3) {};
-
-		\draw[bend left=20] (Pk-2H-1) to (Pk-1H);
-		\draw[bend left=20] (Pk0) to (Pk+11);
-		\draw[bend left=20] (Pk+11) to (Pk+22);
-
-		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-H}{1}\)}] (Pk-H1) at (0,3) {};
-		\node[label={center:\(\dots\)}] (dots40) at (-.5, 3) {};
-		\node[label={center:\(\dots\)}] (dots41) at (.5, 3) {};
-		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k+H}{H}\)}] (Pk+HH) at (6,3) {};
-		\node[label={center:\(\dots\)}] (dots42) at (5.5, 3) {};
-		\node[label={center:\(\dots\)}] (dots43) at (6.5, 3) {};
-
-		% Dashed rectangle
-		\draw[dashed, thick, rounded corners] (2.6,-.25) rectangle (3.4,3.5);
-
-	\end{tikzpicture}
+	\input{Fig/Tikz_Sources/tikz_covariance_evolution_matteo.tex}
 	\caption{Schematic representation of the continuity and initialization equations associated with the $H+1$ propagated instances of the covariance matrix introduced in problem~\eqref{eq:DOCPopenloop}.
@@
 	\definecolor{mygreen}{RGB}{0,150,0}
-	\begin{tikzpicture}[
-		smallnode/.style={circle, draw, minimum size=3mm, inner sep=0pt},
-		r_smallnode/.style={circle, draw=red, minimum size=3mm, inner sep=0pt},
-		g_smallnode/.style={circle, draw=mygreen, minimum size=3mm, inner sep=0pt},
-		r_hatnode/.style={
-			circle, draw=red, pattern=north east lines, pattern color=red,
-			minimum size=3mm, inner sep=0pt},
-		g_hatnode/.style={
-			circle, draw=mygreen, pattern=north east lines, pattern color=mygreen,
-			minimum size=3mm, inner sep=0pt},
-		every path/.style={->, thick},
-		x=1.4cm, y=1.2cm
-		]
-
-		% ROW j = 0 (y=3)
-		\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k-3}^0\)}] (P00) at (1,3) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k-2}^1\)}] (P11) at (2,3) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k-1}^2\)}] (P22) at (3,3) {};
-		\node[g_hatnode, label={[label distance=-2pt]above:\(\bP_k^3\)}] (P33) at (4,3) {};
-		\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k+1}^0\)}] (P44) at (5,3) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+2}^1\)}] (P55) at (6,3) {};
-
-		% ROW j = 1 (y=2)
-		\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k-2}^0\)}] (P10) at (2,2) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k-1}^1\)}] (P21) at (3,2) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\bP_k^2\)}] (P32) at (4,2) {};
-		\node[g_hatnode, label={[label distance=-2pt]above:\(\bP_{k+1}^3\)}] (P43) at (5,2) {};
-		\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k+2}^0\)}] (P54) at (6,2) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+3}^1\)}] (P65) at (7,2) {};
-
-		% ROW j = 2 (y=1)
-		\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k-1}^0\)}] (P20) at (3,1) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\bP_k^1\)}] (P31) at (4,1) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+1}^2\)}] (P42) at (5,1) {};
-		\node[g_hatnode, label={[label distance=-2pt]above:\(\bP_{k+2}^3\)}] (P53) at (6,1) {};
-		\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k+3}^0\)}] (P64) at (7,1) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+4}^1\)}] (P75) at (8,1) {};
-
-		% ROW j = 3 (y=0)
-		\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_k^0\)}] (P30) at (4,0) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+1}^1\)}] (P41) at (5,0) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+2}^2\)}] (P52) at (6,0) {};
-		\node[g_hatnode, label={[label distance=-2pt]above:\(\bP_{k+3}^3\)}] (P63) at (7,0) {};
-		\node[r_hatnode, label={[label distance=-2pt]above:\(\bP_{k+4}^0\)}] (P74) at (8,0) {};
-		\node[smallnode, label={[label distance=-2pt]above:\(\bP_{k+5}^1\)}] (P85) at (9,0) {};
-
-		% ARROWS
-		\draw (P00) -- (P11);
-		\draw (P11) -- (P22);
-		\draw (P22) -- (P33);
-		\draw (P22) -- (P33);
-
-		\draw (P44) -- (P55);
-		\draw[dashed,->] (P55) -- ++(1,0);
-
-		\draw (P10) -- (P21);
-		\draw (P21) -- (P32);
-		\draw (P32) -- (P43);
-		\draw (P54) -- (P65);
-		\draw[dashed,->] (P65) -- ++(1,0);
-
-		\draw (P20) -- (P31);
-		\draw (P31) -- (P42);
-		\draw (P42) -- (P53);
-		\draw (P64) -- (P75);
-		\draw[dashed,->] (P75) -- ++(1,0);
-
-		\draw (P30) -- (P41);
-		\draw (P41) -- (P52);
-		\draw (P52) -- (P63);
-
-		\draw (P74) -- (P85);
-		\draw[dashed,->] (P85) -- ++(1,0);
-
-		% Dashed box around column k (x=4)
-		\draw[dashed, thick, rounded corners] (3.6,-0.5) rectangle (4.4,3.7);
-
-		% Top axis labels
-		\node at (0, 4.3) {\scriptsize \(\cdots\)};
-		\node at (1, 4.3) {\scriptsize \(k-3\)};
-		\node at (2, 4.3) {\scriptsize \(k-2\)};
-		\node at (3, 4.3) {\scriptsize \(k-1\)};
-		\node (knode) at (4, 4.3) {\scriptsize \(k\)};
-		\node at (5, 4.3) {\scriptsize \(k+1\)};
-		\node at (6, 4.3) {\scriptsize \(k+2\)};
-		\node at (7, 4.3) {\scriptsize \(k+3\)};
-		\node at (8, 4.3) {\scriptsize \(k+4\)};
-		\node at (9, 4.3) {\scriptsize \(\cdots\)};
-
-		% Wiggly arrow from label to top edge of dashed box
-		\draw[->, thick, decorate, decoration={snake, amplitude=0.25mm, segment length=2.5mm}]
-		(knode.south) -- (4, 3.8);
-
-		% Legend box (top right)
-		%\begin{scope}[shift={(6.5,2.5)}]
-		%    \node[r_smallnode] (r) at (0,0) {};
-		%    \node[anchor=west] at (0.4,0) {\scriptsize Initialized matrix \(\bP_k^0\)};
-		%
-		%    \node[g_smallnode] (g) at (0,-0.8) {};
-		%    \node[anchor=west] at (0.4,-0.8) {\scriptsize Most propagated matrix \(\bP_k^3\)};
-		%\end{scope}
-
-	\end{tikzpicture}
-
+	\input{Fig/Tikz_Sources/tikz_covariance_evolution.tex}
 	\caption{Cartoon representation (depicted for $H=3$) of the continuity and initialization equations associated with the $H+1$ propagated instances of the covariance matrix introduced in problem~\eqref{eq:DOCPopenloop}.

### 2025-04-28 stochastic_vehicle_dynamics.tex
[MM] tikz sources folder created

@@
 	\centering
-	\begin{tikzpicture}[scale=2, x = {(0,1cm)}, y = {(-1cm,0)},
-		force/.style={->, thick, red!60!black, >={Latex[length=10pt, width=4pt]}},
-		vec/.style={->, thick, mgreen, >={Latex[length=10pt, width=4pt]}},
-		aux/.style={thin},
-		frame/.style={->,thin,>={Latex}},
-		every node/.style={font=\small, text=black}
-		]
-		\definecolor{mgreen}{RGB}{0,100,0}
-		\begin{scope}[rotate = 25,shift = {(-.3,-2)}]
-			% Coordinates
-			\coordinate (rear) at (0,0);
-			\coordinate (G) at (1.2,0);
-			\coordinate (front) at (2.8,0);
-
-			% Main body line
-			\draw[blue!50!black, thick] (rear) -- (front);
-
-			% Rear wheel
-			\draw[fill=black!60!white, draw=black, rounded corners=2pt]
-			($(rear)+(-.2,.1)$) rectangle ($(rear)+(+.2,-.1)$);
-
-			% Front wheel (rotated around its center)
-			\begin{scope}[shift={(front)}, rotate=25]
-				\draw[fill=black!60!white, draw=black, rounded corners=2pt]
-				(-0.2,.1) rectangle (0.2,-.1);
-			\end{scope}
-
-			% Local axes at rear
-			\draw[force] (rear) -- ++(.5,0) node[above,right] {$X_2$};
-			\draw[force] (rear) -- ++(0,1.3) node[above] {$Y_2$};
-
-			% Local axes at front
-			\draw[force] (front) -- ++(-.3,0) node[right] {$X_1$};
-			\draw[force] (front) -- ++(0,1.5) node[above] {$Y_1$};
-
-			% Velocity vectors at G
-			\draw[vec] (G) -- ++(.6,0) node[left, shift={(-.2,-.1)}] {$u\,\mathbf{i}$};
-			\draw[vec] (G) -- ++(0,.3) node[above] {$v\,\mathbf{j}$};
-
-			% Angular velocity
-			\begin{scope}[rotate = -115]
-				\draw[vec] 
-				(-.1,.8) arc[start angle=0, end angle=50, radius=1]
-				node[midway, right=2pt] {$r\,\mathbf{k}$};
-			\end{scope}
-
-			% G point
-			\fill (G) circle (1pt) node[right] {$G$};
-
-			% Body frame
-			\draw[frame] (G) --++ (1,0) node[right] {$x_b$};
-			\draw[frame] (G) --++ (0,1) node[above] {$y_b$};
-
-			% Distances
-%			\draw[<->,>={Latex}] (3,-1) -- node[left] {$a_1$} (1.5,-1);
-%			\draw[<->,>={Latex}] (1.5,-1) -- node[left] {$a_2$} (0,-1);
-
-			% Steering angle (auxiliary vertical line)
-			\draw[aux] (front) -- ++(.4,0);
-			\draw[aux] (front) -- ++({0.4*cos(25)}, {0.4*sin(25)});
-
-			% Steering angle arc
-			\draw[->,thin,>={Latex}] ([shift={(0.35,0)}]front) arc[start angle=0, end angle=25, radius=0.35];
-			\node[left,shift={(.15,0)}] at ([shift={(0.35,0)}]front) {$\delta$};
-
-			% Psi angle aux
-			\draw[aux] (G) -- ++({0.4*cos(25)}, {-0.4*sin(25)});
-
-			% Psi angle
-			\draw[->,thin,>={Latex}] ([shift={({0.35*cos(25)},{-.35*sin(25)})}]G) arc[start angle=-25, end angle=0, radius=0.35];
-			\node[left,shift={(.2,-.05)}] at ([shift={({0.35*cos(25)},{-.35*sin(25)})}]G) {$\psi$};
-
-		\end{scope}
-		\coordinate (ground) at (0,0);
-		\draw[frame] (ground) --++ (1,0) node[right] {$x$};
-		\draw[frame] (ground) --++ (0,1) node[above] {$y$};
-		\draw[->,thick,>=Latex] (ground) --++ (G);
-%		\draw[->,thick,>=Latex] (ground) --++ (G) node[midway, shift={(0,-.3)}] {$\boldsymbol{p}$};
-
-	\end{tikzpicture}
+	\input{Fig/Tikz_Sources/tikz_vehicle_model.tex}
 	\caption{Single Track model.
@@
 \begin{align}
-& \underset{\bmu(t), \bu(t), \bP(t)}{\text{minimize}} \quad J(\bmu(t), \bu(t), \bP(t)) \label{eq:OCPcost} \\
-& \text{s.t.} \quad \dbmu(t)           = f(\bmu(t), \bu(t)) \label{eq:OCPdyn} \\
-& \phantom{\text{s.t.} \quad} \bmu(0)  = \bmu_0 \label{eq:OCPdynIC} \\
-& \phantom{\text{s.t.} \quad} \dbP(t) = \bA(t) \bP(t) + \bP(t) \bA^T(t) + \bQ(t) \label{eq:OCPdP} \\
-& \phantom{\text{s.t.} \quad} \bP(0)  = \bP_0 \succeq 0 \label{eq:OCPdPIC} \\ %& \phantom{\text{s.t.} \qquad} 0       \geq h_i(\bmu(t), \bu(t))
+	\underset{\bmu(t), \bu(t), \bP(t)}{\text{minimize}} \quad & J(\bmu(t), \bu(t), \bP(t)) \label{eq:OCPcost} \\
+	\text{s.t.} \quad \dbmu(t)           &= f(\bmu(t), \bu(t)) \label{eq:OCPdyn} \\
+	\phantom{\text{s.t.} \quad} \bmu(0)  &= \bmu_0 \label{eq:OCPdynIC} \\
+	\phantom{\text{s.t.} \quad} \dbP(t) &= \bA(t) \bP(t) + \bP(t) \bA^T(t) + \bQ(t) \label{eq:OCPdP} \\
+	\phantom{\text{s.t.} \quad} \bP(0)  &= \bP_0 \succeq 0 \label{eq:OCPdPIC} \\ %& \phantom{\text{s.t.} \qquad} 0       \geq h_i(\bmu(t), \bu(t))
 	%+ \overbrace{\gamma \bigg[\underbrace{\na^T_{\bx} h_i \bP(t) \na_{\bx} h_i}_{\text{variance of $h_i(\bx)$}}\bigg]^{\frac{1}{2}}}^{\text{safety margin}},
-&\phantom{\text{s.t.} \qquad} 0       \geq h_i(\bmu(t), \bu(t))
+	\phantom{\text{s.t.} \qquad} 0&       \geq h_i(\bmu(t), \bu(t))
 	+ \be_i(\bmu(t), \bu(t), \bP(t)),
@@
 	\centering
-	\begin{tikzpicture}[scale = .7]
-		% Parameters
-		\def\outerX{3}   % x radius of outer ellipse
-		\def\outerY{2}   % y radius of outer ellipse
-		\def\gap{0.2}    % distance between outer and inner ellipse
-		\def\innerX{2.68}
-		\def\innerY{1.79}
-
-		% Draw outer ellipse
-		\draw[] (0,0) ellipse [x radius=\outerX, y radius=\outerY];
-
-		% Draw inner ellipse filled with color
-		\fill[black!30] (0,0) ellipse [
-		x radius={\innerX},
-		y radius={\innerY}
-		];
-		\draw[dashed] (0,0) ellipse [
-		x radius={\innerX},
-		y radius={\innerY}
-		];
-
-		% Draw axes
-		\draw[->] (-\outerX-1,0) -- (\outerX+1,0) node[right] {$X_j$};
-		\draw[->] (0,-\outerY-1) -- (0,\outerY+1) node[above] {$Y_j$};
-
-		% Arrows pointing to the ellipses
-		\draw[->,thin] (1.7,2) -- ({\outerX*cos(60)},{\outerY*sin(60)}) node[midway, above right] {$\beta=0$};
-		\draw[->,thin] (2.3,1.7) -- ({(\innerX)*cos(45)},{(\innerY)*sin(45)}) node[midway, right, shift={(0.05,.2)}] {$\beta=\gap$};
-
-		% Arrows pointing to the ellipses
-		\draw[->,thin] ({\outerX+.3},.3) -- (\outerX,0) node[midway, above right,shift={(.05,0)}] {$\mu_x Z_j$};
-		\draw[->,thin] (.3,{\outerY+.3}) -- (0,\outerY) node[midway, above right,shift={(.05,0)}] {$\mu_y Z_j$};
-
-	\end{tikzpicture}
+	\input{Fig/Tikz_Sources/tikz_adherence_constraint.tex}
 	\hfill
-	\begin{tikzpicture}[scale = .7]
-		% Parameters
-		\def\outerX{3}   % x radius of outer ellipse
-		\def\outerY{2}   % y radius of outer ellipse
-		\def\gap{0.2}    % distance between outer and inner ellipse
-		\def\innerX{2.68}
-		\def\innerY{1.79}
-
-		% Draw outer ellipse
-		\draw[] (0,0) ellipse [x radius=\outerX, y radius=\outerY];
-
-		% Draw inner ellipse filled with color
-		\fill[black!30] (0,0) ellipse [
-		x radius={\innerX},
-		y radius={\innerY}
-		];
-		\node[above] at (0,0) {PATH CONSTRAINT FIG};
-		\draw[dashed] (0,0) ellipse [
-		x radius={\innerX},
-		y radius={\innerY}
-		];
-
-		% Draw axes
-		\draw[->] (-\outerX-1,0) -- (\outerX+1,0) node[right] {$X_i$};
-		\draw[->] (0,-\outerY-1) -- (0,\outerY+1) node[above] {$Y_i$};
-
-		% Arrows pointing to the ellipses
-		\draw[->,thin] (1.7,2) -- ({\outerX*cos(60)},{\outerY*sin(60)}) node[midway, above right] {$\beta=0$};
-		\draw[->,thin] (2.3,1.7) -- ({(\innerX)*cos(45)},{(\innerY)*sin(45)}) node[midway, right, shift={(0.05,.2)}] {$\beta=\gap$};
-
-		% Arrows pointing to the ellipses
-		\draw[->,thin] ({\outerX+.3},.3) -- (\outerX,0) node[midway, above right,shift={(.05,0)}] {$\mu_x Z_i$};
-		\draw[->,thin] (.3,{\outerY+.3}) -- (0,\outerY) node[midway, above right,shift={(.05,0)}] {$\mu_y Z_i$};
-
-	\end{tikzpicture}
+	\input{Fig/Tikz_Sources/tikz_stayontrack_constraint.tex}
 	\caption{Graphical representation of the two constraints analyzed in the section.

### 2025-04-28 stochastic_vehicle_dynamics.tex
[MM] Stay-on-track constraint partial

@@
 \subsection{Robust adherence limit constraint formulation}
+\label{sec:adherence_constraint}
 The model employs a tire formulation that neglects combined slip effects, i.e., it does not account for the simultaneous utilization of longitudinal and lateral tyre forces.
@@
 \end{figure}
+
+\subsection{Stay-on-track constraint}
+\label{sec:stayontrack_constraint}
+A stay-on-track constraint is required to ensure that the vehicle remains within the circuit boundaries.
+The formulation adopted introduces an algebraic variable $e$, which satisfies the following constraint:
+\begin{equation}
+	\begin{Bmatrix}
+		x\\y
+	\end{Bmatrix}
+	-\boldsymbol{c} - e\bn=\boldsymbol{0} \label{eq:onplane_constraint}
+\end{equation}
+%\begin{equation}
+%	\bp-\boldsymbol{c}_k - e_k\bn_k=\boldsymbol{0} \label{eq:onplane_constraint}
+%\end{equation}
+where $\boldsymbol{c}$ indicates the position of the centerline, and $\bn$ is the unit normal vector to the centerline.
+Equation~\eqref{eq:onplane_constraint} constraints the CoM of the vehicle to lie on the vertical plane defined by the unit normal vector $\bn$.
+The previously introduced algebraic variable $e$ is further constrained to take values between $e_\textrm{min}$ and $e_\textrm{max}$, which are determined based on the width of the circuit and the vehicle's front and rear tracks.
+
+In order to take into account for the back-off terms, the maximum and minimum values are modified coherently, so that:
+\begin{equation}
+	e_\textrm{min} + \be^\textrm{SOT} \leq e \leq e_\textrm{max} - \be^\textrm{SOT}\label{eq:stayontrack_constraint}
+\end{equation}
+where $\be^\textrm{SOT}$ is computed by considering that the gradient of the constraint w.r.t. the state vector $\na_{\bx}h^\textrm{SOT}$ depends solely on $\bn$.
+%Adding a back-off to this constraint produces the same effect as an increase of the front and rear tracks of the vehicle does.

### 2025-05-02 intro.tex
[MGa] some notes in intro

@@
 
+Some key aspects: The omission of explicit disturbance modeling within the planning framework can critically compromise the satisfaction of operational constraints.
+In particular, time-optimal planning algorithms frequently yield trajectories that lie in close proximity to constraint boundaries, such as those associated with collision avoidance.
+Under such conditions, even small perturbations may induce constraint violations, thereby significantly jeopardizing system safety.
+Although the introduction of fixed, heuristically selected safety margins around constraint sets may offer a nominal safeguard, this strategy does not provide formal assurances of constraint satisfaction under perturbations and often leads to overly conservative and suboptimal solutions.
+
 \subsection{Related work}
 In recent years, extensive research has been conducted on ...
+
 \subsection{Paper's contributions}

### 2025-05-02 intro.tex
[MGa] sketching intro

--- 2025-04-03 intro.tex
@@
 \subsection{Related work}
-In recent years, extensive research has been conducted on ... 
+In recent years, extensive research has been conducted on minimum-lap-time planning for motorsport vehicles.
+Cite: Massaro, Biral, noi...
+with different formulations, vehicle models and solution strategies.
+
+%We try to fill the gap that the omission of explicit disturbance modeling within the planning framework can critically compromise the satisfaction of operational constraints.
+In particular, time-optimal planning algorithms frequently yield trajectories that lie in close proximity to constraint boundaries, such as those associated with collision avoidance.
+Under such conditions, even small perturbations may induce constraint violations, thereby significantly jeopardizing system safety.
+%
+%Although the introduction of fixed, heuristically selected safety margins around constraint sets may offer a nominal safeguard, this strategy does not provide formal assurances of constraint satisfaction under perturbations and often leads to overly conservative and suboptimal solutions.
+%
+%What people do in related fields to consider uncertainty?
+
+%In orbital mechanics, since the state of the system is uncertain, it is important to estimate the collision event between satellites and space debris in probabilistic terms~\cite{Tapley:StatisticalOrbitDetermination:2004}.
+%
+%In robotics it is important, in the presence of uncertainty, to avoid collisions between autonomous mobile platforms and obstacles.
+One of the first contribution where a probabilistic representation of uncertainty was proposed, i.e.
+a chance-constrained approach, is~\cite{Blackmore:ChanceConstrainedOptimalPath:2011}.
+This approach plans the future probabilistic distribution of the vehicle state so that the probability of failure is below a specified threshold.
+%
+%In Model Predictive Control approaches, strategies to include probabilistic safety with respect to collisions have been proposed~\cite{Gao:CollisionfreeMotionPlanning:2023} (Zo-Ro), \cite{Zhang:RobustifiedTimeoptimalPointtopoint:2025} and~\cite{Zhang:RobustifiedTimeoptimalCollisionfree:2024} (Joris).
+The latter consider also a trajectory tracking implementation for an autonomous mobile robot under bounded process noise with pre-computed feedback gain.
+In the Zo-Ro methods, a zero-order based algorithm is presented to also optimize the feedback gain in the OCP for counteracting the growth of state uncertainty.
+%A heuristic MPC-based approach using $n$-steps-ahead uncertainty predictions for constraint-tightening calculations have been proposed in chemical process engineering for the control of an industrial polymerization reactor~\cite{Krog:SimpleFastRobust:2024}.
+%
+%In some cases uncertainty can be attributed to partial knowledge of system parameters.
+A wealth of research on this topic mainly devoted to trajectory planning of Unmanned Aerial Vehicles (UAVs) with minimal state sensitivity has been presented in~\cite{Brault:RobustTrajectoryPlanning:2021} and~\cite{Giordano:TrajectoryGenerationMinimum:2018} \cite{Brault:TubebasedTrajectoryOptimization:} where also closed-loop and input sensitivity in the nonlinear constraints are considered.
+Other research deal with the problem of how to combine closed-loop state sensitivity and input sensitivity-based planning with observability-aware planning.
+This approach is given a multi-step single-objective optimization framework in~\cite{Bohm:COPControlObservabilityaware:2022}.
+
+The omission of explicit disturbance modeling within planning frameworks can critically undermine the satisfaction of operational constraints.
+In particular, time-optimal planning algorithms often yield trajectories that lie in close proximity to constraint boundaries - such as those associated with collision avoidance - where even small perturbations may result in constraint violations, thereby significantly compromising system safety.
+
+While the use of fixed, heuristically defined safety margins around constraint sets may provide a nominal safeguard, such an approach lacks formal guarantees under uncertainty and typically leads to overly conservative and suboptimal solutions.
+This motivates the need for planning methods that explicitly account for uncertainty and provide quantifiable safety assurances.
+
+In several related fields, various strategies have been developed to address planning under uncertainty.
+
+\textit{Orbital Mechanics.} In orbital mechanics, uncertainty in the state estimate necessitates a probabilistic treatment of collision events between satellites and space debris.
+A foundational reference in this context is~\cite{Tapley:StatisticalOrbitDetermination:2004}, where statistical orbit determination techniques are employed to assess and mitigate collision risks.
+
+\textit{Robotics.} In robotic motion planning, uncertainty-aware techniques are crucial to ensure safe navigation in environments populated with obstacles.
+One of the earliest contributions proposing a probabilistic representation of uncertainty is the chance-constrained framework in~\cite{Blackmore:ChanceConstrainedOptimalPath:2011}, which plans over the predicted probabilistic distribution of the system state to ensure that the probability of constraint violation remains below a specified threshold.
+
+\textit{Model Predictive Control (MPC).} Within the MPC paradigm, recent works have incorporated probabilistic safety guarantees.
+Notably, the methods presented in~\cite{Gao:CollisionfreeMotionPlanning:2023},~\cite{Zhang:RobustifiedTimeoptimalPointtopoint:2025}, and~\cite{Zhang:RobustifiedTimeoptimalCollisionfree:2024} introduce stochastic MPC frameworks for autonomous mobile platforms, where process noise is explicitly modeled and closed-loop tracking performance is maintained via either pre-computed or optimized feedback gains.
+The approach in~\cite{Gao:CollisionfreeMotionPlanning:2023} additionally proposes a zero-order optimization scheme to include the feedback gain directly in the optimal control problem, mitigating the growth of uncertainty over the planning horizon.
+
+\textit{Chemical Process Control.} In the domain of chemical engineering, robust MPC approaches have been proposed to address uncertainty in industrial settings.
+For example,~\cite{Krog:SimpleFastRobust:2024} presents a heuristic method based on $n$-step-ahead uncertainty predictions, which are used to compute constraint tightening margins for the control of a polymerization reactor.
+
+\textit{Planning with Parametric Uncertainty.} In some contexts, uncertainty arises from partial or imprecise knowledge of system parameters.
+This has led to a significant body of work on robust trajectory planning, particularly for unmanned aerial vehicles (UAVs), with the aim of minimizing sensitivity to state and input variations.
+For instance,~\cite{Brault:RobustTrajectoryPlanning:2021},~\cite{Giordano:TrajectoryGenerationMinimum:2018}, and~\cite{Brault:TubebasedTrajectoryOptimization:} propose tube-based and sensitivity-aware optimization frameworks that explicitly account for both input and closed-loop state sensitivity in the planning phase.
+Further developments integrate observability-aware planning into a unified multi-step optimization framework, as demonstrated in~\cite{Bohm:COPControlObservabilityaware:2022}.
+
 \subsection{Paper's contributions}
-In this study, we investigate ...
+What we propose. Two new robust uncertainty-aware time-optimal planning strategies for motorsport vehicles.

### 2025-05-05 intro.tex
[MGa] added discussion on LLE
and why they are of limited use in this context

@@
 
-In several related fields, various strategies have been developed to address planning under uncertainty.
-
-\textit{Orbital Mechanics.} In orbital mechanics, uncertainty in the state estimate necessitates a probabilistic treatment of collision events between satellites and space debris.
+In several science fields, different strategies have been developed to address planning under uncertainty.
+In orbital mechanics, uncertainty in the state estimate necessitates a probabilistic treatment of collision events between satellites and space debris.
 A foundational reference in this context is~\cite{Tapley:StatisticalOrbitDetermination:2004}, where statistical orbit determination techniques are employed to assess and mitigate collision risks.
 
-\textit{Robotics.} In robotic motion planning, uncertainty-aware techniques are crucial to ensure safe navigation in environments populated with obstacles.
+In robotic motion planning, uncertainty-aware techniques are crucial to ensure safe navigation in environments populated with obstacles.
 One of the earliest contributions proposing a probabilistic representation of uncertainty is the chance-constrained framework in~\cite{Blackmore:ChanceConstrainedOptimalPath:2011}, which plans over the predicted probabilistic distribution of the system state to ensure that the probability of constraint violation remains below a specified threshold.
 
-\textit{Model Predictive Control (MPC).} Within the MPC paradigm, recent works have incorporated probabilistic safety guarantees.
+Within the Model Predictive Control paradigm, recent works have incorporated probabilistic safety guarantees.
 Notably, the methods presented in~\cite{Gao:CollisionfreeMotionPlanning:2023},~\cite{Zhang:RobustifiedTimeoptimalPointtopoint:2025}, and~\cite{Zhang:RobustifiedTimeoptimalCollisionfree:2024} introduce stochastic MPC frameworks for autonomous mobile platforms, where process noise is explicitly modeled and closed-loop tracking performance is maintained via either pre-computed or optimized feedback gains.
@@
 
-\textit{Chemical Process Control.} In the domain of chemical engineering, robust MPC approaches have been proposed to address uncertainty in industrial settings.
+In the domain of chemical engineering, robust MPC approaches have been proposed to address uncertainty in industrial settings.
 For example,~\cite{Krog:SimpleFastRobust:2024} presents a heuristic method based on $n$-step-ahead uncertainty predictions, which are used to compute constraint tightening margins for the control of a polymerization reactor.
 
-\textit{Planning with Parametric Uncertainty.} In some contexts, uncertainty arises from partial or imprecise knowledge of system parameters.
+In some contexts, uncertainty arises from partial or imprecise knowledge of system parameters.
 This has led to a significant body of work on robust trajectory planning, particularly for unmanned aerial vehicles (UAVs), with the aim of minimizing sensitivity to state and input variations.
@@
 
+%In other approaches, the largest Lyapunov exponent (LLE) is proposed as an indicator of local stability~\cite{Meng:AnalysisGlobalCharacteristics:2022}.
+Lyapunov exponents, equal in number to the system's dimension, characterize the average exponential rate at which nearby trajectories of a nonlinear dynamical system converge or diverge~\cite{Strogatz:NonlinearDynamicsChaos:2019}.
+With the exception of chaotic systems - where a positive LLE may still be associated with bounded motion on a strange attractor - a negative LLE indicates local stability, as nearby trajectories tend to converge.
+%
+%A notable property is that, for trajectories approaching a stable equilibrium, the largest Lyapunov exponent asymptotically converges to the maximum real part of the eigenvalues of the linearized system at the equilibrium point.
+
+Following a different route, the largest Lyapunov exponent (LLE) has been proposed as an indicator of local stability~\cite{Meng:AnalysisGlobalCharacteristics:2022}, as it quantifies the average exponential rate at which nearby trajectories of a nonlinear system converge (negative) or diverge (positive)~\cite{Strogatz:NonlinearDynamicsChaos:2019}.
+Except in chaotic systems where a positive LLE may still correspond to bounded motion on a strange attractor - a negative LLE typically signals local stability.
+Notably, for trajectories converging to a stable equilibrium, the LLE asymptotically approaches the largest real part of the eigenvalues of the linearized system at that point.
+
+This result has been leveraged in the literature to infer whether a system, not yet at equilibrium, is ultimately converging toward one, thereby enabling a localized assessment of stability based on the transient behavior~\cite{Sadri:StabilityAnalysisNonlinear:2013}.
+Applications of this method are particularly relevant in domains such as chaotic vessel motions~\cite{McCue:UseLyapunovExponents:2011} and rotorcraft dynamics~\cite{Tamer:StabilityNonlinearTimeDependent:2016,Cassoni:RotorcraftStabilityAnalysis:2024}, where the primary concern is understanding the long-term evolution of the system and determining its asymptotic behavior.
+
+%However, while Lyapunov-based indicators have been applied in vehicular contexts as well~\cite{Sadri:StabilityAnalysisNonlinear:2013,Meng:AnalysisGlobalCharacteristics:2022}, their relevance here is arguably of reduced interest - not due to their binary nature, but because even identifying asymptotic stability regions offers limited practical insight during planning.
+A trajectory may lie within the basin of attraction of a distant equilibrium, yet the long-term convergence may be of little operational use.
+What matters instead is whether, in the short term, a realistically modeled driver - acting in feedback or planning a bit into the future - can ensure that the vehicle's response to perturbations remains within safety-critical constraints, such as grip limits and collision avoidance.
+This perspective prioritizes short-horizon constraint-satisfying stabilizability over asymptotic notions of stability.
+
+Although Lyapunov-based indicators have been applied in vehicular contexts~\cite{Sadri:StabilityAnalysisNonlinear:2013,Meng:AnalysisGlobalCharacteristics:2022}, their utility in planning is limited by the fact that asymptotic stability - while theoretically informative - often lacks operational relevance.
+A point along the planned trajectory may belong to the basin of attraction of a stable equilibrium, but if that equilibrium lies far ahead in time or outside the operational domain (e.g., off track), its practical relevance is limited.
+
+%A point along a planned trajectory may lie within the basin of attraction of a stable equilibrium, yet if that equilibrium is reached only asymptotically, possibly over a long horizon or outside the intended operational domain (e.g., off track), its practical utility diminishes.
+In contrast, what is of primary interest in our setting is whether perturbations, even in transiently unstable conditions, can be effectively managed -- either through feedback or short-horizon planning -- so that the system remains within safety-critical constraints such as tire force limits and collision avoidance.
+This shifts the focus from long-term convergence toward evaluating whether the planned trajectory permits short-term stabilizability under uncertainty while ensuring constraint satisfaction.
+
 \subsection{Paper's contributions}

### 2025-05-08 intro.tex
[MGa] a better intro

@@
 \subsection{Related work}
-In recent years, extensive research has been conducted on minimum-lap-time planning for motorsport vehicles.
-Cite: Massaro, Biral, noi...
-with different formulations, vehicle models and solution strategies.
+In recent years, significant progress has been made in minimum-lap-time optimization for motorsport vehicles, encompassing fixed- and free-trajectory formulations~\cite{Veneri:FreetrajectoryQuasisteadystateOptimalcontrol:2020, Lovato:ThreedimensionalFixedtrajectoryApproaches:2022, Lovato:ThreedimensionalFreetrajectoryQuasisteadystate:2022}, comparisons between direct and indirect solution techniques~\cite{DalBianco:ComparisonDirectIndirect:2019, Bertolazzi:DirectIndirectApproach:2025}, and serial and parallel solver frameworks~\cite{Biniewicz:QuasisteadystateMinimumLap:2024, Bartali:SchwarzDecompositionParallel:2024, Bartali:ConsensusbasedAlternatingDirection:2024}.
 
-%We try to fill the gap that the omission of explicit disturbance modeling within the planning framework can critically compromise the satisfaction of operational constraints.
-In particular, time-optimal planning algorithms frequently yield trajectories that lie in close proximity to constraint boundaries, such as those associated with collision avoidance.
-Under such conditions, even small perturbations may induce constraint violations, thereby significantly jeopardizing system safety.
-%
-%Although the introduction of fixed, heuristically selected safety margins around constraint sets may offer a nominal safeguard, this strategy does not provide formal assurances of constraint satisfaction under perturbations and often leads to overly conservative and suboptimal solutions.
-%
-%What people do in related fields to consider uncertainty?
+Despite these advances, most planning frameworks still fail to incorporate robustness in the planning phase itself.
+Piccinini et al.~\cite{Piccinini:HowOptimalMinimumtime:2024} provide a direct comparison between an offline minimum-lap-time optimal control problem (MLT-OCP) and an online Artificial Racing Driver (ARD) that controls the very same high-fidelity vehicle model.
+They show that, by leveraging a physics-driven structure and a novel g-g-v performance constraint, ARD can achieve lap times within a few milliseconds of the offline benchmark and generalize to unseen circuits even under unmodeled mass variations.
+However, their work remains focused on quantifying the execution gap - how ARD mitigates local tracking errors - rather than on embedding disturbance handling directly into the trajectory planner.
 
-%In orbital mechanics, since the state of the system is uncertain, it is important to estimate the collision event between satellites and space debris in probabilistic terms~\cite{Tapley:StatisticalOrbitDetermination:2004}.
-%
-%In robotics it is important, in the presence of uncertainty, to avoid collisions between autonomous mobile platforms and obstacles.
-One of the first contribution where a probabilistic representation of uncertainty was proposed, i.e.
-a chance-constrained approach, is~\cite{Blackmore:ChanceConstrainedOptimalPath:2011}.
-This approach plans the future probabilistic distribution of the vehicle state so that the probability of failure is below a specified threshold.
-%
-%In Model Predictive Control approaches, strategies to include probabilistic safety with respect to collisions have been proposed~\cite{Gao:CollisionfreeMotionPlanning:2023} (Zo-Ro), \cite{Zhang:RobustifiedTimeoptimalPointtopoint:2025} and~\cite{Zhang:RobustifiedTimeoptimalCollisionfree:2024} (Joris).
-The latter consider also a trajectory tracking implementation for an autonomous mobile robot under bounded process noise with pre-computed feedback gain.
-In the Zo-Ro methods, a zero-order based algorithm is presented to also optimize the feedback gain in the OCP for counteracting the growth of state uncertainty.
-%A heuristic MPC-based approach using $n$-steps-ahead uncertainty predictions for constraint-tightening calculations have been proposed in chemical process engineering for the control of an industrial polymerization reactor~\cite{Krog:SimpleFastRobust:2024}.
-%
-%In some cases uncertainty can be attributed to partial knowledge of system parameters.
-A wealth of research on this topic mainly devoted to trajectory planning of Unmanned Aerial Vehicles (UAVs) with minimal state sensitivity has been presented in~\cite{Brault:RobustTrajectoryPlanning:2021} and~\cite{Giordano:TrajectoryGenerationMinimum:2018} \cite{Brault:TubebasedTrajectoryOptimization:} where also closed-loop and input sensitivity in the nonlinear constraints are considered.
-Other research deal with the problem of how to combine closed-loop state sensitivity and input sensitivity-based planning with observability-aware planning.
-This approach is given a multi-step single-objective optimization framework in~\cite{Bohm:COPControlObservabilityaware:2022}.
-
-The omission of explicit disturbance modeling within planning frameworks can critically undermine the satisfaction of operational constraints.
-In particular, time-optimal planning algorithms often yield trajectories that lie in close proximity to constraint boundaries - such as those associated with collision avoidance - where even small perturbations may result in constraint violations, thereby significantly compromising system safety.
-
+The omission of explicit disturbance modeling at the planning stage can critically undermine constraint satisfaction: time-optimal planners typically produce trajectories that closely approach safety boundaries (e.g., collision avoidance or tire-grip limits), where even minor perturbations may lead to violations and thus severely compromise system safety.
 While the use of fixed, heuristically defined safety margins around constraint sets may provide a nominal safeguard, such an approach lacks formal guarantees under uncertainty and typically leads to overly conservative and suboptimal solutions.
@@
 
-In several science fields, different strategies have been developed to address planning under uncertainty.
+In diverse fields, a variety of strategies has been proposed to tackle planning under uncertainty.
 In orbital mechanics, uncertainty in the state estimate necessitates a probabilistic treatment of collision events between satellites and space debris.
 A foundational reference in this context is~\cite{Tapley:StatisticalOrbitDetermination:2004}, where statistical orbit determination techniques are employed to assess and mitigate collision risks.
-
 In robotic motion planning, uncertainty-aware techniques are crucial to ensure safe navigation in environments populated with obstacles.
 One of the earliest contributions proposing a probabilistic representation of uncertainty is the chance-constrained framework in~\cite{Blackmore:ChanceConstrainedOptimalPath:2011}, which plans over the predicted probabilistic distribution of the system state to ensure that the probability of constraint violation remains below a specified threshold.
-
 Within the Model Predictive Control paradigm, recent works have incorporated probabilistic safety guarantees.
@@
 The approach in~\cite{Gao:CollisionfreeMotionPlanning:2023} additionally proposes a zero-order optimization scheme to include the feedback gain directly in the optimal control problem, mitigating the growth of uncertainty over the planning horizon.
-
 In the domain of chemical engineering, robust MPC approaches have been proposed to address uncertainty in industrial settings.
@@
 Further developments integrate observability-aware planning into a unified multi-step optimization framework, as demonstrated in~\cite{Bohm:COPControlObservabilityaware:2022}.
-
-%In other approaches, the largest Lyapunov exponent (LLE) is proposed as an indicator of local stability~\cite{Meng:AnalysisGlobalCharacteristics:2022}.
-Lyapunov exponents, equal in number to the system's dimension, characterize the average exponential rate at which nearby trajectories of a nonlinear dynamical system converge or diverge~\cite{Strogatz:NonlinearDynamicsChaos:2019}.
-With the exception of chaotic systems - where a positive LLE may still be associated with bounded motion on a strange attractor - a negative LLE indicates local stability, as nearby trajectories tend to converge.
-%
-%A notable property is that, for trajectories approaching a stable equilibrium, the largest Lyapunov exponent asymptotically converges to the maximum real part of the eigenvalues of the linearized system at the equilibrium point.
 
@@
 This perspective prioritizes short-horizon constraint-satisfying stabilizability over asymptotic notions of stability.
-
 Although Lyapunov-based indicators have been applied in vehicular contexts~\cite{Sadri:StabilityAnalysisNonlinear:2013,Meng:AnalysisGlobalCharacteristics:2022}, their utility in planning is limited by the fact that asymptotic stability - while theoretically informative - often lacks operational relevance.
@@
 %A point along a planned trajectory may lie within the basin of attraction of a stable equilibrium, yet if that equilibrium is reached only asymptotically, possibly over a long horizon or outside the intended operational domain (e.g., off track), its practical utility diminishes.
-In contrast, what is of primary interest in our setting is whether perturbations, even in transiently unstable conditions, can be effectively managed -- either through feedback or short-horizon planning -- so that the system remains within safety-critical constraints such as tire force limits and collision avoidance.
-This shifts the focus from long-term convergence toward evaluating whether the planned trajectory permits short-term stabilizability under uncertainty while ensuring constraint satisfaction.
+In our setting, the key question is whether perturbations -- even during transiently unstable phases -- can be handled via feedback or short-horizon replanning to keep the vehicle within safety-critical limits, such as tire forces and collision boundaries.
+This reframes the problem: rather than long-term convergence, we evaluate whether the planned trajectory ensures short-term stabilizability under uncertainty while maintaining constraint satisfaction.
 
 \subsection{Paper's contributions}
-What we propose. Two new robust uncertainty-aware time-optimal planning strategies for motorsport vehicles.
+We tackle the challenge of embedding robustness against disturbances directly into minimum-lap-time planning and propose two complementary methods.
+In the first one, the open-loop worst-case covariance propagation, at each discretization point, we propagate the state covariance forward over a fixed horizon and tighten all path constraints against the maximum covariance growth.
+By back-offing constraints using this \emph{worst-case} covariance, the resulting trajectory maximizes robustness without relying on feedback during planning.
+
+In the second one, the closed-loop covariance-aware planning, we integrate a time-varying LQR feedback policy into the planning process to realistically tame uncertainty growth.
+First, a nominal time-optimal trajectory is computed under deterministic dynamics.
+Second, an LQR controller is designed to stabilize that nominal path.
+Finally, we re-solve the planning problem - together with the feedback gains - so that the closed-loop covariance propagation satisfies tightened constraints via a Lyapunov-based formulation.
+
+Together, these two approaches strike a balance between computational simplicity and realistic, feedback-informed robustness, laying the groundwork for safe, high-performance lap-time planning under uncertainty.

### 2025-05-09 intro.tex
[MGa] paper's contributions and organization

@@
 \subsection{Related work}
-In recent years, significant progress has been made in minimum-lap-time optimization for motorsport vehicles, encompassing fixed- and free-trajectory formulations~\cite{Veneri:FreetrajectoryQuasisteadystateOptimalcontrol:2020, Lovato:ThreedimensionalFixedtrajectoryApproaches:2022, Lovato:ThreedimensionalFreetrajectoryQuasisteadystate:2022}, comparisons between direct and indirect solution techniques~\cite{DalBianco:ComparisonDirectIndirect:2019, Bertolazzi:DirectIndirectApproach:2025}, and serial and parallel solver frameworks~\cite{Biniewicz:QuasisteadystateMinimumLap:2024, Bartali:SchwarzDecompositionParallel:2024, Bartali:ConsensusbasedAlternatingDirection:2024}.
+In recent years, comprehensive analyses have been conducted on minimum-lap-time optimization for motorsport vehicles, covering fixed- and free-trajectory formulations~\cite{Veneri:FreetrajectoryQuasisteadystateOptimalcontrol:2020, Lovato:ThreedimensionalFixedtrajectoryApproaches:2022, Lovato:ThreedimensionalFreetrajectoryQuasisteadystate:2022}, comparing direct and indirect solution techniques~\cite{DalBianco:ComparisonDirectIndirect:2019, Bertolazzi:DirectIndirectApproach:2025}, and contrasting serial and parallel solver frameworks~\cite{Biniewicz:QuasisteadystateMinimumLap:2024, Bartali:SchwarzDecompositionParallel:2024, Bartali:ConsensusbasedAlternatingDirection:2024}.
 
@@
 In diverse fields, a variety of strategies has been proposed to tackle planning under uncertainty.                                          
-In orbital mechanics, uncertainty in the state estimate necessitates a probabilistic treatment of collision events between satellites and space debris.
+In orbital mechanics, uncertainty in state estimation necessitates a probabilistic framework for modeling interactions and potential close approaches among natural and artificial celestial bodies.
 A foundational reference in this context is~\cite{Tapley:StatisticalOrbitDetermination:2004}, where statistical orbit determination techniques are employed to assess and mitigate collision risks.
 In robotic motion planning, uncertainty-aware techniques are crucial to ensure safe navigation in environments populated with obstacles.
-One of the earliest contributions proposing a probabilistic representation of uncertainty is the chance-constrained framework in~\cite{Blackmore:ChanceConstrainedOptimalPath:2011}, which plans over the predicted probabilistic distribution of the system state to ensure that the probability of constraint violation remains below a specified threshold.
+One of the earliest contributions proposing a probabilistic representation of uncertainty is the chance-constrained framework in~\cite{Blackmore:ChanceConstrainedOptimalPath:2011}, which plans over the predicted distribution of the system state to ensure that the probability of constraint violation remains below a specified threshold.
 Within the Model Predictive Control paradigm, recent works have incorporated probabilistic safety guarantees.
@@
 
-Following a different route, the largest Lyapunov exponent (LLE) has been proposed as an indicator of local stability~\cite{Meng:AnalysisGlobalCharacteristics:2022}, as it quantifies the average exponential rate at which nearby trajectories of a nonlinear system converge (negative) or diverge (positive)~\cite{Strogatz:NonlinearDynamicsChaos:2019}.
-Except in chaotic systems where a positive LLE may still correspond to bounded motion on a strange attractor - a negative LLE typically signals local stability.
-Notably, for trajectories converging to a stable equilibrium, the LLE asymptotically approaches the largest real part of the eigenvalues of the linearized system at that point.
-
+Following an alternative approach, the largest Lyapunov exponent (LLE) has been proposed as an indicator of local stability~\cite{Meng:AnalysisGlobalCharacteristics:2022}, since it quantifies the average exponential rate at which infinitesimal perturbations off a fiducial trajectory grow or decay -- thus providing a direct measure of how deviations propagate through the system (negative values denote convergence toward the reference trajectory, while positive values indicate divergence)~\cite{Strogatz:NonlinearDynamicsChaos:2019}.
+%Except in chaotic systems where a positive LLE may still correspond to bounded motion on a strange attractor - a negative LLE typically signals local stability. 
+Notably, for trajectories converging to a stable equilibrium, the (negative) LLE asymptotically approaches the real part of the largest eigenvalues of the linearized system at that point.
 This result has been leveraged in the literature to infer whether a system, not yet at equilibrium, is ultimately converging toward one, thereby enabling a localized assessment of stability based on the transient behavior~\cite{Sadri:StabilityAnalysisNonlinear:2013}.
-Applications of this method are particularly relevant in domains such as chaotic vessel motions~\cite{McCue:UseLyapunovExponents:2011} and rotorcraft dynamics~\cite{Tamer:StabilityNonlinearTimeDependent:2016,Cassoni:RotorcraftStabilityAnalysis:2024}, where the primary concern is understanding the long-term evolution of the system and determining its asymptotic behavior.
+Applications of this method are particularly relevant to study chaotic vessel motions~\cite{McCue:UseLyapunovExponents:2011} and rotorcraft dynamics~\cite{Tamer:StabilityNonlinearTimeDependent:2016,Cassoni:RotorcraftStabilityAnalysis:2024}, where the primary concern is understanding the long-term evolution of the system and determining its asymptotic behavior.
 
@@
 This perspective prioritizes short-horizon constraint-satisfying stabilizability over asymptotic notions of stability.
-Although Lyapunov-based indicators have been applied in vehicular contexts~\cite{Sadri:StabilityAnalysisNonlinear:2013,Meng:AnalysisGlobalCharacteristics:2022}, their utility in planning is limited by the fact that asymptotic stability - while theoretically informative - often lacks operational relevance.
-A point along the planned trajectory may belong to the basin of attraction of a stable equilibrium, but if that equilibrium lies far ahead in time or outside the operational domain (e.g., off track), its practical relevance is limited.
+Although Lyapunov-based indicators have been applied in vehicular contexts as well~\cite{Sadri:StabilityAnalysisNonlinear:2013,Meng:AnalysisGlobalCharacteristics:2022}, their utility in planning is limited by the fact that asymptotic stability -- while theoretically informative -- often lacks operational relevance.
+A point along the planned trajectory may belong to the basin of attraction of a stable equilibrium, but if that equilibrium lies far ahead in time or outside the operational domain (e.g., off track), its practical relevance is questionable.
 
 %A point along a planned trajectory may lie within the basin of attraction of a stable equilibrium, yet if that equilibrium is reached only asymptotically, possibly over a long horizon or outside the intended operational domain (e.g., off track), its practical utility diminishes.
-In our setting, the key question is whether perturbations -- even during transiently unstable phases -- can be handled via feedback or short-horizon replanning to keep the vehicle within safety-critical limits, such as tire forces and collision boundaries.
+In our setting, the key question is whether perturbations occurring in unstable segments of the nominal trajectory can be handled via feedback or short-horizon replanning to keep the vehicle within safety-critical limits, such as tire forces and collision boundaries.
 This reframes the problem: rather than long-term convergence, we evaluate whether the planned trajectory ensures short-term stabilizability under uncertainty while maintaining constraint satisfaction.
 
-\subsection{Paper's contributions}
+\subsection{Paper's contributions and organization}
 We tackle the challenge of embedding robustness against disturbances directly into minimum-lap-time planning and propose two complementary methods.
-In the first one, the open-loop worst-case covariance propagation, at each discretization point, we propagate the state covariance forward over a fixed horizon and tighten all path constraints against the maximum covariance growth.
+In the first one -- the open-loop worst-case covariance propagation -- at each discretization point, we propagate the state covariance forward over a fixed horizon and tighten all path constraints against the maximum covariance growth.
 By back-offing constraints using this \emph{worst-case} covariance, the resulting trajectory maximizes robustness without relying on feedback during planning.
 
-In the second one, the closed-loop covariance-aware planning, we integrate a time-varying LQR feedback policy into the planning process to realistically tame uncertainty growth.
+In the second one -- the closed-loop covariance-aware planning, we integrate a time-varying LQR feedback policy into the planning process to realistically tame uncertainty growth.
 First, a nominal time-optimal trajectory is computed under deterministic dynamics.
@@
 Finally, we re-solve the planning problem - together with the feedback gains - so that the closed-loop covariance propagation satisfies tightened constraints via a Lyapunov-based formulation.
+Together, these two approaches strike a balance between computational simplicity and realistic, feedback-informed robustness, laying the groundwork for safe, high-performance lap-time planning under uncertainty.
 
-Together, these two approaches strike a balance between computational simplicity and realistic, feedback-informed robustness, laying the groundwork for safe, high-performance lap-time planning under uncertainty.
+The rest of the paper is organized as follows. 
+In Section~\ref{sec:svdf} we introduce our stochastic dynamics framework, recalling the single-track vehicle model with nonlinear tires used for planning, and formulate both the continuous-time planning problem and its discrete-time counterpart via direct-collocation.
+Here, we also derive the probabilistic constraint back-off formulation.
+ Section~\ref{sec:open_loop_planning} then outlines the open-loop worst-case covariance-propagation planning approach, while Section~\ref{sec:closed_loop_planning} presents the closed-loop, covariance-aware planner.
+In Section~\ref{sec:performance_comparison} we compare both methods on a representative track scenario, reporting adhesion and collision-avoidance results, parameter-sensitivity studies, and performance trade-offs.
+Finally, Section~\ref{sec:conclusions} draws conclusions and offers directions for future work.

### 2025-05-09 performance_comparison.tex
[MGa] paper's contributions and organization

@@
 \label{sec:performance_comparison}
-
-In this section, some of the new vehicle architectures are compared to the traditional vehicle to enlighten the different driving styles that lead to a faster or a slower lap.
-The track selected for the comparisons is Siena, described previously and shown in Figure~\ref{fig:siena_track}.
-Siena track presents low-speed curves in which dynamic camber control (DCC) architecture or four independent steering wheels (AS) impact substantially.
-The track is run clockwise from the start point, represented by the bold red line in Figure~\ref{fig:siena_track}.
-The thin red lines are spaced by a $\De\al=0.1$, and the black lines, also spaced by $\De\al=0.1$, are shifted from the red ones by $\al_s=0.05$.
-Additionally, the plot shows the enumeration of the turns, from 1 to 7, which will be useful for further observations.
-
-First of all, a traditional vehicle (RWD FS) is compared with an AIWD FS vehicle, capable of applying four independent torques to the wheels.
-As already mentioned, this feature provides the highest improvement in performance.
-In the analyzed case, the vehicle is able to apply not only four driving torques, each one completely independent of the others, but also four independent braking torques.
-This is possible if we assume the vehicle has four separate oil lines or a different braking system.
-For this type of vehicle, usually driven by a human, this design is never applied due to the driver's inability to manage four control inputs, such as pedal forces.
-The decision to maintain the four independent braking torques is made here considering that the vehicles under study are not (yet) existing models and can be thought of as new architectures envisioned for future autonomous vehicle competitions.
-In this case, having more inputs would not be a problem unless some limits in the real-time computing power of the ECU are encountered --- a situation not discussed here.
-In this work, in fact, we focused on identifying and analyzing these solutions, which serve as a benchmark for maximizing vehicle performance, while intentionally postponing their implementation in a real vehicle to future designs.
-%(Questa parte sotto e' poco chiara.
-Quello che facciamo noi fornisce un riferimento su come si dovrebbe guidare.
-E' chiaro che poi ti serve un controllore in feedback per fare tracking oppure per fare una pianificazione dinamica con re-planning tipo MPC che gestisce le discrepanze.
-Tuttavia questi sono aspetti che nella analisi corrente non sono considerati perche' andrebbero valutati in seconda battuta.
-Pero' non terrei il tono un po' negativo che traspare da questa parte di testo.
-Direi semplicemente che questi aspetti sarebbero da analizzare in 'future work' e per 'mettere a terra' una soluzione che noi analizziamo a livello funzionale e di potenzialita'.)
-%In general, in worth bearing in mind that these solutions cannot be implemented ``as is'' for feed-forward control for multiple reasons.
-The most important is that the errors arising from the modeling process and discretization lead to a solution that is almost correct, and thus applying the resulting inputs would generate a different evolution of the system, resulting in a significant final cumulative error.
-Therefore, having a controller would be mandatory, and the high number of control channels would call for appropriate analyses.
-%The results of this case will justify the decision to separately compare RWD and AWD vehicles.
-
-\begin{figure}[]
-	\centering
-	\includegraphics[scale = .85]{Images/DrivingStyles/siena_track.pdf}
-	\caption{Siena Circuit.
-The bold red line indicates the starting point, while the other red lines are equally spaced by a $\De\al=0.1$.
-The black lines are also spaced like the red ones but are shifted by $\al_s=0.05$.
-The track is run in a clockwise direction.
-The curves, consisting of 2 to the left and 5 to the right, are sequentially numbered based on the order of encounter.}
-	\label{fig:siena_track}
-\end{figure}
-
-\subsection{All Independent Wheel Drive, Forward Steering Vehicle - AIWD FS}
-\label{sec:awdfs}
-To perform a valid comparison between vehicles with different powertrain, they are assumed to have the same maximum (and maximum negative, i.e.
-braking) power.
-While the RWD vehicle, modeled with an open differential, will have identical torques on the driving wheels to generate power, the AIWD vehicle will  optimally distribute the power among four wheels, maintaining the same total value of the RWD vehicle.
-
-For most turns, the two vehicles follow different trajectories, with the AIWD staying significantly closer to the inside boundary of the track during turns, as it can be observed in Figure~\ref{fig:rwdfs_awdfs_uayey}, third panel.
-See also the accompanying video~\cite{video:AWDFS:2024}.
-The third plot, in fact, depicts the evolution of the variable $e_y (m)$, which represents the lateral distance, described in $\sref{S}$ components, of the CoM with respect to the center-line, at the corresponding $\al$.
-In a left turn, positive values of $e_y$ correspond to points that are closer to the inner boundary.
-Vice versa, in a right turn, negative values correspond to points that are closer to the inner boundary.
-Each of the three plots shown in Figure~\ref{fig:rwdfs_awdfs_uayey} includes green dash-dot lines marking the turns, while the forward speed plot (first panel) additionally displays the turn numbers, with labels placed over double-ended arrows.
-The enumeration follows the one described in Figure~\ref{fig:siena_track}, and each label indicates a letter that identifies the direction of the curve (L for left turns, R for right turns).
-Again from Figure~\ref{fig:rwdfs_awdfs_uayey}, third plot, we can observe that the AIWD vehicle begins to move toward the inner boundary of the track before the RWD vehicle.
-This behavior is evident in turn 2, which starts at $\al=0.2$, where the AIWD vehicle (blue line) reaches the inner boundary before the RWD (red line), with a difference of $\de\al=0.01$.
-On this track, this difference in curvilinear abscissa corresponds to 12.9\,m.
-Consequently, as the vehicles approach turn 2, the AIWD vehicle reaches the inner boundary approximately 13\,m ahead of the RWD vehicle, indicating a significantly different driving style.
-%Di cosa stai parlando? dove devo guardare? a quale $\al$? non si segue...
-%(quale e' la turn 2? cosa deve guardare il lettore? cosa sono le 1R, 2R, 3L sopra nel primo grafico? non sono mai state introdotte o spiegate)
-\begin{figure}[h]
-	\centering
-	\includegraphics[scale = .8]{Images/DrivingStyles/rwdfsaiwdfs_p.pdf}
-	\caption{
-		Forward speed $u$, lateral acceleration in body components $a_y$ and $e_y$ evolution for RWD FS and AIWD FS vehicle, in red and blue respectively, over curvilinear abscissa $\al$.
-The quantity $e_y$ represents the lateral distance, in $\sref{S}$ reference, of the CoM with respect to the center-line at the corresponding $\al$.
-		%Positive values in a left turn correspond to points that are closer to the inner limit, and vice versa in a right turn, positive values correspond to points that are closer to the outer limit.
-		The turns are indicated in the forward speed plot (first panel) with the corresponding number (see Figure~\ref{fig:siena_track}) and a letter that identifies the direction (L for left turns and R for right turns).
-The more positive the values, the closer the trajectory to the left edge of the track, while the more negative the values, the closer the trajectory to the right edge.}
-	\label{fig:rwdfs_awdfs_uayey}
-\end{figure}
-
-Still referring to Figure~\ref{fig:rwdfs_awdfs_uayey}, in particular to the first plot, which shows the forward speed $u$ over the curvilinear abscissa $\al$, we notice that, during cornering, the AIWD vehicle (blue line) approaches the point of minimum speed earlier than the RWD vehicle (red line).
-The combination of this trend, and the previously analyzed tendency to remain closer to the inner boundary while cornering, allows the AIWD vehicle to exit with higher forward speed than the RWD vehicle.
-This affects positively the lap time.
-
-In the second plot of Figure~\ref{fig:rwdfs_awdfs_uayey}, which reports the lateral acceleration $a_y$ over the curvilinear abscissa $\al$, we can notice that during cornering the AIWD vehicle (blue line) maintains a higher mean absolute value of the lateral acceleration.
-
-In particular, traveling the chicane composed by turn 3 and 4, starting at $\al = 0.27$, the lateral acceleration $a_y$ of the AIWD FS vehicle (blue line) is greater (in magnitude) than the lateral acceleration of the RWD FS vehicle (red line), except for the small range, in which $\al\in[0.27, 0.28]$.
-
-Still referring to the second plot of Figure~\ref{fig:rwdfs_awdfs_uayey}, and  focusing on turn 4, starting approximately at $\al = 0.33$, the lateral acceleration $a_y$ of the AIWD vehicle has smaller variation than the lateral acceleration of the RWD vehicle, maintaining almost a constant value during cornering, denoting a smoother behavior.
-%This behavior can be explained by considering that having four driving torques allows the vehicle to achieve a higher absolute yaw moment, and hence a higher yaw rate $r$, which results in higher lateral acceleration $a_y$.
-In fact if we consider, for simplicity, a 2D vehicle model for handling, as the one described in~\cite[Chapter~3.2]{Guiggiani:book:2023}, with the hypothesis listed in the introduction of the chapter, is possible to express the lateral acceleration as follows:
-%\begin{equation}\label{eq:ayguiggiani}
-%a_y = \dot{v} + ur,
-%\end{equation}
-%where $u$, $v$ and $r$ are respectively, longitudinal and lateral velocity, and yaw rate.
-
-%Apart from the different trajectory, the AWD vehicle pushes the wheels close to the grip limit not only in the lateral direction but also in the longitudinal direction, thereby generating greater tangential forces.
-
-%Qui non si capisce dove vuoi arrivare.
-Il ragionamento (o la spiegazione associata) non e' chiaro.
-Hai richiamato la formula della $a_y$ per cosa?
-Cosa aumenta diminuisce nei due termini?
-Non si capisce, prova a rileggerti cercando di desumere il messaggio solo da cio' che e' scritto.
-%Considering this evidence, we can affirm that the two vehicles have significantly different driving styles, and thus the effects of the added features could be misinterpreted.
-
-\subsection{Rear Wheel Drive, All Steering Vehicle - RWD AS}
-We can observe in Figure \ref{fig:RWD_AS_toes}, first plot, the road wheel angles $\de_{i,j}$ of the four wheels with respect to the curvilinear abscissa $\al$ for the RWD AS vehicle.
-It is worth recalling that, during cornering, this vehicle can exhibit different road wheel angle values for the wheels of the same axle.
-
-In the first plot, the color code is such that wheels of the same axle share the color (red for the front axle, blue for the rear axle), while wheels of the same side share the same line style (solid line for the left side, dashed line for the right side).
-In the second plot, the forward speed $u$ with respect to the curvilinear abscissa $\al$ is depicted.
-In both plots, the green vertical lines mark the start and end point of the turns.
-In the lower plot, the turns are labeled with a progressive number, according to Figure~\ref{fig:siena_track} and a letter indicating its direction (L for left-hand turns and R for right-hand turns).
-
-Although a slight penalization is applied to the cost function to avoid discordant wheel angles on the same axle during turns, the independent behavior of the wheels allows for better control of ground forces and their direction, enhancing turning performance.
-This is especially noticeable in turn 5 (right turn), where the left wheels steer more than the right, and in turn 6 (left turn), where the right wheels steer more than the left.
-See also the accompanying video~\cite{video:RWDAS:2024}.
-Generally, the outer wheel, which bears a greater vertical load due to lateral transfer, has a larger road wheel angle thus generating more lateral force, while the inner wheel steers less since it contributes less to the total lateral force.
-%\begin{figure}[h]
-%	\centering
-%	\includegraphics[scale = .65]{Images/DrivingStyles/rwdfs_rwdas_toeForces.pdf}
-%	\caption{Toe angle $\de$, $F_x$ and $F_y$ forces for the front left wheel of a RWD FS and a RWD AS vehicle, in red and blue respectively.
-Dashed lines in the forces plots indicate the maximum and minimum force reachable.}
-%	\label{fig:rwdfs_rwdas_toeForces}
-%\end{figure}
-\begin{figure}[h]
-	\centering
-	\includegraphics[scale = .8]{Images/DrivingStyles/RWD_AS_toes.pdf}
-	\caption{Road wheel angle $\de$ of all wheels (above), and forward speed $u$ (below), evolution over curvilinear abscissa $\al$.
-As reported in the legend, wheels of the same axle share the same color, while wheels of the same side (left/right) share the same line style.
-The green dash-dot lines represent the starting and ending point of turns.
-The turns are indicated in the forward speed plot with the corresponding number (see Figure~\ref{fig:siena_track}) and a letter that identifies the direction (L for left turns and R for right turns).}
-	\label{fig:RWD_AS_toes}
-\end{figure}
-
-Another aspect worth noting is the behavior of the rear wheels in relation to the front ones.
-In fact, the rear wheels steer in phase with the front wheels for most of the corners and for most of the corner entrances, except for some cases where, in the slower part of the turn, the rear wheels steer in counterphase with the front wheels.
-In particular, referring again to Figure~\ref{fig:RWD_AS_toes}, it is possible to observe that in turn 2-5-7 the front and rear road wheel angles are concordant, while in the remaining turns (1-3-4-6), the road wheel angles are discordant for most part of the turn.
-This behavior is generally consistent with Guiggiani's analysis of MAP's (Map of Achievable Performance) for a Single-Track model of vehicle.
-In fact, in~\cite[Section~6.8.1]{Guiggiani:book:2023} the author affirms that a desired behavior is to have a vehicle where the $\be$ angle spans a small range.
-This can be achieved by having in-phase rear steering at high forward speeds and counter-phase rear steering at low forward speeds.
-
-Although a general rule like this may not hold true for every corner on the track, it nonetheless represents the desired trend to optimize performance, which is here confirmed by quantitative lapsim outputs.
-It also may be worthwhile to point out that a human driver can drive differently with respect to the optimizer, due to the fact that it lacks the perfect view of the track beforehand which, on the other hand, is profitably exploited by the optimizer.
-%This in also confirmed if we observe that the amplitude of the beta angle, slightly reduced in many turns.
-%\begin{figure}[h]
-%	\centering
-%	\includegraphics[scale = .65]{Images/DrivingStyles/rwdas_toe.pdf}
-%	\caption{Toe angle $\de$ for the front and rear left wheel, and forward speed $u$ for a RWD AS vehicle.
-(For the toe angle, only the left side is plotted for clarity and readability, considering that the wheels on the same axle steer in-phase with each other)}
-%	\label{fig:rwdas_toe}
-%\end{figure}
-
-%The last aspect we discuss here is the approach to the turns.
-The AS vehicle brakes less than the FS vehicle and approaches the corners with a higher forward speed.
-This behavior can be explained by considering that, with better control of the ground forces, the vehicle reaches the grip limit at higher velocities and lateral accelerations.
-Not only does the higher limit contribute to this, but having four steered wheels also generates a higher resistance force.
-This becomes clear when we decompose the four lateral forces of the wheels in the vehicle reference system.
-
-\subsection{Rear Wheel Drive, Forward Steering, Dynamic Camber Control - RWD FS DCC}
-The ability to adjust the camber of each wheel during cornering allows the vehicle to maximize the ground forces even more.
-In fact, in a traditional vehicle, the position and orientation of the wheels during a turn are determined by the passive motion of the suspension, in addition to the steering system for the front wheels.
-This means that, even if the suspension is well designed, the position of the wheels will not be always the best for performance and could be improved in each turn.
-
-The first plot of Figure~\ref{fig:rwdfs_rwdfsdcc_gammaFyFR} shows the camber angles $\ga$ of the front right wheel (FR) over the curvilinear abscissa $\al$ for a RWD FS vehicle (red line) and for a RWD FS DCC vehicle (blue line), respectively.
-The second plot of Figure~\ref{fig:rwdfs_rwdfsdcc_gammaFyFR} shows the lateral force $F_y$ of the front right wheel (FR) for the RWD FS vehicle (red solid line) and for the RWD FS DCC vehicle (blue solid line).
-Furthermore the maximum and minimum achievable lateral forces are shown in dashed lines and following the colors of the solid lines (RWD FS in red, RWD FS DCC in blue).
-Both plots show green dash-dot lines to delimit turns, while the first plot, in addition, reports labels with turn numbers (that follow the enumeration of Figure~\ref{fig:siena_track}), and an indication of the turn direction (L for left turns, R for right turns).
-
-It is possible to note that, during cornering, e.g.
-turn 3, the camber tends to settle around a value of $-7.5$~degrees which contributes to increasing the lateral force by 128 N.
-This corresponds to a 7.6\% increase w.r.t.
-the baseline, which can be considered a significant achievement (see the zoomed area in the second plot at turn 3, where $\al\in [0.26, 0.33]$).
-See also the accompanying video~\cite{video:RWDFSDCC:2024}.
-
-Similarly, in turn 6, starting at $\al = 0.69$, it can be observed that the maximum lateral force $F_y$ achievable is higher for the RWD FS DCC vehicle, implying a better wheel positioning relative to the road.
-We remind that, according to the convention of positive lateral forces $F_y$ pointing to the left of the vehicle, ``the higher (more positive) the better'' is valid if we analyze a left turn, such as turn 3 or 6.
-Otherwise, in a right turn, an improvement on force limits would correspond to a lower minimum for achievable lateral force.
-For this reason, considering here the front right (FR) wheel, we focus only on left turns.
-%In the Figure \ref{fig:rwdfs_rwdfsdcc_gammaFyFR}, showing the camber angle $\ga$ evolution (above) and the lateral force $F_y$ (below) of the front right wheel over the curvilinear abscissa $\al$, the limits, indicated by dashed lines and coded with the same color of the actual forces, are higher than those of the traditional vehicle, implying a better wheel positioning relative to the road.
-This is more evident in the left turns, in which this wheel has a higher vertical load.
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[scale = .8]{Images/DrivingStyles/rwdfs_rwdfsdcc_gammaFyFR.pdf}
-	\caption{Camber angle $\ga$ (first plot), and lateral force $F_y$ (second plot) over curvilinear abscissa $\al$, for the front right wheel (FR) of a RWD FS (red line) and a RWD FS DCC (blue line) vehicle.
-The dashed lines represent the upper and lower limit of the forces for each vehicle, coded by the same color.
-The green dash-dot lines represent the starting and ending point of turns.
-The turns are indicated in the forward speed plot with the number and the letter that identifies the direction (L for left turns and R for right turns).
-The zoomed area reports the section with $\al\in [0.26, 0.33]$ to highlight the improvement of $\simeq 130$\,N in the lateral force achievable by Dynamic Camber Control (DCC).}
-	\label{fig:rwdfs_rwdfsdcc_gammaFyFR}
-\end{figure}
-
-%The vertical loads from the optimization, combined with the tyre model, are used to analyze the behavior of the camber angle during cornering.
-In particular, the goal is to verify that the value assumed by the camber angle guarantees the maximum lateral force.
-In Figure~\ref{fig:dcc_lateralforce_gammaspan}, a zoomed-view of the lateral force $F_y$ of the front right wheel (FR) in first part of the turn 3 ($\al\in[0.268, 0.286]$) over the curvilinear abscissa $\al$ is presented.
-The solid curve in black represents the optimal lateral force derived from the optimization, obtained with a value of the camber angle $\gamma^*(\al)$, varying with the curvilinear abscissa.
-The dashed red curve represents the potential force obtained with the minimum value of the added camber $q_{hk,\text{max}}=9$~degrees (see Section~\ref{sec:cambersystemmodeling}).
-The other colored lines represent the lateral forces obtained with the same vertical load $F_z(\al)$ derived by the optimization, and varying the total camber angle $\ga\in[\ga^*(\al)-\De\ga,\ga^*(\al)+\De\ga]$, with $\De\ga = 4$ degrees.
-As we can observe, by varying $\ga$ such that $\ga = \ga^*(\al) + \de\ga$, with the value of $\de\ga$ coded by colors in the bar on the right side, the lateral force presents a peak that, especially in the second part of the plot, doesn't correspond to $\de$ equal to the limit value $\pm\De\ga$.
-After the first part of the curve, which presents a delay due to the fact that we command the second derivative of $q_{hk}$, the solution, adjusting the additional camber, coasts the maximum value of the force.
-
-%\begin{figure}[h]
-%	\centering
-%	\includegraphics[scale = .65]{Images/DrivingStyles/dcc_lateralforce_gammaspan}
-%	\caption{Reachable lateral forces $F_y$ over curvilinear abscissa $\al$ when spanning the camber angle $\ga\in[\ga^*(\al)-\De\ga,\ga^*(\al)+\De\ga]$, equivalent to a window of $\pm 4$ degrees centered on the optimal value $\ga^*(\al)$, which varies with curvilinear abscissa $\al$.
-The black line indicates the optimal value of the force, while the red dashed line indicates the force obtained with the maximum, or minimum, depending on the sign, value of the additional camber $\De\ga$.
-The colored lines represents level curves with constant $\De\ga$, but, in general, different $\ga$.}
-%	\label{fig:dcc_lateralforce_gammaspan}
-%\end{figure}
-
-For brevity, we only mention that the RWD AS DCC vehicle combines the advantages of the previous two, thereby guaranteeing the highest performance.
-Similarly to the RWD AS vehicle, it steers more with the outer wheels thus exploiting the lateral load transfers; at the same time, as registered for the RWD FS DCC vehicle, it adjusts the camber during cornering to settle around the value of maximum ground force available.
-
-%In Figure~\ref{fig:rwdfs_rwdasdcc}, upper plot, the evolution of the road wheel angles $\de_{i,j}$ with respect to the curvilinear abscissa $\al$ is depicted using the same conventions of Figure~\ref{fig:RWD_AS_toes}.
-%The lower plot shows the lateral force $F_y$ evolution over the curvilinear abscissa $\al$ and the force bounds, with the same layout of Figure~\ref{fig:rwdfs_rwdfsdcc_gammaFyFR}, for a RWD FS and RWD AS DCC vehicle, depicted in red and blue, respectively.
-%
-%With reference to the upper plot, we can observe the behavior noted for the RWD AS vehicle, which steers more with the outer wheels thus exploiting the lateral load transfers.
-%Similarly to the RWD FS DCC, this vehicle adjusts the camber during cornering to settle around the value of maximum ground force available.
-Still referring to turn 3, as done in the zoomed area of Figure~\ref{fig:rwdfs_rwdfsdcc_gammaFyFR}, the maximum increase of the force grows now up to 12.1\%.
-These two effects combined also allow the vehicle to navigate corners with higher driving torque on the rear wheels or, in general, less braking torque on all four wheels, thus resulting in a higher forward speed.
-%\begin{figure}[h]
-%	\centering
-%	\includegraphics[scale = .8]{Images/DrivingStyles/rwdfs_rwdasdcc.pdf}
-%	\caption{Road wheel angles $\de_{i,j}$ of all wheels of a RWD AS DCC vehicle over curvilinear abscissa $\al$ (upper plot), and lateral force $F_y$ over curvilinear abscissa $\al$ (lower plot), for the front right wheel (FR) of a RWD FS (red line) and a RWD AS DCC (blue line) vehicle.
-The dashed lines represent the upper and lower limit of the forces for each vehicle, coded by the same color.}
-%	\label{fig:rwdfs_rwdasdcc}
-%\end{figure}
-
-\section{AIWD 0S DCC: in depth analysis}
-This section aims to showcase that on a track with characteristics like the N\"{u}rburgring section used or Siena, whose results we present here, a vehicle able of applying four independent acceleration torques (AIWD) and dynamically controlling the camber (DCC) has a very significant advantage in terms of lap time and cornering performance.
-
-To stress the power of these features, a vehicle with such characteristics but \emph{without steering abilities} (0S, zero steering) has been considered.
-See also the accompanying video~\cite{video:AWDOSDCC:2024}.
-Cornering is still possible if the vehicles applies different torques to the wheels on the same axle, as we can observe in Figure~\ref{fig:awd0sdcc_deltaT}.
-Here, the evolution of the torques applied to the wheels $T_{i,j}$ over the curvilinear abscissa $\al$ for the front axle (upper plot) and the rear axle (lower plot) is illustrated.
-The left wheels (FL and RL) are plotted with red lines, while the right wheels (FR and RR) are plotted with blue lines.
-Similarly to the previous figures, the green dash-dot lines in both plots delimit the turns, while the labels in the first plot denote the turn number, according to Figure~\ref{fig:siena_track}, and a letter which discriminates the direction (L for left-hand turns, R for right-hand turns).
-
-During turns 2-5-6-7, the torque of the inner wheel settles to zero, as is possible to observe in the first panel of Figure~\ref{fig:awd0sdcc_deltaT}.
-Here, the inner wheel starts a free rolling motion, with no applied torque by the motor.
-
-Otherwise a different behavior emerges in turn 4, starting around $\al=0.33$.
-Here, the two wheels experience torques that are approximately mirror-symmetric with respect to the horizontal line $T=100$\,Nm, with the inner torque (FR, blue line) that assumes negative values closely to $\al=0.33$, without losing contact to the ground.
-\begin{figure}[h]
-	\centering
-	\includegraphics[scale = .8]{Images/DrivingStyles/awd0sdcc_deltaT.pdf}
-	\caption{Wheel torques $T_{i,j}$ evolution over curvilinear abscissa $\al$, for the front wheels (above) and for the rear wheels (below) of a AIWD 0S DCC vehicle (0S indicates incapability of steering).
-The red lines refer to the left side while the blue lines refer to the right side.
-As for the previous plots, the green dash-dot lines indicates the turns, labeled in the first plot by the number and a letter L, for left turns, or R, for right turns.}
-	\label{fig:awd0sdcc_deltaT}
-\end{figure}
-
-While the AIWD FS DCC vehicle completes a lap in 51.40 seconds, the AIWD 0S DCC vehicle does so in 51.64 seconds -- only 0.24 seconds slower but 3.66 seconds faster than a traditional RWD FS vehicle.
-This is achieved by properly adjusting camber and applying four independent torques, thus compensating for the lack of steering.
-The torque vectoring logic, used in both AIWD FS DCC and AIWD 0S DCC, plays a key role in optimizing turning maneuvers.
-By simply adjusting the intervention to avoid steering, the lap is completed with a 0.47\% increase in lap time compared to the AIWD FS DCC.
-It is worth mentioning in passing that the model implements power limits as total values, neglecting potential motor saturation, and may lead to an overestimation of performance.
-Therefore, this analysis should not be taken as a direct guideline for designing a AIWD 0S DCC vehicle but an informative case study highlighting factors relevant to cornering and steering.
-
-%While an AWD FS DCC vehicle completes a lap in 51.40\,s, this vehicle (here named AWD 0S DCC) completes a lap in 51.64\,s, thus with only 0.24\,s more, but with 3.66\,s less than a traditional vehicle (RWD FS).
-This is possible because having the ability of adjust the camber and apply 4 independent torques compensate for the lack of steering.
-In particular the torque vectoring logic, used also for AWD FS DCC, provide the most for the turning maneuver, and hence, simply adjusting this intervention to avoid steering the vehicle completes the lap without significant problems.
-It's important to underline that, in the model, the power limits are implemented as total values, hence neglecting potential saturation of power for individual motors.
-This approach has been employed to compare different traction configurations (RWD/AWD); however, it may lead to an overestimation of achievable conditions.
-Nonetheless, this analysis should not be used as a guideline for designing a real AWD 0S DCC vehicle but rather serves as an informative case study that highlights various factors during cornering, including those related to steering.

### 2025-05-09 intro.tex
[MGa] small mod

@@
 Second, an LQR controller is designed to stabilize that nominal path.
-Finally, we re-solve the planning problem - together with the feedback gains - so that the closed-loop covariance propagation satisfies tightened constraints via a Lyapunov-based formulation.
-Together, these two approaches strike a balance between computational simplicity and realistic, feedback-informed robustness, laying the groundwork for safe, high-performance lap-time planning under uncertainty.
+Finally, we re-solve the planning problem - together with the feedback gains - so that, along this new nominal trajectory, the closed-loop covariance propagation satisfies tightened constraints via a Lyapunov-based formulation.
+These two approaches offer a balance between computational simplicity and realistic, feedback-informed robustness, laying the groundwork for safe, high-performance lap-time planning under uncertainty in the motorsport context.
 

### 2025-05-12 closed_loop_planning.tex
[MGa] mutually exclusive throttle and braking command
how to handle

@@
 
-Given the nominal trajectory $(\hbmu_k, \hbu_k)$ over $k=0,\ldots, N$, planned at step (i), and introduced the deviation variables $\bbmu_k=\bmu_k - \hbmu_k$ and $\bbu = \bu_k - \hbu_k$, we design a linear state-feedback law $\bbu_k= -\hbK_k \bbmu_k$, or equivalently $\bu_k = \hbu_k - \hbK_k (\bmu_k - \hbmu_k)$, to mitigate disturbances while tracking the nominal trajectory.
+Given the nominal trajectory $(\hbmu_k, \hbu_k)$ over $k=0,\ldots, N$, planned at step (i), and introduced the deviation variables $\bbmu_k=\bmu_k - \hbmu_k$ and $\bbu_k = \bu_k - \hbu_k$, we design a linear state-feedback law $\bbu_k= -\hbK_k \bbmu_k$, or equivalently $\bu_k = \hbu_k - \hbK_k (\bmu_k - \hbmu_k)$, to mitigate disturbances while tracking the nominal trajectory.
 
@@
 
+\subsection{Handling mutually exclusive throttle and braking commands}
+In our formulation, the state vector is defined as $\bx=\big[u\;v\;r\;x_{G}\;y_{G}\;\psi \big]^T\in\bbR^6$, while the control input vector is $\bu=\big[F_x^a\;F_x^b\;\de \big]^T\in\bbR^3$.
+The longitudinal acceleration and braking forces, $F_x^a \geq 0$ and $F_x^b \leq 0$, respectively, are not independent but are subject to the \emph{complementarity constraint} $F_x^a F_x^b = 0$, which enforces mutual exclusivity between throttle and braking at any instant.
+
+For convenience, we denote $u_1 = F_x^a \geq 0$ and $u_2 = F_x^b \leq 0$, so that the complementarity condition becomes $u_1 u_2 = 0$.
+Beyond the modeling constraint itself, this condition also imposes structural implications on admissible control variations, particularly relevant in the closed-loop formulation.
+Specifically, in a nominal \emph{acceleration} phase ($\hat{u}_1 > 0, \hat{u}_2 = 0$), both \(\delta u_1 \lesseqgtr 0\) are permitted, but only variations $\delta u_2 < 0$ are physically meaningful -- since a positive brake increment ($\delta u_2 > 0$) would unphysically contribute to propulsion.
+Conversely, in a \emph{braking} phase ($\hat{u}_2 < 0$, $\hat{u}_1 = 0$), $\delta u_2 \lesseqgtr 0$ are admissible, but only $\delta u_1 > 0$ should be allowed, as negative throttle increments ($\delta u_1 < 0$) would again introduce non-physical behavior.
+
+To handle this, we introduce a \emph{minimal control vector} \(\tilde{\bu} = [F_x\;\; \delta]^T \in \mathbb{R}^2\), where \(F_x\) represents the net longitudinal force.
+The full control vector $\bu$ is then recovered via a state-dependent linear transformation $
+\bu = \bH(\tilde{\bu}) \tilde{\bu}$ with
+\begin{equation} \label{eq:minimalcontrol}
+\bH(\tilde{\bu}) =
+\begin{bmatrix}
+\Ipos{F_x} & 0 \\
+\Ineg{F_x} & 0 \\
+0 & 1
+\end{bmatrix}, \quad
+\Ipos{x} :=
+\begin{cases}
+1, & \text{if } x \geq 0 \\
+0, & \text{if } x < 0
+\end{cases}, \quad
+\Ineg{x} :=
+\begin{cases}
+0, & \text{if } x \geq 0 \\
+1, & \text{if } x < 0
+\end{cases}.
+\end{equation}
+
+In this formulation, \(\Ipos{x}\) and \(\Ineg{x}\) denote the \emph{positive-} and \emph{negative-part indicator functions}, respectively.
+Importantly, control variations also respect this mapping: $\delta\bu = \bH(\tilde{\bu}) \delta\tilde{\bu}$.
+Consequently, the control input matrix can be expressed as \(\bB \bH = \tilde{\bB}\), and the state-feedback gain as \(\bH \tilde{\bK} = \bK\), such that the resulting feedback law satisfies \(\bB \bK \delta\bx = \tilde{\bB} \tilde{\bK} \delta\bx\) and $\de \tbu = \tbK \de \bx$.
+
+To correctly enforce the mutual exclusivity within the LQR synthesis via Riccati recursion, the state-feedback gain must be computed with respect to the minimal control input.
+Specifically, the control input matrix $\tbB$ is employed and the gain $\tilde{\bK} \in \mathbb{R}^{2 \times 6}$ is sought for which operates on the independent variation \(\delta\tilde{\bu}\).
+This is structured as
+\[
+\tilde{\bK} = \begin{bmatrix} \tilde{\bK}_F^T & \tilde{\bK}_\delta^T \end{bmatrix}^T, \quad \tilde{\bK}_F, \tilde{\bK}_\delta \in \mathbb{R}^{1 \times 6}.
+\]
+The extended gain $\bK$ is automatically recovered via $\bK = \bH \tbK$ and, depending on the operating mode, it will result in the following expressions:
+\begin{itemize}
+  \item during \emph{acceleration}: \(\bK = \begin{bmatrix} \tilde{\bK}_F^T & \mathbf{0}^T & \tilde{\bK}_\delta^T \end{bmatrix}^T\),
+  \item during \emph{braking}: \(\bK = \begin{bmatrix} \mathbf{0}^T & \tilde{\bK}_F^T & \tilde{\bK}_\delta^T \end{bmatrix}^T\).
+\end{itemize}
+This construction ensures consistency with the complementarity constraint, while maintaining a tractable structure for controller synthesis.
+
+A similar consideration applies when solving problem~\eqref{eq:ROCP} in Sec.~\ref{sec:CLrobustified}, where the complementarity structure must again be taken into account.
+In this case, it is sufficient to treat the elements of $\tbK_k$ as the minimal decision variables and reconstruct the full gain matrix $\bK_k$ via the mapping $\bK_k = \bH_k^{\text{S}} \tbK_k$, where
+\begin{equation} \label{eq:minimalcontrolopt}
+\bH^{\text{S}}(\tilde{\bu}) =
+\begin{bmatrix}
+\SIpos{u_1+u_2} & 0 \\
+\SIneg{u_1+u_2} & 0 \\
+0 & 1
+\end{bmatrix}. 
+\end{equation}
+To ensure differentiability and numerical smoothness in the optimization problem~\eqref{eq:ROCP}, we employ smoothed approximations of the indicator functions $\Ipos{x}$ and $\Ineg{x}$, denoted as $\SIpos{x}$ and $\SIneg{x}$ in~\eqref{eq:minimalcontrolopt}, respectively.
+These are defined as:
+\begin{equation} \label{eq:smoothedindicator}
+\SIpos{x} :=
+\frac{1}{2} \left(1 + \tanh(\chi x)\right), \quad \SIneg{x} := \frac{1}{2}\left(1 - \tanh(\chi x) \right),
+\end{equation}
+where the parameter $\chi>0$ controls the sharpness of the transition.
+Smaller values of $\chi$ yield smoother, more relaxed approximations, while larger values make the transition steeper and the function increasingly resemble a hard switching indicator.
+
+  %However, longitudinal acceleration and braking forces $F_x^a \geq 0$ and $F_x^b \leq 0$ are not independent, but are subjected to the complementarity constraint $F_x^a F_x^b=0$, meaning that are mutually exclusive.
+Introduced for simplicity $u_1=F_x^a\geq 0$ and $u_2 = F_x^b\leq 0$, at any instant in time it must hold that $u_1 u_2 = 0$.
+What is important is also that such constraints have also an implication on the values that control variations associated to closed-loop types of correction can assume.
+More in details, in a condition of acceleration (reference acceleration $\hat{u}_1>0$ and $\hat{u}_2 = 0$) $\de u_1\lesseqgtr 0$ but only $\de u_2 <0$ is allowed, otherwise a brake variation $\de u_2 > 0$ would unphysically help the vehicle to accelerate.
+Similarly, in a braking condition (reference braking $\hat{u}_2 < 0$ and $\hat{u}_1 = 0$) $\de u_2 \lesseqgtr 0$ but only $\de u_1 > 0$, otherwise a variation of acceleration $\de u_1 < 0$ would unphysically help the vehicle to brake.
+This brings about the necessity to properly \emph{under-actuate} the system in order account for this behavior.
+This can be accomplished by introducing the minimal control $\tilde{\bu}=[F_x, \de]^T$ such that full control with mutually exclusive acceleration/braking can be recovered as $\bu = \bH(\tbu) \tbu$ where
+%\begin{align}\label{eq:minimalcontrol}
+%\bH(\tbu) =
+%\begin{bmatrix}
+%\Ipos{F_x} & 0\\
+%\Ineg{F_x} & 0\\
+%       0 & 1
+%\end{bmatrix}, \, \text{with} \quad
+%\Ipos{x} :=
+%\begin{cases}
+%1, & \text{if } x \geq 0 \\
+%0, & \text{if } x < 0
+%\end{cases}, \quad \Ineg{x} :=
+%\begin{cases}
+%0, & \text{if } x \geq 0 \\
+%1, & \text{if } x < 0
+%\end{cases}.
+%\end{align}
+%In~\eqref{eq:minimalcontrol} we introduced the positive- and negative-part indicator functions $\Ipos{x}$ and $\Ineg{x}$, respectively.
+It is worth observing that the extended control variation $\de \bu = \bH(\tbu) \de \tbu$.
+% These following positions can be made $\bB \bH = \tbB$, $\bH \tbK = \bK$, such that $\bB \bK \de \bx = \tbB \tbK \de \bx$.
+% It is worth observing that in order to account for the mutually exclusive condition in the synthesis of the LQR controller via the Riccati recursion, the state feedback gain matrix $\tbK\in\bbR^{2\times 6}$ must be employed.
+This allows us to follow the classical recursion by operating on the fully independent $\de \tbu$ and it will result in the calculation of the minimal state-feedback gain matrix $\tbK=\big[ \tbK_{F}^T\, \tbK_{\de}^T\big]^T$ with $\tbK_{F}\in\bbR^{1\times 6}$ and $\tbK_{\de}\in\bbR^{1\times 6}$.
+Then, in the extended controls it will be possible to recover $\bK$ such that during acceleration $\bK=\big[ \tbK_{F}^T\, \bzero^T\, \tbK_{\de}^T\big]^T$  and, during braking, $\bK=\big[ \bzero^T\, \tbK_{F}^T\, \tbK_{\de}^T\big]^T$.
+
 %To mark the difference with~\eqref{eq:OCPdP}, besides using P,CL to denote closed-loop dynamics, a new symbol $\tbP$ is introduced to the denote the covariance matrix in this case.

### 2025-05-13 closed_loop_planning.tex
[MGa] to handle this...

@@
 
-To handle this, we introduce a \emph{minimal control vector} \(\tilde{\bu} = [F_x\;\; \delta]^T \in \mathbb{R}^2\), where \(F_x\) represents the net longitudinal force.
+To ensure physical consistency while preserving a tractable control structure, we introduce a \emph{minimal control vector} \(\tilde{\bu} = [F_x\;\; \delta]^T \in \mathbb{R}^2\), where \(F_x\) represents the net longitudinal force.
 The full control vector $\bu$ is then recovered via a state-dependent linear transformation $
@@
 \end{itemize}
-This construction ensures consistency with the complementarity constraint, while maintaining a tractable structure for controller synthesis.
+This formulation preserves consistency with the complementarity constraint while retaining a structure amenable to controller synthesis.
 

### 2025-05-19 performance_comparison.tex
[MM] First draft telemetries comparison

@@
 \label{sec:performance_comparison}
+
+\begin{figure}
+	\centering
+	\includegraphics{Fig/track.eps}
+	\caption{Sector of the Catalunya circuit considered in the analysis, corresponding to the curvilinear abscissa interval $\left[0.70, 0.77\right]$.
+Checkpoints are indicated by labels and are uniformly spaced by $\Delta\al=0.01$.
+This sector includes two distinct corners: a low-speed turn from $\al = 0.72$ to $\al = 0.73$, and a high-speed turn from $\al = 0.75$ to $\al = 0.76$, allowing for the evaluation of vehicle behavior across different dynamic scenarios.}
+	\label{fig:track}
+\end{figure}
+
+\subsection{Robustified constraints comparison}
+It is of interest to analyze how a robustified constraint---or a combination of the two described in Sections~\ref{sec:adherence_constraint} and~\ref{sec:stayontrack_constraint}---influences the driving style.
+This analysis compares the nominal trajectory, obtained without robustified constraints, with those resulting from a robustified stay-on-track constraint (denoted as SoT), a robustified adherence limit constraint (denoted as Adh), and a scenario where both constraints are robustified (denoted as SoT+Adh).
+All the optimization are obtained with the open-loop method described in Section~\ref{sec:open_loop_planning}, setting $H=4$, and $\ga^\textrm{SoT}=\ga^\textrm{Adh}=1.28$, corresponding to a 90\% probability of meeting both constraints.
+
+Figure~\ref{fig:ol_telemetries} presents the variation of four signals with respect to the nominal trajectory, for the three previously mentioned cases.
+The panels display the longitudinal speed $u$ (first panel), wheel steering angle $\delta$ (second panel), total longitudinal force $F_x$ (third panel), and lateral deviation of the center of mass from the track centerline $e$ (fourth panel).
+Except for the $u$ panel, each plot includes dashed black lines that indicate the sign of the nominal signal-whether it is positive, negative, or numerically zero.
+These sign indicators are represented as three distinct levels---high, zero, and low---corresponding to positive, zero, and negative values of the nominal signal, respectively.
+The levels are scaled appropriately in each plot to ensure clarity of visualization.
+
+\begin{figure}
+	\centering
+	\includegraphics{Fig/ol_telemetries.eps}
+	\caption{}
+	\label{fig:ol_telemetries}
+\end{figure}
+
+From the first panel, it can be observed that all configurations maintain a lower mean value of the longitudinal speed compared to the nominal trajectory, resulting in an increased sector time.
+The two configurations incorporating the robustified stay-on-track constraint exhibit, as expected, a significantly altered CoM trajectory.
+This behavior explains the higher longitudinal speed observed around $\alpha = 0.71$, which is followed by a sharp reduction.
+
+The second and fourth panels clearly show that the configurations with the robustified stay-on-track constraint tend to follow a path closer to the centerline.
+In particular, for most of the sector, the variation in lateral displacement and the lateral displacement $e$ itself exhibit opposite signs, indicating a corrective behavior of the steering angle $\de$ that pulls the vehicle toward the centerline.

### 2025-05-21 performance_comparison.tex
[MM] constraints name changed and some results description

@@
 	\centering
-	\includegraphics{Fig/track.eps}
+	\includegraphics{Fig/track.pdf}
 	\caption{Sector of the Catalunya circuit considered in the analysis, corresponding to the curvilinear abscissa interval $\left[0.70, 0.77\right]$.
-Checkpoints are indicated by labels and are uniformly spaced by $\Delta\al=0.01$.
+Checkpoints are indicated by labels and are uniformly spaced by $\De\al=0.01$.
 This sector includes two distinct corners: a low-speed turn from $\al = 0.72$ to $\al = 0.73$, and a high-speed turn from $\al = 0.75$ to $\al = 0.76$, allowing for the evaluation of vehicle behavior across different dynamic scenarios.}
@@
 
+\subsection{Open loop parameters sensitivity}
+\label{sec:ol_param_sensitivity}
+In this section, we explore the open-loop approach to gain deeper insight into the influence of key parameters on the results.
+For this analysis, we consider optimizations in which only the stay-on-track constraint is robustified.
+This choice is motivated by the fact that the effects of the stay-on-track constraint are more visually evident compared to those of the adherence limit constraint, thereby facilitating a more immediate understanding of their impact.
+
+The first parameter analyzed is the factor $\ga^\textrm{OTC}$, which multiplies the standard deviation of the constraint, $\sigma^\textrm{OTC}$, to determine the total back-off term $\be^\textrm{OTC}$.
+This parameter directly influences the probability of satisfying the stay-on-track constraint: higher values of $\ga^\textrm{OTC}$ correspond to more conservative (i.e., robust) behavior.
+Specifically, $\ga^\textrm{OTC} = 0$ yields a satisfaction probability of 50\%, while $\ga^\textrm{OTC} = 3$ corresponds to 99\%.
+For the purposes of this analysis, we examine the effect of varying $\ga^\textrm{OTC}$ within this range.
+
+The second parameter examined is the initial value of the covariance matrix, $\bP_0$.
+In addition to the baseline value $\bar{\bP}_0$, two alternative values are considered: $\bP_0 = 4\bar{\bP}_0$ and $\bP_0 = \frac{1}{4}\bar{\bP}_0$.
+These correspond, respectively, to doubling and halving the initial uncertainty associated with the state vector.
+
+\begin{figure}
+	\centering
+	\includegraphics{Fig/gamma_sensitivity.pdf}
+	\hfill
+	\includegraphics{Fig/Pzero_sensitivity.pdf}
+	\caption{}
+	\label{fig:ol_sensitivities}
+\end{figure}
+
 \subsection{Robustified constraints comparison}
-It is of interest to analyze how a robustified constraint---or a combination of the two described in Sections~\ref{sec:adherence_constraint} and~\ref{sec:stayontrack_constraint}---influences the driving style.
-This analysis compares the nominal trajectory, obtained without robustified constraints, with those resulting from a robustified stay-on-track constraint (denoted as SoT), a robustified adherence limit constraint (denoted as Adh), and a scenario where both constraints are robustified (denoted as SoT+Adh).
-All the optimization are obtained with the open-loop method described in Section~\ref{sec:open_loop_planning}, setting $H=4$, and $\ga^\textrm{SoT}=\ga^\textrm{Adh}=1.28$, corresponding to a 90\% probability of meeting both constraints.
+It is of interest to analyze how a robustified constraint---or a combination of the two described in Sections~\ref{sec:adherence_constraint} and~\ref{sec:ontrack_constraint}---influences the driving style.
+This analysis compares the nominal trajectory, obtained without robustified constraints, with those resulting from a robustified stay-on-track constraint (denoted as OTC), a robustified adherence limit constraint (denoted as FLC), and a scenario where both constraints are robustified (denoted as OTC+FLC).
+All the optimization are obtained with the open-loop method described in Section~\ref{sec:open_loop_planning}, setting $H=4$, and $\ga^\textrm{OTC}=\ga^\textrm{FLC}=1.28$, corresponding to a 90\% probability of meeting both constraints.
 
 Figure~\ref{fig:ol_telemetries} presents the variation of four signals with respect to the nominal trajectory, for the three previously mentioned cases.
-The panels display the longitudinal speed $u$ (first panel), wheel steering angle $\delta$ (second panel), total longitudinal force $F_x$ (third panel), and lateral deviation of the center of mass from the track centerline $e$ (fourth panel).
+The panels display the longitudinal speed $u$ (first panel), wheel steering angle $\de$ (second panel), total longitudinal force $F_x$ (third panel), and lateral deviation of the center of mass from the track centerline $e$ (fourth panel).
 Except for the $u$ panel, each plot includes dashed black lines that indicate the sign of the nominal signal-whether it is positive, negative, or numerically zero.
@@
 	\centering
-	\includegraphics{Fig/ol_telemetries.eps}
+	\includegraphics{Fig/ol_telemetries.pdf}
 	\caption{}
@@
 The two configurations incorporating the robustified stay-on-track constraint exhibit, as expected, a significantly altered CoM trajectory.
-This behavior explains the higher longitudinal speed observed around $\alpha = 0.71$, which is followed by a sharp reduction.
+This behavior explains the higher longitudinal speed observed around $\al = 0.71$, which is followed by a sharp reduction.
 
@@
 In particular, for most of the sector, the variation in lateral displacement and the lateral displacement $e$ itself exhibit opposite signs, indicating a corrective behavior of the steering angle $\de$ that pulls the vehicle toward the centerline.
+
+It can be interesting to notice that, in the third panel, only the configuration with both the constraints robustified needs to reduce the longitudinal force during the high speed turn, in the curvilinear abscissa interval $\al\in\left[0.75, 0.76\right]$, to meet both the constraints.
+The vehicle constrained by the stay-on-track condition is forced to follow a wider trajectory, which entails greater lateral acceleration and, consequently, a higher lateral force demand---resulting in increased tire utilization.
+Due to the limitations imposed by the robust adherence constraint, it is necessary to reduce the available accelerating force.
+
+\begin{figure}
+	\centering
+	\includegraphics{Fig/ol_saturation.pdf}
+	\caption{}
+	\label{fig:ol_saturation}
+\end{figure}
+
+\subsection{Simulations with noise realization}
+This analysis aims to provide empirical validation of the robust trajectories obtained using the methods proposed in this work.
+For conciseness, the comparison is limited to a nominal trajectory and a trajectory computed using the closed-loop method with only the on-track constraint robustified.
+An LQR controller, designed according to the procedure outlined in Section~\ref{sec:LQR}, is implemented along the nominal trajectory.
+In contrast, the robust trajectory is tracked using the controller directly obtained from the optimization process.
+
+\begin{figure}
+	\centering
+	\includegraphics{Fig/olcl_traj_strings.pdf}
+	\caption{}
+	\label{fig:traj_strings}
+\end{figure}

### 2025-05-21 stochastic_vehicle_dynamics.tex
[MGa] paper's contributions and organization
[MM] constraints name changed and some results description

@@
-\section{TMP NAME}
-\label{sec:tmp}
+\section{Stochastic vehicle dynamic framework}
+\label{sec:svdf}
 
@@
 
-\subsection{Robust adherence limit constraint formulation}
+\subsection{Robust friction limit constraint formulation}
 \label{sec:adherence_constraint}
@@
 \begin{equation}
-	h^\textrm{A}_j(\bx,\bu) = \left[ \left( \frac{X_j(\bx,\bu)}{\mu_{x,j}} \right)^2+ \left( \frac{Y_j(\bx,\bu)}{\mu_{y,j}}\right)^2\right]\frac{1}{Z_j^2(\bx,\bu)}-1\leq0\qquad j=1,2.\label{eq:adherence_constraint}
+	h^\textrm{FLC}_j(\bx,\bu) = \left[ \left( \frac{X_j(\bx,\bu)}{\mu_{x,j}} \right)^2+ \left( \frac{Y_j(\bx,\bu)}{\mu_{y,j}}\right)^2\right]\frac{1}{Z_j^2(\bx,\bu)}-1\leq0\qquad j=1,2.\label{eq:adherence_constraint}
 \end{equation}
-where the superscript A denotes adherence, and the subscript $j$ indicates the axle index, with $j=1$ referring to the front axle and $j=2$ to the rear axle.
+where the superscript FLC denotes friction limit constraint, and the subscript $j$ indicates the axle index, with $j=1$ referring to the front axle and $j=2$ to the rear axle.
 Here, $X_j$ and $Y_j$ denote the longitudinal and lateral components of the in-plane ground forces, respectively, as illustrated in Figure~\ref{fig:vehicle_model}, while $Z_j$ represents the total vertical load acting on the axle.
@@
 
-To better clarify the effect of the back-off term on the constraint we can suppose to have a combination of $\bx$, $\bu$, $P$ and $\ga^\textrm{A}_j$ so that $\be^\textrm{A}_j=0.2$.
-This implies that the semi-axes of the ellipse are reduced by a factor $\sqrt{1-\be^\textrm{A}_j} = \sqrt{0.8}$, as shown in the left panel Figure~\ref{fig:robust_constraints}, where the ellipse in dashed line delimits the available region with the back-off applied.
+To better clarify the effect of the back-off term on the constraint we can suppose to have a combination of $\bx$, $\bu$, $P$ and $\ga^\textrm{FLC}_j$ so that $\be^\textrm{FLC}_j=0.2$.
+This implies that the semi-axes of the ellipse are reduced by a factor $\sqrt{1-\be^\textrm{FLC}_j} = \sqrt{0.8}$, as shown in the left panel Figure~\ref{fig:robust_constraints}, where the ellipse in dashed line delimits the available region with the back-off applied.
 This allows the optimizer to determine the most appropriate trade-off between the longitudinal force $X_j$ and the lateral force $Y_j$ --- often by reducing both components to some extent --- based on the specific requirements of the manoeuvre.
@@
 	\caption{Graphical representation of the two constraints analyzed in the section.
-The left panel shows the adherence constraint in the plane $X_j-Y_j$.
+The left panel shows the friction limit constraint in the plane $X_j-Y_j$.
 The ellipse in solid line, whose semi-axes are given by $\mu_{x,j}Z_j$ and $\mu_{y,j}Z_j$, represents the border of the available region for the base constraint, without back-off.
-The ellipse in dashed line delimits the available region when a back-off $\be^\textrm{A}_j=0.2$ is present.
-This ellipse has its semi-axes reduced by a factor $\sqrt{1-\be^\textrm{A}_j}$ w.r.t.
+The ellipse in dashed line delimits the available region when a back-off $\be^\textrm{FLC}_j=0.2$ is present.
+This ellipse has its semi-axes reduced by a factor $\sqrt{1-\be^\textrm{FLC}_j}$ w.r.t.
 the one in solid line.
@@
 
-\subsection{Stay-on-track constraint}
-\label{sec:stayontrack_constraint}
-A stay-on-track constraint is required to ensure that the vehicle remains within the circuit boundaries.
+\subsection{On-track constraint}
+\label{sec:ontrack_constraint}
+An on-track constraint is required to ensure that the vehicle remains within the circuit boundaries.
 The formulation adopted introduces an algebraic variable $e$, which satisfies the following constraint:
@@
 \begin{equation}
-	e_\textrm{min} + \be^\textrm{SOT} \leq e \leq e_\textrm{max} - \be^\textrm{SOT}\label{eq:stayontrack_constraint}
+	e_\textrm{min} + \be^\textrm{OTC} \leq e \leq e_\textrm{max} - \be^\textrm{OTC}\label{eq:ontrack_constraint}
 \end{equation}
-where $\be^\textrm{SOT}$ is computed by considering that the gradient of the constraint w.r.t. the state vector $\na_{\bx}h^\textrm{SOT}$ depends solely on $\bn$.
+where $\be^\textrm{OTC}$ is computed by considering that the gradient of the constraint w.r.t. the state vector $\na_{\bx}h^\textrm{OTC}$ depends solely on $\bn$.
 %Adding a back-off to this constraint produces the same effect as an increase of the front and rear tracks of the vehicle does.

### 2025-05-21 performance_comparison.tex
[MM] definitive constraints acronym

@@
 
-The first parameter analyzed is the factor $\ga^\textrm{OTC}$, which multiplies the standard deviation of the constraint, $\sigma^\textrm{OTC}$, to determine the total back-off term $\be^\textrm{OTC}$.
-This parameter directly influences the probability of satisfying the stay-on-track constraint: higher values of $\ga^\textrm{OTC}$ correspond to more conservative (i.e., robust) behavior.
-Specifically, $\ga^\textrm{OTC} = 0$ yields a satisfaction probability of 50\%, while $\ga^\textrm{OTC} = 3$ corresponds to 99\%.
-For the purposes of this analysis, we examine the effect of varying $\ga^\textrm{OTC}$ within this range.
+The first parameter analyzed is the factor $\ga^\textrm{TLC}$, which multiplies the standard deviation of the constraint, $\sigma^\textrm{TLC}$, to determine the total back-off term $\be^\textrm{TLC}$.
+This parameter directly influences the probability of satisfying the stay-on-track constraint: higher values of $\ga^\textrm{TLC}$ correspond to more conservative (i.e., robust) behavior.
+Specifically, $\ga^\textrm{TLC} = 0$ yields a satisfaction probability of 50\%, while $\ga^\textrm{TLC} = 3$ corresponds to 99\%.
+For the purposes of this analysis, we examine the effect of varying $\ga^\textrm{TLC}$ within this range.
 
@@
 \subsection{Robustified constraints comparison}
-It is of interest to analyze how a robustified constraint---or a combination of the two described in Sections~\ref{sec:adherence_constraint} and~\ref{sec:ontrack_constraint}---influences the driving style.
-This analysis compares the nominal trajectory, obtained without robustified constraints, with those resulting from a robustified stay-on-track constraint (denoted as OTC), a robustified adherence limit constraint (denoted as FLC), and a scenario where both constraints are robustified (denoted as OTC+FLC).
-All the optimization are obtained with the open-loop method described in Section~\ref{sec:open_loop_planning}, setting $H=4$, and $\ga^\textrm{OTC}=\ga^\textrm{FLC}=1.28$, corresponding to a 90\% probability of meeting both constraints.
+It is of interest to analyze how a robustified constraint---or a combination of the two described in Sections~\ref{sec:FLC} and~\ref{sec:TLC}---influences the driving style.
+This analysis compares the nominal trajectory, obtained without robustified constraints, with those resulting from a robustified track limit constraint (denoted as TLC), a robustified friction limit constraint (denoted as FLC), and a scenario where both constraints are robustified (denoted as TLC+FLC).
+All the optimization are obtained with the open-loop method described in Section~\ref{sec:open_loop_planning}, setting $H=4$, and $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$, corresponding to a 90\% probability of meeting both constraints.
 

### 2025-05-21 stochastic_vehicle_dynamics.tex
[MM] definitive constraints acronym

@@
 \subsection{Robust friction limit constraint formulation}
-\label{sec:adherence_constraint}
+\label{sec:FLC}
 The model employs a tire formulation that neglects combined slip effects, i.e., it does not account for the simultaneous utilization of longitudinal and lateral tyre forces.
@@
 
-\subsection{On-track constraint}
-\label{sec:ontrack_constraint}
+\subsection{Robust track limit constraint formulation}
+\label{sec:TLC}
 An on-track constraint is required to ensure that the vehicle remains within the circuit boundaries.
@@
 \begin{equation}
-	e_\textrm{min} + \be^\textrm{OTC} \leq e \leq e_\textrm{max} - \be^\textrm{OTC}\label{eq:ontrack_constraint}
+	e_\textrm{min} + \be^\textrm{TLC} \leq e \leq e_\textrm{max} - \be^\textrm{TLC}\label{eq:TLC}
 \end{equation}
-where $\be^\textrm{OTC}$ is computed by considering that the gradient of the constraint w.r.t. the state vector $\na_{\bx}h^\textrm{OTC}$ depends solely on $\bn$.
+where $\be^\textrm{TLC}$ is computed by considering that the gradient of the constraint w.r.t. the state vector $\na_{\bx}h^\textrm{TLC}$ depends solely on $\bn$.
 %Adding a back-off to this constraint produces the same effect as an increase of the front and rear tracks of the vehicle does.

### 2025-05-21 performance_comparison.tex
[MM] some constraint references were not updated

@@
 In this section, we explore the open-loop approach to gain deeper insight into the influence of key parameters on the results.
-For this analysis, we consider optimizations in which only the stay-on-track constraint is robustified.
-This choice is motivated by the fact that the effects of the stay-on-track constraint are more visually evident compared to those of the adherence limit constraint, thereby facilitating a more immediate understanding of their impact.
+For this analysis, we consider optimizations in which only the track limit constraint is robustified.
+This choice is motivated by the fact that the effects of the track limit constraint are more visually evident compared to those of the friction limit constraint, thereby facilitating a more immediate understanding of their impact.
 
 The first parameter analyzed is the factor $\ga^\textrm{TLC}$, which multiplies the standard deviation of the constraint, $\sigma^\textrm{TLC}$, to determine the total back-off term $\be^\textrm{TLC}$.
-This parameter directly influences the probability of satisfying the stay-on-track constraint: higher values of $\ga^\textrm{TLC}$ correspond to more conservative (i.e., robust) behavior.
+This parameter directly influences the probability of satisfying the track limit constraint: higher values of $\ga^\textrm{TLC}$ correspond to more conservative (i.e., robust) behavior.
 Specifically, $\ga^\textrm{TLC} = 0$ yields a satisfaction probability of 50\%, while $\ga^\textrm{TLC} = 3$ corresponds to 99\%.
@@
 From the first panel, it can be observed that all configurations maintain a lower mean value of the longitudinal speed compared to the nominal trajectory, resulting in an increased sector time.
-The two configurations incorporating the robustified stay-on-track constraint exhibit, as expected, a significantly altered CoM trajectory.
+The two configurations incorporating the robustified track limit constraint exhibit, as expected, a significantly altered CoM trajectory.
 This behavior explains the higher longitudinal speed observed around $\al = 0.71$, which is followed by a sharp reduction.
 
-The second and fourth panels clearly show that the configurations with the robustified stay-on-track constraint tend to follow a path closer to the centerline.
+The second and fourth panels clearly show that the configurations with the robustified track limit constraint tend to follow a path closer to the centerline.
 In particular, for most of the sector, the variation in lateral displacement and the lateral displacement $e$ itself exhibit opposite signs, indicating a corrective behavior of the steering angle $\de$ that pulls the vehicle toward the centerline.
@@
 It can be interesting to notice that, in the third panel, only the configuration with both the constraints robustified needs to reduce the longitudinal force during the high speed turn, in the curvilinear abscissa interval $\al\in\left[0.75, 0.76\right]$, to meet both the constraints.
-The vehicle constrained by the stay-on-track condition is forced to follow a wider trajectory, which entails greater lateral acceleration and, consequently, a higher lateral force demand---resulting in increased tire utilization.
-Due to the limitations imposed by the robust adherence constraint, it is necessary to reduce the available accelerating force.
+The vehicle constrained by the track limit condition is forced to follow a wider trajectory, which entails greater lateral acceleration and, consequently, a higher lateral force demand---resulting in increased tire utilization.
+Due to the limitations imposed by the robust friction limit constraint, it is necessary to reduce the available accelerating force.
 
@@
 This analysis aims to provide empirical validation of the robust trajectories obtained using the methods proposed in this work.
-For conciseness, the comparison is limited to a nominal trajectory and a trajectory computed using the closed-loop method with only the on-track constraint robustified.
+For conciseness, the comparison is limited to a nominal trajectory and a trajectory computed using the closed-loop method with only the track limit constraint robustified.
 An LQR controller, designed according to the procedure outlined in Section~\ref{sec:LQR}, is implemented along the nominal trajectory.

### 2025-05-22 performance_comparison.tex
[MM] some captions

@@
 	\includegraphics{Fig/Pzero_sensitivity.pdf}
-	\caption{}
+	\caption{Effect of $\ga^\textrm{TLC}$ (left panel) and $P_0$ (right panel) on optimal trajectory.
+The parameter $\ga^\textrm{TLC}$ span in the range $\left[0,3\right]$ corresponding to a probability to meet the track limit constraint of 50\% and 99\%, respectively.
+In the left panel, in addition to the baseline value $\bar{\bP}_0$ (black line), two alternative values are considered: $\bP_0 = 4\bar{\bP}_0$ (blue line) and $\bP_0 = \frac{1}{4}\bar{\bP}_0$ (red line).
+These correspond, respectively, to doubling and halving the initial uncertainty associated with the state vector.}
 	\label{fig:ol_sensitivities}
@@
 It is of interest to analyze how a robustified constraint---or a combination of the two described in Sections~\ref{sec:FLC} and~\ref{sec:TLC}---influences the driving style.
-This analysis compares the nominal trajectory, obtained without robustified constraints, with those resulting from a robustified track limit constraint (denoted as TLC), a robustified friction limit constraint (denoted as FLC), and a scenario where both constraints are robustified (denoted as TLC+FLC).
+This analysis compares the not-robust trajectory, obtained without robustified constraints, with those resulting from a robustified track limit constraint (denoted as TLC), a robustified friction limit constraint (denoted as FLC), and a scenario where both constraints are robustified (denoted as TLC+FLC).
 All the optimization are obtained with the open-loop method described in Section~\ref{sec:open_loop_planning}, setting $H=4$, and $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$, corresponding to a 90\% probability of meeting both constraints.
 
-Figure~\ref{fig:ol_telemetries} presents the variation of four signals with respect to the nominal trajectory, for the three previously mentioned cases.
+Figure~\ref{fig:ol_telemetries} presents the variation of four signals with respect to the not-robust trajectory, for the three previously mentioned cases.
 The panels display the longitudinal speed $u$ (first panel), wheel steering angle $\de$ (second panel), total longitudinal force $F_x$ (third panel), and lateral deviation of the center of mass from the track centerline $e$ (fourth panel).
-Except for the $u$ panel, each plot includes dashed black lines that indicate the sign of the nominal signal-whether it is positive, negative, or numerically zero.
-These sign indicators are represented as three distinct levels---high, zero, and low---corresponding to positive, zero, and negative values of the nominal signal, respectively.
+Except for the $u$ panel, each plot includes dashed black lines that indicate the sign of the reference signal-whether it is positive, negative, or numerically zero.
+These sign indicators are represented as three distinct levels---high, zero, and low---corresponding to positive, zero, and negative values of the reference signal, respectively.
 The levels are scaled appropriately in each plot to ensure clarity of visualization.
@@
 	\includegraphics{Fig/ol_telemetries.pdf}
-	\caption{}
+	\caption{Variation of four signals w.r.t.
+the not-robust trajectory for the three configurations TLC (green line), FLC (red line) and TLC+FLC (blue line).
+The panels display the longitudinal speed $u$ (first panel), wheel steering angle $\de$ (second panel), total longitudinal force $F_x$ (third panel), and lateral deviation of the center of mass from the track centerline $e$ (fourth panel), over the curvilinear abscissa $\al$.
+The dashed lines in the last three panels indicate the sign of the reference signal, allowing one to determine whether a positive variation with respect to the baseline corresponds to an increase or a decrease in the absolute value.
+This distinction is particularly important for quantities that depend on the direction of the turn, such as the wheel steering angle $\de$ and the lateral deviation $e$, while in the case of $F_x$, it indicates whether the vehicle is accelerating or braking.}
 	\label{fig:ol_telemetries}
@@
 
-From the first panel, it can be observed that all configurations maintain a lower mean value of the longitudinal speed compared to the nominal trajectory, resulting in an increased sector time.
+From the first panel, it can be observed that all configurations maintain a lower mean value of the longitudinal speed compared to the reference trajectory, resulting in an increased sector time.
 The two configurations incorporating the robustified track limit constraint exhibit, as expected, a significantly altered CoM trajectory.
@@
 	\includegraphics{Fig/ol_saturation.pdf}
-	\caption{}
+	\caption{Tyre usage for front axle (first panel) and rear axle (second panel) over curvilinear abscissa $\al$.
+The plotted lines represent the quantity $\left[ \left( \frac{X_j(\bx,\bu)}{\mu_{x,j}} \right)^2+ \left( \frac{Y_j(\bx,\bu)}{\mu_{y,j}}\right)^2\right]\frac{1}{Z_j^2(\bx,\bu)}$, where $j=1$ denotes the front axle and $j=2$ the rear axle.
+The dashed black line corresponds to the reference (not-robust) configuration, while the green, red, and blue lines represent the robust configurations TLC, FLC, and TLC+FLC, respectively.}
 	\label{fig:ol_saturation}
@@
 For conciseness, the comparison is limited to a nominal trajectory and a trajectory computed using the closed-loop method with only the track limit constraint robustified.
-An LQR controller, designed according to the procedure outlined in Section~\ref{sec:LQR}, is implemented along the nominal trajectory.
+An LQR controller, designed according to the procedure outlined in Section~\ref{sec:LQR}, is implemented along the not-robust trajectory.
 In contrast, the robust trajectory is tracked using the controller directly obtained from the optimization process.
@@
 	\includegraphics{Fig/olcl_traj_strings.pdf}
-	\caption{}
+	\caption{Simulations with random initial conditions and Gaussian noise realizations.
+In the left panel, the reference trajectory is the non-robust one, tracked using an LQR controller designed around it.
+In the right panel, the reference is a robust trajectory obtained using the closed-loop optimization method, and it is tracked using the corresponding optimized controller produced by the same method.}
 	\label{fig:traj_strings}

### 2025-05-23 performance_comparison.tex
[MM] few changes

@@
 
-Figure~\ref{fig:ol_telemetries} presents the variation of four signals with respect to the not-robust trajectory, for the three previously mentioned cases.
+Figure~\ref{fig:ol_telemetries} presents the variation of four signals with respect to the nominal trajectory, for the three previously mentioned cases.
 The panels display the longitudinal speed $u$ (first panel), wheel steering angle $\de$ (second panel), total longitudinal force $F_x$ (third panel), and lateral deviation of the center of mass from the track centerline $e$ (fourth panel).
@@
 	\includegraphics{Fig/ol_telemetries.pdf}
-	\caption{Variation of four signals w.r.t.
-the not-robust trajectory for the three configurations TLC (green line), FLC (red line) and TLC+FLC (blue line).
-The panels display the longitudinal speed $u$ (first panel), wheel steering angle $\de$ (second panel), total longitudinal force $F_x$ (third panel), and lateral deviation of the center of mass from the track centerline $e$ (fourth panel), over the curvilinear abscissa $\al$.
+	\caption{Evolution of the longitudinal speed u, wheel steering angle $\de$, longitudinal force $F_x$, and lateral deviation $e$ are reported for three robustification scenarios: TLC (Track Limit Constraints), FLC (Friction Limit Constraint), and TLC+FLC (both constraints enforced concurrently).
+Each profile is shown relative to the baseline (nominal) trajectory to highlight the effect of each constraint-handling strategy.
 The dashed lines in the last three panels indicate the sign of the reference signal, allowing one to determine whether a positive variation with respect to the baseline corresponds to an increase or a decrease in the absolute value.
@@
 	\includegraphics{Fig/ol_saturation.pdf}
-	\caption{Tyre usage for front axle (first panel) and rear axle (second panel) over curvilinear abscissa $\al$.
+	\caption{Tire usage for front axle (first panel) and rear axle (second panel) over curvilinear abscissa $\al$.
 The plotted lines represent the quantity $\left[ \left( \frac{X_j(\bx,\bu)}{\mu_{x,j}} \right)^2+ \left( \frac{Y_j(\bx,\bu)}{\mu_{y,j}}\right)^2\right]\frac{1}{Z_j^2(\bx,\bu)}$, where $j=1$ denotes the front axle and $j=2$ the rear axle.
-The dashed black line corresponds to the reference (not-robust) configuration, while the green, red, and blue lines represent the robust configurations TLC, FLC, and TLC+FLC, respectively.}
+The dashed black line corresponds to the reference (nominal) configuration, while the green, red, and blue lines represent the robust configurations TLC, FLC, and TLC+FLC, respectively.}
 	\label{fig:ol_saturation}
@@
 For conciseness, the comparison is limited to a nominal trajectory and a trajectory computed using the closed-loop method with only the track limit constraint robustified.
-An LQR controller, designed according to the procedure outlined in Section~\ref{sec:LQR}, is implemented along the not-robust trajectory.
+An LQR controller, designed according to the procedure outlined in Section~\ref{sec:LQR}, is implemented along the nominal trajectory.
 In contrast, the robust trajectory is tracked using the controller directly obtained from the optimization process.

### 2025-05-23 stochastic_vehicle_dynamics.tex
[MM] some constraint references were not updated
[MM] few changes

@@
 \label{sec:FLC}
-The model employs a tire formulation that neglects combined slip effects, i.e., it does not account for the simultaneous utilization of longitudinal and lateral tyre forces.
-As a result, the optimal control problem must include an additional constraint to ensure that the total ground reaction forces remain within the bounds of the tyre's adherence ellipse.
+The model employs a tire formulation that neglects combined slip effects, i.e., it does not account for the simultaneous utilization of longitudinal and lateral tire forces.
+As a result, the optimal control problem must include an additional constraint to ensure that the total ground reaction forces remain within the bounds of the tire's adherence ellipse.
 
@@
 \label{sec:TLC}
-An on-track constraint is required to ensure that the vehicle remains within the circuit boundaries.
+An track limit constraint is required to ensure that the vehicle remains within the circuit boundaries.
 The formulation adopted introduces an algebraic variable $e$, which satisfies the following constraint:

### 2025-05-24 performance_comparison.tex
[MM] some words on sensitivity

@@
 
-The second parameter examined is the initial value of the covariance matrix, $\bP_0$.
+Setting $\ga^\textrm{TLC} = 0$ leads to the nominal solution.
+In this case, the back-off term $\be^\textrm{TLC}$ becomes zero, meaning that the constraint is not robustified and thus coincides with the original, non-robust formulation.
+
+Observing the left panel of Figure~\ref{fig:ol_sensitivities}, it is evident that the trajectories corresponding to configurations with lower values of $\ga^\textrm{TLC}$ tend to travel closer to the track boundaries.
+In contrast, the trajectory associated with $\ga^\textrm{TLC} = 3$---represented by the light green line---remains significantly farther from the edges.
+This behavior is consistent with the increased conservativeness introduced by higher values of $\ga^\textrm{TLC}$, which amplify the back-off term $\be^\textrm{TLC}$ and thereby enforce a larger safety margin from the track limits.
+Chasing a trajectory associated with a larger safety margin leads to higher sector times; spanning $\ga^\textrm{TLC}$ in the range $\left[0,3\right]$ results in sector times that vary from 12.135\,s to 13.162\,s.
+
+The second parameter examined is the initial value of the covariance matrix, $\bP_0$. From this matrix, the standard deviation of each component of the mean state vector can be computed as:
+\begin{equation}
+	\sigma_i = \sqrt{e_i^T \bP_0 e_i},\label{eq:sigma_from_P}
+\end{equation}
+where $e_i$ is the unit vector with a 1 in the $i$-th position and zeros elsewhere.
+These standard deviations, $\sigma_i$, are used to evaluate the probability of the state vector remaining within a given interval.
+For example, considering the first element of the state vector, $\mu_1$, the interval $\left[\mu_1 - \sigma_1,\ \mu_1 + \sigma_1\right]$ corresponds to a 68\% probability of containing the true state, while the interval $\left[\mu_1 - 3\sigma_1,\ \mu_1 + 3\sigma_1\right]$ corresponds to approximately 99\%.
+
 In addition to the baseline value $\bar{\bP}_0$, two alternative values are considered: $\bP_0 = 4\bar{\bP}_0$ and $\bP_0 = \frac{1}{4}\bar{\bP}_0$.
-These correspond, respectively, to doubling and halving the initial uncertainty associated with the state vector.
+These correspond, respectively, to doubling and halving the initial uncertainty associated with the state vector, as it's possible to verify from \eqref{eq:sigma_from_P}.
 

### 2025-05-30 performance_comparison.tex
[MM] some words on the results

@@
 These correspond, respectively, to doubling and halving the initial uncertainty associated with the state vector, as it's possible to verify from \eqref{eq:sigma_from_P}.
+For simplicity, a diagonal $\bar{\bP}_0$ has been chosen for this study.
+In this case, $\bar{\bP}_0$ is constructed by assigning the standard deviations of the elements of the state vector---collected in the vector $\bar{\boldsymbol{\sigma}}_0$---and placing their squared values along the diagonal.
+The values of $\bar{\boldsymbol{\sigma}}_0$ has been chosen accordingly to the range of variation of the associated variable.
+In particular, assuming the state vector is ordered as described in Section~\ref{sec:vehicle_model}, we selected $\bar{\boldsymbol{\sigma}}_0 = \left[0.1\,\textrm{m/s}, 0.01\,\textrm{m/s}, 0.01\,\textrm{rad/s}, 1\,\textrm{m}, 1\,\textrm{m}, 0.0175\,\textrm{rad}\right]^T$.
 
+The results of the analysis conducted on this parameter are shown in the right panel of Figure~\ref{fig:ol_sensitivities} where the black line corresponds to the baseline, obtained setting $\bP_0 = \bar{\bP}_0$, while the red and the blue lines corresponds to configurations with $\bP_0 = \frac{1}{4}\bar{\bP}_0$ and $\bP_0 = 4\bar{\bP}_0$, respectively.
+Obviously, using a larger initial covariance matrix results in a greater covariance matrix after $H$ steps, which in turn leads to a higher back-off term.
+This is reflected in the blue line, which travels closer to the centerline w.r.t.
+the other two lines.
 \begin{figure}

### 2025-05-31 performance_comparison.tex
[MM] inserted \Delta in variations plot and some comments added

@@
 
+\begin{figure}
+	\centering
+	\includegraphics{Fig/gamma_sensitivity.pdf}
+	\hfill
+	\includegraphics{Fig/Pzero_sensitivity.pdf}
+	\caption{Effect of $\ga^\textrm{TLC}$ (left panel) and $P_0$ (right panel) on optimal trajectory.
+The parameter $\ga^\textrm{TLC}$ span in the range $\left[0,3\right]$ corresponding to a probability to meet the track limit constraint of 50\% and 99\%, respectively.
+In the left panel, in addition to the baseline value $\bar{\bP}_0$ (black line), two alternative values are considered: $\bP_0 = 4\bar{\bP}_0$ (blue line) and $\bP_0 = \frac{1}{4}\bar{\bP}_0$ (red line).
+These correspond, respectively, to doubling and halving the initial uncertainty associated with the state vector.}
+	\label{fig:ol_sensitivities}
+\end{figure}
+
 The second parameter examined is the initial value of the covariance matrix, $\bP_0$. From this matrix, the standard deviation of each component of the mean state vector can be computed as:
@@
 the other two lines.
-\begin{figure}
-	\centering
-	\includegraphics{Fig/gamma_sensitivity.pdf}
-	\hfill
-	\includegraphics{Fig/Pzero_sensitivity.pdf}
-	\caption{Effect of $\ga^\textrm{TLC}$ (left panel) and $P_0$ (right panel) on optimal trajectory.
-The parameter $\ga^\textrm{TLC}$ span in the range $\left[0,3\right]$ corresponding to a probability to meet the track limit constraint of 50\% and 99\%, respectively.
-In the left panel, in addition to the baseline value $\bar{\bP}_0$ (black line), two alternative values are considered: $\bP_0 = 4\bar{\bP}_0$ (blue line) and $\bP_0 = \frac{1}{4}\bar{\bP}_0$ (red line).
-These correspond, respectively, to doubling and halving the initial uncertainty associated with the state vector.}
-	\label{fig:ol_sensitivities}
-\end{figure}
 
@@
 Figure~\ref{fig:ol_telemetries} presents the variation of four signals with respect to the nominal trajectory, for the three previously mentioned cases.
-The panels display the longitudinal speed $u$ (first panel), wheel steering angle $\de$ (second panel), total longitudinal force $F_x$ (third panel), and lateral deviation of the center of mass from the track centerline $e$ (fourth panel).
-Except for the $u$ panel, each plot includes dashed black lines that indicate the sign of the reference signal-whether it is positive, negative, or numerically zero.
+The panels display variations on the longitudinal speed $\De u$ (first panel), wheel steering angle $\De \de$ (second panel), total longitudinal force $\De F_x$ (third panel), and lateral deviation of the center of mass from the track centerline $\De e$ (fourth panel).
+Except for the $\De u$ panel, each plot includes dashed black lines that indicate the sign of the reference signal-whether it is positive, negative, or numerically zero.
 These sign indicators are represented as three distinct levels---high, zero, and low---corresponding to positive, zero, and negative values of the reference signal, respectively.
@@
 	\includegraphics{Fig/ol_telemetries.pdf}
-	\caption{Evolution of the longitudinal speed u, wheel steering angle $\de$, longitudinal force $F_x$, and lateral deviation $e$ are reported for three robustification scenarios: TLC (Track Limit Constraints), FLC (Friction Limit Constraint), and TLC+FLC (both constraints enforced concurrently).
+	\caption{Evolution of the longitudinal speed $\De u$, wheel steering angle $\De \de$, longitudinal force $\De F_x$, and lateral deviation $\De e$ are reported for three robustification scenarios: TLC (Track Limit Constraints), FLC (Friction Limit Constraint), and TLC+FLC (both constraints enforced concurrently).
 Each profile is shown relative to the baseline (nominal) trajectory to highlight the effect of each constraint-handling strategy.
@@
 From the first panel, it can be observed that all configurations maintain a lower mean value of the longitudinal speed compared to the reference trajectory, resulting in an increased sector time.
-The two configurations incorporating the robustified track limit constraint exhibit, as expected, a significantly altered CoM trajectory.
+Specifically, the configuration with the robustified TLC increases the sector time by 1.66\%, while the configuration with the robustified FLC and with both TLC and FLC robustified increase the sector time by a 0.79\% and 2.57\%, respectively.
+The two configurations incorporating the robustified track limit constraint (blue and green lines) exhibit, as expected, a significantly altered CoM trajectory.
 This behavior explains the higher longitudinal speed observed around $\al = 0.71$, which is followed by a sharp reduction.
@@
 The second and fourth panels clearly show that the configurations with the robustified track limit constraint tend to follow a path closer to the centerline.
-In particular, for most of the sector, the variation in lateral displacement and the lateral displacement $e$ itself exhibit opposite signs, indicating a corrective behavior of the steering angle $\de$ that pulls the vehicle toward the centerline.
+In particular, for most of the sector, the variation in lateral displacement $\De e$ and the lateral displacement $e$ itself exhibit opposite signs, indicating a corrective behavior of the steering angle $\de$ that pulls the vehicle toward the centerline.
 
-It can be interesting to notice that, in the third panel, only the configuration with both the constraints robustified needs to reduce the longitudinal force during the high speed turn, in the curvilinear abscissa interval $\al\in\left[0.75, 0.76\right]$, to meet both the constraints.
+The third panel, which shows the variation of the total longitudinal force $\De F_x$, is consistent with the $\De u$ panel.
+In particular, the configurations with the robustified TLC exhibit a higher longitudinal force while approaching Turn 1, in the curvilinear abscissa interval $\al\in\left[0.707, 0.712\right]$, which results in a higher longitudinal speed.
+Subsequently, these two configurations apply a lower longitudinal force during braking compared to the nominal case, indicating more intense braking.
+This behavior explains the sharp reduction in longitudinal speed observed thereafter.
+
+It can be interesting to notice that, in the third panel, only the configuration with both the constraints robustified (blue line) needs to reduce the longitudinal force during the high speed turn, in the curvilinear abscissa interval $\al\in\left[0.75, 0.76\right]$, to meet both the constraints.
 The vehicle constrained by the track limit condition is forced to follow a wider trajectory, which entails greater lateral acceleration and, consequently, a higher lateral force demand---resulting in increased tire utilization.

### 2025-06-02 performance_comparison.tex
[MM] intro to saturation
still no nice name for the quantity

@@
 From the first panel, it can be observed that all configurations maintain a lower mean value of the longitudinal speed compared to the reference trajectory, resulting in an increased sector time.
-Specifically, the configuration with the robustified TLC increases the sector time by 1.66\%, while the configuration with the robustified FLC and with both TLC and FLC robustified increase the sector time by a 0.79\% and 2.57\%, respectively.
+Specifically, the configuration with the robustified TLC increases the sector time by 1.66\%, while the configuration with the robustified FLC and with both TLC and FLC robustified increase the sector time by 0.79\% and 2.57\%, respectively.
 The two configurations incorporating the robustified track limit constraint (blue and green lines) exhibit, as expected, a significantly altered CoM trajectory.
@@
 
+After analyzing the variations of four key signals with respect to the nominal solution, we now turn our attention to the evaluation of axle saturation, which provides insight into how close each configuration operates to the tire grip limits.
+Figure~\ref{fig:ol_saturation} presents the quantity $\left[ \left( \frac{X_j(\bx,\bu)}{\mu_{x,j}} \right)^2+ \left( \frac{Y_j(\bx,\bu)}{\mu_{y,j}}\right)^2\right]\frac{1}{Z_j^2(\bx,\bu)}$, where $j=1$ refers to the front axle (first panel) and $j=2$ the rear axle (second panel), plotted for the three previously introduced configurations (color-coded as before) and for the nominal solution (black dashed line).
+This quantity indicates how close each configuration operates to the friction limit; it equals 1 when the point $\left(X_j, Y_j\right)$ lies exactly on the friction ellipse defined by $Z_j$.
+
 \begin{figure}
@@
 \end{figure}
+
+As expected, the two configurations with the robustified friction limit constraint (red and blue lines) never reach a value of 1, due to the presence of the back-off term, which enforces a safety margin from the friction limit.
 

### 2025-06-03 performance_comparison.tex
[MM] tmp for Martino

@@
 
-Figure~\ref{fig:ol_telemetries} presents the variation of four signals with respect to the nominal trajectory, for the three previously mentioned cases.
-The panels display variations on the longitudinal speed $\De u$ (first panel), wheel steering angle $\De \de$ (second panel), total longitudinal force $\De F_x$ (third panel), and lateral deviation of the center of mass from the track centerline $\De e$ (fourth panel).
-Except for the $\De u$ panel, each plot includes dashed black lines that indicate the sign of the reference signal-whether it is positive, negative, or numerically zero.
-These sign indicators are represented as three distinct levels---high, zero, and low---corresponding to positive, zero, and negative values of the reference signal, respectively.
-The levels are scaled appropriately in each plot to ensure clarity of visualization.
-
 \begin{figure}
@@
 \end{figure}
+
+Figure~\ref{fig:ol_telemetries} presents the variation of four signals with respect to the nominal trajectory, for the three previously mentioned cases.
+The panels display variations on the longitudinal speed $\De u$ (first panel), wheel steering angle $\De \de$ (second panel), total longitudinal force $\De F_x$ (third panel), and lateral deviation of the center of mass from the track centerline $\De e$ (fourth panel).
+Except for the $\De u$ panel, each plot includes dashed black lines that indicate the sign of the reference signal-whether it is positive, negative, or numerically zero.
+These sign indicators are represented as three distinct levels---high, zero, and low---corresponding to positive, zero, and negative values of the reference signal, respectively.
+The levels are scaled appropriately in each plot to ensure clarity of visualization.
 
@@
 As expected, the two configurations with the robustified friction limit constraint (red and blue lines) never reach a value of 1, due to the presence of the back-off term, which enforces a safety margin from the friction limit.
+From this plot, it can be observed that the configurations with the robustified track limit constraint (green and blue lines) follow significantly different trajectories compared to the nominal solution, resulting in distinct ground force demands.
+This behavior is noticeable in turn 1 and even more clearly in turn 2.
+Specifically, these two configurations begin to experience high axle loads approximately 12\,m before the nominal trajectory and sustain them for an additional 4.6\,m (TLC, green line) and 9.2\,m (TLC+FLC, blue line) beyond the nominal reference.
 

### 2025-06-03 performance_comparison.tex
[mg] sec 5 intro

@@
 \label{sec:performance_comparison}
+
+In this section we present some results obtained using the described robust planning frameworks. All planned trajectories refer to the sector of the Catalunya circuit shown in Figure~\ref{fig:track}. 
+The entire circuit is parametrized by the curvilinear abscissa $\alpha \in [0,1]$, and the selected sector corresponds to the interval $\left[0.70, 0.77\right]$.
+Checkpoints are indicated by labels and are uniformly spaced by $\De\al=0.01$.
+This sector includes two distinct corners: a low-speed turn from $\al = 0.72$ to $\al = 0.73$, and a high-speed turn from $\al = 0.75$ to $\al = 0.76$, allowing for the evaluation of vehicle behavior across different dynamic scenarios.
+
+We begin by presenting a selection of results obtained with the open-loop planning approach, which allows for a clearer illustration of the method and its underlying mechanisms.
+This approach is used as a reference to analyze the influence of key design parameters and of the different types of constraints on the resulting trajectories.
+A comparison is then carried out between the open-loop and closed-loop strategies.
+Finally, the planned trajectories are compared against those followed in simulations with noise realizations, providing empirical validation of the proposed framework.
+
+% Contour plot dei LAP time in funzione dei due gamma, lo mettiamo o no? Eventialmente nel comparison open/closed
 
@@
 
-\subsection{Open loop parameters sensitivity}
+\subsection{Open-loop parameters sensitivity}
 \label{sec:ol_param_sensitivity}
-In this section, we explore the open-loop approach to gain deeper insight into the influence of key parameters on the results.
-For this analysis, we consider optimizations in which only the track limit constraint is robustified.
-This choice is motivated by the fact that the effects of the track limit constraint are more visually evident compared to those of the friction limit constraint, thereby facilitating a more immediate understanding of their impact.
+To gain deeper insight into how key parameters affect the planning outcome, we focus here on the open-loop formulation.
+%In this section, we explore the open-loop approach to gain deeper insight into the influence of key parameters on the results.
+At this stage, only the track limit constraint is robustified. 
+%For this analysis, we consider optimizations in which only the track limit constraint is robustified. 
+This choice is motivated by the fact that its effects are more visually evident than those of the friction limit constraint, thereby facilitating a more immediate understanding of their impact.
 
-The first parameter analyzed is the factor $\ga^\textrm{TLC}$, which multiplies the standard deviation of the constraint, $\sigma^\textrm{TLC}$, to determine the total back-off term $\be^\textrm{TLC}$.
+The first parameter considered is the confidence level of constraint satisfaction. 
+We recall that the factor $\ga^\textrm{TLC}$ acts as a tuning knob, multiplying the standard deviation of the constraint, $\sigma^\textrm{TLC}$, to determine the total back-off term $\be^\textrm{TLC}$.
 This parameter directly influences the probability of satisfying the track limit constraint: higher values of $\ga^\textrm{TLC}$ correspond to more conservative (i.e., robust) behavior. 
-Specifically, $\ga^\textrm{TLC} = 0$ yields a satisfaction probability of 50\%, while $\ga^\textrm{TLC} = 3$ corresponds to 99\%.
-For the purposes of this analysis, we examine the effect of varying $\ga^\textrm{TLC}$ within this range.
+Specifically, $\ga^\textrm{TLC} = 0$ yields a satisfaction probability of 50\% and leads to the nominal solution with no back-off, while $\ga^\textrm{TLC} = 3$ corresponds to a confidence level of 99\%.
+For the purposes of this analysis, we examine the effect of varying $\ga^\textrm{TLC}$ across this range.
 
-Setting $\ga^\textrm{TLC} = 0$ leads to the nominal solution.
+%Setting $\ga^\textrm{TLC} = 0$ leads to the nominal solution.
 In this case, the back-off term $\be^\textrm{TLC}$ becomes zero, meaning that the constraint is not robustified and thus coincides with the original, non-robust formulation.
@@
 	\caption{Effect of $\ga^\textrm{TLC}$ (left panel) and $P_0$ (right panel) on optimal trajectory.
-The parameter $\ga^\textrm{TLC}$ span in the range $\left[0,3\right]$ corresponding to a probability to meet the track limit constraint of 50\% and 99\%, respectively.
-In the left panel, in addition to the baseline value $\bar{\bP}_0$ (black line), two alternative values are considered: $\bP_0 = 4\bar{\bP}_0$ (blue line) and $\bP_0 = \frac{1}{4}\bar{\bP}_0$ (red line).
+The parameter $\ga^\textrm{TLC}$ spans in the range $\left[0,3\right]$ corresponding to a probability to meet the track limit constraint ranging from 50\% to 99\%.
+In the right panel, in addition to the baseline value $\bar{\bP}_0$ (black line), two alternative values are considered: $\bP_0 = 4\bar{\bP}_0$ (blue line) and $\bP_0 = \frac{1}{4}\bar{\bP}_0$ (red line).
 These correspond, respectively, to doubling and halving the initial uncertainty associated with the state vector.}

### 2025-06-03 performance_comparison.tex
[MM] alignment
[mg] subsec 5.1

@@
 
-The second parameter examined is the initial value of the covariance matrix, $\bP_0$. From this matrix, the standard deviation of each component of the mean state vector can be computed as:
-\begin{equation}
-	\sigma_i = \sqrt{e_i^T \bP_0 e_i},\label{eq:sigma_from_P}
-\end{equation}
-where $e_i$ is the unit vector with a 1 in the $i$-th position and zeros elsewhere.
-These standard deviations, $\sigma_i$, are used to evaluate the probability of the state vector remaining within a given interval.
-For example, considering the first element of the state vector, $\mu_1$, the interval $\left[\mu_1 - \sigma_1,\ \mu_1 + \sigma_1\right]$ corresponds to a 68\% probability of containing the true state, while the interval $\left[\mu_1 - 3\sigma_1,\ \mu_1 + 3\sigma_1\right]$ corresponds to approximately 99\%.
+The second parameter examined is the initial value of the covariance matrix, $\bP_0$. 
+We recall that at each grid point a new prediction horizon is initialized, and the corresponding matrix $\bP$ at its starting point, i.e., $\bP^0_k$, is set to $\bP_0$. 
+The diagonal elements of $\bP$ are the variances $\text{var}(x_i)$ of each state variables, while the off-diagonal elements encode their covariances $\text{cov}(x_i,x_j)$.
+Larger diagonal values indicate higher uncertainty in the corresponding states, and nonzero off-diagonal terms imply mutual dependence between them.
+For simplicity, a diagonal $\bP_0$ is used in this study, that is, for each prediction horizon we assign an initial standard deviation $\sigma_i = \sqrt{\text{var}(x_i)}$ to each state and assume all initial covariances to be zero.
+
+To provide a concrete interpretation, consider the first state variable $x_1$ with mean $\mu_1$ and standard deviation $\sigma_1$.
+Under the Gaussian assumption, the true value of $x_1$ lies within the interval $\left[\mu_1 - \sigma_1,\ \mu_1 + \sigma_1\right]$ with approximately 68\% probability, and within the interval $\left[\mu_1 - 3\sigma_1,\ \mu_1 + 3\sigma_1\right]$ with approximately 99\% probability.
+
+We assign a baseline initial covariance matrix $\bP_0 = \bar{\bP}_0$ by selecting standard deviations $\bar{\sigma}_i$ for each state variable.
+These standard deviations have been chosen based on the typical range of variation of each state.
+Assuming the state vector is ordered as described in Section~\ref{sec:vehicle_model}, the values used are $\bar{\boldsymbol{\sigma}}~=~\left[0.1\,\textrm{m/s}, 0.01\,\textrm{m/s}, 0.01\,\textrm{rad/s}, 1\,\textrm{m}, 1\,\textrm{m}, 0.0175\,\textrm{rad}\right]^T$.
 
 In addition to the baseline value $\bar{\bP}_0$, two alternative values are considered: $\bP_0 = 4\bar{\bP}_0$ and $\bP_0 = \frac{1}{4}\bar{\bP}_0$.
-These correspond, respectively, to doubling and halving the initial uncertainty associated with the state vector, as it's possible to verify from \eqref{eq:sigma_from_P}.
-For simplicity, a diagonal $\bar{\bP}_0$ has been chosen for this study.
-In this case, $\bar{\bP}_0$ is constructed by assigning the standard deviations of the elements of the state vector---collected in the vector $\bar{\boldsymbol{\sigma}}_0$---and placing their squared values along the diagonal.
-The values of $\bar{\boldsymbol{\sigma}}_0$ has been chosen accordingly to the range of variation of the associated variable.
-In particular, assuming the state vector is ordered as described in Section~\ref{sec:vehicle_model}, we selected $\bar{\boldsymbol{\sigma}}_0 = \left[0.1\,\textrm{m/s}, 0.01\,\textrm{m/s}, 0.01\,\textrm{rad/s}, 1\,\textrm{m}, 1\,\textrm{m}, 0.0175\,\textrm{rad}\right]^T$.
-
-The results of the analysis conducted on this parameter are shown in the right panel of Figure~\ref{fig:ol_sensitivities} where the black line corresponds to the baseline, obtained setting $\bP_0 = \bar{\bP}_0$, while the red and the blue lines corresponds to configurations with $\bP_0 = \frac{1}{4}\bar{\bP}_0$ and $\bP_0 = 4\bar{\bP}_0$, respectively.
-Obviously, using a larger initial covariance matrix results in a greater covariance matrix after $H$ steps, which in turn leads to a higher back-off term.
+These correspond, respectively, to doubling and halving the initial standard deviations $\sigma_i$ associated with the state vector.
+A comparison between the trajectories obtained using different values of $\bP_0$ is shown in the right panel of Figure~\ref{fig:ol_sensitivities}.
+The black line corresponds to the baseline case with $\bP_0 = \bar{\bP}_0$, while the red and the blue lines represent the cases with $\bP_0 = \frac{1}{4}\bar{\bP}_0$ and $\bP_0 = 4\bar{\bP}_0$, respectively.
+As reasonably expected, using a larger initial covariance matrix results in a greater covariance matrix after $H$ steps, which in turn leads to a higher back-off term.
 This is reflected in the blue line, which travels closer to the centerline w.r.t.
@@
 For conciseness, the comparison is limited to a nominal trajectory and a trajectory computed using the closed-loop method with only the track limit constraint robustified.
-An LQR controller, designed according to the procedure outlined in Section~\ref{sec:LQR}, is implemented along the nominal trajectory.
+An LQR controller, designed according to the procedure outlined in Section~\ref{sec:LQR}, is implemented to stabilize the nominal trajectory.
 In contrast, the robust trajectory is tracked using the controller directly obtained from the optimization process.

### 2025-06-04 performance_comparison.tex
[mg] subsec 5.2

@@
 Assuming the state vector is ordered as described in Section~\ref{sec:vehicle_model}, the values used are $\bar{\boldsymbol{\sigma}}~=~\left[0.1\,\textrm{m/s}, 0.01\,\textrm{m/s}, 0.01\,\textrm{rad/s}, 1\,\textrm{m}, 1\,\textrm{m}, 0.0175\,\textrm{rad}\right]^T$.
-
 In addition to the baseline value $\bar{\bP}_0$, two alternative values are considered: $\bP_0 = 4\bar{\bP}_0$ and $\bP_0 = \frac{1}{4}\bar{\bP}_0$.
 These correspond, respectively, to doubling and halving the initial standard deviations $\sigma_i$ associated with the state vector.
+
 A comparison between the trajectories obtained using different values of $\bP_0$ is shown in the right panel of Figure~\ref{fig:ol_sensitivities}.
@@
 As reasonably expected, using a larger initial covariance matrix results in a greater covariance matrix after $H$ steps, which in turn leads to a higher back-off term.
-This is reflected in the blue line, which travels closer to the centerline w.r.t.
-the other two lines.
+This is reflected in the blue line, which travels closer to the centerline with respect to the other two lines.
 
 \subsection{Robustified constraints comparison}
-It is of interest to analyze how a robustified constraint---or a combination of the two described in Sections~\ref{sec:FLC} and~\ref{sec:TLC}---influences the driving style.
-This analysis compares the not-robust trajectory, obtained without robustified constraints, with those resulting from a robustified track limit constraint (denoted as TLC), a robustified friction limit constraint (denoted as FLC), and a scenario where both constraints are robustified (denoted as TLC+FLC).
-All the optimization are obtained with the open-loop method described in Section~\ref{sec:open_loop_planning}, setting $H=4$, and $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$, corresponding to a 90\% probability of meeting both constraints.
+%It is of interest to analyze how a robustified constraint---or a combination of the two described in Sections~\ref{sec:FLC} and~\ref{sec:TLC}---influences the driving style. 
+We now investigate how the inclusion of different kinds of robustified constraints---individually or in combination, as described in Sections~\ref{sec:FLC} and~\ref{sec:TLC}---influences the driving style.
+We compare the nominal feed-forward optimal trajectory, obtained without robustified constraints, with those resulting from a robustified track limit constraint (denoted as TLC), a robustified friction limit constraint (denoted as FLC), and a scenario where both constraints are robustified (denoted as TLC+FLC).
+All the optimization are obtained with the open-loop method described in Section~\ref{sec:open_loop_planning}, setting $H=4$, and $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$, corresponding to a 90\% probability of satisfying both constraints.
 
@@
 
-Figure~\ref{fig:ol_telemetries} presents the variation of four signals with respect to the nominal trajectory, for the three previously mentioned cases.
-The panels display variations on the longitudinal speed $\De u$ (first panel), wheel steering angle $\De \de$ (second panel), total longitudinal force $\De F_x$ (third panel), and lateral deviation of the center of mass from the track centerline $\De e$ (fourth panel).
-Except for the $\De u$ panel, each plot includes dashed black lines that indicate the sign of the reference signal-whether it is positive, negative, or numerically zero.
-These sign indicators are represented as three distinct levels---high, zero, and low---corresponding to positive, zero, and negative values of the reference signal, respectively.
-The levels are scaled appropriately in each plot to ensure clarity of visualization.
+%---Delta key signals
 
-From the first panel, it can be observed that all configurations maintain a lower mean value of the longitudinal speed compared to the reference trajectory, resulting in an increased sector time.
-Specifically, the configuration with the robustified TLC increases the sector time by 1.66\%, while the configuration with the robustified FLC and with both TLC and FLC robustified increase the sector time by 0.79\% and 2.57\%, respectively.
-The two configurations incorporating the robustified track limit constraint (blue and green lines) exhibit, as expected, a significantly altered CoM trajectory.
-This behavior explains the higher longitudinal speed observed around $\al = 0.71$, which is followed by a sharp reduction.
+The comparison between the nominal case and the three robustified cases is illustrated in Figure~\ref{fig:ol_telemetries}, in terms of the deviation of selected signals from to their nominal counterparts.
+The panels display variations on the longitudinal speed $\De u$ (first panel), wheel steering angle $\De \de$ (second panel), total longitudinal force $\De F_x$ (third panel), and lateral deviation of the center of mass (CoM) from the track centerline $\De e$ (fourth panel).
+Except for the $\De u$ panel, each plot includes dashed black lines that indicate the sign of the nominal signal. 
+These indicators are shown at three distinct vertical levels---upper, central, and lower---corresponding respectively to positive, near-zero, and negative values of the nominal signal.
 
-The second and fourth panels clearly show that the configurations with the robustified track limit constraint tend to follow a path closer to the centerline.
+%--------1 Considerazioni generali sul settore
+
+From the first panel, it can be observed that all robustified configurations, on average, show a lower longitudinal speed compared to the nominal trajectory, resulting in an increased sector time.
+Specifically, the configuration with the robustified TLC increases the sector time by 1.66\%, while that with the robustified FLC leads to a 0.79\% increase.
+When both TLC and FLC are robustified, the sector time increases by 2.57\%.
+The two configurations incorporating the robustified TLC (blue and green lines) exhibit, as expected, a significantly altered CoM trajectory.
+Ideed, the second and fourth panels clearly show that the TLC and TLC+FLC configurations tend to follow a path closer to the centerline.
 In particular, for most of the sector, the variation in lateral displacement $\De e$ and the lateral displacement $e$ itself exhibit opposite signs, indicating a corrective behavior of the steering angle $\de$ that pulls the vehicle toward the centerline.
 
-The third panel, which shows the variation of the total longitudinal force $\De F_x$, is consistent with the $\De u$ panel.
-In particular, the configurations with the robustified TLC exhibit a higher longitudinal force while approaching Turn 1, in the curvilinear abscissa interval $\al\in\left[0.707, 0.712\right]$, which results in a higher longitudinal speed.
+%--------2 Sharp turn
+
+This shift in trajectory and longitudinal speed becomes especially evident when approaching the sharp turn at $\al \in [0.707, 0.712]$.
+A noticeable increase in longitudinal speed is observed around $\alpha = 0.71$ for the TLC configuration  in the $\De u$ panel.
+This peak is followed by a sharp drop in speed, indicating intense braking.
+The TLC+FLC configuration exhibits a qualitatively similar trend in $\Delta u$, although the peak is significantly lower and the subsequent speed reduction is clearly stronger.
+This behavior is consistent with the variation of the total longitudinal force $\Delta F_x$ shown in the third panel.
+In particular, the TLC and TLC+FLC configurations exhibit a higher longitudinal force while approaching the sharp turn, in the curvilinear abscissa interval $\al\in\left[0.707, 0.712\right]$, resulting in a higher longitudinal speed.
 Subsequently, these two configurations apply a lower longitudinal force during braking compared to the nominal case, indicating more intense braking.
@@
 
-It can be interesting to notice that, in the third panel, only the configuration with both the constraints robustified (blue line) needs to reduce the longitudinal force during the high speed turn, in the curvilinear abscissa interval $\al\in\left[0.75, 0.76\right]$, to meet both the constraints.
-The vehicle constrained by the track limit condition is forced to follow a wider trajectory, which entails greater lateral acceleration and, consequently, a higher lateral force demand---resulting in increased tire utilization.
-Due to the limitations imposed by the robust friction limit constraint, it is necessary to reduce the available accelerating force.
+%--------3 High speed turn
 
-After analyzing the variations of four key signals with respect to the nominal solution, we now turn our attention to the evaluation of axle saturation, which provides insight into how close each configuration operates to the tire grip limits.
-Figure~\ref{fig:ol_saturation} presents the quantity $\left[ \left( \frac{X_j(\bx,\bu)}{\mu_{x,j}} \right)^2+ \left( \frac{Y_j(\bx,\bu)}{\mu_{y,j}}\right)^2\right]\frac{1}{Z_j^2(\bx,\bu)}$, where $j=1$ refers to the front axle (first panel) and $j=2$ the rear axle (second panel), plotted for the three previously introduced configurations (color-coded as before) and for the nominal solution (black dashed line).
-This quantity indicates how close each configuration operates to the friction limit; it equals 1 when the point $\left(X_j, Y_j\right)$ lies exactly on the friction ellipse defined by $Z_j$.
+Later in the sector, a further deviation in driving style appears during the high-speed turn, specifically around $\alpha \in [0.75, 0.76]$.
+As shown in the $\Delta F_x$ panel, the combined configuration TLC+FLC needs to reduce the longitudinal force within this section to meet both the constraints. 
+This adaptation is necessary because the TLC forces the vehicle to follow a wider trajectory, which entails greater lateral acceleration and, consequently, higher lateral force demand.
+To satisfy the FLC under the resulting increase in tire utilization, the available accelerating force must be reduced.
+Based on the nominal force sign indicator, this reduction is achieved by partially lifting the throttle.
+
+%---Axle saturation
+After analyzing the variations of these four key signals with respect to the nominal solution, we now turn our attention to the evaluation of axle saturation, which provides insight into how close each configuration operates to the tire grip limits.
+In accordance with the FLC formulation introduced in~\ref{eq:adherence_constraint}, we define the \emph{axle saturation ratio} $S_j$ as
+\begin{equation}
+	S_j = \frac{ \left( \frac{X_j(\bx,\bu)}{\mu_{x,j}} \right)^2+ \left( \frac{Y_j(\bx,\bu)}{\mu_{y,j}}\right)^2}{Z_j^2(\bx,\bu)}.
+\end{equation}
+This quantity indicates how close each configuration operates to the friction limit; $S_j = 1$ when the point $\left(X_j, Y_j\right)$ lies exactly on the friction ellipse, i.e., under full saturation.
+Figure~\ref{fig:ol_saturation} shows the values of $S_j$ for the front axle ($j=1$, first panel) and the rear axle ($j=2$, rear panel) for the previously introduced configurations: the nominal case (black dashed line) and the three robustified cases (color-coded as before).
 
@@
 
-As expected, the two configurations with the robustified friction limit constraint (red and blue lines) never reach a value of 1, due to the presence of the back-off term, which enforces a safety margin from the friction limit.
-From this plot, it can be observed that the configurations with the robustified track limit constraint (green and blue lines) follow significantly different trajectories compared to the nominal solution, resulting in distinct ground force demands.
+First, we observe that the FLC and TLC+FLC configurations (red and blue lines) never reach full saturation, i.e., $S_j=1$, due to the presence of the back-off term, which enforces a safety margin from the friction limit.
+Second, we point out how the different trajectories followed by the TLC and TLC+FLC configurations (green and blue lines), compared to the nominal and FLC cases, entail distinct ground force demands, which are clearly visible here in terms of the $S_j$ ratio.
+This behavior emerges around the sharp turn and becomes even more evident in the high-speed turn. 
+Specifically, in the interval $\alpha \in [0.745, 0.762]$, the TLC and TLC+FLC configurations begin to experience high axle loads approximately 12\,m before the nominal trajectory and sustain them for an additional 4.6\,m (TLC, green line) and 9.2\,m (TLC+FLC, blue line) beyond the nominal reference.
+
+%As already observed, the configurations with the robustified track limit constraint (green and blue lines) follow significantly different paths compared to the nominal solution.
+This deviation results in distinct ground force demands, clearly visible in the saturation index.
+%From this plot, it can be observed that the configurations with the robustified track limit constraint (green and blue lines) follow significantly different trajectories compared to the nominal solution, resulting in distinct ground force demands.
 This behavior is noticeable in turn 1 and even more clearly in turn 2.
 Specifically, these two configurations begin to experience high axle loads approximately 12\,m before the nominal trajectory and sustain them for an additional 4.6\,m (TLC, green line) and 9.2\,m (TLC+FLC, blue line) beyond the nominal reference.
+
+\subsection{Comparison between open-loop and closed-loop approaches}
 

### 2025-06-04 performance_comparison.tex
[mg] subsec 5.3 ol VS cl

@@
 After analyzing the variations of these four key signals with respect to the nominal solution, we now turn our attention to the evaluation of axle saturation, which provides insight into how close each configuration operates to the tire grip limits.
-In accordance with the FLC formulation introduced in~\ref{eq:adherence_constraint}, we define the \emph{axle saturation ratio} $S_j$ as
+In accordance with the FLC formulation introduced in~\eqref{eq:adherence_constraint}, we define the \emph{axle saturation ratio} $S_j$ as
 \begin{equation}
@@
 
-\subsection{Comparison between open-loop and closed-loop approaches}
+\subsection{Open-loop vs Closed-loop evaluation}
+We now compare the open-loop planning with the closed-loop approach.
+In particular, we consider open-loop and closed-loop trajectories computed using both the robustified TLC and FLC, with $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$.
+As before, the open-loop planning is carried out with a prediction horizon of $H=4$ steps.
+
+For the TLC, the back-off term $\be^{TLC}$ in~\eqref{eq:TLC} admits a clear geometrical interpretation: it corresponds to the amount by which the track boundaries are tightened to meet the probabilistic constraint, or, equivalently, to a virtual extension of the vehicle's width.
+According to the latter interpretation, Figure~\ref{fig:olcl_traj_bands} shows two highlighted bands---red for the open-loop case and cyan for the closed-loop case---corresponding to the area swept by the back-off-extended-vehicle.
+Ultimately, the TLC enforces precisely that these bands remain entirely within the actual track limits.
+The comparison between the two regions illustrates how the additional conservativeness of the open-loop approach stems from larger back-off terms, i.e., a wider band that must stay within the track boundaries.
+The closed-loop method, on the other hand, benefits from the LQR-based feedback, which allows it to maintain smaller back-off terms and thus a narrower band, requiring a reduced safety margin from the track limits.
+
+\begin{figure}
+	\centering
+	\includegraphics{Fig/olcl_traj_bands.pdf}
+	\caption{Lorem ipsum.}
+	\label{fig:olcl_traj_bands}
+\end{figure}
 

### 2025-06-04 performance_comparison.tex
[mg] subsec 5.4

@@
 
-\subsection{Open-loop vs Closed-loop evaluation}
+\subsection{Open-loop vs closed-loop robust planning}
 We now compare the open-loop planning with the closed-loop approach.
@@
 The closed-loop method, on the other hand, benefits from the LQR-based feedback, which allows it to maintain smaller back-off terms and thus a narrower band, requiring a reduced safety margin from the track limits.
-
 \begin{figure}
@@
 \subsection{Simulations with noise realization}
-This analysis aims to provide empirical validation of the robust trajectories obtained using the methods proposed in this work.
+Finally, we provide an empirical validation of the closed-loop robustified planning by simulating the evolution of the controlled vehicle with random initial conditions and including noise realizations.
+The random values for the initial state and the noise follow Gaussian distributions with covariance matrices $\bP_0$ and $\bQ$, respectively, as introduced in~\eqref{eq:dP}.
+This analysis cannot be performed for the open-loop approach, as it does not compute a stabilizing controller---resulting in trajectories that would significantly deviate from the track limits beyond the prediction horizon.
+For visual clarity, the comparison is limited to trajectories with the robustified TLC, but analogous considerations apply to the robustified FLC as well.
+
+The left panel of Figure~\ref{fig:traj_strings} shows random evolutions (red lines) for an LQR-controlled vehicle stabilized to follow the nominal optimal trajectory (black line).
+No robust constraints are introduced, and the controller implements the feedback policy from step 2 of Subsection~\ref{sec:LQR}.
+The right panel, instead, shows the random evolutions (red lines) for the robustified LQR-based controller obtained in the step 3 of the closed-loop approach described in Subsection~\ref{sec:CLrobustified}.
+The corresponding robustified mean trajectory is shown in black.
+The comparison between the panels shows that the robustified controller is significantly more effective at satisfying the TLC.
+Indeed, even the most lateral trajectories remain within the track boundaries.
+
+%---------------------------------
+%ORIGINAL VERSION
+
+%This analysis aims to provide empirical validation of the robust trajectories obtained using the methods proposed in this work.
 For conciseness, the comparison is limited to a nominal trajectory and a trajectory computed using the closed-loop method with only the track limit constraint robustified.
-An LQR controller, designed according to the procedure outlined in Section~\ref{sec:LQR}, is implemented to stabilize the nominal trajectory.
+%An LQR controller, designed according to the procedure outlined in Section~\ref{sec:LQR}, is implemented to stabilize the nominal trajectory.
 In contrast, the robust trajectory is tracked using the controller directly obtained from the optimization process.

### 2025-06-04 performance_comparison.tex
[mg] sec 5 review

@@
 This analysis cannot be performed for the open-loop approach, as it does not compute a stabilizing controller---resulting in trajectories that would significantly deviate from the track limits beyond the prediction horizon.
-For visual clarity, the comparison is limited to trajectories with the robustified TLC, but analogous considerations apply to the robustified FLC as well.
+%For visual clarity, the comparison is limited to trajectories with the robustified TLC, but analogous considerations apply to the robustified FLC as well.
 
@@
 The corresponding robustified mean trajectory is shown in black.
-The comparison between the panels shows that the robustified controller is significantly more effective at satisfying the TLC.
-Indeed, even the most lateral trajectories remain within the track boundaries.
+The comparison between the panels shows that the robustified controller is significantly more effective at satisfying the TLC, resulting in safer and more reliable trajectory tracking.
 

### 2025-06-04 Images/test0.tex
[mg] general reviews

\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt

\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
	%uncomment if require: \path (0,470); %set diagram left start at 0, and has height of 470

	%Curve Lines [id:da3902337050237109] 
	\draw    (131,68.6) .. controls (115,164.6) and (172,202.6) .. (201,210.6) .. controls (230,218.6) and (301.88,256.97) .. (321,278.6) .. controls (340.13,300.22) and (365,360.6) .. (467,351.6) ;
	%Curve Lines [id:da9978493675513382] 
	\draw    (224,42.6) .. controls (208,138.6) and (250,136.6) .. (279,144.6) .. controls (308,152.6) and (368.88,190.97) .. (388,212.6) .. controls (407.13,234.22) and (417,285.6) .. (519,276.6) ;
	%Curve Lines [id:da9577580542618884] 
	\draw [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ] [dash pattern={on 4.5pt off 4.5pt}]  (205,56.6) .. controls (189,152.6) and (231,150.6) .. (260,158.6) .. controls (289,166.6) and (349.88,204.97) .. (369,226.6) .. controls (388.13,248.22) and (405,304.6) .. (507,295.6) ;
	%Curve Lines [id:da8960204057983273] 
	\draw [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ] [dash pattern={on 4.5pt off 4.5pt}]  (155,69.6) .. controls (139,165.6) and (191,186.6) .. (220,194.6) .. controls (249,202.6) and (324.88,251.97) .. (344,273.6) .. controls (363.13,295.22) and (362,332.6) .. (469,338.6) ;
	%Curve Lines [id:da018938305669631506] 
	\draw [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ] [dash pattern={on 0.84pt off 2.51pt}]  (176,61.6) .. controls (160,157.6) and (205,165.6) .. (234,173.6) .. controls (263,181.6) and (325.88,218.97) .. (345,240.6) .. controls (364.13,262.22) and (383,331.6) .. (485,322.6) ;
	%Straight Lines [id:da3684832566848931] 
	\draw    (288.9,137.15) -- (274.98,159.6) ;
	\draw [shift={(273.4,162.15)}, rotate = 301.8] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (8.93,-4.29) -- (0,0) -- (8.93,4.29) -- cycle    ;
	%Straight Lines [id:da19558075317267187] 
	\draw    (245,102.6) -- (233.97,120.59) ;
	\draw [shift={(232.4,123.15)}, rotate = 301.51] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (8.93,-4.29) -- (0,0) -- (8.93,4.29) -- cycle    ;

	% Text Node
	\draw (229,83.4) node [anchor=north west][inner sep=0.75pt]    {$\beta =0$};
	% Text Node
	\draw (281.5,114.4) node [anchor=north west][inner sep=0.75pt]    {$\beta =0.1$};

\end{tikzpicture}

### 2025-06-04 Images/test1.tex
[mg] general reviews

\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt

\begin{tikzpicture}[x=0.75pt,y=0.75pt,scale=0.5]
	%uncomment if require: \path (0,470); %set diagram left start at 0, and has height of 470

	%Curve Lines [id:da3902337050237109] 
	\draw    (209.6,68.4) .. controls (145.12,155.42) and (212.32,229.5) .. (271.54,259.19) .. controls (275.38,261.12) and (279.19,262.86) .. (282.93,264.4) .. controls (344.27,289.73) and (346.27,313.07) .. (354.93,333.73) ;
	%Straight Lines [id:da3684832566848931] 
	\draw    (336.93,189.73) -- (316.45,213.94) ;
	\draw [shift={(314.52,216.23)}, rotate = 310.23] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (8.93,-4.29) -- (0,0) -- (8.93,4.29) -- cycle    ;
	%Straight Lines [id:da19558075317267187] 
	\draw    (299.6,121.07) -- (263.8,137.25) ;
	\draw [shift={(261.07,138.48)}, rotate = 335.68] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (8.93,-4.29) -- (0,0) -- (8.93,4.29) -- cycle    ;
	%Straight Lines [id:da7674435216299973] 
	\draw    (354.93,333.73) -- (432.27,319.07) ;
	%Curve Lines [id:da6299814455818392] 
	\draw    (277.6,91.07) .. controls (238.93,153.73) and (274.93,179.73) .. (322.27,203.73) .. controls (369.6,227.73) and (424.27,267.07) .. (432.27,319.07) ;
	%Straight Lines [id:da357006401572867] 
	\draw    (209.6,68.4) -- (277.6,91.07) ;
	%Curve Lines [id:da08014732331041308] 
	\draw  [dash pattern={on 4.5pt off 4.5pt}]  (264.27,87.07) .. controls (225.6,149.73) and (260.27,187.07) .. (309.6,212.4) .. controls (358.93,237.73) and (410.27,269.73) .. (418.27,321.73) ;
	%Curve Lines [id:da8971212017686895] 
	\draw  [dash pattern={on 4.5pt off 4.5pt}]  (221.6,73.07) .. controls (151.6,161.73) and (245.6,240.4) .. (296.27,258.4) .. controls (346.93,276.4) and (357.6,305.73) .. (369.6,331.73) ;
	%Curve Lines [id:da47999458459142474] 
	\draw  [dash pattern={on 3.75pt off 3pt on 1.5pt off 3pt}]  (243.6,79.73) .. controls (182.93,173.07) and (272.93,223.73) .. (321.6,246.4) .. controls (370.27,269.07) and (384.27,297.07) .. (393.6,326.4) ;

	% Text Node
	\draw (285.67,100.73) node [anchor=north west][inner sep=0.75pt]    {$\beta =0$};
	% Text Node
	\draw (321.5,168.4) node [anchor=north west][inner sep=0.75pt]    {$\beta =0.1$};

\end{tikzpicture}

### 2025-06-04 closed_loop_planning.tex
[mg] general reviews

@@
 
-The approach is based on the three following steps: (i) A nominal time-optimal trajectory is planned: this will be referred to as nominal feed-forward optimal trajectory.
+The approach is based on the three following steps. 
+\begin{enumerate}[label=\roman*)]
+\item A nominal time-optimal trajectory is planned: this will be referred to as nominal feed-forward optimal trajectory.
 This represents the trajectory that an expert driver would follow in an ideal scenario without uncertainty.
@@
 
-(ii) A closed-loop controller mimicking the driver's action is computed which stabilizes the nominal feed-forward optimal trajectory computed at step (i).
+\item A closed-loop controller mimicking the driver's action is computed to stabilize the nominal feed-forward optimal trajectory obtained at step (i).
 We assume a discrete time-varying Linear Quadratic Regulator (LQR) controller is a good approximation of how an expert driver can track a prescribed trajectory.
 
-(iii) In the last step the stochastic framework is reintroduced and the time-optimal planning problem is robustified using the Lyapunov framework.
+\item In the last step the stochastic framework is reintroduced and the time-optimal planning problem is robustified using the Lyapunov framework.
 The main idea is to \emph{re-plan} a mean trajectory of a stochastic system such that, a time-varying LQR controller that aims at stabilizing it, is able to properly tame the propagation of the uncertainty.
@@
 As a matter of fact, they need to explicitly account for the changed requirements associated with stabilizing a trajectory different from the nominal one, which must now satisfy the robustified constraints.
+\end{enumerate}
+
 \subsection{Step 1 -- Nominal feed-forward time-optimal trajectory planning}
 \label{sec:nominalFF}
-Step (i) consists in planning a nominal feed-forward trajectory.
+Step (i) consists of planning a nominal feed-forward trajectory.
 This assumes deterministic system dynamics, where $\hbmu_k$, $\hbu_k$, and $\hbz_k$ denote differential states, controls, and algebraic states, respectively, for \( k = 0, \ldots, N \).
@@
 As the third and last step, a closed-loop robustified planning problem is set up.
-Its aim is the definition of a mean trajectory of a stochastic system such that, working in tandem with state-feedback controller that aims at stabilizing it, is able to mitigate the perturbations and reduce the propagation of uncertainty.
+Its aim is the definition of a mean trajectory of a stochastic system which, working in tandem with a state-feedback controller that aims at stabilizing it, is able to mitigate the perturbations and reduce the propagation of uncertainty.
 This helps to guarantee that the robustified path constraints can be satisfied over the entire planning horizon.
@@
 To correctly enforce the mutual exclusivity within the LQR synthesis via Riccati recursion, the state-feedback gain must be computed with respect to the minimal control input.
-Specifically, the control input matrix $\tbB$ is employed and the gain $\tilde{\bK} \in \mathbb{R}^{2 \times 6}$ is sought for which operates on the independent variation \(\delta\tilde{\bu}\).
+Specifically, the control input matrix $\tbB$ is employed, and the gain $\tilde{\bK} \in \mathbb{R}^{2 \times 6}$, which operates on the independent variation \(\delta\tilde{\bu}\), is to be determined.
 This is structured as

### 2025-06-04 intro.tex
[mg] general reviews

@@
 One of the earliest contributions proposing a probabilistic representation of uncertainty is the chance-constrained framework in~\cite{Blackmore:ChanceConstrainedOptimalPath:2011}, which plans over the predicted distribution of the system state to ensure that the probability of constraint violation remains below a specified threshold.
-Within the Model Predictive Control paradigm, recent works have incorporated probabilistic safety guarantees.
+
+Within the Model Predictive Control (MPC) paradigm, recent works have incorporated probabilistic safety guarantees.
 Notably, the methods presented in~\cite{Gao:CollisionfreeMotionPlanning:2023},~\cite{Zhang:RobustifiedTimeoptimalPointtopoint:2025}, and~\cite{Zhang:RobustifiedTimeoptimalCollisionfree:2024} introduce stochastic MPC frameworks for autonomous mobile platforms, where process noise is explicitly modeled and closed-loop tracking performance is maintained via either pre-computed or optimized feedback gains.
@@
 
-Following an alternative approach, the largest Lyapunov exponent (LLE) has been proposed as an indicator of local stability~\cite{Meng:AnalysisGlobalCharacteristics:2022}, since it quantifies the average exponential rate at which infinitesimal perturbations off a fiducial trajectory grow or decay -- thus providing a direct measure of how deviations propagate through the system (negative values denote convergence toward the reference trajectory, while positive values indicate divergence)~\cite{Strogatz:NonlinearDynamicsChaos:2019}.
+Following an alternative approach, the largest Lyapunov exponent (LLE) has been proposed as an indicator of local stability~\cite{Meng:AnalysisGlobalCharacteristics:2022}.
+The LLE quantifies the average exponential rate at which infinitesimal perturbations off a fiducial trajectory grow or decay---thus providing a direct measure of how deviations propagate through the system.
+\textcolor{red}{(negative values denote convergence toward the reference trajectory, while positive values indicate divergence)~\cite{Strogatz:NonlinearDynamicsChaos:2019}. 
 %Except in chaotic systems where a positive LLE may still correspond to bounded motion on a strange attractor - a negative LLE typically signals local stability. 
 Notably, for trajectories converging to a stable equilibrium, the (negative) LLE asymptotically approaches the real part of the largest eigenvalues of the linearized system at that point.
-This result has been leveraged in the literature to infer whether a system, not yet at equilibrium, is ultimately converging toward one, thereby enabling a localized assessment of stability based on the transient behavior~\cite{Sadri:StabilityAnalysisNonlinear:2013}.
-Applications of this method are particularly relevant to study chaotic vessel motions~\cite{McCue:UseLyapunovExponents:2011} and rotorcraft dynamics~\cite{Tamer:StabilityNonlinearTimeDependent:2016,Cassoni:RotorcraftStabilityAnalysis:2024}, where the primary concern is understanding the long-term evolution of the system and determining its asymptotic behavior.
-
+This result has been leveraged in the literature to infer whether a system, not yet at equilibrium, is ultimately converging toward one, thereby enabling a localized assessment of stability based on the transient behavior~\cite{Sadri:StabilityAnalysisNonlinear:2013}.} Applications of this method are particularly relevant to study chaotic vessel motions~\cite{McCue:UseLyapunovExponents:2011} and rotorcraft dynamics~\cite{Tamer:StabilityNonlinearTimeDependent:2016,Cassoni:RotorcraftStabilityAnalysis:2024}, where the primary concern is understanding the long-term evolution of the system and determining its asymptotic behavior.
 %However, while Lyapunov-based indicators have been applied in vehicular contexts as well~\cite{Sadri:StabilityAnalysisNonlinear:2013,Meng:AnalysisGlobalCharacteristics:2022}, their relevance here is arguably of reduced interest - not due to their binary nature, but because even identifying asymptotic stability regions offers limited practical insight during planning.
@@
 We tackle the challenge of embedding robustness against disturbances directly into minimum-lap-time planning and propose two complementary methods.
-In the first one -- the open-loop worst-case covariance propagation -- at each discretization point, we propagate the state covariance forward over a fixed horizon and tighten all path constraints against the maximum covariance growth.
-By back-offing constraints using this \emph{worst-case} covariance, the resulting trajectory maximizes robustness without relying on feedback during planning.
+In the first one---the open-loop \emph{horizon-based} covariance propagation---at each discretization point, we propagate the state covariance forward over a fixed horizon and tighten all path constraints against the maximum covariance growth.
+By back-offing constraints using this worst-case covariance, the resulting trajectory maximizes robustness without relying on feedback during planning.
 
@@
 Second, an LQR controller is designed to stabilize that nominal path.
-Finally, we re-solve the planning problem - together with the feedback gains - so that, along this new nominal trajectory, the closed-loop covariance propagation satisfies tightened constraints via a Lyapunov-based formulation.
+Finally, we re-solve the planning problem - together with the feedback gains - so that, along this new robust trajectory, the closed-loop covariance propagation satisfies tightened constraints via a Lyapunov-based formulation.
 These two approaches offer a balance between computational simplicity and realistic, feedback-informed robustness, laying the groundwork for safe, high-performance lap-time planning under uncertainty in the motorsport context.
@@
 Here, we also derive the probabilistic constraint back-off formulation.
- Section~\ref{sec:open_loop_planning} then outlines the open-loop worst-case covariance-propagation planning approach, while Section~\ref{sec:closed_loop_planning} presents the closed-loop, covariance-aware planner.
-In Section~\ref{sec:performance_comparison} we compare both methods on a representative track scenario, reporting adhesion and collision-avoidance results, parameter-sensitivity studies, and performance trade-offs.
+Section~\ref{sec:open_loop_planning} then outlines the open-loop horizon-based covariance-propagation planning approach, while Section~\ref{sec:closed_loop_planning} presents the closed-loop, covariance-aware planner.
+In Section~\ref{sec:performance_comparison} we compare both methods on a representative track scenario, reporting parameter-sensitivity studies, analyzing the influence of different sets of constraints, and performance trade-offs.
 Finally, Section~\ref{sec:conclusions} draws conclusions and offers directions for future work.

### 2025-06-04 open_loop_planning.tex
[mg] general reviews

@@
 \label{sec:open_loop_planning}
-In this section, we address the problem of planning under uncertainty for a stochastic system, where the covariance matrix is propagated in a feed-forward fashion for a fixed prediction horizon of $H$ steps at each grid point $k$ of the discretized trajectory.
+In this section, we address the problem of planning under uncertainty for a stochastic system. 
+At each grid point $k$ of the discretized trajectory the covariance matrix is propagated in a feed-forward fashion for a fixed prediction horizon of $H$ steps.
 In the planning problem, we conservatively evaluate the maximum propagation of the covariance, that is, the matrix evolved over $H$ steps, to ensure the tightest possible enforcement of constraints.
-This maximizes the robustness of the resulting solution, as the constraint backoff is evaluated in the worst-case scenario.
+This maximizes the robustness of the resulting solution, as the constraint back-off is evaluated in the worst-case scenario within the prediction horizon.
 
@@
 The repeated initialization of the covariance matrix at each step is necessary to prevent its unbounded growth, as we assume that the system evolves without feedback control.
-In the absence of feedback, the uncertainty on the position and orientation of the body frame accumulates along the track, which may severely and unrealistically degrade the conditioning of the covariance matrix.
-For this reason, the propagation horizon in this open-loop setting is a critical design parameter that must be bounded to realistically capture the evolution of disturbances occurring prior to any corrective driver response.
+In the absence of feedback, the uncertainty on the position and orientation of the body frame accumulates along the track, leading to a severe and unrealistic degradation of the covariance matrix conditioning.
+For this reason, the propagation horizon in this open-loop setting is a critical design parameter.
+We bound the number of steps $H$ to realistically capture the evolution of disturbances occurring prior to any corrective driver response.
 

### 2025-06-04 stochastic_vehicle_dynamics.tex
[mg] general reviews

@@
 The vehicle model employed in this study is a Single Track model featuring nonlinear axle characteristics.
-The lateral forces $Y_1$ and $Y_2$ are computed using a Pacejka's Magic Formula, which depends on the total vertical load acting on the axle and the theoretical lateral slip evaluated at the midpoint of the axle.
+A schematic representation of the Single Track model is shown in Figure~\ref{fig:vehicle_model}.
+The lateral forces $Y_1$ and $Y_2$ are computed using a Pacejka's Magic Formula, which depends on the total vertical load acting on the axle and the axle's apparent slip angle.
 The longitudinal forces $X_1$ and $X_2$ are provided as inputs to the system.
@@
 
-The state vector $\bx$ has six components and includes the three kinematic quantities $u$, $v$, and $r$, along with the position and orientation of the barycentrical reference frame.
+The state vector $\bx$ has six components and includes the three kinematic quantities $u$, $v$, and $r$, along with the position $\left(x_G, y_G\right)$ and orientation $\psi$ of the barycentrical reference frame.
 As illustrated in Figure~\ref{fig:vehicle_model} $u$ and $v$ represent the longitudinal and lateral velocity of the CoM w.r.t its reference frame, while $r$ represents vehicle's yaw rate.
@@
 where $\bu(t)$ are the deterministic control inputs, $\bw(t)$ is additive Gaussian white noise with zero mean and known covariance $\bQ(t)$, i.e.
-$\bw(t) \sim \calN(\bzero, \bQ(t))$, and the state $\bx(t)$ results in a Gaussian distribution with mean $\bmu(t)$ and covariance $\bP(t)$, so that $\bx(t) \sim \calN(\bmu(t), \bP(t))$.
+$\bw(t) \sim \calN(\bzero, \bQ(t))$, and $\bx(t)$ is the state vector.
+Assuming a \emph{first-order} approximation for the disturbance propagation rule, $\bx(t)$ results in a Gaussian distribution with mean $\bmu(t)$ and covariance $\bP(t)$, so that $\bx(t)~\sim~\calN(\bmu(t), \bP(t))$.
 
@@
 \end{align}
-In~\eqref{eq:dP}, $\bP_0$ is the \emph{initial} state covariance and $\bA(t)$ is the usual shorthand notation for the Jacobian along the mean trajectory $\bmu(t)$, that is $\bA(t)=\bJ(\bmu(t),\bu(t))$, where $\bJ(\bx, \bu) = \frac{\pd f(\bx, \bu)}{\pd \bx}$.
+In~\eqref{eq:dP}, $\bP_0$ is the \emph{initial} state covariance and $\bA(t)$ is the usual shorthand notation for the Jacobian along the mean trajectory $\bmu(t)$, that is $\bA(t)=\frac{\pd f(\bmu, \bu)}{\pd \bmu}$.
+%\bJ(\bmu(t),\bu(t))$, where $\bJ(\bx, \bu) = \frac{\pd f(\bx, \bu)}{\pd \bx}$.
 
@@
 This matrix encodes the evolution from $t_0$ to $t$ of a perturbation w.r.t.
-to a fiducial trajectory -- the mean $\bmu(t)$.
+to the mean trajectory $\bmu(t)$.
 In symbols, $\barbx(t) = \bPhi(t,t_0) \barbx(t_0)$, with $\barbx(\cdot) =\bx(\cdot) - \bmu(\cdot) $.
@@
 $\be_i = \gamma \sig_i$.
-The coefficient $\gamma = \Phi^{-1}(p)$ is the quantile function, where $\Phi(z)=\Prob \{Z\leq z\}$ is the cdf of a standard normal distribution $Z \sim \calN(0,1)$, and acts as a \emph{tuning knob}: the greater the confidence level $p$ required, the higher the gain $\ga$\footnote{For example, with $p=0.84$ $\ga = 1.0$, with $p=0.97$ $\ga = 2.0$, with $p=0.99$ $\ga = 3.0$}.
+The coefficient $\gamma = \Phi^{-1}(p)$ is the quantile function, where $\Phi(z)=\Prob \{Z\leq z\}$ is the cumulative distribution function (cdf) of a standard normal distribution $Z \sim \calN(0,1)$, and acts as a \emph{tuning knob}: the greater the confidence level $p$ required, the higher the gain $\ga$\footnote{For example, with $p=0.84$ $\ga = 1.0$, with $p=0.97$ $\ga = 2.0$, with $p=0.99$ $\ga = 3.0$}.
 The term $\sig_i = \big[\na^T_{\bx} h_i(\bmu) \bP(t) \na_{\bx} h_i(\bmu)\big]^{\frac{1}{2}}$ represents the standard deviation of the constraint $h_i(\bx)$ linearized around the mean, i.e.
@@
 \label{sec:discretization}
-
 The nonlinear optimal control problem~\eqref{eq:OCP} can be discretized by applying a suitable collocation integrator obtaining the following nonlinear program (NLP)
@@
 The $\bz_k$'s are introduced as direct handles for physically meaningful quantities such contact forces. Similarly, $\bP_k$ represents the covariance matrix at the $k$-th node.
-Following the direct collocation approach, both the mean and covariance state trajectories are approximated, in the $k$-th interval, with polynomials $\pi_k(\tau)$, defined on the unit interval $\tau\in[0,1]$, and then scaled to match the width $h_k$ of the corresponding time step.
+Following the direct collocation approach, both the mean and covariance state trajectories are approximated, in the $k$-th interval, with polynomials $\pi_k(\tau)$, defined on the unit interval $\tau\in[0,1]$, and then scaled to match the width $\nu_k$ of the corresponding time step.
 On the unit interval, we select $d$ collocation points $\tau_1, \ldots, \tau_d$, associated with as many collocation states.
@@
 
-To better clarify the effect of the back-off term on the constraint we can suppose to have a combination of $\bx$, $\bu$, $P$ and $\ga^\textrm{FLC}_j$ so that $\be^\textrm{FLC}_j=0.2$.
+To better clarify the effect of the back-off term on the constraint let us consider a combination of $\bx$, $\bu$, $P$ and $\ga^\textrm{FLC}_j$ so that $\be^\textrm{FLC}_j=0.2$.
 This implies that the semi-axes of the ellipse are reduced by a factor $\sqrt{1-\be^\textrm{FLC}_j} = \sqrt{0.8}$, as shown in the left panel Figure~\ref{fig:robust_constraints}, where the ellipse in dashed line delimits the available region with the back-off applied.

### 2025-06-05 performance_comparison.tex
[MGa] subsec 5.1

@@
 	\includegraphics{Fig/track.pdf}
-	\caption{Sector of the Catalunya circuit considered in the analysis, corresponding to the curvilinear abscissa interval $\left[0.70, 0.77\right]$.
-Checkpoints are indicated by labels and are uniformly spaced by $\De\al=0.01$.
-This sector includes two distinct corners: a low-speed turn from $\al = 0.72$ to $\al = 0.73$, and a high-speed turn from $\al = 0.75$ to $\al = 0.76$, allowing for the evaluation of vehicle behavior across different dynamic scenarios.}
+	\caption{The analyzed Catalunya circuit sector spans $\alpha \in [0.70, 0.77]$, with labeled checkpoints every $\Delta\alpha = 0.01$.
+It contains two corners -- a low-speed turn ($\alpha = 0.72-0.73$) and a high-speed turn ($\alpha = 0.75-0.76$) -- providing contrasting dynamic conditions for vehicle behavior evaluation.}
 	\label{fig:track}
@@
 
-\subsection{Open-loop parameters sensitivity}
+\subsection{Open-loop parameter sensitivity}
 \label{sec:ol_param_sensitivity}
@@
 Observing the left panel of Figure~\ref{fig:ol_sensitivities}, it is evident that the trajectories corresponding to configurations with lower values of $\ga^\textrm{TLC}$ tend to travel closer to the track boundaries.
-In contrast, the trajectory associated with $\ga^\textrm{TLC} = 3$---represented by the light green line---remains significantly farther from the edges.
+In contrast, the trajectory associated with $\ga^\textrm{TLC} = 3$\,--- represented by the light green line \,--- remains significantly farther from the edges.
 This behavior is consistent with the increased conservativeness introduced by higher values of $\ga^\textrm{TLC}$, which amplify the back-off term $\be^\textrm{TLC}$ and thereby enforce a larger safety margin from the track limits.
@@
 	\includegraphics{Fig/Pzero_sensitivity.pdf}
-	\caption{Effect of $\ga^\textrm{TLC}$ (left panel) and $P_0$ (right panel) on optimal trajectory.
+\caption{Effect of the TLC penalty coefficient $\gamma^{\mathrm{TLC}}$ (left panel) and the initial covariance $\mathbf P_{0}$ (right panel) on the optimal trajectory.
+Raising $\gamma^{\mathrm{TLC}}$ from 0 to 3 increases the probability of satisfying the track-limit constraints from 50 \% to 99 \%.
+The right panel compares the nominal covariance $\bar{\mathbf P}_{0}$ (black) with $4\bar{\mathbf P}_{0}$ (blue line) and $\tfrac14\,\bar{\mathbf P}_{0}$ (red line), i.e., twice and half the nominal state uncertainty.
+}
+%\caption{Effect of $\ga^\textrm{TLC}$ (left panel) and $\bP_0$ (right panel) on optimal trajectory.
 The parameter $\ga^\textrm{TLC}$ spans in the range $\left[0,3\right]$ corresponding to a probability to meet the track limit constraint ranging from 50\% to 99\%.
@@
 
-The second parameter examined is the initial value of the covariance matrix, $\bP_0$. 
+The second parameter examined is the initial value of the covariance matrix $\bP_0$.
 We recall that at each grid point a new prediction horizon is initialized, and the corresponding matrix $\bP$ at its starting point, i.e., $\bP^0_k$, is set to $\bP_0$.
-The diagonal elements of $\bP$ are the variances $\text{var}(x_i)$ of each state variables, while the off-diagonal elements encode their covariances $\text{cov}(x_i,x_j)$.
+The diagonal elements of $\bP$ are the variances $\Var{x_i}$  of each state variables, while the off-diagonal elements encode their covariances $\Cov{x_i}{x_j}$.
 Larger diagonal values indicate higher uncertainty in the corresponding states, and nonzero off-diagonal terms imply mutual dependence between them.
-For simplicity, a diagonal $\bP_0$ is used in this study, that is, for each prediction horizon we assign an initial standard deviation $\sigma_i = \sqrt{\text{var}(x_i)}$ to each state and assume all initial covariances to be zero.
+For simplicity, a diagonal $\bP_0$ is used in this study, that is, for each prediction horizon we assign an initial standard deviation $\sig_{x_i}$ to each state and assume all initial covariances $\Cov{x_i}{x_j}$  to be zero.
 
-To provide a concrete interpretation, consider the first state variable $x_1$ with mean $\mu_1$ and standard deviation $\sigma_1$.
-Under the Gaussian assumption, the true value of $x_1$ lies within the interval $\left[\mu_1 - \sigma_1,\ \mu_1 + \sigma_1\right]$ with approximately 68\% probability, and within the interval $\left[\mu_1 - 3\sigma_1,\ \mu_1 + 3\sigma_1\right]$ with approximately 99\% probability.
+ Assuming the state vector is ordered as described in Section~\ref{sec:vehicle_model}, the baseline initial covariance matrix is set to $\bP_0=\diag(\bbsig^2)$, with
+\begin{align}
+\bar{\boldsymbol{\sigma}}~=~\left[0.1\,\textrm{m/s}, 0.01\,\textrm{m/s}, 0.01\,\textrm{rad/s}, 1\,\textrm{m}, 1\,\textrm{m}, 0.0175\,\textrm{rad}\right]^T.
+\end{align}
+These standard deviations have been chosen based on the typical range of variation of each state.
+To provide a concrete interpretation at the starting condition, the assumption on first state $u \sim \calN(\mu_u,\sig_u^2)$, with $\sig_u=0.1$ m/s, implies that
+$\Pr\{\mu_u-\sig_u \leq u \leq \mu_u+\sig_u\}\approx 68\%$ and $\Pr\{\mu_u-3\sig_u \leq u \leq \mu_u+3\sig_u\}\approx 99\%$.
 
-We assign a baseline initial covariance matrix $\bP_0 = \bar{\bP}_0$ by selecting standard deviations $\bar{\sigma}_i$ for each state variable.
-These standard deviations have been chosen based on the typical range of variation of each state.
-Assuming the state vector is ordered as described in Section~\ref{sec:vehicle_model}, the values used are $\bar{\boldsymbol{\sigma}}~=~\left[0.1\,\textrm{m/s}, 0.01\,\textrm{m/s}, 0.01\,\textrm{rad/s}, 1\,\textrm{m}, 1\,\textrm{m}, 0.0175\,\textrm{rad}\right]^T$.
 In addition to the baseline value $\bar{\bP}_0$, two alternative values are considered: $\bP_0 = 4\bar{\bP}_0$ and $\bP_0 = \frac{1}{4}\bar{\bP}_0$.
-These correspond, respectively, to doubling and halving the initial standard deviations $\sigma_i$ associated with the state vector.
+These correspond, respectively, to doubling and halving the initial standard deviations $\sig_{x_i}$ associated with each state vector.
 

### 2025-06-05 conclusions.tex
[mg] conclusions

@@
 
-In this work we investigated the potential of some new technologies in empowering next generation autonomous race vehicles.
-In particular, the upgrades considered here were: four independent steering wheels (All Steering - AS), four independent wheel torques (All Independent Wheel Drive - AIWD), and independent control of the four camber angles (Dynamic Camber Control - DCC).
+This work has proposed and compared two robust planning strategies for minimum-time trajectory optimization in motorsport scenarios under uncertainty.
+The first approach, based on open-loop horizon-based covariance propagation, relies on a worst-case analysis of disturbance growth over a finite horizon to conservatively back off safety constraints.
+The second, a closed-loop covariance-aware method, incorporates a time-varying LQR feedback policy into the planning process, yielding a more realistic estimate of uncertainty evolution and tighter constraint enforcement.
 
-Firstly, we discussed the impact of enabling these control inputs on the vehicle performance, in relation to the mass increase associated to the new components needed for the systems.
-The results confirmed that the vehicle endowed with all the features (All Independent Wheel Drive, All Steering, Dynamic Camber Control - AIWD AS DCC) is the most performant and forgiving to added mass.
-Compared to a traditional vehicle (Rear Wheel Drive, Forward Steering - RWD FS), its advantage in lap time persists until 123.5\,kg of added mass.
+Quantitative results on a representative circuit segment have highlighted the trade-offs between robustness and performance.
+Open-loop planning, while conservative, ensures feasibility even in the absence of feedback, but incurs a higher performance cost due to larger safety margins.
+The closed-loop method achieves comparable safety guarantees with reduced conservativeness and improved lap-time performance, thanks to its ability to explicitly model the stabilizing action of the driver.
+Finally, random simulations with noise realization performed adopting the closed-loop approach proved the effectiveness of the robustified constraints.
+A LQR-based controller tracking a robust trajectory planned with probabilistic constraint tightening exhibits significantly fewer track violations compared to one tracking the nominal optimal trajectory.
 
-Then, lap times are presented for the baseline (Rear Wheel Drive, Forward Steering - RWD FS), and for vehicles with single upgrades added, keeping constant the total mass.
-In particular, the configuration analyzed are: Rear Wheel Drive, All Steering - RWD AS, All Independent Wheel Drive, Forward Steering - AIWD FS, and Rear Wheel Drive, Forward Steering, Dynamic Camber Control - RWD FS DCC.
-From these results it is evident that the single feature that improves the performance the most is the AIWD FS in Siena circuit, and the RWD FS DCC in N\"{u}rburgring circuit.
-
-Focusing on trajectories, we can highlight the following features.
-The All Independent Wheel Drive, Forward Steering (AIWD FS) vehicle, being able to perform torque vectoring, travels turns closer to the inner boundary of the track, compared to a traditional vehicle, and exits with an higher forward speed.
-
-A Rear Wheel Drive, All Steering (RWD AS) vehicle experiences in-phase rear steering for high speed turns, and counter-phase rear steering for low speed turns, hence improving cornering in each condition.
-
-Being able to control the camber dynamically, a Rear Wheel Drive Vehicle, Forward Steering, Dynamic Camber Control (RWD FS DCC) vehicle, positions each wheel w.r.t.
-the road to raise the force limits, thus maximizing the ground forces.
-
-To highlight the impact of the torque vectoring and camber control, a vehicle with these characteristics but \emph{without steering abilities} (All Independent Wheel Drive, Zero Steering, Dynamic Camber Control - AIWD 0S DCC) has been analyzed.
-This vehicle completes a lap in Siena 3.66 seconds faster than a traditional vehicle (RWD FS), and corners effectively by applying higher torques on outer wheels than the inner wheels.
-
-Introducing a temperature-dependent tyre model, we observed that the least affected vehicle is the traditional RWD FS, which stresses the wheels the least and thus loses only 11 ms due to the drop in grip factor.
-In contrast, All Independent Wheel Drive (AIWD) vehicles tend to stress all four wheels much more.
-As a result, when the tire thermal model in the MLTP takes into account the reduction in grip with increasing temperature, AIWD vehicles are penalized more than RWD FS vehicles, losing approximately 100 ms each in the considered sector.
-
-\section{Future Work}
-
-While this study provides a broad overview of the possibilities and implications of introducing certain features to a new generation of autonomous vehicles, some aspects require further investigation.
-
-A more complete tire model, which could account for tread wear and the consequent inevitable performance drop due to tire usage, should be implemented.
-To achieve this, either a testing campaign on the actual modeled tires to fine-tune the parameters of the wear-dependent phenomena, or a new dataset of fully characterized tires, is essential.
-
-The introduction of realistic aerodynamic maps could indeed lead to a refinement of the model, enhancing the accuracy of the vehicle dynamics representation.
-
-Modeling a non-developable 3D track that incorporates bumps and holes would represent a significant advancement in fully understanding the potential and limitations of these features.
-This approach would enhance the physical realism of the road-vehicle interaction, providing deeper insights into its dynamics.
-
-Additionally, future developments could include an evaluation of design solutions for the various subsystems that control the additional degrees of freedom.
-These solutions must, however, take into account their practical feasibility for real-world systems, considering current technological limitations in sensors and actuators, as well as the challenges posed by their integration.
+Overall, the proposed framework enables the design of minimum-time trajectories that are not only fast but also certified to be safe under realistic disturbances.
+By embedding robustness into the planning stage, the gap between theoretical optimality and real-world executability is narrowed.
+Future work will explore the generalization to more accurate vehicle models and experimental validation of the methodology.

### 2025-06-05 performance_comparison.tex
[MGa] first pass till the end

@@
 %It is of interest to analyze how a robustified constraint---or a combination of the two described in Sections~\ref{sec:FLC} and~\ref{sec:TLC}---influences the driving style.
-We now investigate how the inclusion of different kinds of robustified constraints---individually or in combination, as described in Sections~\ref{sec:FLC} and~\ref{sec:TLC}---influences the driving style.
+We now investigate how the inclusion of different kinds of robustified constraints --- individually or in combination, as described in Sections~\ref{sec:FLC} and~\ref{sec:TLC} --- influences the driving style.
 We compare the nominal feed-forward optimal trajectory, obtained without robustified constraints, with those resulting from a robustified track limit constraint (denoted as TLC), a robustified friction limit constraint (denoted as FLC), and a scenario where both constraints are robustified (denoted as TLC+FLC).
@@
 	\includegraphics{Fig/ol_telemetries.pdf}
-	\caption{Evolution of the longitudinal speed $\De u$, wheel steering angle $\De \de$, longitudinal force $\De F_x$, and lateral deviation $\De e$ are reported for three robustification scenarios: TLC (Track Limit Constraints), FLC (Friction Limit Constraint), and TLC+FLC (both constraints enforced concurrently).
+\caption{Variations of longitudinal speed $\Delta u$, steering angle $\Delta\delta$, longitudinal force $\Delta F_x$, and lateral deviation $\Delta e$ relative to the nominal trajectory for three schemes --- TLC (track-limit), FLC (friction-limit), and TLC+FLC (both).
+Dashed lines in the last three panels mark the sign of the reference signal, indicating whether a positive variation increases or decreases the absolute value (crucial for the turn-dependent variables $\delta$ and $e$ and for distinguishing traction from braking in $F_x$).}
+
+%\caption{Evolution of the longitudinal speed $\De u$, wheel steering angle $\De \de$, longitudinal force $\De F_x$, and lateral deviation $\De e$ are reported for three robustification scenarios: TLC (Track Limit Constraints), FLC (Friction Limit Constraint), and TLC+FLC (both constraints enforced concurrently).
 Each profile is shown relative to the baseline (nominal) trajectory to highlight the effect of each constraint-handling strategy.
@@
 Except for the $\De u$ panel, each plot includes dashed black lines that indicate the sign of the nominal signal.
-These indicators are shown at three distinct vertical levels---upper, central, and lower---corresponding respectively to positive, near-zero, and negative values of the nominal signal.
+These indicators are shown at three distinct vertical levels --- upper, central, and lower --- corresponding respectively to positive, near-zero, and negative values of the nominal signal.
 
@@
 
-This shift in trajectory and longitudinal speed becomes especially evident when approaching the sharp turn at $\al \in [0.707, 0.712]$.
+This shift in trajectory and longitudinal speed becomes especially evident when approaching the sharp turn for $\al \in [0.707, 0.712]$.
 A noticeable increase in longitudinal speed is observed around $\alpha = 0.71$ for the TLC configuration  in the $\De u$ panel.
@@
 This behavior is consistent with the variation of the total longitudinal force $\Delta F_x$ shown in the third panel.
-In particular, the TLC and TLC+FLC configurations exhibit a higher longitudinal force while approaching the sharp turn, in the curvilinear abscissa interval $\al\in\left[0.707, 0.712\right]$, resulting in a higher longitudinal speed.
+\textcolor{red}{In particular, the TLC and TLC+FLC configurations exhibit a higher longitudinal force while approaching the sharp turn, in the curvilinear abscissa interval $\al\in\left[0.707, 0.712\right]$, resulting in a higher longitudinal speed.
 Subsequently, these two configurations apply a lower longitudinal force during braking compared to the nominal case, indicating more intense braking.
-This behavior explains the sharp reduction in longitudinal speed observed thereafter.
+This behavior explains the sharp reduction in longitudinal speed observed thereafter.}
 
@@
 Later in the sector, a further deviation in driving style appears during the high-speed turn, specifically around $\alpha \in [0.75, 0.76]$.
-As shown in the $\Delta F_x$ panel, the combined configuration TLC+FLC needs to reduce the longitudinal force within this section to meet both the constraints.
+As shown in the third panel ($\Delta F_x$), the combined configuration TLC+FLC needs to reduce the longitudinal force within this sector to meet both the constraints.
 This adaptation is necessary because the TLC forces the vehicle to follow a wider trajectory, which entails greater lateral acceleration and, consequently, higher lateral force demand.
@@
 %---Axle saturation
-After analyzing the variations of these four key signals with respect to the nominal solution, we now turn our attention to the evaluation of axle saturation, which provides insight into how close each configuration operates to the tire grip limits.
-In accordance with the FLC formulation introduced in~\eqref{eq:adherence_constraint}, we define the \emph{axle saturation ratio} $S_j$ as
-\begin{equation}
-	S_j = \frac{ \left( \frac{X_j(\bx,\bu)}{\mu_{x,j}} \right)^2+ \left( \frac{Y_j(\bx,\bu)}{\mu_{y,j}}\right)^2}{Z_j^2(\bx,\bu)}.
-\end{equation}
-This quantity indicates how close each configuration operates to the friction limit; $S_j = 1$ when the point $\left(X_j, Y_j\right)$ lies exactly on the friction ellipse, i.e., under full saturation.
-Figure~\ref{fig:ol_saturation} shows the values of $S_j$ for the front axle ($j=1$, first panel) and the rear axle ($j=2$, rear panel) for the previously introduced configurations: the nominal case (black dashed line) and the three robustified cases (color-coded as before).
+After analyzing the variations of these four key signals with respect to the nominal solution, we now turn our attention to the evaluation of axle saturation $S_j(\bx,\bu)$ defined in~\eqref{eq:axle_saturation}, which provides insight into how close each configuration operates to the tire grip limits.
+
+Figure~\ref{fig:ol_saturation} shows the values of $S_j(\bx,\bu)$ for the front axle ($j=1$, first panel) and the rear axle ($j=2$, second panel) for the previously introduced configurations: the nominal case (black dashed line) and the three robustified cases TLC, FLC and TLC+FLC (color-coded as before).
 
@@
 	\includegraphics{Fig/ol_saturation.pdf}
-	\caption{Tire usage for front axle (first panel) and rear axle (second panel) over curvilinear abscissa $\al$.
-The plotted lines represent the quantity $\left[ \left( \frac{X_j(\bx,\bu)}{\mu_{x,j}} \right)^2+ \left( \frac{Y_j(\bx,\bu)}{\mu_{y,j}}\right)^2\right]\frac{1}{Z_j^2(\bx,\bu)}$, where $j=1$ denotes the front axle and $j=2$ the rear axle.
-The dashed black line corresponds to the reference (nominal) configuration, while the green, red, and blue lines represent the robust configurations TLC, FLC, and TLC+FLC, respectively.}
+	\caption{Axle saturation $S_j(\bx,\bu)$ for front axle (j=1, first panel) and rear axle (j=2, second panel) with respect to the curvilinear abscissa $\al$.
+The dashed black line corresponds to the reference (nominal) configuration, while the green, red, and blue lines represent the robustified configurations TLC, FLC, and TLC+FLC, respectively.}
 	\label{fig:ol_saturation}
@@
 
-First, we observe that the FLC and TLC+FLC configurations (red and blue lines) never reach full saturation, i.e., $S_j=1$, due to the presence of the back-off term, which enforces a safety margin from the friction limit.
-Second, we point out how the different trajectories followed by the TLC and TLC+FLC configurations (green and blue lines), compared to the nominal and FLC cases, entail distinct ground force demands, which are clearly visible here in terms of the $S_j$ ratio.
+Firstly, we observe that the FLC and TLC+FLC configurations (red and blue lines) never reach full saturation ($S_j=1)$ due to the presence of the back-off term, which enforces a safety margin from the friction limit.
+Secondly, we point out how the different trajectories followed by the TLC and TLC+FLC configurations (green and blue lines), compared to the nominal and FLC cases, entail distinct ground force demands, which are clearly visible here in terms of the $S_j$ ratio.
 This behavior emerges around the sharp turn and becomes even more evident in the high-speed turn.
@@
 The left panel of Figure~\ref{fig:traj_strings} shows random evolutions (red lines) for an LQR-controlled vehicle stabilized to follow the nominal optimal trajectory (black line).
-No robust constraints are introduced, and the controller implements the feedback policy from step 2 of Subsection~\ref{sec:LQR}.
-The right panel, instead, shows the random evolutions (red lines) for the robustified LQR-based controller obtained in the step 3 of the closed-loop approach described in Subsection~\ref{sec:CLrobustified}.
+No robust constraints are introduced, and the controller implements the feedback policy from Step 2 of Subsection~\ref{sec:LQR}.
+The right panel, instead, shows the random evolutions (red lines) for the robustified LQR-based controller obtained in the Step 3 of the closed-loop approach described in Subsection~\ref{sec:CLrobustified}.
 The corresponding robustified mean trajectory is shown in black.
@@
 	\caption{Simulations with random initial conditions and Gaussian noise realizations.
-In the left panel, the reference trajectory is the non-robust one, tracked using an LQR controller designed around it.
-In the right panel, the reference is a robust trajectory obtained using the closed-loop optimization method, and it is tracked using the corresponding optimized controller produced by the same method.}
+In the left panel, the reference trajectory (black curve) is the non-robust one, tracked using an LQR controller designed around it as per Step 2 of Subsection~\ref{sec:LQR}.
+In the right panel, the reference (black curve) is a robust trajectory obtained using the closed-loop optimization method as per Step 3 of Subsection~\ref{sec:CLrobustified}, which is tracked using the corresponding optimized controller.}
 	\label{fig:traj_strings}

### 2025-06-05 stochastic_vehicle_dynamics.tex
[MGa] first pass till the end

@@
 
-The base version of the constraint for each axle can be expressed as a function of $\bx$ and $\bu$, as follows:
+The base version of the constraint for each axle can be expressed as the following inequality depending on $\bx$ and $\bu$
 \begin{equation}
-	h^\textrm{FLC}_j(\bx,\bu) = \left[ \left( \frac{X_j(\bx,\bu)}{\mu_{x,j}} \right)^2+ \left( \frac{Y_j(\bx,\bu)}{\mu_{y,j}}\right)^2\right]\frac{1}{Z_j^2(\bx,\bu)}-1\leq0\qquad j=1,2.\label{eq:adherence_constraint}
+	h^\textrm{FLC}_j(\bx,\bu) =  S_j(\bx,\bu) - 1 \leq 0, \qquad{(j=1,2)}
+\label{eq:adherence_with_axle_saturation}
 \end{equation}
-where the superscript FLC denotes friction limit constraint, and the subscript $j$ indicates the axle index, with $j=1$ referring to the front axle and $j=2$ to the rear axle.
-Here, $X_j$ and $Y_j$ denote the longitudinal and lateral components of the in-plane ground forces, respectively, as illustrated in Figure~\ref{fig:vehicle_model}, while $Z_j$ represents the total vertical load acting on the axle.
-As can be seen from~\eqref{eq:adherence_constraint}, the points $\left(X_j,Y_j\right)$ are constrained to lie within an ellipse whose semi-axes are given by $\mu_{x,j}Z_j$ and $\mu_{y,j}Z_j$.
-The constraint, without the back-off term, is represented in the left panel of Figure~\ref{fig:robust_constraints} by the solid line ellipse in the plane $X_j-Y_j$.
+where we introduced the \emph{axle saturation ratio} $S_j(\bx,\bu)$ as follows
+\begin{equation}
+	S_j(\bx,\bu) = \frac{ \left( \frac{X_j(\bx,\bu)}{\mu_{x,j}} \right)^2+ \left( \frac{Y_j(\bx,\bu)}{\mu_{y,j}}\right)^2}{Z_j^2(\bx,\bu)}.
+\label{eq:axle_saturation}
+\end{equation}
+In~\eqref{eq:adherence_with_axle_saturation}, the superscript FLC denotes friction limit constraint, and $j=1,2$ refer the front and rear axle, respectively.
+In~\eqref{eq:axle_saturation}, $X_j$ and $Y_j$ denote the longitudinal and lateral components of the in-plane ground forces, respectively, as illustrated in Figure~\ref{fig:vehicle_model}, while $Z_j$ represents the total vertical load acting on the axle.
+
+From~\eqref{eq:axle_saturation} it is evident that $S_j(\bx,\bu)\in [0,1]$ indicates how close each configuration operates to the friction limit: $S_j=0$ when the overall grip demand is zero, while $S_j = 1$ when the point $\left(X_j, Y_j\right)$ lies exactly on the friction ellipse, i.e., under full saturation.
+The points $\left(X_j,Y_j\right)$ are constrained to lie within an ellipse whose semi-axes are given by $\mu_{x,j}Z_j$ and $\mu_{y,j}Z_j$.
+The constraint, without the back-off term, is represented in the left panel of Figure~\ref{fig:robust_constraints} by the solid line ellipse in the plane $X_jY_j$.
 

### 2025-06-05 intro.tex

@@
 
-As the automotive industry advances towards higher levels of automation, ...
+Minimum lap time optimization is a fundamental tool in the motorsport field, enabling the synthesis of optimal trajectories and control profiles that push performance to its limits.
+These optimal references are widely used both offline, for vehicle setup and strategy development, and online, as feedforward inputs to advanced driver-assistance and autonomous systems.
+However, despite their optimality under nominal assumptions, trajectories produced by state-of-the-art planners often lie critically close to physical and safety constraints---such as tire friction limits and collision boundaries---rendering them extremely sensitive to disturbances and modeling inaccuracies.
+Consequently, these ideal references may prove difficult or unsafe to follow, even for expert drivers, limiting their practical usability.
+
+This fragility highlights a critical gap: current minimum lap time formulations rarely embed robustness explicitly. 
+Consequently, resulting trajectories lack reliability under uncertainties arising from mismatches between modeling assumptions and real operating conditions. 
+Addressing this shortcoming is essential to bridge the gap between theoretical optimality and real-world feasibility in high-performance motorsport applications.
 

### 2025-06-06 intro.tex
[MM] New plot inserted

@@
 The LLE quantifies the average exponential rate at which infinitesimal perturbations off a fiducial trajectory grow or decay---thus providing a direct measure of how deviations propagate through the system.
-\textcolor{red}{(negative values denote convergence toward the reference trajectory, while positive values indicate divergence)~\cite{Strogatz:NonlinearDynamicsChaos:2019}. 
-%Except in chaotic systems where a positive LLE may still correspond to bounded motion on a strange attractor - a negative LLE typically signals local stability. 
-Notably, for trajectories converging to a stable equilibrium, the (negative) LLE asymptotically approaches the real part of the largest eigenvalues of the linearized system at that point.
-This result has been leveraged in the literature to infer whether a system, not yet at equilibrium, is ultimately converging toward one, thereby enabling a localized assessment of stability based on the transient behavior~\cite{Sadri:StabilityAnalysisNonlinear:2013}.} Applications of this method are particularly relevant to study chaotic vessel motions~\cite{McCue:UseLyapunovExponents:2011} and rotorcraft dynamics~\cite{Tamer:StabilityNonlinearTimeDependent:2016,Cassoni:RotorcraftStabilityAnalysis:2024}, where the primary concern is understanding the long-term evolution of the system and determining its asymptotic behavior.
+Applications of this method are particularly relevant to study chaotic vessel motions~\cite{McCue:UseLyapunovExponents:2011} and rotorcraft dynamics~\cite{Tamer:StabilityNonlinearTimeDependent:2016,Cassoni:RotorcraftStabilityAnalysis:2024}, where the primary concern is understanding the long-term evolution of the system and determining its asymptotic behavior.
 %However, while Lyapunov-based indicators have been applied in vehicular contexts as well~\cite{Sadri:StabilityAnalysisNonlinear:2013,Meng:AnalysisGlobalCharacteristics:2022}, their relevance here is arguably of reduced interest - not due to their binary nature, but because even identifying asymptotic stability regions offers limited practical insight during planning.

### 2025-06-06 open_loop_planning.tex
[MM] New plot inserted

@@
 	\input{Fig/Tikz_Sources/tikz_covariance_evolution.tex}
-	\caption{Cartoon representation (depicted for $H=3$) of the continuity and initialization equations associated with the $H+1$ propagated instances of the covariance matrix introduced in problem~\eqref{eq:DOCPopenloop}.
+	\caption{Schematic representation (depicted for $H=3$) of the continuity and initialization equations associated with the $H+1$ propagated instances of the covariance matrix introduced in problem~\eqref{eq:DOCPopenloop}.
 The dashed rectangle highlights the $k$-th discretization step, at which the robust constraints are enforced.

### 2025-06-06 performance_comparison.tex
[MM] New plot inserted

@@
 
-\subsection{Robustified constraints comparison}
-%It is of interest to analyze how a robustified constraint---or a combination of the two described in Sections~\ref{sec:FLC} and~\ref{sec:TLC}---influences the driving style.
-We now investigate how the inclusion of different kinds of robustified constraints --- individually or in combination, as described in Sections~\ref{sec:FLC} and~\ref{sec:TLC} --- influences the driving style.
-We compare the nominal feed-forward optimal trajectory, obtained without robustified constraints, with those resulting from a robustified track limit constraint (denoted as TLC), a robustified friction limit constraint (denoted as FLC), and a scenario where both constraints are robustified (denoted as TLC+FLC).
-All the optimization are obtained with the open-loop method described in Section~\ref{sec:open_loop_planning}, setting $H=4$, and $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$, corresponding to a 90\% probability of satisfying both constraints.
+\subsection{Open-Loop vs. Closed-Loop comparison}
 
-\begin{figure}
+\begin{figure}[t]
 	\centering
 	\includegraphics{Fig/ol_telemetries.pdf}
-\caption{Variations of longitudinal speed $\Delta u$, steering angle $\Delta\delta$, longitudinal force $\Delta F_x$, and lateral deviation $\Delta e$ relative to the nominal trajectory for three schemes --- TLC (track-limit), FLC (friction-limit), and TLC+FLC (both).
-Dashed lines in the last three panels mark the sign of the reference signal, indicating whether a positive variation increases or decreases the absolute value (crucial for the turn-dependent variables $\delta$ and $e$ and for distinguishing traction from braking in $F_x$).}
+	\caption{Open-Loop method.
+Variations of longitudinal speed $\Delta u$, steering angle $\Delta\delta$, longitudinal force $\Delta F_x$, and lateral deviation $\Delta e$ relative to the nominal trajectory for three schemes --- TLC (track-limit), FLC (friction-limit), and TLC+FLC (both).
+	Dashed lines mark the sign of the reference signal (positive, near-zero and negative).}
 
@@
 
+\begin{figure}[t]
+	\centering
+	\includegraphics[scale = 1]{Fig/ol_saturation.pdf}
+	\caption{Open-Loop.
+Axle saturation $S_j(\bx,\bu)$ for front axle (j=1, first panel) and rear axle (j=2, second panel) with respect to the curvilinear abscissa $\al$.
+Nominal configuration in dashed black, TLC in green, FLC in red, TLC+FLC in blue.
+	%The dashed black line corresponds to the reference (nominal) configuration, while the green, red, and blue lines represent the robustified configurations TLC, FLC, and TLC+FLC, respectively.
+	}
+	\label{fig:ol_saturation}
+\end{figure}
+
+%It is of interest to analyze how a robustified constraint---or a combination of the two described in Sections~\ref{sec:FLC} and~\ref{sec:TLC}---influences the driving style.
+We now investigate how the inclusion of different kinds of robustified constraints --- individually or in combination, as described in Sections~\ref{sec:FLC} and~\ref{sec:TLC} --- influences the driving style.
+We compare the nominal feed-forward optimal trajectory, obtained without robustified constraints, with those resulting from a robustified track limit constraint (denoted as TLC), a robustified friction limit constraint (denoted as FLC), and a scenario where both constraints are robustified (denoted as TLC+FLC).
+
+First we show results obtained with the open-loop method described in Section~\ref{sec:open_loop_planning}, setting $H=4$, and $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$, corresponding to a 90\% probability of satisfying both constraints.
+%All the optimization are obtained with the open-loop method described in Section~\ref{sec:open_loop_planning}, setting $H=4$, and $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$, corresponding to a 90\% probability of satisfying both constraints.
 %---Delta key signals
-
 The comparison between the nominal case and the three robustified cases is illustrated in Figure~\ref{fig:ol_telemetries}, in terms of the deviation of selected signals from to their nominal counterparts.
@@
 This behavior is consistent with the variation of the total longitudinal force $\Delta F_x$ shown in the third panel.
-\textcolor{red}{In particular, the TLC and TLC+FLC configurations exhibit a higher longitudinal force while approaching the sharp turn, in the curvilinear abscissa interval $\al\in\left[0.707, 0.712\right]$, resulting in a higher longitudinal speed.
-Subsequently, these two configurations apply a lower longitudinal force during braking compared to the nominal case, indicating more intense braking.
+%\textcolor{red}{In particular, the TLC and TLC+FLC configurations exhibit a higher longitudinal force while approaching the sharp turn, in the curvilinear abscissa interval $\al\in\left[0.707, 0.712\right]$, resulting in a higher longitudinal speed.
+%Subsequently, these two configurations apply a lower longitudinal force during braking compared to the nominal case, indicating more intense braking.
 This behavior explains the sharp reduction in longitudinal speed observed thereafter.}
@@
 
-\begin{figure}
-	\centering
-	\includegraphics{Fig/ol_saturation.pdf}
-	\caption{Axle saturation $S_j(\bx,\bu)$ for front axle (j=1, first panel) and rear axle (j=2, second panel) with respect to the curvilinear abscissa $\al$.
-The dashed black line corresponds to the reference (nominal) configuration, while the green, red, and blue lines represent the robustified configurations TLC, FLC, and TLC+FLC, respectively.}
-	\label{fig:ol_saturation}
-\end{figure}
-
 Firstly, we observe that the FLC and TLC+FLC configurations (red and blue lines) never reach full saturation ($S_j=1)$ due to the presence of the back-off term, which enforces a safety margin from the friction limit.
@@
 This behavior emerges around the sharp turn and becomes even more evident in the high-speed turn.
-Specifically, in the interval $\alpha \in [0.745, 0.762]$, the TLC and TLC+FLC configurations begin to experience high axle loads approximately 12\,m before the nominal trajectory and sustain them for an additional 4.6\,m (TLC, green line) and 9.2\,m (TLC+FLC, blue line) beyond the nominal reference.
+Specifically, in the interval $\al \in [0.745, 0.762]$, the TLC and TLC+FLC configurations begin to experience high axle loads approximately 12\,m before the nominal trajectory and sustain them for an additional 4.6\,m (TLC, green line) and 9.2\,m (TLC+FLC, blue line) beyond the nominal reference.
 
@@
 
-\subsection{Open-loop vs closed-loop robust planning}
-We now compare the open-loop planning with the closed-loop approach.
-In particular, we consider open-loop and closed-loop trajectories computed using both the robustified TLC and FLC, with $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$.
+\begin{figure}[t]
+	\centering
+	\includegraphics{Fig/cl_telemetries.pdf}
+	\caption{Closed-Loop method.
+Variations of longitudinal speed $\De u$, steering angle $\De\de$, longitudinal force $\De F_x$, and lateral deviation $\De e$ relative to the nominal trajectory for three schemes --- TLC (track-limit), FLC (friction-limit), and TLC+FLC (both).
+		Dashed lines mark the sign of the reference signal (positive, near-zero and negative).}
+
+	%\caption{Evolution of the longitudinal speed $\De u$, wheel steering angle $\De \de$, longitudinal force $\De F_x$, and lateral deviation $\De e$ are reported for three robustification scenarios: TLC (Track Limit Constraints), FLC (Friction Limit Constraint), and TLC+FLC (both constraints enforced concurrently).
+Each profile is shown relative to the baseline (nominal) trajectory to highlight the effect of each constraint-handling strategy.
+The dashed lines in the last three panels indicate the sign of the reference signal, allowing one to determine whether a positive variation with respect to the baseline corresponds to an increase or a decrease in the absolute value.
+This distinction is particularly important for quantities that depend on the direction of the turn, such as the wheel steering angle $\de$ and the lateral deviation $e$, while in the case of $F_x$, it indicates whether the vehicle is accelerating or braking.}
+	\label{fig:cl_telemetries}
+\end{figure}
+
+\begin{figure}[b!]
+	\centering
+	\includegraphics[scale = 1]{Fig/cl_saturation.pdf}
+	\caption{Closed-Loop.
+Axle saturation $S_j(\bx,\bu)$ for front axle (j=1, first panel) and rear axle (j=2, second panel) with respect to the curvilinear abscissa $\al$.
+Nominal configuration in dashed black, TLC in green, FLC in red, TLC+FLC in blue.
+		%The dashed black line corresponds to the reference (nominal) configuration, while the green, red, and blue lines represent the robustified configurations TLC, FLC, and TLC+FLC, respectively.
+	}
+	\label{fig:cl_saturation}
+\end{figure}
+
+An analogous analysis has been conducted using the closed-loop approach described in Section~\ref{sec:closed_loop_planning}, setting $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$.
+The same key signals displayed for the previous case are shown in Figure~\ref{fig:cl_telemetries}, while the axle saturation ratios $S_j$ are illustrated in Figure~\ref{fig:cl_saturation}.
+From one side, the results appear very similar to those obtained with the open-loop method.
+While the closed-loop method offers a more rigorous framework by enabling seamless covariance propagation over the entire lap, the observed agreement indicates that open-loop planning remains a valid and effective strategy, as long as the prediction horizon is appropriately selected.
+
+On the other hand, the trajectories are also sufficiently different to appreciate some subtle but relevant distinctions.
+For instance, sector time increments w.r.t.
+the nominal case of corresponding configurations are lower than the ones arising from the open-loop method: 1.39\% for the TLC, 0.66\% for the FLC, and 2.07\% for the TLC+FLC.
+Moreover, in both TLC and TLC+FLC configurations, the magnitude of $\De\de$ is appreciably smaller in the closed-loop case compared to the open-loop one, especially during the high-speed turn. 
+Additionally, the TLC+FLC configuration no longer exhibits the throttle release in $\De F_x$ observed in the open-loop case.
+
+As for the axle saturation, both methods lead to qualitatively similar evolutions, but slight differences can be observed within the high-speed turn interval $\al \in [0.745, 0.762]$.
+In this case, the TLC and TLC+FLC configurations begin to experience high axle loads approximately 10\,m before the nominal trajectory and sustain them for an additional 4.5\,m beyond the nominal reference.
+This is achieved through the closed-loop method's capability to \emph{steer} and tame the covariance matrix via the dynamic action of the controller.
+
+%\subsection{Open-loop vs closed-loop robust planning}
+Finally, we compare the back-off influence on the open-loop and closed-loop trajectories, using the robustified TLC with a higher $\ga^\textrm{TLC}=3.0$ for better visualization.
 As before, the open-loop planning is carried out with a prediction horizon of $H=4$ steps.
+The back-off term $\be^{TLC}$ in~\eqref{eq:TLC} admits a clear geometrical interpretation: it corresponds to the amount by which the track boundaries are tightened to meet the probabilistic constraint, or, equivalently, to a virtual extension of the vehicle's width.
 
-For the TLC, the back-off term $\be^{TLC}$ in~\eqref{eq:TLC} admits a clear geometrical interpretation: it corresponds to the amount by which the track boundaries are tightened to meet the probabilistic constraint, or, equivalently, to a virtual extension of the vehicle's width.
 According to the latter interpretation, Figure~\ref{fig:olcl_traj_bands} shows two highlighted bands---red for the open-loop case and cyan for the closed-loop case---corresponding to the area swept by the back-off-extended-vehicle.
@@
 \subsection{Simulations with noise realization}
-Finally, we provide an empirical validation of the closed-loop robustified planning by simulating the evolution of the controlled vehicle with random initial conditions and including noise realizations.
+To provide an empirical validation of the closed-loop robustified planning we simulated the evolution of the controlled vehicle with random initial conditions and included noise realizations.
 The random values for the initial state and the noise follow Gaussian distributions with covariance matrices $\bP_0$ and $\bQ$, respectively, as introduced in~\eqref{eq:dP}.

### 2025-06-06 stochastic_vehicle_dynamics.tex
[MM] New plot inserted

@@
 	\input{Fig/Tikz_Sources/tikz_vehicle_model.tex}
-	\caption{Single Track model.
+	%\caption{Single Track model.
 The ground force acting on each axle is decomposed into longitudinal and lateral components, indicated by red arrows.
@@
 the fixed frame.}
+	\caption{Single Track model.}
 	\label{fig:vehicle_model}

### 2025-06-09 MAIN.tex
[MM] merged figure versions
[MGa] sketching intro
[MGa] added discussion on LLE
and why they are of limited use in this context
[mg] general reviews
[MM] New plot inserted
[MGa] abstract and mods to conclusions

@@
 \usepackage{fancyhdr}
+\usepackage{enumitem}
+\usepackage{lipsum}
 
@@
 
+\usepackage{etoolbox} % for conditional blocks
+
 \DeclareMathOperator{\mf}{F_{\text{MF}}}
+
+\DeclareMathOperator{\diag}{diag}
+
+% version 1 (next 4 lines)
+%\DeclareMathOperator{\VarOp}{Var}
+%\DeclareMathOperator{\CovOp}{Cov}
+%\newcommand\Var[1]{\VarOp(#1)}
+%\newcommand\Cov[2]{\CovOp(#1,#2)}
+
+% version 2 (next 2 lines)
+\newcommand\Var[1]{\sigma_{#1}^2}
+\newcommand\Cov[2]{\sigma_{#1\,#2}}
 
@@
 \begin{abstract}
-This paper analyzes ...
+This paper presents a disturbance-aware framework that embeds robustness into minimum-lap-time trajectory optimization for motorsport.
+Two formulations are introduced.
+(i) Open-loop, horizon-based covariance propagation uses worst-case uncertainty growth over a finite window to tighten tire-friction and track-limit constraints.
+(ii) Closed-loop, covariance-aware planning incorporates a time-varying LQR feedback law in the optimizer, providing a feedback-consistent estimate of disturbance attenuation and enabling sharper yet reliable constraint tightening.
+Both methods yield reference trajectories for human or artificial drivers: in autonomous applications the modelled controller can replicate the on-board implementation, while for human driving accuracy increases with the extent to which the driver can be approximated by the assumed time-varying LQR policy.
+Computational tests on a representative Barcelona-Catalunya sector show that both schemes meet the prescribed safety probability, yet the closed-loop variant incurs smaller lap-time penalties than the more conservative open-loop solution, while the nominal (non-robust) trajectory remains infeasible under the same uncertainties.
+By accounting for uncertainty growth and feedback action during planning, the proposed framework delivers trajectories that are both performance-optimal and probabilistically safe, advancing minimum-time optimization toward real-world deployment in high-performance motorsport and autonomous racing.
 \end{abstract}
@@
 \input{stochastic_vehicle_dynamics}
-\input{marco_open_loop_planning}
+\input{open_loop_planning}
 %for back-off calculations
@@
 \bibliographystyle{tfnlm}
-\bibliography{biblio}
+\bibliography{Disturbance-Aware-Planning}
 

### 2025-06-09 conclusions.tex
[MGa] abstract and mods to conclusions

@@
 
-Quantitative results on a representative circuit segment have highlighted the trade-offs between robustness and performance.
+Both schemes deliver reference trajectories suitable for human or artificial drivers.
+In autonomous applications the modelled controller can replicate the on-board implementation, providing highly accurate disturbance predictions; for human driving, fidelity improves with the extent to which the driver's behaviour can be approximated by the assumed time-varying LQR policy.
+
+Quantitative results on a representative Barcelona-Catalunya sector highlight the trade-offs between robustness and performance.
 Open-loop planning, while conservative, ensures feasibility even in the absence of feedback, but incurs a higher performance cost due to larger safety margins.
 The closed-loop method achieves comparable safety guarantees with reduced conservativeness and improved lap-time performance, thanks to its ability to explicitly model the stabilizing action of the driver.
-Finally, random simulations with noise realization performed adopting the closed-loop approach proved the effectiveness of the robustified constraints.
+Finally, random simulations with noise realization, performed adopting the closed-loop approach, prove the effectiveness of the robustified constraints.
 A LQR-based controller tracking a robust trajectory planned with probabilistic constraint tightening exhibits significantly fewer track violations compared to one tracking the nominal optimal trajectory.
 
-Overall, the proposed framework enables the design of minimum-time trajectories that are not only fast but also certified to be safe under realistic disturbances.
-By embedding robustness into the planning stage, the gap between theoretical optimality and real-world executability is narrowed.
+By embedding uncertainty propagation and feedback action directly in the planning stage, the proposed framework yields trajectories that are simultaneously fast and probabilistically safe, advancing minimum-time optimisation toward practical deployment in high-performance motorsport and autonomous racing.
 Future work will explore the generalization to more accurate vehicle models and experimental validation of the methodology.

### 2025-06-09 performance_comparison.tex
[MGa] abstract and mods to conclusions

@@
 In this section we present some results obtained using the described robust planning frameworks. All planned trajectories refer to the sector of the Catalunya circuit shown in Figure~\ref{fig:track}.
-The entire circuit is parametrized by the curvilinear abscissa $\alpha \in [0,1]$, and the selected sector corresponds to the interval $\left[0.70, 0.77\right]$.
+The entire circuit is parametrized by the curve parameter $\alpha \in [0,1]$, and the selected sector corresponds to the interval $\left[0.70, 0.77\right]$.
 Checkpoints are indicated by labels and are uniformly spaced by $\De\al=0.01$.
@@
 Specifically, $\ga^\textrm{TLC} = 0$ yields a satisfaction probability of 50\% and leads to the nominal solution with no back-off, while $\ga^\textrm{TLC} = 3$ corresponds to a confidence level of 99\%.
-For the purposes of this analysis, we examine the effect of varying $\ga^\textrm{TLC}$ across this range.
+For the purposes of this analysis, we examine the effect of varying $\ga^\textrm{TLC}$ in this range.
 
@@
 
-Observing the left panel of Figure~\ref{fig:ol_sensitivities}, it is evident that the trajectories corresponding to configurations with lower values of $\ga^\textrm{TLC}$ tend to travel closer to the track boundaries.
+Observing the left panel of Figure~\ref{fig:ol_sensitivities}, it is evident that the trajectories corresponding to configurations with lower values of $\ga^\textrm{TLC}$ yield smaller lateral margins with respect to the track boundaries.
 In contrast, the trajectory associated with $\ga^\textrm{TLC} = 3$\,--- represented by the light green line \,--- remains significantly farther from the edges.
@@
 We recall that at each grid point a new prediction horizon is initialized, and the corresponding matrix $\bP$ at its starting point, i.e., $\bP^0_k$, is set to $\bP_0$.
-The diagonal elements of $\bP$ are the variances $\Var{x_i}$  of each state variables, while the off-diagonal elements encode their covariances $\Cov{x_i}{x_j}$.
+The diagonal elements of $\bP$ are the variances $\Var{x_i}$  of each state variable, while the off-diagonal elements encode their covariances $\Cov{x_i}{x_j}$.
 Larger diagonal values indicate higher uncertainty in the corresponding states, and nonzero off-diagonal terms imply mutual dependence between them.
@@
 The black line corresponds to the baseline case with $\bP_0 = \bar{\bP}_0$, while the red and the blue lines represent the cases with $\bP_0 = \frac{1}{4}\bar{\bP}_0$ and $\bP_0 = 4\bar{\bP}_0$, respectively.
-As reasonably expected, using a larger initial covariance matrix results in a greater covariance matrix after $H$ steps, which in turn leads to a higher back-off term.
-This is reflected in the blue line, which travels closer to the centerline with respect to the other two lines.
+As reasonably expected, using a larger initial covariance matrix results in an inflated covariance matrix that reflects increased uncertainty after $H$ steps, which in turn leads to a higher back-off term.
+This is associated to the blue line, which travels closer to the centerline with respect to the other two lines.
 
@@
 Except for the $\De u$ panel, each plot includes dashed black lines that indicate the sign of the nominal signal.
-These indicators are shown at three distinct vertical levels --- upper, central, and lower --- corresponding respectively to positive, near-zero, and negative values of the nominal signal.
+These indicators are shown at three distinct vertical levels --- high, medium, and low --- corresponding respectively to positive, near-zero, and negative values of the nominal signal.
 
@@
 This peak is followed by a sharp drop in speed, indicating intense braking.
-The TLC+FLC configuration exhibits a qualitatively similar trend in $\Delta u$, although the peak is significantly lower and the subsequent speed reduction is clearly stronger.
+The TLC+FLC configuration demonstrates a qualitatively similar trend in $\Delta u$, although the peak is significantly lower and the subsequent speed reduction is clearly stronger.
 This behavior is consistent with the variation of the total longitudinal force $\Delta F_x$ shown in the third panel.
@@
 Later in the sector, a further deviation in driving style appears during the high-speed turn, specifically around $\alpha \in [0.75, 0.76]$.
-As shown in the third panel ($\Delta F_x$), the combined configuration TLC+FLC needs to reduce the longitudinal force within this sector to meet both the constraints.
+As shown in the third panel ($\Delta F_x$), in the combined configuration TLC+FLC a reduction in the longitudinal force within this sector is needed to meet both constraints.
 This adaptation is necessary because the TLC forces the vehicle to follow a wider trajectory, which entails greater lateral acceleration and, consequently, higher lateral force demand.
@@
 %---Axle saturation
-After analyzing the variations of these four key signals with respect to the nominal solution, we now turn our attention to the evaluation of axle saturation $S_j(\bx,\bu)$ defined in~\eqref{eq:axle_saturation}, which provides insight into how close each configuration operates to the tire grip limits.
+After analyzing the variations of these four key signals with respect to the nominal solution, we now turn our attention to the evaluation of axle saturation $S_j(\bx,\bu)$ defined in Eq.~\eqref{eq:axle_saturation}, which provides insight into how close each configuration operates to the tire grip limits.
 
@@
 Firstly, we observe that the FLC and TLC+FLC configurations (red and blue lines) never reach full saturation ($S_j=1)$ due to the presence of the back-off term, which enforces a safety margin from the friction limit.
-Secondly, we point out how the different trajectories followed by the TLC and TLC+FLC configurations (green and blue lines), compared to the nominal and FLC cases, entail distinct ground force demands, which are clearly visible here in terms of the $S_j$ ratio.
+Secondly, we point out how the different trajectories followed by the TLC and TLC+FLC configurations (green and blue lines), compared to the nominal (black) and FLC case (red), entail distinct ground force demands, which are clearly visible here in terms of the $S_j$ ratio.
 This behavior emerges around the sharp turn and becomes even more evident in the high-speed turn.
@@
 In this case, the TLC and TLC+FLC configurations begin to experience high axle loads approximately 10\,m before the nominal trajectory and sustain them for an additional 4.5\,m beyond the nominal reference.
-This is achieved through the closed-loop method's capability to \emph{steer} and tame the covariance matrix via the dynamic action of the controller.
+This is achieved through the closed-loop method's capability to \emph{steer} and \emph{tame} the covariance matrix propagation via the dynamic action of the controller.
 
 %\subsection{Open-loop vs closed-loop robust planning}
-Finally, we compare the back-off influence on the open-loop and closed-loop trajectories, using the robustified TLC with a higher $\ga^\textrm{TLC}=3.0$ for better visualization.
+Finally, we compare the influence of the back-off terms on the open-loop and closed-loop trajectories, in the robustified TLC scenario with the high $\ga^\textrm{TLC}=3.0$ value for better visualization.
 As before, the open-loop planning is carried out with a prediction horizon of $H=4$ steps.
-The back-off term $\be^{TLC}$ in~\eqref{eq:TLC} admits a clear geometrical interpretation: it corresponds to the amount by which the track boundaries are tightened to meet the probabilistic constraint, or, equivalently, to a virtual extension of the vehicle's width.
+The back-off term $\be^{TLC}$ in Eq.~\eqref{eq:TLC} admits an interesting geometrical interpretation: it corresponds to the amount by which the track boundaries are (not uniformly) tightened to meet the probabilistic constraint, or, equivalently, to a virtual extension of the vehicle's width.
 
@@
 \subsection{Simulations with noise realization}
-To provide an empirical validation of the closed-loop robustified planning we simulated the evolution of the controlled vehicle with random initial conditions and included noise realizations.
-The random values for the initial state and the noise follow Gaussian distributions with covariance matrices $\bP_0$ and $\bQ$, respectively, as introduced in~\eqref{eq:dP}.
+To provide an empirical validation of the closed-loop robustified planning we simulated the evolution of the controlled vehicle with 100 random initial conditions and included noise realizations.
+The random values for the initial state and the noise follow a Gaussian distribution with covariance matrices $\bP_0$ and $\bQ$, in both cases, as introduced in Eq.~\eqref{eq:dP}.
 This analysis cannot be performed for the open-loop approach, as it does not compute a stabilizing controller---resulting in trajectories that would significantly deviate from the track limits beyond the prediction horizon.

### 2025-06-09 performance_comparison.tex
[MGa] Fig.10 caption

@@
 	\includegraphics{Fig/olcl_traj_bands.pdf}
-	\caption{Lorem ipsum.}
+	\caption{Back-off envelopes for the open-loop (red) and closed-loop (cyan) robust planners.
+Each shaded band represents the area swept by the back-off-extended vehicle and must lie entirely within the actual track boundaries.
+The open-loop approach yields a wider envelope due to worst-case growth assumptions, whereas the closed-loop design, aided by time-varying LQR controller, maintains a narrower margin.}
 	\label{fig:olcl_traj_bands}

### 2025-06-09 Fig/TikZ_Sources/tikz_stayontrack_constraint.tex
[MM] TLC figure added and relative caption shortened

@@
-\begin{tikzpicture}[scale = .7]
-	% Parameters
-	\def\outerX{3}   % x radius of outer ellipse
-	\def\outerY{2}   % y radius of outer ellipse
-	\def\gap{0.2}    % distance between outer and inner ellipse
-	\def\innerX{2.68}
-	\def\innerY{1.79}
+\begin{tikzpicture}[scale = 1.1]
+	\tikzset{sref/.style={
+			draw = blue,
+			%thick, 
+			->
+			%, >={Latex[length=5pt, width=3pt]}
+	},
+	elat/.style={
+		red!60!black,
+		%thick,
+		->
+		%, >={Latex[length=5pt, width=3pt]}
+	},
+	elimits/.style={
+	->
+	%, >={Latex[length=5pt, width=3pt]}
+	},
+	nplane/.style={
+		very thin,
+		draw=black
+	}
+}
+%% --------------------- TRACK -----------------------
+% FILLED AREA
+\draw  [draw opacity=0][fill = black!30] (-1.430,1.400) ..
+controls (-1.295,1.353) and (0.079,1.126) ..
+(1.413,1.259) ..
+controls (2.446,1.339) and (3.866,1.619) ..
+(4.460,1.742) ..
+controls (4.513,1.493) and (4.737,0.363) ..
+(4.810,0.012) ..
+controls (3.786,-0.174) and (3.079,-0.348) ..
+(1.613,-0.501) ..
+controls (0.219,-0.621) and (-1.014,-0.508) ..
+(-1.810,-0.320) ..
+controls (-1.736,-0.005) and (-1.466,1.235) ..
+(-1.430,1.400) -- cycle ;
+% CENTERLINE
+\draw    (-1.616,0.536) .. controls (0.594,0.038) and (2.894,0.528) .. (4.634,0.878) ;
+% E_MAX 
+\draw    (-1.350,1.740) .. controls (0.500,1.302) and (2.930,1.782) .. (4.390,2.092) ;
+% E_MAX - BO
+\draw[dashed]    (-1.430,1.400) .. controls (0.820,0.942) and (2.990,1.462) .. (4.460,1.742) ;
+% E_MIN 
+\draw    (-1.900,-0.660) .. controls (0.550,-1.218) and (2.990,-0.688) .. (4.870,-0.338) ;
+% E_MIN - BO 
+\draw[dashed]    (-1.810,-0.320) .. controls (0.700,-0.858) and (2.930,-0.338) .. (4.810,0.012) ;
+% NORMAL PLANE 1
+\draw  [nplane]  (-0.114,1.573) -- (-0.250,-0.884) ;
+% NORMAL PLANE 2 
+\draw  [nplane]  (3.110,1.826) -- (3.480,-0.584) ;
+% CENTER POINTS' POSITIONS
+%\draw    (-0.182,0.344) -- (3.295,0.621) ;
 
-	% Draw outer ellipse
-	\draw[] (0,0) ellipse [x radius=\outerX, y radius=\outerY];
+%% --------------------- REF AND DISTANCES -----------------------
+% X-AXIS 1
+%\draw [sref] (-0.182,0.344) -- (0.097,0.330);
+% Y-AXIS 1
+%\draw [sref]   (-0.182,0.344) -- (-0.169,0.617) ;
+% E-LAT 1
+\draw [elat]   (-0.290,0.357) -- (-0.260,0.832) node[midway,black,xshift=-4,yshift=-2] {$e$};;
+% X-AXIS 2
+%\draw [sref]   (3.295,0.621) -- (3.571,0.661);
+% Y-AXIS 2
+%\draw [sref]   (3.295,0.621) -- (3.254,0.890) ;
+% E-LAT 2
+\draw[elat]   (3.387,0.641) -- (3.326,1.042) ;
 
-	% Draw inner ellipse filled with color
-	\fill[black!30] (0,0) ellipse [
-	x radius={\innerX},
-	y radius={\innerY}
-	];
-	\node[above] at (0,0) {PATH CONSTRAINT FIG};
-	\draw[dashed] (0,0) ellipse [
-	x radius={\innerX},
-	y radius={\innerY}
-	];
+%% --------------------- VEHICLES -----------------------
+% VEHICLE'S COM POSITIONS 
+%\draw  (-0.160,0.826) -- (3.241,1.019) ;
+\fill (-0.160,0.826) circle (1pt) ;
+\fill (3.241,1.019) circle (1pt) ;
+% TRAJECTORY
+%\draw[dashdotted]    (-1.407,0.809) ..
+controls (-1.007,1.109) and (-0.560,0.526) ..
+(-0.160,0.826) ..
+controls (0.240,1.126) and (0.890,1.084) ..
+(1.496,1.130) ..
+controls (2.103,1.177) and (2.666,1.255) ..
+(3.241,1.019) ..
+controls (3.816,0.784) and (4.183,0.944) ..
+(4.296,1.057) ;
+\draw[dashdotted]    (-1.460,0.281) ..
+controls (-0.767,0.347) and (-0.540,0.514) ..
+(-0.160,0.826) ..
+controls (0.220,1.137) and (0.886,1.113) ..
+(1.496,1.130) ..
+controls (2.107,1.147) and (2.715,1.258) ..
+(3.241,1.019) ..
+controls (3.767,0.781) and (4.053,0.721) ..
+(4.513,0.734) ;
+% VEHICLE 1
+% CHASSIS 1 
+\draw    (-0.642,0.927) -- (0.322,0.724) ;
+% FRONT WHEEL 1
+\draw  [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ,fill opacity=0.6 ] (0.176,0.862) ..
+controls (0.186,0.881) and (0.209,0.888) ..
+(0.228,0.879) -- (0.501,0.742) ..
+controls (0.521,0.733) and (0.528,0.709) ..
+(0.519,0.690) -- (0.467,0.587) ..
+controls (0.457,0.568) and (0.434,0.560) ..
+(0.415,0.569) -- (0.142,0.706) ..
+controls (0.122,0.716) and (0.115,0.739) ..
+(0.124,0.758) -- cycle ;
+% REAR WHEEL 1
+\draw  [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ,fill opacity=0.6 ] (-0.818,1.022) ..
+controls (-0.814,1.042) and (-0.793,1.056) ..
+(-0.772,1.052) -- (-0.473,0.991) ..
+controls (-0.452,0.987) and (-0.438,0.967) ..
+(-0.442,0.946) -- (-0.465,0.833) ..
+controls (-0.469,0.812) and (-0.490,0.798) ..
+(-0.511,0.802) -- (-0.810,0.863) ..
+controls (-0.831,0.867) and (-0.845,0.887) ..
+(-0.841,0.908) -- cycle ;
+% VEHICLE 2
+% CHASSIS 2 
+\draw    (2.772,0.870) -- (3.710,1.169) ;
+% FRONT WHEEL 2
+\draw  [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ,fill opacity=0.6 ] (3.515,1.217) ..
+controls (3.514,1.238) and (3.531,1.256) ..
+(3.552,1.257) -- (3.857,1.273) ..
+controls (3.879,1.274) and (3.897,1.258) ..
+(3.898,1.237) -- (3.904,1.121) ..
+controls (3.905,1.100) and (3.889,1.082) ..
+(3.868,1.081) -- (3.562,1.065) ..
+controls (3.541,1.063) and (3.523,1.080) ..
+(3.521,1.101) -- cycle ;
+% REAR WHEEL 2
+\draw  [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ,fill opacity=0.6 ] (2.572,0.865) ..
+controls (2.565,0.886) and (2.576,0.907) ..
+(2.597,0.914) -- (2.887,1.009) ..
+controls (2.908,1.016) and (2.929,1.005) ..
+(2.936,0.985) -- (2.972,0.875) ..
+controls (2.979,0.854) and (2.968,0.833) ..
+(2.948,0.826) -- (2.657,0.731) ..
+controls (2.637,0.724) and (2.615,0.735) ..
+(2.608,0.755) -- cycle ;
 
-	% Draw axes
-	\draw[->] (-\outerX-1,0) -- (\outerX+1,0) node[right] {$X_i$};
-	\draw[->] (0,-\outerY-1) -- (0,\outerY+1) node[above] {$Y_i$};
-
-	% Arrows pointing to the ellipses
-	\draw[->,thin] (1.7,2) -- ({\outerX*cos(60)},{\outerY*sin(60)}) node[midway, above right] {$\beta=0$};
-	\draw[->,thin] (2.3,1.7) -- ({(\innerX)*cos(45)},{(\innerY)*sin(45)}) node[midway, right, shift={(0.05,.2)}] {$\beta=\gap$};
-
-	% Arrows pointing to the ellipses
-	\draw[->,thin] ({\outerX+.3},.3) -- (\outerX,0) node[midway, above right,shift={(.05,0)}] {$\mu_x Z_i$};
-	\draw[->,thin] (.3,{\outerY+.3}) -- (0,\outerY) node[midway, above right,shift={(.05,0)}] {$\mu_y Z_i$};
+%% --------------------- E-MAX/E-MIN LABELS -----------------------
+% E-MAX
+\draw[elimits]    (0.090,1.692) -- (0.345,1.562) node[above,xshift = -16,yshift = 1] {$e_\textrm{max}$};
+% E-MIN
+\draw[elimits]   (0.275,-1.023) node[below,yshift = 3] {$e_\textrm{min}$} -- (0.520,-0.903) ;
+% E-MAX - BO
+%\draw[elimits]     (0.850,1.372) -- (1.105,1.242) node[above,xshift = -16,yshift = -2] {$e_\textrm{max}-\be$};
+\draw[elimits]     (1.6,1.8) node[above,yshift = -4,xshift = 5] {$e_\textrm{max}-\be$} -- (1.105,1.242) ;
+% E-MIN - BO
+%\draw[elimits]     (0.865,-0.654) -- (1.110,-0.534) ;
+\draw[elimits]     (1.8,-1.0) node[below,yshift = 8,xshift = 11] {$e_\textrm{min}+\be$} -- (1.110,-0.534) ;
 

### 2025-06-09 stochastic_vehicle_dynamics.tex
[MM] TLC figure added and relative caption shortened

@@
 	\centering
-	\input{Fig/Tikz_Sources/tikz_adherence_constraint.tex}
+	\adjustbox{valign=c}{\input{Fig/Tikz_Sources/tikz_adherence_constraint.tex}}
 	\hfill
-	\input{Fig/Tikz_Sources/tikz_stayontrack_constraint.tex}
-	\caption{Graphical representation of the two constraints analyzed in the section.
+	\adjustbox{valign=c}{\input{Fig/Tikz_Sources/tikz_stayontrack_constraint.tex}}
+	\caption{Graphical representation of the two constraints analyzed in this section: the friction limit constraint in the $X_j$-$Y_j$ plane (left panel), and the track limit constraint in the $x$-$y$ plane (right panel).
+In both figures, the filled gray area bounded by a dashed line indicates the accessible region in the presence of a back-off term.
+	} 
+		%Graphical representation of the two constraints analyzed in the section.
 The left panel shows the friction limit constraint in the plane $X_j-Y_j$.
@@
 This ellipse has its semi-axes reduced by a factor $\sqrt{1-\be^\textrm{FLC}_j}$ w.r.t.
-the one in solid line.
-	}
+the one in solid line.}
 	\label{fig:robust_constraints}

### 2025-06-11 intro.tex
[MGa] curvilinear abscissa -> curvilinear parameter

@@
 This has led to a significant body of work on robust trajectory planning, particularly for unmanned aerial vehicles (UAVs), with the aim of minimizing sensitivity to state and input variations.
-For instance,~\cite{Brault:RobustTrajectoryPlanning:2021},~\cite{Giordano:TrajectoryGenerationMinimum:2018}, and~\cite{Brault:TubebasedTrajectoryOptimization:} propose tube-based and sensitivity-aware optimization frameworks that explicitly account for both input and closed-loop state sensitivity in the planning phase.
+For instance,~\cite{Brault:RobustTrajectoryPlanning:2021} and~\cite{Giordano:TrajectoryGenerationMinimum:2018} propose tube-based and sensitivity-aware optimization frameworks that explicitly account for both input and closed-loop state sensitivity in the planning phase.
 Further developments integrate observability-aware planning into a unified multi-step optimization framework, as demonstrated in~\cite{Bohm:COPControlObservabilityaware:2022}.

### 2025-06-11 performance_comparison.tex
[MGa] curvilinear abscissa -> curvilinear parameter

@@
 
-In this section we present some results obtained using the described robust planning frameworks. All planned trajectories refer to the sector of the Catalunya circuit shown in Figure~\ref{fig:track}.
-The entire circuit is parametrized by the curve parameter $\alpha \in [0,1]$, and the selected sector corresponds to the interval $\left[0.70, 0.77\right]$.
+In this section we present some results obtained using the described robust planning frameworks. All planned trajectories refer to the sector of the Catalunya circuit shown in Fig.~\ref{fig:track}.
+The entire circuit is parameterized by the curvilinear parameter $\alpha \in [0,1]$, and the selected sector corresponds to the interval $\left[0.70, 0.77\right]$.
 Checkpoints are indicated by labels and are uniformly spaced by $\De\al=0.01$.
@@
 
-Observing the left panel of Figure~\ref{fig:ol_sensitivities}, it is evident that the trajectories corresponding to configurations with lower values of $\ga^\textrm{TLC}$ yield smaller lateral margins with respect to the track boundaries.
+Observing the left panel of Fig.~\ref{fig:ol_sensitivities}, it is evident that the trajectories corresponding to configurations with lower values of $\ga^\textrm{TLC}$ yield smaller lateral margins with respect to the track boundaries.
 In contrast, the trajectory associated with $\ga^\textrm{TLC} = 3$\,--- represented by the light green line \,--- remains significantly farther from the edges.
@@
 
-A comparison between the trajectories obtained using different values of $\bP_0$ is shown in the right panel of Figure~\ref{fig:ol_sensitivities}.
+A comparison between the trajectories obtained using different values of $\bP_0$ is shown in the right panel of Fig.~\ref{fig:ol_sensitivities}.
 The black line corresponds to the baseline case with $\bP_0 = \bar{\bP}_0$, while the red and the blue lines represent the cases with $\bP_0 = \frac{1}{4}\bar{\bP}_0$ and $\bP_0 = 4\bar{\bP}_0$, respectively.
@@
 	\caption{Open-Loop.
-Axle saturation $S_j(\bx,\bu)$ for front axle (j=1, first panel) and rear axle (j=2, second panel) with respect to the curvilinear abscissa $\al$.
+Axle saturation $S_j(\bx,\bu)$ for front axle (j=1, first panel) and rear axle (j=2, second panel) with respect to the curvilinear parameter $\al$.
 Nominal configuration in dashed black, TLC in green, FLC in red, TLC+FLC in blue.
@@
 %---Delta key signals
-The comparison between the nominal case and the three robustified cases is illustrated in Figure~\ref{fig:ol_telemetries}, in terms of the deviation of selected signals from to their nominal counterparts.
+The comparison between the nominal case and the three robustified cases is illustrated in Fig.~\ref{fig:ol_telemetries}, in terms of the deviation of selected signals from to their nominal counterparts.
 The panels display variations on the longitudinal speed $\De u$ (first panel), wheel steering angle $\De \de$ (second panel), total longitudinal force $\De F_x$ (third panel), and lateral deviation of the center of mass (CoM) from the track centerline $\De e$ (fourth panel).
@@
 	\caption{Closed-Loop.
-Axle saturation $S_j(\bx,\bu)$ for front axle (j=1, first panel) and rear axle (j=2, second panel) with respect to the curvilinear abscissa $\al$.
+Axle saturation $S_j(\bx,\bu)$ for front axle (j=1, first panel) and rear axle (j=2, second panel) with respect to the curvilinear parameter $\al$.
 Nominal configuration in dashed black, TLC in green, FLC in red, TLC+FLC in blue.
@@
 An analogous analysis has been conducted using the closed-loop approach described in Section~\ref{sec:closed_loop_planning}, setting $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$.
-The same key signals displayed for the previous case are shown in Figure~\ref{fig:cl_telemetries}, while the axle saturation ratios $S_j$ are illustrated in Figure~\ref{fig:cl_saturation}.
+The same key signals displayed for the previous case are shown in Fig.~\ref{fig:cl_telemetries}, while the axle saturation ratios $S_j$ are illustrated in Fig.~\ref{fig:cl_saturation}.
 From one side, the results appear very similar to those obtained with the open-loop method.
@@
 
-According to the latter interpretation, Figure~\ref{fig:olcl_traj_bands} shows two highlighted bands---red for the open-loop case and cyan for the closed-loop case---corresponding to the area swept by the back-off-extended-vehicle.
+According to the latter interpretation, Fig.~\ref{fig:olcl_traj_bands} shows two highlighted bands---red for the open-loop case and cyan for the closed-loop case---corresponding to the area swept by the back-off-extended-vehicle.
 Ultimately, the TLC enforces precisely that these bands remain entirely within the actual track limits.
@@
 
-The left panel of Figure~\ref{fig:traj_strings} shows random evolutions (red lines) for an LQR-controlled vehicle stabilized to follow the nominal optimal trajectory (black line).
+The left panel of Fig.~\ref{fig:traj_strings} shows random evolutions (red lines) for an LQR-controlled vehicle stabilized to follow the nominal optimal trajectory (black line).
 No robust constraints are introduced, and the controller implements the feedback policy from Step 2 of Subsection~\ref{sec:LQR}.
@@
 	\includegraphics{Fig/olcl_traj_strings.pdf}
-	\caption{Simulations with random initial conditions and Gaussian noise realizations.
+	\caption{100 simulations (each panel) with random initial conditions and Gaussian noise realizations.
 In the left panel, the reference trajectory (black curve) is the non-robust one, tracked using an LQR controller designed around it as per Step 2 of Subsection~\ref{sec:LQR}.

### 2025-06-11 stochastic_vehicle_dynamics.tex
[MGa] curvilinear abscissa -> curvilinear parameter

@@
 The vehicle model employed in this study is a Single Track model featuring nonlinear axle characteristics.
-A schematic representation of the Single Track model is shown in Figure~\ref{fig:vehicle_model}.
+A schematic representation of the Single Track model is shown in Fig.~\ref{fig:vehicle_model}.
 The lateral forces $Y_1$ and $Y_2$ are computed using a Pacejka's Magic Formula, which depends on the total vertical load acting on the axle and the axle's apparent slip angle.
@@
 The state vector $\bx$ has six components and includes the three kinematic quantities $u$, $v$, and $r$, along with the position $\left(x_G, y_G\right)$ and orientation $\psi$ of the barycentrical reference frame.
-As illustrated in Figure~\ref{fig:vehicle_model} $u$ and $v$ represent the longitudinal and lateral velocity of the CoM w.r.t its reference frame, while $r$ represents vehicle's yaw rate.
+As illustrated in Fig.~\ref{fig:vehicle_model} $u$ and $v$ represent the longitudinal and lateral velocity of the CoM w.r.t its reference frame, while $r$ represents vehicle's yaw rate.
 
@@
 The backoff terms stem for a linearization around the mean of the original \emph{chance constraint} on $\bx(t)$ expressed by $\Prob \{h_i(\bx) \leq 0\}\geq p$, where $p$ is the confidence level of constraint satisfaction.
-Explicitly the backoff terms are given by
+Explicitly, the backoff terms are given by
 $\be_i = \gamma \sig_i$.
@@
 \end{subequations}
-To perform the discretization, the track centerline is parameterized using a curvilinear abscissa $\alpha \in [0,1]$, and uniformly sampled at $N+1$ points $\alpha_0, \ldots, \alpha_N$.
+To perform the discretization, the track centerline is parameterized using a curvilinear parameter $\alpha \in [0,1]$, and uniformly sampled at $N+1$ points $\alpha_0, \ldots, \alpha_N$.
 Accordingly, $\bmu_k$ denotes the mean state at grid node $\alpha_k$, while the controls $\bu_k$ and the algebraic variables $\bz_k$ are assumed to be piecewise constant over each interval $[\al_k, \al_{k+1}]$.
@@
 In~\eqref{eq:adherence_with_axle_saturation}, the superscript FLC denotes friction limit constraint, and $j=1,2$ refer the front and rear axle, respectively.
-In~\eqref{eq:axle_saturation}, $X_j$ and $Y_j$ denote the longitudinal and lateral components of the in-plane ground forces, respectively, as illustrated in Figure~\ref{fig:vehicle_model}, while $Z_j$ represents the total vertical load acting on the axle.
+In~\eqref{eq:axle_saturation}, $X_j$ and $Y_j$ denote the longitudinal and lateral components of the in-plane ground forces, respectively, as illustrated in Fig.~\ref{fig:vehicle_model}, while $Z_j$ represents the total vertical load acting on the axle.
 
@@
 The points $\left(X_j,Y_j\right)$ are constrained to lie within an ellipse whose semi-axes are given by $\mu_{x,j}Z_j$ and $\mu_{y,j}Z_j$.
-The constraint, without the back-off term, is represented in the left panel of Figure~\ref{fig:robust_constraints} by the solid line ellipse in the plane $X_jY_j$.
+The constraint, without the back-off term, is represented in the left panel of Fig.~\ref{fig:robust_constraints} by the solid line ellipse in the $X_jY_j$-plane.
 
 To better clarify the effect of the back-off term on the constraint let us consider a combination of $\bx$, $\bu$, $P$ and $\ga^\textrm{FLC}_j$ so that $\be^\textrm{FLC}_j=0.2$.
-This implies that the semi-axes of the ellipse are reduced by a factor $\sqrt{1-\be^\textrm{FLC}_j} = \sqrt{0.8}$, as shown in the left panel Figure~\ref{fig:robust_constraints}, where the ellipse in dashed line delimits the available region with the back-off applied.
+This implies that the semi-axes of the ellipse are reduced by a factor $\sqrt{1-\be^\textrm{FLC}_j} = \sqrt{0.8}$, as shown in the left panel of Fig.~\ref{fig:robust_constraints}, where the ellipse in dashed line delimits the available region with the back-off applied.
 This allows the optimizer to determine the most appropriate trade-off between the longitudinal force $X_j$ and the lateral force $Y_j$ --- often by reducing both components to some extent --- based on the specific requirements of the manoeuvre.
@@
 	\adjustbox{valign=c}{\input{Fig/Tikz_Sources/tikz_stayontrack_constraint.tex}}
-	\caption{Graphical representation of the two constraints analyzed in this section: the friction limit constraint in the $X_j$-$Y_j$ plane (left panel), and the track limit constraint in the $x$-$y$ plane (right panel).
+	\caption{Graphical representation of the two constraints analyzed in this section: the friction limit constraint in the $X_jY_j$ plane (left panel), and the track limit constraint in the $xy$ plane (right panel).
 In both figures, the filled gray area bounded by a dashed line indicates the accessible region in the presence of a back-off term.

### 2025-06-11 performance_comparison.tex
[MM] Some software detail added

--- 2025-06-09 performance_comparison.tex
@@
 This sector includes two distinct corners: a low-speed turn from $\al = 0.72$ to $\al = 0.73$, and a high-speed turn from $\al = 0.75$ to $\al = 0.76$, allowing for the evaluation of vehicle behavior across different dynamic scenarios.
+
+The chosen sector has been discretized with 140 intervals, corresponding to one grid point each 0.43\,m. 
+The optimization problems are formulate within the CasADi+MATLAB framework~\cite{Andersson:MPC:2019}, and solved using the interior point method implemented in IPOPT~\cite{Wachter:MP:2006}.
 
@@
 
+% Calculus times 
+We provide here some details on the software implementation and computation times to complement the comparison between the two proposed methods.
+All optimizations were performed on a machine equipped with an Intel Core i9-13980HX processor and 32\,Gb of RAM.
+Only the solver time is reported for each configuration, as the problem build time is negligible.
+The nominal problem involves 4200 optimization variables and is solved in 7.5\,s.
+The open-loop configuration, using a prediction horizon of $H=4$ as adopted in this study, includes 48300 variables and requires 128\,s to solve.
+The closed-loop method first solves the nominal problem, then performs the Step~2 in 0.5\,s, and finally formulates an additional problem with 14700 variables, which is solved in 33\,s, resulting in a total computation time of 40\,s.
+
 \subsection{Simulations with noise realization}

### 2025-06-11 closed_loop_planning.tex
[MM] typos

@@
 This assumes deterministic system dynamics, where $\hbmu_k$, $\hbu_k$, and $\hbz_k$ denote differential states, controls, and algebraic states, respectively, for \( k = 0, \ldots, N \).
-To avoid redundancy, we note that the NLP to be solved is a special case of~\eqref{eq:DOCP}, where~\eqref{eq:DOCPdP},~\eqref{eq:DOCPdPIC}, and the backoff terms are omitted, since the goal is to compute a nominal trajectory.
+To avoid redundancy, we note that the NLP to be solved is a special case of~\eqref{eq:DOCP}, where~\eqref{eq:DOCPdP},~\eqref{eq:DOCPdPIC}, and the back-off terms are omitted, since the goal is to compute a nominal trajectory.
 
@@
 To distinguish this case from~\eqref{eq:DOCP}, we use the symbol \( \tbP \) to denote the closed-loop (CL) covariance matrix, whose discrete evolution corresponds to a collocation-based approximation of the continuous-time dynamics $\dot{\tbP}(t) = \tbA(t) \tbP(t) + \tbP(t)\tbA(t)^T + \bQ(t)$, with $\tbA(t)=\bA(t)-\bB(t)\bK(t)$.
-It is worth noting that, accordingly, in equation~\eqref{eq:ROCPconstraints} the backoff term $\be_i$ is a function of closed-loop uncertainty $\tbP_k$.
+It is worth noting that, accordingly, in equation~\eqref{eq:ROCPconstraints} the back-off term $\be_i$ is a function of closed-loop uncertainty $\tbP_k$.
 Another key aspect is that the state-feedback matrix $\bK_k$ is also treated as a decision variable.

### 2025-06-11 open_loop_planning.tex
[MGa] curvilinear abscissa -> curvilinear parameter
[MM] typos

@@
 Equation~\eqref{eq:DOCPdPinitopenloop} defines the initialization of the appropriate covariance matrix version at step $k$, while Eq.~\eqref{eq:DOCPconstraintsopenloop} specifies that the back-off term in the constraints is computed using the most propagated version, $\bP^H_k$.
-A schematic illustration of the management of the $H+1$ covariance matrix instances is provided in Figure~\ref{fig:DOCPgrid}.
+A schematic illustration of the management of the $H+1$ covariance matrix instances is provided in Fig.~\ref{fig:DOCPgrid}.
 The dashed rectangle highlights the $k$-th discretization step, where the constraints are evaluated.
@@
 Each node corresponds to a specific covariance matrix, where the subscript denotes the current grid index and the superscript indicates the number of propagation steps since initialization.
-At each step, two key instances are emphasized: the matrix to be initialized at that step (red node), and the matrix that has been propagated over $H$ steps (green node), which is used in the evaluation of the constraint backoff.}
+At each step, two key instances are emphasized: the matrix to be initialized at that step (red node), and the matrix that has been propagated over $H$ steps (green node), which is used in the evaluation of the constraint back-off.}
 	%\caption{Schematic representation of the continuity and initialization Equations for the $H+1$ versions of the covariance matrix introduced in the discretized OCP.

### 2025-06-11 stochastic_vehicle_dynamics.tex
[MM] typos

@@
 \begin{align}\label{eq:P_STM}
-\bP(t) = \bPhi(t,t_0) \bP_0 \bPhi(t,t_0)+\int_{t_0}^{t} \bPhi(t,\tau) \bQ(\tau) \bPhi^T(t,\tau) \dd \tau \quad \bP(0)=\bP_0,
+\bP(t) = \bPhi(t,t_0) \bP_0 \bPhi^T(t,t_0)+\int_{t_0}^{t} \bPhi(t,\tau) \bQ(\tau) \bPhi^T(t,\tau) \dd \tau \quad \bP(0)=\bP_0,
 \end{align}
@@
 The mean dynamics is expressed by~\eqref{eq:OCPdyn} with~\eqref{eq:OCPdynIC}, and the covariance dynamics is expressed by~\eqref{eq:OCPdP} with~\eqref{eq:OCPdPIC}.
-In eq.~\eqref{eq:OCPconstraints} the \emph{backoff terms} $\be_i$ account for the disturbances and serve the purpose of obtaining deterministic safety margins directly on $\bmu(t)$.
+In eq.~\eqref{eq:OCPconstraints} the \emph{back-off terms} $\be_i$ account for the disturbances and serve the purpose of obtaining deterministic safety margins directly on $\bmu(t)$.
 $\calI$ is the set of indices defining the inequality constraints.
-The backoff terms stem for a linearization around the mean of the original \emph{chance constraint} on $\bx(t)$ expressed by $\Prob \{h_i(\bx) \leq 0\}\geq p$, where $p$ is the confidence level of constraint satisfaction.
-Explicitly, the backoff terms are given by
+The back-off terms stem from a linearization around the mean of the original \emph{chance constraint} on $\bx(t)$ expressed by $\Prob \{h_i(\bx) \leq 0\}\geq p$, where $p$ is the confidence level of constraint satisfaction.
+Explicitly, the back-off terms are given by
 $\be_i = \gamma \sig_i$.
@@
 \label{sec:TLC}
-An track limit constraint is required to ensure that the vehicle remains within the circuit boundaries.
+A track limit constraint is required to ensure that the vehicle remains within the circuit boundaries.
 The formulation adopted introduces an algebraic variable $e$, which satisfies the following constraint:
@@
 
-In order to take into account for the back-off terms, the maximum and minimum values are modified coherently, so that:
+In order to take into account the back-off terms, the maximum and minimum values are modified coherently, so that:
 \begin{equation}

### 2025-06-11 performance_comparison.tex
[MGa] some amendments to num res comments

--- 2025-06-11 performance_comparison.tex
@@
 Finally, the planned trajectories are compared against those followed in simulations with noise realizations, providing empirical validation of the proposed framework.
+
+The chosen sector was discretized into 140 spatial intervals ($\De s$ = 0.43 m).
+Both the open- and closed-loop disturbance-aware planning problems were transcribed on this grid in the CasADi-MATLAB environment~\cite{Andersson:MPC:2019}, and the resulting nonlinear programs were solved with IPOPT's interior-point algorithm~\cite{Wachter:MP:2006}.
 
@@
 
+% Solver times
+Solver times (build time is negligible) were measured on a laptop equipped with an Intel Core i9-13980HX with 32 GB RAM.
+The reference (nominal) plan, described in Sec.~\ref{sec:nominalFF}, contains 4 200 decision variables and is solved in 7.5 s.
+The open-loop robust plan (prediction horizon H = 4) described in Sec.~\ref{sec:open_loop_planning}, expands the model to 48 300 variables, which IPOPT solves in 128 s.
+The closed-loop robust plan described in Sec.~\ref{sec:closed_loop_planning} first finds the nominal plan (4 200 vars, 7.5 s), executes a policy-search step (Step 2, Sec.~\ref{sec:LQR}, 20.5 s), and then solves an additional robust problem with 14 700 variables (Step 3, Sec.~\ref{sec:CLrobustified}, 33 s), giving a total of 41 s.
+
 \subsection{Simulations with noise realization}

### 2025-06-12 performance_comparison.tex
[MM] Step 2 runtime 20.5->0.5 (typo)

@@
 Solver times (build time is negligible) were measured on a laptop equipped with an Intel Core i9-13980HX with 32 GB RAM.
-The reference (nominal) plan, described in Sec.~\ref{sec:nominalFF}, contains 4 200 decision variables and is solved in 7.5 s.
-The open-loop robust plan (prediction horizon H = 4) described in Sec.~\ref{sec:open_loop_planning}, expands the model to 48 300 variables, which IPOPT solves in 128 s.
-The closed-loop robust plan described in Sec.~\ref{sec:closed_loop_planning} first finds the nominal plan (4 200 vars, 7.5 s), executes a policy-search step (Step 2, Sec.~\ref{sec:LQR}, 20.5 s), and then solves an additional robust problem with 14 700 variables (Step 3, Sec.~\ref{sec:CLrobustified}, 33 s), giving a total of 41 s.
+The reference (nominal) plan, described in Sec.~\ref{sec:nominalFF}, contains 4\,200 decision variables and is solved in 7.5\,s.
+The open-loop robust plan (prediction horizon H = 4) described in Sec.~\ref{sec:open_loop_planning}, expands the model to 48\,300 variables, which IPOPT solves in 128\,s.
+The closed-loop robust plan described in Sec.~\ref{sec:closed_loop_planning} first finds the nominal plan (4\,200 vars, 7.5\,s), executes a policy-search step (Step 2, Sec.~\ref{sec:LQR}, 0.5\,s), and then solves an additional robust problem with 14\,700 variables (Step 3, Sec.~\ref{sec:CLrobustified}, 33\,s), giving a total of 41\,s.
 

### 2025-06-12 conclusions.tex
[MGa] curvilinear abscissa -> curvilinear parameter
[MM] add sentence in conclusion pros/cons

@@
 
-This work has proposed and compared two robust planning strategies for minimum-time trajectory optimization in motorsport scenarios under uncertainty.
+This work proposes and compares two robust planning strategies for minimum-time trajectory optimization in motorsport scenarios under uncertainty.
 The first approach, based on open-loop horizon-based covariance propagation, relies on a worst-case analysis of disturbance growth over a finite horizon to conservatively back off safety constraints.
@@
 The closed-loop method achieves comparable safety guarantees with reduced conservativeness and improved lap-time performance, thanks to its ability to explicitly model the stabilizing action of the driver.
+
+Comparing the computational costs of the two approaches, we observe that the closed-loop method demands significantly more programming effort but results in substantially lower total computation time.
+In contrast, the open-loop approach is simpler to implement but incurs higher solution times due to the larger number of optimization variables.
+
 Finally, random simulations with noise realization, performed adopting the closed-loop approach, prove the effectiveness of the robustified constraints.

### 2025-06-16 conclusions.tex
[mg] minor edits

@@
 
-By embedding uncertainty propagation and feedback action directly in the planning stage, the proposed framework yields trajectories that are simultaneously fast and probabilistically safe, advancing minimum-time optimisation toward practical deployment in high-performance motorsport and autonomous racing.
+By embedding uncertainty propagation and feedback action directly in the planning stage, the proposed framework yields trajectories that are simultaneously fast and probabilistically safe, advancing minimum-time optimization toward practical deployment in high-performance motorsport and autonomous racing.
 Future work will explore the generalization to more accurate vehicle models and experimental validation of the methodology.

### 2025-06-16 performance_comparison.tex
[mg] minor edits

@@
 
-We begin by presenting a selection of results obtained with the open-loop planning approach, which allows for a clearer illustration of the method and its underlying mechanisms.
-This approach is used as a reference to analyze the influence of key design parameters and of the different types of constraints on the resulting trajectories.
-A comparison is then carried out between the open-loop and closed-loop strategies.
+First we present selected results from the open-loop planning, which helps illustrate the method and its underlying mechanism. 
+%We begin by presenting a selection of results obtained with the open-loop planning approach, which allows for a clearer illustration of the method and its underlying mechanisms. 
+This serves as a reference to assess the impact of key design parameters and constraint types on planned trajectories.
+We then compare open- and closed-loop strategies, and finally validate the framework by comparing planned trajectories and simulations with noise realization.
+%A comparison is then carried out between the open-loop and closed-loop strategies.
 Finally, the planned trajectories are compared against those followed in simulations with noise realizations, providing empirical validation of the proposed framework.
@@
 \label{sec:ol_param_sensitivity}
-To gain deeper insight into how key parameters affect the planning outcome, we focus here on the open-loop formulation.
-%In this section, we explore the open-loop approach to gain deeper insight into the influence of key parameters on the results.
+To better understand the effect of key parameters, we focus on the open-loop formulation with robustified track limit constraint. This choice favors clearer visual interpretation.
+%To gain deeper insight into how key parameters affect the planning outcome, we focus here on the open-loop formulation.
 At this stage, only the track limit constraint is robustified.
-%For this analysis, we consider optimizations in which only the track limit constraint is robustified.
 This choice is motivated by the fact that its effects are more visually evident than those of the friction limit constraint, thereby facilitating a more immediate understanding of their impact.
-
 The first parameter considered is the confidence level of constraint satisfaction.
@@
 Specifically, $\ga^\textrm{TLC} = 0$ yields a satisfaction probability of 50\% and leads to the nominal solution with no back-off, while $\ga^\textrm{TLC} = 3$ corresponds to a confidence level of 99\%.
-For the purposes of this analysis, we examine the effect of varying $\ga^\textrm{TLC}$ in this range.
+%For the purposes of this analysis, we examine the effect of varying $\ga^\textrm{TLC}$ in this range.
 
@@
 
-Observing the left panel of Fig.~\ref{fig:ol_sensitivities}, it is evident that the trajectories corresponding to configurations with lower values of $\ga^\textrm{TLC}$ yield smaller lateral margins with respect to the track boundaries.
-In contrast, the trajectory associated with $\ga^\textrm{TLC} = 3$\,--- represented by the light green line \,--- remains significantly farther from the edges.
+As shown in the left panel of Fig.~\ref{fig:ol_sensitivities}, trajectories with lower values of $\ga^\textrm{TLC}$ (darker blue) yield smaller lateral margins with respect to the track boundaries.
+In contrast, the trajectory with $\ga^\textrm{TLC} = 3$ (light green) remains noticeably farther from the edges.
 This behavior is consistent with the increased conservativeness introduced by higher values of $\ga^\textrm{TLC}$, which amplify the back-off term $\be^\textrm{TLC}$ and thereby enforce a larger safety margin from the track limits.
-Chasing a trajectory associated with a larger safety margin leads to higher sector times; spanning $\ga^\textrm{TLC}$ in the range $\left[0,3\right]$ results in sector times that vary from 12.135\,s to 13.162\,s.
+Larger safety margins lead to higher sector times; spanning $\ga^\textrm{TLC}$ in the range $\left[0,3\right]$ results in sector times that vary from 12.135\,s to 13.162\,s.
 
@@
 The two configurations incorporating the robustified TLC (blue and green lines) exhibit, as expected, a significantly altered CoM trajectory.
-Ideed, the second and fourth panels clearly show that the TLC and TLC+FLC configurations tend to follow a path closer to the centerline.
+Indeed, the second and fourth panels clearly show that the TLC and TLC+FLC configurations tend to follow a path closer to the centerline.
 In particular, for most of the sector, the variation in lateral displacement $\De e$ and the lateral displacement $e$ itself exhibit opposite signs, indicating a corrective behavior of the steering angle $\de$ that pulls the vehicle toward the centerline.
@@
 \end{figure}
-
 \begin{figure}[b!]
 	\centering
-	\includegraphics[scale = 1]{Fig/cl_saturation.pdf}
+	\includegraphics[scale = 1.005]{Fig/cl_saturation.pdf}
 	\caption{Closed-Loop.
@@
 %\subsection{Open-loop vs closed-loop robust planning}
-Finally, we compare the influence of the back-off terms on the open-loop and closed-loop trajectories, in the robustified TLC scenario with the high $\ga^\textrm{TLC}=3.0$ value for better visualization.
-As before, the open-loop planning is carried out with a prediction horizon of $H=4$ steps.
-The back-off term $\be^{TLC}$ in Eq.~\eqref{eq:TLC} admits an interesting geometrical interpretation: it corresponds to the amount by which the track boundaries are (not uniformly) tightened to meet the probabilistic constraint, or, equivalently, to a virtual extension of the vehicle's width.
-
-According to the latter interpretation, Fig.~\ref{fig:olcl_traj_bands} shows two highlighted bands---red for the open-loop case and cyan for the closed-loop case---corresponding to the area swept by the back-off-extended-vehicle.
-Ultimately, the TLC enforces precisely that these bands remain entirely within the actual track limits.
-The comparison between the two regions illustrates how the additional conservativeness of the open-loop approach stems from larger back-off terms, i.e., a wider band that must stay within the track boundaries.
-The closed-loop method, on the other hand, benefits from the LQR-based feedback, which allows it to maintain smaller back-off terms and thus a narrower band, requiring a reduced safety margin from the track limits.
-\begin{figure}
+\begin{figure}[b!]
 	\centering
@@
 \end{figure}
+Finally, we compare the influence of the back-off terms on the open-loop and closed-loop trajectories, in the robustified TLC scenario with the high $\ga^\textrm{TLC}=3.0$ value for better visualization.
+As before, the open-loop planning is carried out with a prediction horizon of $H=4$ steps.
+The back-off term $\be^{TLC}$ in Eq.~\eqref{eq:TLC} admits an interesting geometrical interpretation: it corresponds to the amount by which the track boundaries are (not uniformly) tightened to meet the probabilistic constraint, or, equivalently, to a virtual extension of the vehicle's width.
+
+%According to the latter interpretation, Fig.~\ref{fig:olcl_traj_bands} shows two highlighted bands---red for the open-loop case and cyan for the closed-loop case---corresponding to the area swept by the back-off-extended-vehicle.
+Ultimately, the TLC enforces precisely that these bands remain entirely within the actual track limits.
+The comparison between the two regions illustrates how the additional conservativeness of the open-loop approach stems from larger back-off terms, i.e., a wider band that must stay within the track boundaries.
+The closed-loop method, on the other hand, benefits from the LQR-based feedback, which allows it to maintain smaller back-off terms and thus a narrower band, requiring a reduced safety margin from the track limits.
 
@@
 The random values for the initial state and the noise follow a Gaussian distribution with covariance matrices $\bP_0$ and $\bQ$, in both cases, as introduced in Eq.~\eqref{eq:dP}.
-This analysis cannot be performed for the open-loop approach, as it does not compute a stabilizing controller---resulting in trajectories that would significantly deviate from the track limits beyond the prediction horizon.
+This analysis is not applicable the open-loop approach, as it does not compute a stabilizing controller---resulting in trajectories that would significantly deviate from the track limits beyond the prediction horizon.
 %For visual clarity, the comparison is limited to trajectories with the robustified TLC, but analogous considerations apply to the robustified FLC as well.
+\begin{figure}[t!]
+	\centering
+	\includegraphics{Fig/olcl_traj_strings.pdf}
+	\caption{100 simulations (each panel) with random initial conditions and Gaussian noise realizations.
+In the left panel, the reference trajectory (black curve) is the non-robust one, tracked using an LQR controller designed around it as per Step 2 of Subsection~\ref{sec:LQR}.
+In the right panel, the reference (black curve) is a robust trajectory obtained using the closed-loop optimization method as per Step 3 of Subsection~\ref{sec:CLrobustified}, which is tracked using the corresponding optimized controller.}
+	\label{fig:traj_strings}
+\end{figure}
 
@@
 In contrast, the robust trajectory is tracked using the controller directly obtained from the optimization process.
-
-\begin{figure}
-	\centering
-	\includegraphics{Fig/olcl_traj_strings.pdf}
-	\caption{100 simulations (each panel) with random initial conditions and Gaussian noise realizations.
-In the left panel, the reference trajectory (black curve) is the non-robust one, tracked using an LQR controller designed around it as per Step 2 of Subsection~\ref{sec:LQR}.
-In the right panel, the reference (black curve) is a robust trajectory obtained using the closed-loop optimization method as per Step 3 of Subsection~\ref{sec:CLrobustified}, which is tracked using the corresponding optimized controller.}
-	\label{fig:traj_strings}
-\end{figure}

### 2025-06-16 count_words.py

@@ -1,0 +1,33 @@
+import PyPDF2
+import re
+from collections import Counter
+from nltk.corpus import stopwords
+
+import nltk
+nltk.download('stopwords')git 
+
+# === Settings ===
+PDF_PATH = 'MAIN.pdf'   # <-- metti qui il nome del tuo file PDF
+MIN_WORD_LENGTH = 4
+TOP_N = 20
+
+# === Stopwords inglesi ===
+stop_words = set(stopwords.words('english'))
+
+# === Estrai testo dal PDF ===
+with open(PDF_PATH, 'rb') as file:
+    reader = PyPDF2.PdfReader(file)
+    text = ""
+    for page in reader.pages:
+        text += page.extract_text() or ""
+
+# === Pulizia e conteggio parole ===
+words = re.findall(r'\b[a-zA-Z]{%d,}\b' % MIN_WORD_LENGTH, text.lower())
+filtered_words = [w for w in words if w not in stop_words]
+counts = Counter(filtered_words)
+most_common = counts.most_common(TOP_N)
+
+# === Output ===
+print(f"\nTop {TOP_N} words (min {MIN_WORD_LENGTH} letters, no common words):\n")
+for word, freq in most_common:
+    print(f"{freq:4d}  {word}")

### 2025-06-16 count_words.py
[mg] frequent words counter

@@ -4,7 +4,7 @@ from collections import Counter
 from nltk.corpus import stopwords
 
 import nltk
-nltk.download('stopwords')git 
+nltk.download('stopwords')
 
 # === Settings ===
 PDF_PATH = 'MAIN.pdf'   # <-- metti qui il nome del tuo file PDF

### 2025-06-16 performance_comparison.tex
[mg] fixes

@@
 \end{figure}
+
 Finally, we compare the influence of the back-off terms on the open-loop and closed-loop trajectories, in the robustified TLC scenario with the high $\ga^\textrm{TLC}=3.0$ value for better visualization.
@@
 
-%According to the latter interpretation, Fig.~\ref{fig:olcl_traj_bands} shows two highlighted bands---red for the open-loop case and cyan for the closed-loop case---corresponding to the area swept by the back-off-extended-vehicle.
+According to the latter interpretation, Fig.~\ref{fig:olcl_traj_bands} shows two highlighted bands---red for the open-loop case and cyan for the closed-loop case---corresponding to the area swept by the back-off-extended-vehicle.
 Ultimately, the TLC enforces precisely that these bands remain entirely within the actual track limits.
@@
 \subsection{Simulations with noise realization}
-To provide an empirical validation of the closed-loop robustified planning we simulated the evolution of the controlled vehicle with 100 random initial conditions and included noise realizations.
-The random values for the initial state and the noise follow a Gaussian distribution with covariance matrices $\bP_0$ and $\bQ$, in both cases, as introduced in Eq.~\eqref{eq:dP}.
-This analysis is not applicable the open-loop approach, as it does not compute a stabilizing controller---resulting in trajectories that would significantly deviate from the track limits beyond the prediction horizon.
-%For visual clarity, the comparison is limited to trajectories with the robustified TLC, but analogous considerations apply to the robustified FLC as well.
 \begin{figure}[t!]
@@
 \end{figure}
+To provide an empirical validation of the closed-loop robustified planning we simulated the evolution of the controlled vehicle with 100 random initial conditions and included noise realizations.
+The random values for the initial state and the noise follow a Gaussian distribution with covariance matrices $\bP_0$ and $\bQ$, in both cases, as introduced in Eq.~\eqref{eq:dP}.
+This analysis is not applicable the open-loop approach, as it does not compute a stabilizing controller---resulting in trajectories that would significantly deviate from the track limits beyond the prediction horizon.
+%For visual clarity, the comparison is limited to trajectories with the robustified TLC, but analogous considerations apply to the robustified FLC as well.
 

### 2025-06-16 intro.tex
[MGa] abbrv. Sec. where necessary
come commas and some amendments to .bib

@@
 The rest of the paper is organized as follows.
-In Section~\ref{sec:svdf} we introduce our stochastic dynamics framework, recalling the single-track vehicle model with nonlinear tires used for planning, and formulate both the continuous-time planning problem and its discrete-time counterpart via direct-collocation.
+In Sec.~\ref{sec:svdf} we introduce our stochastic dynamics framework, recalling the single-track vehicle model with nonlinear tires used for planning, and formulate both the continuous-time planning problem and its discrete-time counterpart via direct-collocation.
 Here, we also derive the probabilistic constraint back-off formulation.
-Section~\ref{sec:open_loop_planning} then outlines the open-loop horizon-based covariance-propagation planning approach, while Section~\ref{sec:closed_loop_planning} presents the closed-loop, covariance-aware planner.
-In Section~\ref{sec:performance_comparison} we compare both methods on a representative track scenario, reporting parameter-sensitivity studies, analyzing the influence of different sets of constraints, and performance trade-offs.
-Finally, Section~\ref{sec:conclusions} draws conclusions and offers directions for future work.
+Sec.~\ref{sec:open_loop_planning} then outlines the open-loop horizon-based covariance-propagation planning approach, while Sec.~\ref{sec:closed_loop_planning} presents the closed-loop, covariance-aware planner.
+In Sec.~\ref{sec:performance_comparison} we compare both methods on a representative track scenario, reporting parameter-sensitivity studies, analyzing the influence of different sets of constraints, and performance trade-offs.
+Finally, Sec.h~\ref{sec:conclusions} draws conclusions and offers directions for future work.

### 2025-06-16 open_loop_planning.tex
[MGa] abbrv. Sec. where necessary
come commas and some amendments to .bib

@@
 \section{Open-loop planning via H-steps ahead predictions}
-\label{sec:open_loop_planning}
-In this section, we address the problem of planning under uncertainty for a stochastic system.
+c In this section, we address the problem of planning under uncertainty for a stochastic system.
 At each grid point $k$ of the discretized trajectory the covariance matrix is propagated in a feed-forward fashion for a fixed prediction horizon of $H$ steps.
@@
 
-This approach requires modifications to the formulation~\eqref{eq:DOCP} introduced in Section~\ref{sec:discretization}.
+This approach requires modifications to the formulation~\eqref{eq:DOCP} introduced in Sec.~\ref{sec:discretization}.
 In particular, Eqs.~\eqref{eq:DOCPdP}, \eqref{eq:DOCPdPIC}, and \eqref{eq:DOCPconstraints} are replaced by Eqs.~\eqref{eq:DOCPdPopenloop}, \eqref{eq:DOCPdPinitopenloop}, and \eqref{eq:DOCPconstraintsopenloop}.

### 2025-06-16 performance_comparison.tex
[MGa] abbrv. Sec. where necessary
come commas and some amendments to .bib

@@
 
- Assuming the state vector is ordered as described in Section~\ref{sec:vehicle_model}, the baseline initial covariance matrix is set to $\bP_0=\diag(\bbsig^2)$, with
+ Assuming the state vector is ordered as described in Sec.~\ref{sec:vehicle_model}, the baseline initial covariance matrix is set to $\bP_0=\diag(\bbsig^2)$, with
 \begin{align}
@@
 %It is of interest to analyze how a robustified constraint---or a combination of the two described in Sections~\ref{sec:FLC} and~\ref{sec:TLC}---influences the driving style.
-We now investigate how the inclusion of different kinds of robustified constraints --- individually or in combination, as described in Sections~\ref{sec:FLC} and~\ref{sec:TLC} --- influences the driving style.
+We now investigate how the inclusion of different kinds of robustified constraints --- individually or in combination, as described in Sec.~\ref{sec:FLC} and~\ref{sec:TLC} --- influences the driving style.
 We compare the nominal feed-forward optimal trajectory, obtained without robustified constraints, with those resulting from a robustified track limit constraint (denoted as TLC), a robustified friction limit constraint (denoted as FLC), and a scenario where both constraints are robustified (denoted as TLC+FLC).
 
-First we show results obtained with the open-loop method described in Section~\ref{sec:open_loop_planning}, setting $H=4$, and $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$, corresponding to a 90\% probability of satisfying both constraints.
+First we show results obtained with the open-loop method described in Sec.~\ref{sec:open_loop_planning}, setting $H=4$, and $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$, corresponding to a 90\% probability of satisfying both constraints.
 %All the optimization are obtained with the open-loop method described in Section~\ref{sec:open_loop_planning}, setting $H=4$, and $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$, corresponding to a 90\% probability of satisfying both constraints.
@@
 
-An analogous analysis has been conducted using the closed-loop approach described in Section~\ref{sec:closed_loop_planning}, setting $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$.
+An analogous analysis has been conducted using the closed-loop approach described in Sec.~\ref{sec:closed_loop_planning}, setting $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$.
 The same key signals displayed for the previous case are shown in Fig.~\ref{fig:cl_telemetries}, while the axle saturation ratios $S_j$ are illustrated in Fig.~\ref{fig:cl_saturation}.

### 2025-06-16 stochastic_vehicle_dynamics.tex
[MGa] abbrv. Sec. where necessary
come commas and some amendments to .bib

@@
 The state vector $\bx$ has six components and includes the three kinematic quantities $u$, $v$, and $r$, along with the position $\left(x_G, y_G\right)$ and orientation $\psi$ of the barycentrical reference frame.
-As illustrated in Fig.~\ref{fig:vehicle_model} $u$ and $v$ represent the longitudinal and lateral velocity of the CoM w.r.t its reference frame, while $r$ represents vehicle's yaw rate.
+As illustrated in Fig.~\ref{fig:vehicle_model}, $u$ and $v$ represent the longitudinal and lateral velocity of the CoM w.r.t its reference frame, while $r$ represents vehicle's yaw rate.
 

### 2025-09-28 Rebuttal/Rebuttal.tex
[MGa] added rebuttal template

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Thin Formal Letter
% LaTeX Template
% Version 1.11 (8/12/12)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% WikiBooks (http://en.wikibooks.org/wiki/LaTeX/Letters)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass{letter}

\newcommand\setItemnumber[1]{\setcounter{enumi}{\numexpr#1-1\relax}}

\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathrsfs}
\usepackage{color,soul}
\usepackage{hyperref}

\setulcolor{red}
\sethlcolor{yellow}
\renewcommand\ul[1]{#1} % <<<< uncomment to remove 'underlined text'
\renewcommand\hl[1]{#1} % <<<< uncomment to remove 'highlighted text'

%\usepackage{caption}
%\usepackage{float}

\input{mymacros}

\DeclareMathOperator{\diag}{diag}

% Adjust margins for aesthetics
\addtolength{\voffset}{-0.5in}
\addtolength{\hoffset}{-0.3in}
\addtolength{\textheight}{2cm}

%\longindentation=0pt % Un-commenting this line will push the closing "Sincerely," to the left of the page

%----------------------------------------------------------------------------------------
%	YOUR NAME & ADDRESS SECTION
%----------------------------------------------------------------------------------------

\signature{Martino Gulisano, Matteo Masoni, Marco Gabiccini and Massimo Guiggiani} % Your name for the signature at the bottom

\address{Martino Gulisano$^{1}$, Matteo Masoni$^{1}$, Marco Gabiccini$^{1,2}$, Massimo Guiggiani \\[1em]
1. Dipartimento di Ingegneria Civile e Industriale, Universit\`{a} di Pisa, Pisa, Italy\\
2. Centro di Ricerca ``E. Piaggio'', Universit\`{a} di Pisa, Pisa, Italy}
 % Your address and phone number

%----------------------------------------------------------------------------------------

\begin{document}

%----------------------------------------------------------------------------------------
%	ADDRESSEE SECTION
%----------------------------------------------------------------------------------------

\begin{letter}{Professor Manfred Ploechl\\
Co-Editor\\
Vehicle System Dynamics} % Name/title of the addressee

%----------------------------------------------------------------------------------------
%	LETTER CONTENT SECTION
%----------------------------------------------------------------------------------------

\textbf{Paper - Disturbance-aware minimum-time planning strategies for motorsport vehicles with probabilistic safety certificates}

\opening{Dear Professor Ploechl,}

thank you very much for the kind response. We appreciate the opportunity to revise our paper entitled ``Disturbance-aware minimum-time planning strategies for motorsport vehicles with probabilistic safety certificates'' for the Vehicle System Dynamics Journal: please find the revised version accompanying this letter.

Observations and suggestions made by the reviewers have been carefully analysed and corrections to the paper have been made consistently. In the following, answers to the reviewers' comments are listed.

For the reviewers' convenience we included to the rebuttal (as a separate PDF file) also a version of the paper where the modifications and amendments are highlighted in \textcolor{red}{red colored text} and new text is highlighted in \textcolor{blue}{blue colored text}.
%The amended version of the paper (in black colored text) is provided separately through the website as well.
The amended version of the paper is provided separately through the website.

\textbf{Reviewer 1}

\begin{enumerate}

\item
\textit{Thanks for the interesting contribution.
The paper is well written and the proposed method is very interesting, especially considering that there are just few contributions applying robust optimal control concepts to vehicle dynamics problems.
Nothing to suggest: I believe the paper is ready to be published.}

\vspace{2mm}

\textbf{Answer:} We are grateful for the reviewer's positive comments and recommendation for publication. Although no changes were requested, we have made some changes as requested by the second reviewer and performed a final proofreading pass to ensure clarity and consistency throughout the manuscript.

\hrulefill
\end{enumerate}

\textbf{Reviewer 2}

The paper presents a novel disturbance-aware framework for minimum-time planning of road vehicles.
The manuscript is generally well written, and its contribution to the topic is valuable.
This reviewer suggests that the paper be accepted for publication after addressing the following minor concerns.

\begin{enumerate}

\item
\textit{Sec. 2.1, page 4: the framework is demostrated using a single-track model. Although relatively simple (e.g. 6 states), this reviewer acknowledges that the model is typically sufficient to capture the essential dynamics of cars and can therefore be used effectively to illustrate the approach. However, in minimum-lap-time of motorsport vehicles, more complex models (including e.g. out-of-plane dynamics etc.) are often employed. The concern of this reviewer is whether the proposed framework is scalable to more complex vehicle model. With a single-track model (6 states), the resulting stochastic OCP has 26 states. For instance, how does the OCP given in Eq. (7) scale when the vehicle model is made more complex?
}

\vspace{2mm}

\textbf{Answer:}

We thank the reviewer ...

\hrulefill

\item
\textit{Sec. 2.1, page 4: the vehicle model employs a magic formula to calculate lateral forces, while the longitudinal forces are treated as inputs. However, the force coupling is introduced using a standard friction ellipse. To this reviewer, such an approach appears mixed: typically, the magic formula is applied to both lateral and longitudinal directions (e.g. with longitudinal slip as an input). without relying on the friction ellipse as a constraint, since the tyre coupling is implicitly accounted for in the magic formula. Some comments on this point would be appreciated.}

\vspace{2mm}

\textbf{Answer:}

We thank the reviewer ...

\hrulefill

\item
\textit{Sec. 2.3, page 7: as stated by the authors, the positive-definiteness of the covariance matrix P is ensured through a sufficiently fine discretization. The authors should provide some guidelines on how to tune the mesh or collocation points to guarantee this property.
}

\vspace{2mm}

\textbf{Answer:}

We thank the reviewer ...

\hrulefill

\item
\textit{Sec. 2.4, page 7: as shown by the authors, considering the robustness of the friction limit is relatively straightforward when using the friction-ellipse constraint. How can this aspect be addressed when tyre coupling is implicitly captured by the magic formula (i.e. when no explicit friction-ellipse constraint is imposed)?}

\vspace{2mm}

\textbf{Answer:}

We sincerely thank the reviewer ...

\hrulefill

\item
\textit{Sec. 2.5, page 8: eq. (11) introduces an algebraic variable e, representing the lateral position of the vehicle with respect to the centreline. Why not use e as a vehicle state instead of x, y, thereby avoiding algebraic varibles?}

\vspace{2mm}

\textbf{Answer:}

We thank ...

\hrulefill

\item
\textit{Sec. 4.4, page 13: eq (20) introduces smoothed approaximation functions to ensure differentiatibility, which depend on a sharpness parameter $\chi$. Which are the typical values used in the authours' implementation?}

\vspace{2mm}

\textbf{Answer:}

We appreciate the reviewer's request \ul{for more details on these technical aspects.}

\hrulefill

\item
\textit{Sec. 5.1, page 16: in the equation $P_0=\diag(\sigma^2)$ (just before eq. (21)) the operation $\sigma^2$ should be interpreted pointwise. This reviewer suggests that this clarification should be explicitly stated in the manuscript.}

\vspace{2mm}

\textbf{Answer:}

As correctly pointed out, ...

\hrulefill

\item
\textit{The paper is interesting and generally well written. In the related work section maybe a brief comment on https://doi.org/10.1177/0954407013516102 and its relation with the proposed approaches could be interesting, being one of the earliest example of investigation of robustness of lap time solution nominal solutions.}

\vspace{2mm}

\textbf{Answer:}

 As correctly pointed out, ...

\hrulefill

\end{enumerate}

We hope we have appropriately clarified all the issues raised and modified the paper accordingly. Thank you for your time and consideration.

%Comunque il revisore 2 è un demente spocchioso ed il 3 braccia rubate all'agricoltura.

\vspace{2\parskip} % Extra whitespace for aesthetics
\closing{Sincerely,}
\vspace{2\parskip} % Extra whitespace for aesthetics

%\ps{P.S. You can find additional information attached to this letter.} % Postscript text, comment this line to remove it

%\encl{Copyright permission form} % Enclosures with the letter, comment this line to remove it

%----------------------------------------------------------------------------------------
\end{letter}
\end{document}

### 2025-09-30 Rebuttal/Rebuttal.tex
[mg] Rev2 - Ans 1,3,6,7 - Ans 2 incomplete

@@
 \textit{Sec.
-2.1, page 4: the framework is demostrated using a single-track model.
+2.1, page 4: the framework is demonstrated using a single-track model.
 Although relatively simple (e.g.
@@
 
-We thank the reviewer ...
+We thank the reviewer for raising this interesting point of discussion.
+The scalability of the problem when dealing with more complex vehicle models is a challenging aspect of the procedure: in fact, considering a $n$-states vehicle model, the resulting stochastic OCP system has $n+\frac{n(n+1)}{2}$ states.
+
+The simplified model used in the current implementation fits well to the serial approach of the OCP, since computational times and total number of decision variables are still sufficiently small.
+
+In further works we are actually planning to adopt more complex models and to parallelize the problem solution on multiple cores.
+This approach has been explored by our research group in previous works [...], where the nominal minumim lap-time problem was solved for a 14-DoF's multibody model.
+
+Finally, since the reviewer has pointed out this specific feature, we would like to remark that, although the out-of-plane dynamics is neglected in our model, we took into account longitudinal load transfers.
 
@@
 \textbf{Answer:}
+(INCOMPLETE)
+We thank the reviewer for the valuable suggestion.
 
-We thank the reviewer ...
+To introduce the framework, we chose to adopt a simplified model with the longitudinal force $F_x$ as one of the inputs and to constraint the lateral forces 
+Typically, single track models use longitudinal speed $u$ and steering wheel angle $\delta$ as inputs [Guigg.]. Since we included the longitudinal dynamics.
 
@@
 
-We thank the reviewer ...
+We thank the reviewer for giving the opportunity to clarify this point.
+Drawing on the experience from our previous works, we found that a discretization grid for the track of one point per meter is sufficient to obtain accurate results while keeping the computational time within reasonable limits.
+We observed that the same grid size also preserved the positive-definiteness of the covariance matrix $\bP$.
 
@@
 
-We appreciate the reviewer's request \ul{for more details on these technical aspects.}
+We appreciate the reviewer's request \ul{for more details on these technical aspects.} In this context, we chose a value of $\chi=10$ as the sharpness parameter, and we now explicitly state it in the paper.
+This value proved to be a good trade-off between a steep transition and the smoothness required by the solver to achieve convergence.
+In general, $\chi$ must be tuned according to the range of the corresponding variable.
+In this case, the longitudinal force $F_x$ is on the order of $10^4$\,N.
 
@@
 
-As correctly pointed out, ...
+As correctly pointed out, our notation in this equation was not sufficiently clear.
+We have now chosen to use $P_0=\diag(\sigma)^2$, since matrix powers are defined for square matrices.
+The result remains the same even if the operator is interpreted elementwise, because the matrix is diagonal.
 
 \hrulefill
+
+\end{enumerate}
+
+\textbf{Reviewer 3}
+
+\begin{enumerate}
 

### 2025-09-30 Rebuttal/Rebuttal.tex
[mg] Rev 2 - Ans 2

@@
 \textbf{Answer:}
-(INCOMPLETE)
 We thank the reviewer for the valuable suggestion. 
+To introduce the framework, we adopted a simplified vehicle model with the longitudinal force $F_x$ and the steering angle $\delta$ as inputs, based on the natural correspondence to the control actions a driver can apply.
+Indeed, our choice follows this simple control-oriented structure.
 
-To introduce the framework, we chose to adopt a simplified model with the longitudinal force $F_x$ as one of the inputs and to constraint the lateral forces 
-Typically, single track models use longitudinal speed $u$ and steering wheel angle $\delta$ as inputs [Guigg.]. Since we included the longitudinal dynamics.
+We acknowledge that the Magic Formula can also be applied in both lateral and longitudinal directions with implicit tyre coupling; incorporating this fully coupled formulation is indeed a valid approach and will be considered in future extensions of the framework.
+For the present study, the friction ellipse provides a convenient and well-understood way to capture tyre coupling while keeping the control structure aligned with driver inputs.
 

### 2025-10-01 Rebuttal/Rebuttal.tex
[MGa] Reviewer 3 on rebuttal only

@@
 
- As correctly pointed out, ...
+We thank the reviewer for drawing our attention to this work, which, regrettably, had escaped our initial literature search despite its clear relevance.
+In the revised manuscript, we now cite it in the Introduction (and Related Work) and include a brief summary of its key contributions and their relation to our approach, highlighting its role as an early study on robustness in lap-time optimal solutions.
 

### 2025-10-01 intro.tex
[MGa] mods to .bib
+ submitted manuscript
[MGa] added comment to Cole in text

@@
 Despite these advances, most planning frameworks still fail to incorporate robustness in the planning phase itself.
-Piccinini et al.~\cite{Piccinini:HowOptimalMinimumtime:2024} provide a direct comparison between an offline minimum-lap-time optimal control problem (MLT-OCP) and an online Artificial Racing Driver (ARD) that controls the very same high-fidelity vehicle model.
+A notable exception is~\cite{Timings:RobustLaptimeSimulation:2014a}, which first plans a nominal trajectory via MPC and then employs a feedback MPC to counteract disturbance realizations inferred from road-surface roughness via a ride model.
+An inspiring aspect is the explicit trade-off between drivability and the control effort required to stay close to the nominal path.
+While there are many similarities with our setting, the key differences are methodological: their approach uses a two-level, MPC-based scheme with discrete disturbance realizations in the robust layer, whereas our lap-time simulation/planning is posed as a single large-scale NLP in which robustness to (Gaussian) disturbances is embedded directly within one optimization problem.
+
+Other contributions like Piccinini et al.~\cite{Piccinini:HowOptimalMinimumtime:2024} provide a direct comparison between an offline minimum-lap-time optimal control problem (MLT-OCP) and an online Artificial Racing Driver (ARD) that controls the very same high-fidelity vehicle model.
 They show that, by leveraging a physics-driven structure and a novel g-g-v performance constraint, ARD can achieve lap times within a few milliseconds of the offline benchmark and generalize to unseen circuits even under unmodeled mass variations.
@@
 In Sec.~\ref{sec:performance_comparison} we compare both methods on a representative track scenario, reporting parameter-sensitivity studies, analyzing the influence of different sets of constraints, and performance trade-offs.
-Finally, Sec.h~\ref{sec:conclusions} draws conclusions and offers directions for future work.
+Finally, Sec.~\ref{sec:conclusions} draws conclusions and offers directions for future work.

### 2025-10-01 intro.tex
[MGa] mods to comments to Cole

@@
 A notable exception is~\cite{Timings:RobustLaptimeSimulation:2014a}, which first plans a nominal trajectory via MPC and then employs a feedback MPC to counteract disturbance realizations inferred from road-surface roughness via a ride model.
-An inspiring aspect is the explicit trade-off between drivability and the control effort required to stay close to the nominal path.
+An inspiring aspect is the trade-off between vehicle/driver performance and the control effort demanded of the driver to stay close to the nominal path despite disturbances, which essentially defines \emph{drivability}.
 While there are many similarities with our setting, the key differences are methodological: their approach uses a two-level, MPC-based scheme with discrete disturbance realizations in the robust layer, whereas our lap-time simulation/planning is posed as a single large-scale NLP in which robustness to (Gaussian) disturbances is embedded directly within one optimization problem.

### 2025-10-01 Rebuttal/Rebuttal.tex
[MGa] first complete pass

@@
 In further works we are actually planning to adopt more complex models and to parallelize the problem solution on multiple cores.
-This approach has been explored by our research group in previous works [...], where the nominal minumim lap-time problem was solved for a 14-DoF's multibody model.
+This approach has been explored by our research group in these previous works:
 
-Finally, since the reviewer has pointed out this specific feature, we would like to remark that, although the out-of-plane dynamics is neglected in our model, we took into account longitudinal load transfers.
+\begin{itemize}
+  \item https://doi.org/10.1007/s11044-023-09932-6
+  \item https://doi.org/10.1080/00423114.2024.2407175,
+\end{itemize}
+
+where the nominal minimum lap-time problem was solved for a 14-DoF multibody model.
+
+Finally, since the reviewer has pointed out this specific feature, we would like to remark that, although the out-of-plane dynamics is neglected in our model, we took at least into account the (steady-state) longitudinal load transfers.
 
@@
 
-We sincerely thank the reviewer ...
+We sincerely thank the reviewer for the stimulating question.
+When tyre coupling is modeled implicitly via the Magic Formula (MF) in combined slip, the friction boundary is not an explicit ellipse but a nonlinear frontier emerging from
+\begin{equation}
+F_x = G_{xa}(\alpha,\kappa,\ldots)\,F_{x0}(\kappa),
+\;
+F_y = G_{yk}(\alpha,\kappa,\ldots)\,F_{y0}(\alpha) + SV_{yk}(\alpha,\kappa,\ldots),
+\label{eq:MF-combined}
+\end{equation}
+where $F_{x0},F_{y0}$ are the pure‐slip curves (with peak factors $D_x,D_y$ depending on $F_z,\gamma,\ldots$), $G_{xa},G_{yk}$ are the normalized “hill–shaped” coupling weights (equal to $1$ in pure slip), and $SV_{yk}$ is the shift induced by ply–steer/conicity effects/rolling resistance shift, see e.g.
+[Pacejka 2012, 3rd edition, p.166].
+To retain the same \emph{robustness to friction variations} that is straightforward with an explicit friction ellipse, we could introduce a \emph{control-level} constraint built from the instantaneous MF capacities (a sort of \emph{breathing ellipse}).
+
+\vspace{4pt} 
+Given force references $(F_x^{\mathrm{ref}},F_y^{\mathrm{ref}})$ produced at the generic iteration of the NLP solver, we could define the instantaneous directional limits from the MF model at the current $(\alpha,\kappa)$:
+\begin{equation}
+F_{x,\lim} \;=\; D_x\,G_{xa}(\alpha,\kappa,\ldots),
+\qquad
+F_{y,\lim} \;=\; D_y\,G_{yk}(\alpha,\kappa,\ldots).
+\label{eq:limits}
+\end{equation}
+Of course, since $SV_{yk}$ is an offset, we could work with the lateral \emph{effective} command $F_{y,\mathrm{eff}}^{\mathrm{ref}} := F_y^{\mathrm{ref}} - SV_{yk}$.
+Then the new \emph{axle saturation ratio} $S$ would read
+\begin{equation}
+S=
+\left(\frac{F_x^{\mathrm{ref}}}{F_{x,\lim}}\right)^2
++
+\left(\frac{F_{y,\mathrm{eff}}^{\mathrm{ref}}}{F_{y,\lim}}\right)^2,
+\label{eq:util}
+\end{equation}
+and robustness to friction uncertainty would be enforced by the same considerations that have been proposed in Sec. 2.4 of the paper.
+It is worth observing that this \emph{breathing ellipse} would not add additional physics: the MF already encodes tyre coupling.
+It would just provide a control–level robustness margin in analogy to the classical friction‐ellipse constraint, but \emph{coherent} with the instantaneous MF capacities $D\,G$ (and with $SV_{yk}$ excluded from the lateral capacity).
 
@@
 
-We thank ...
+We thank the reviewer for the insightful question.
+The main motivation for our current formulation is that the vehicle dynamics are not embedded in the track frame.
+In earlier work we did adopt a track-embedded type formulation (see A reduced-order Lie-group–based race car model for systematic trajectory optimization on 3D tracks, Meccanica 58:1869–1883, 2023, doi:10.1007/s11012-023-01708-8.) where $e$ ($q_2$ in the mentioned paper) would emerge naturally as a parametrization coordinate.
+
+Within a direct-collocation + IPOPT pipeline, keeping $(x,y)$ as states and also adding $e$ as algebraic variable, yields Jacobians that are sparser and typically less involved algebraically: the track geometry (centerline curvature/banking and their derivatives) enters only the algebraic relation for $e$, rather than propagating through the full state dynamics at every collocation point.
+By contrast, promoting $e$ to a state in a track-embedded (Frenet-Serret type) model couples the dynamics to space-varying geometry terms, which tends to densify the KKT system and degrade conditioning.
+A further practical advantage is that many constraints and features (e.g., global track limits, obstacles, multi-vehicle interactions) are naturally expressed in inertial coordinates and remain simpler when $(x,y)$ are states.
 
@@
 
-We appreciate the reviewer's request \ul{for more details on these technical aspects.} In this context, we chose a value of $\chi=10$ as the sharpness parameter, and we now explicitly state it in the paper.
-This value proved to be a good trade-off between a steep transition and the smoothness required by the solver to achieve convergence.
+We appreciate the reviewer's request \ul{for more details on these technical aspects.} In this context, we chose a value of $\chi=10$ as the sharpness parameter, and we have explicitly stated it in the paper.
+This value proved to be a good trade-off between a steep transition and the smoothness required by the solver to enhance convergence.
 In general, $\chi$ must be tuned according to the range of the corresponding variable.

### 2025-10-01 performance_comparison.tex
[mg] Paper edits Rev2-Ans1,3,6,7 + bold P pag 8 + old comments cleaning

@@
 First we present selected results from the open-loop planning, which helps illustrate the method and its underlying mechanism.
-%We begin by presenting a selection of results obtained with the open-loop planning approach, which allows for a clearer illustration of the method and its underlying mechanisms.
 This serves as a reference to assess the impact of key design parameters and constraint types on planned trajectories.
 We then compare open- and closed-loop strategies, and finally validate the framework by comparing planned trajectories and simulations with noise realization.
-%A comparison is then carried out between the open-loop and closed-loop strategies.
-Finally, the planned trajectories are compared against those followed in simulations with noise realizations, providing empirical validation of the proposed framework.
 
@@
 Both the open- and closed-loop disturbance-aware planning problems were transcribed on this grid in the CasADi-MATLAB environment~\cite{Andersson:MPC:2019}, and the resulting nonlinear programs were solved with IPOPT's interior-point algorithm~\cite{Wachter:MP:2006}.
-
-% Contour plot dei LAP time in funzione dei due gamma, lo mettiamo o no? Eventialmente nel comparison open/closed
 
@@
 To better understand the effect of key parameters, we focus on the open-loop formulation with robustified track limit constraint. This choice favors clearer visual interpretation.
-%To gain deeper insight into how key parameters affect the planning outcome, we focus here on the open-loop formulation.
-At this stage, only the track limit constraint is robustified.
-This choice is motivated by the fact that its effects are more visually evident than those of the friction limit constraint, thereby facilitating a more immediate understanding of their impact.
 The first parameter considered is the confidence level of constraint satisfaction.
@@
 Specifically, $\ga^\textrm{TLC} = 0$ yields a satisfaction probability of 50\% and leads to the nominal solution with no back-off, while $\ga^\textrm{TLC} = 3$ corresponds to a confidence level of 99\%.
-%For the purposes of this analysis, we examine the effect of varying $\ga^\textrm{TLC}$ in this range.
-
-%Setting $\ga^\textrm{TLC} = 0$ leads to the nominal solution.
-In this case, the back-off term $\be^\textrm{TLC}$ becomes zero, meaning that the constraint is not robustified and thus coincides with the original, non-robust formulation.
 
@@
 }
-%\caption{Effect of $\ga^\textrm{TLC}$ (left panel) and $\bP_0$ (right panel) on optimal trajectory.
-The parameter $\ga^\textrm{TLC}$ spans in the range $\left[0,3\right]$ corresponding to a probability to meet the track limit constraint ranging from 50\% to 99\%.
-In the right panel, in addition to the baseline value $\bar{\bP}_0$ (black line), two alternative values are considered: $\bP_0 = 4\bar{\bP}_0$ (blue line) and $\bP_0 = \frac{1}{4}\bar{\bP}_0$ (red line).
-These correspond, respectively, to doubling and halving the initial uncertainty associated with the state vector.}
+
 	\label{fig:ol_sensitivities}
@@
 
- Assuming the state vector is ordered as described in Sec.~\ref{sec:vehicle_model}, the baseline initial covariance matrix is set to $\bP_0=\diag(\bbsig^2)$, with
+ Assuming the state vector is ordered as described in Sec.~\ref{sec:vehicle_model}, the baseline initial covariance matrix is set to $\bP_0=\diag(\bbsig)^2$, with
 \begin{align}
@@
 
-	%\caption{Evolution of the longitudinal speed $\De u$, wheel steering angle $\De \de$, longitudinal force $\De F_x$, and lateral deviation $\De e$ are reported for three robustification scenarios: TLC (Track Limit Constraints), FLC (Friction Limit Constraint), and TLC+FLC (both constraints enforced concurrently).
-Each profile is shown relative to the baseline (nominal) trajectory to highlight the effect of each constraint-handling strategy.
-The dashed lines in the last three panels indicate the sign of the reference signal, allowing one to determine whether a positive variation with respect to the baseline corresponds to an increase or a decrease in the absolute value.
-This distinction is particularly important for quantities that depend on the direction of the turn, such as the wheel steering angle $\de$ and the lateral deviation $e$, while in the case of $F_x$, it indicates whether the vehicle is accelerating or braking.}
 	\label{fig:ol_telemetries}
@@
 Nominal configuration in dashed black, TLC in green, FLC in red, TLC+FLC in blue.
-	%The dashed black line corresponds to the reference (nominal) configuration, while the green, red, and blue lines represent the robustified configurations TLC, FLC, and TLC+FLC, respectively.
 	}
@@
 
-%It is of interest to analyze how a robustified constraint---or a combination of the two described in Sections~\ref{sec:FLC} and~\ref{sec:TLC}---influences the driving style.
 We now investigate how the inclusion of different kinds of robustified constraints --- individually or in combination, as described in Sec.~\ref{sec:FLC} and~\ref{sec:TLC} --- influences the driving style.
@@
 First we show results obtained with the open-loop method described in Sec.~\ref{sec:open_loop_planning}, setting $H=4$, and $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$, corresponding to a 90\% probability of satisfying both constraints.
-%All the optimization are obtained with the open-loop method described in Section~\ref{sec:open_loop_planning}, setting $H=4$, and $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$, corresponding to a 90\% probability of satisfying both constraints.
+
 %---Delta key signals
@@
 This behavior is consistent with the variation of the total longitudinal force $\Delta F_x$ shown in the third panel.
-%\textcolor{red}{In particular, the TLC and TLC+FLC configurations exhibit a higher longitudinal force while approaching the sharp turn, in the curvilinear abscissa interval $\al\in\left[0.707, 0.712\right]$, resulting in a higher longitudinal speed.
-%Subsequently, these two configurations apply a lower longitudinal force during braking compared to the nominal case, indicating more intense braking.
-This behavior explains the sharp reduction in longitudinal speed observed thereafter.}
 
@@
 
-%As already observed, the configurations with the robustified track limit constraint (green and blue lines) follow significantly different paths compared to the nominal solution.
-This deviation results in distinct ground force demands, clearly visible in the saturation index.
-%From this plot, it can be observed that the configurations with the robustified track limit constraint (green and blue lines) follow significantly different trajectories compared to the nominal solution, resulting in distinct ground force demands.
-This behavior is noticeable in turn 1 and even more clearly in turn 2.
-Specifically, these two configurations begin to experience high axle loads approximately 12\,m before the nominal trajectory and sustain them for an additional 4.6\,m (TLC, green line) and 9.2\,m (TLC+FLC, blue line) beyond the nominal reference.
-
 \begin{figure}[t]
@@
 
-	%\caption{Evolution of the longitudinal speed $\De u$, wheel steering angle $\De \de$, longitudinal force $\De F_x$, and lateral deviation $\De e$ are reported for three robustification scenarios: TLC (Track Limit Constraints), FLC (Friction Limit Constraint), and TLC+FLC (both constraints enforced concurrently).
-Each profile is shown relative to the baseline (nominal) trajectory to highlight the effect of each constraint-handling strategy.
-The dashed lines in the last three panels indicate the sign of the reference signal, allowing one to determine whether a positive variation with respect to the baseline corresponds to an increase or a decrease in the absolute value.
-This distinction is particularly important for quantities that depend on the direction of the turn, such as the wheel steering angle $\de$ and the lateral deviation $e$, while in the case of $F_x$, it indicates whether the vehicle is accelerating or braking.}
 	\label{fig:cl_telemetries}
@@
 Nominal configuration in dashed black, TLC in green, FLC in red, TLC+FLC in blue.
-		%The dashed black line corresponds to the reference (nominal) configuration, while the green, red, and blue lines represent the robustified configurations TLC, FLC, and TLC+FLC, respectively.
 	}
@@
 
-%\subsection{Open-loop vs closed-loop robust planning}
 \begin{figure}[b!]
@@
 This analysis is not applicable the open-loop approach, as it does not compute a stabilizing controller---resulting in trajectories that would significantly deviate from the track limits beyond the prediction horizon.
-%For visual clarity, the comparison is limited to trajectories with the robustified TLC, but analogous considerations apply to the robustified FLC as well.
 
@@
 The comparison between the panels shows that the robustified controller is significantly more effective at satisfying the TLC, resulting in safer and more reliable trajectory tracking.
-
-%---------------------------------
-%ORIGINAL VERSION
-
-%This analysis aims to provide empirical validation of the robust trajectories obtained using the methods proposed in this work.
-For conciseness, the comparison is limited to a nominal trajectory and a trajectory computed using the closed-loop method with only the track limit constraint robustified.
-%An LQR controller, designed according to the procedure outlined in Section~\ref{sec:LQR}, is implemented to stabilize the nominal trajectory.
-In contrast, the robust trajectory is tracked using the controller directly obtained from the optimization process.

### 2025-10-01 stochastic_vehicle_dynamics.tex
[mg] Paper edits Rev2-Ans1,3,6,7 + bold P pag 8 + old comments cleaning

@@
 	\input{Fig/Tikz_Sources/tikz_vehicle_model.tex}
-	%\caption{Single Track model.
-The ground force acting on each axle is decomposed into longitudinal and lateral components, indicated by red arrows.
-The longitudinal components are part of the input vector $\bu$, encompassing both acceleration and braking forces, while the lateral components are computed using Pacejka's Magic Formula based on the vertical load and slip angle at each axle.
-The figure also depicts the fixed reference frame and the body reference frame, with the origin located at the vehicle's center of mass (CoM).
-The state variables $u$ and $v$ represent the longitudinal and lateral velocities of the CoM w.r.t.
-the body frame, while $r$ denotes the yaw rate.
-The angle $\psi$ is used to describe the orientation of the body frame w.r.t.
-the fixed frame.}
 	\caption{Single Track model.}
@@
 	\phantom{\text{s.t.} \quad} \bP(0)  &= \bP_0 \succeq 0 \label{eq:OCPdPIC} \\ %& \phantom{\text{s.t.} \qquad} 0       \geq h_i(\bmu(t), \bu(t))
-	%+ \overbrace{\gamma \bigg[\underbrace{\na^T_{\bx} h_i \bP(t) \na_{\bx} h_i}_{\text{variance of $h_i(\bx)$}}\bigg]^{\frac{1}{2}}}^{\text{safety margin}},
 	\phantom{\text{s.t.} \qquad} 0&       \geq h_i(\bmu(t), \bu(t))
@@
 of random variable $h_i(\bmu)+\na^T_{\bx} h_i(\bmu) (\bx - \bmu)$, and follows immediately from the propagation rule of covariance.
+
+It is worth noting that scaling to more complex vehicle models increases the number of decision variables, which grows roughly quadratically with the vehicle state dimension due to the covariance matrix $\bP$.
+Parallelization may therefore be needed to keep computational times tractable.
 
@@
 In certain cases~\cite{Gillis:PracticalMethodsApproximate:2015}, positive-definiteness-preserving Lyapunov discretization schemes are used in~\eqref{eq:DOCPdP}.
-However, for sufficiently fine discretization, we found that integrating only the lower triangular part of $\bP(t)$ in~\eqref{eq:OCPdPIC}  (and reconstructing its strictly upper part accordingly), ensures that $\bP_k$ remain symmetric and positive definite when the initial $\bP_0$ is symmetric and positive definite.
+However, for sufficiently fine discretization (one point per meter of track), we found that integrating only the lower triangular part of $\bP(t)$ in~\eqref{eq:OCPdPIC}  (and reconstructing its strictly upper part accordingly), ensures that $\bP_k$ remain symmetric and positive definite when the initial $\bP_0$ is symmetric and positive definite.
 In all our tests direct collocation with cubic polynomial state representations and Gauss-Legendre collocation points provided accurate results.
@@
 
-To better clarify the effect of the back-off term on the constraint let us consider a combination of $\bx$, $\bu$, $P$ and $\ga^\textrm{FLC}_j$ so that $\be^\textrm{FLC}_j=0.2$.
+To better clarify the effect of the back-off term on the constraint let us consider a combination of $\bx$, $\bu$, $\bP$ and $\ga^\textrm{FLC}_j$ so that $\be^\textrm{FLC}_j=0.2$.
 This implies that the semi-axes of the ellipse are reduced by a factor $\sqrt{1-\be^\textrm{FLC}_j} = \sqrt{0.8}$, as shown in the left panel of Fig.~\ref{fig:robust_constraints}, where the ellipse in dashed line delimits the available region with the back-off applied.
 This allows the optimizer to determine the most appropriate trade-off between the longitudinal force $X_j$ and the lateral force $Y_j$ --- often by reducing both components to some extent --- based on the specific requirements of the manoeuvre.
-
-%The gradient of the constraint w.r.t.
-the state vector $\bx$, denoted by $\na_{\bx}h^\textrm{A}_j(\bmu)$, is used to evaluate the standard deviation $\sigma^\textrm{A}_j$ associated with the constraint.
-This gradient can be computed using the chain rule, as follows:
-%\begin{equation}
-%	\na_{\bx}h^\textrm{A}_j(\bmu,\bu) =
-%	\frac{\pd h^\textrm{A}_j}{\pd X_j}\at_{\bmu,\bu}\na_{\bx}X_j(\bmu,\bu) +
-%	\frac{\pd h^\textrm{A}_j}{\pd Y_j}\at_{\bmu,\bu}\na_{\bx}Y_j(\bmu,\bu) +
-%	\frac{\pd h^\textrm{A}_j}{\pd Z_j}\at_{\bmu,\bu}\na_{\bx}Z_j(\bmu,\bu)
-%\end{equation}
-%where the first element is zero because $X_j$ is a function of the input vector $\bu$ only.
 
@@
 	}
-		%Graphical representation of the two constraints analyzed in the section.
-The left panel shows the friction limit constraint in the plane $X_j-Y_j$.
-The ellipse in solid line, whose semi-axes are given by $\mu_{x,j}Z_j$ and $\mu_{y,j}Z_j$, represents the border of the available region for the base constraint, without back-off.
-The ellipse in dashed line delimits the available region when a back-off $\be^\textrm{FLC}_j=0.2$ is present.
-This ellipse has its semi-axes reduced by a factor $\sqrt{1-\be^\textrm{FLC}_j}$ w.r.t.
-the one in solid line.}
 	\label{fig:robust_constraints}
@@
 \end{equation}
-%\begin{equation}
-%	\bp-\boldsymbol{c}_k - e_k\bn_k=\boldsymbol{0} \label{eq:onplane_constraint}
-%\end{equation}
 where $\boldsymbol{c}$ indicates the position of the centerline, and $\bn$ is the unit normal vector to the centerline.
@@
 where $\be^\textrm{TLC}$ is computed by considering that the gradient of the constraint w.r.t. the state vector $\na_{\bx}h^\textrm{TLC}$ depends solely on $\bn$.
-%Adding a back-off to this constraint produces the same effect as an increase of the front and rear tracks of the vehicle does.
